### 2013-10-30 analysis.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

% Attributions for Poverty in Post-Socialist Countries
% Markus Kainu & Mikko Niemelä
% Jan 2013

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>

******

<center>
<table>
<tr>
<td>[Etusivu](index.html)</td>
<td>[Analysis](analysis.html)</td>
<td>[Article](article2013.html)</td>
</tr>
</table>
</center>

*******

# Analysis

### 2013-10-30 article2013.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

% Attributions for Poverty in Post-Socialist Countries[^*]
% Markus Kainu[^**];Mikko Niemelä[^***]
% March 2013

[^*]:*Work in progress. Please, do not cite without permission of authors*

[^**]:PhD Student. Aleksanteri institute, University of Helsinki & Department of Social Research, University of Turku <a href="mailto:markuskainu@gmail.com">markuskainu@gmail.com</a>

[^***]:Senior Researcher. Social Insurance Institution of Finland, KELA a<a href="mailto:mikko.niemela@kela.fi">mikko.niemela@kela.fi</a>

<link href="http://markuskainu.fi/material/css/article.css" rel="stylesheet" type="text/css" title="compact"></link>
<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(knitr)
knit("article2013.Rmd")
system("~/.cabal/bin/pandoc article2013.md --number-sections -o article2013.pdf")
system("~/.cabal/bin/pandoc -s article2013.md --toc --number-sections -o article2013.html")
system("~/.cabal/bin/pandoc -f markdown -t docx article2013.md -o article2013.docx --standalone --toc --number-sections")

# jos multilevelit mukana
#system("~/.cabal/bin/pandoc article2013.md --number-sections --template=tex_templates/dcolumn.latex -o article2013.tex")
#system("perl -pi -e 's/footnotesize/tiny/g' article2013.tex")
#system("pdflatex article2013.tex")

lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions
```

**Abstract**

*Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries. The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions. As country-level determinants, paper analyses the changes in country's economic performance during the period of global financial crises, income inequality and the state of democracy. As individual-level determinants, the paper tests to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.*

\newpage

# Introduction

As Joseph Schumpeter (1942: 12) has once argued, "attitudes are coins that do not readily melt". This means that normative expectations, justice beliefs or welfare perceptions are very hard to change. In this way, welfare attitudes can be associated with policy changes by highlighting public sentiments, which are normative background assumptions or ideas that constrain decision-making and institutional change by limiting the range of programme that decision-making elites are likely to perceive as acceptable and legitimate to both their constituents and themselves (Campbell, 2004: 96-100). On the other hand, "the worlds on welfare" literature on welfare attitudes has in turn emphasised that type of welfare policy affects public attitudes (Andress and Heien, 2001; Arts and Gelissen, 2001; Jaeger, 2006; Svallfors, 1997). In regard to public perceptions of the causes of poverty, in the selective welfare-policy dominated liberal regimes explanations of poverty are more often based on individual morality that they are in the Continental and Nordic welfare states – where external and structural explanations prevail (Albrekt Larsen, 2006: 47-56; Alesina and Gleaser, 2004).

The focus of this study is on the public perceptions of the causes of poverty in post-socialist countries. The development of post-socialist welfare state has received abundant scholarly attention over last years (e.g. Cerami and Vanhuysse, 2009, Inglot, 2008, Haggard and Kauffman, 2008; Cook, 2007; Aidukaite, 2004; Deacon 1992). As the decades under socialist regime have had an undeniable influence on societies, these studies have emphasised the role of institutional differences within region as well as pre-socialist historical origins of the welfare states. Now after 20 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka, 2009, Tomka, 2006, Cerami and Vanhuysse, 2009, Inglot, 2008).

Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasised clear differences between Western and Eastern European countries (Gallie and Paugam, 2002; Kallio and Niemelä, 2010; Lepianka, 2007; van Oorschot and Halman, 2000). East Europeans are more inclined to blame either the individual behaviour or the flaws of the economic system of country than their Western counterparts. However, there is a shortage of comparative studies on attributions for poverty in particular focusing on the East Central European countries or on the whole post-socialist space. There is therefore a need for a detailed and more up-to-date analysis of the public perceptions of the causes of poverty in post-socialist countries.

Hence, utilising The Life in Transition Survey (LiTS) conducted jointly by the European Bank for Reconstruction and Development (EBRD) and the World Bank in 2010, the purpose of the article is to examine cross-country differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries. The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions.

# Prior findings on cross-national differences

Pioneering empirical studies on lay explanations of poverty emerged at the early 1970s when Joe R. Feagin (1972; 1975) examined Americans perceptions of the causes of poverty. He categorised reasons for poverty into three basic categories: individualistic, structural and fatalistic. Later, many factor analytic studies have given empirical support to Feagin’s categorisation (e.g. Feather, 1974; Furnham, 1982; Hunt, 1996; Morçöl, 1997; Nasser et al., 2002; Niemelä, 2008). Yet some of the recent studies have expanded our understanding by incorporating more contemporary beliefs into the attributional scales. These recent works emphasise especially the relevance of cultural attributions for poverty (Bullock et al., 2003; Cozzarelli et al., 2001).

van Oorschot and Halman (2000) have suggested a four-tier typology (Table 1) in which individual blame emphasises internal factors, such as undesirable behaviour of the poor, moral looseness and laziness. Individual fate type of explanation emphasises that poverty can be caused by uncontrollable and inescapable internal factors. These are factors such as personal misfortune, illness or just bad luck. Social fate points out that poverty is caused by societal factors, but these factors are due to impersonal and unavoidable processes such as economic recessions. Finally, social blame sees poverty as the result of processes induced and controlled by the actions of certain groups and parties in society. Therefore, these groups and parties can be blamed for poverty.

![Dimensions and types of explanations of poverty](figure/nelikentta.png)

Prior empirical findings on attributions for poverty in Europe have shown that there are some differences between the West and the East of Europe. Western Europeans are more inclined to support structural explanations for poverty than their Eastern counterparts. On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä, 2010; Kreidl, 2000; Lepianka, 2007; Lepianka et al., 2010; van Oorschot and Halman, 2000). Earlier results suggest, however, that there are variations both within and across groups of countries. Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä, 2010). Also Lepianka et al. (2010) come to the conclusion that further investigation is required due to rather large variation across countries. They showed that individual fate type of explanation is supported relatively less frequently in the East of Europe than in Western European countries. Regarding the variation within the Eastern European countries, they found that in some countries individual blame is strongly endorsed: especially the citizens of Czech Republic attribute poverty in internal terms (see also Lepianka, 2007). On the other hand, there are countries like Lithuania, Poland, Romania and Croatia in which social blame type of explanation is strongly supported. Finally, social fate type of explanation – living in poverty is explained as simply being part of modern progress – is endorsed especially in Russia.

National case studies from Estonia and Russia have given support to findings of cross-national studies. There is strong support for the idea that the poor have only themselves to blame for their poverty. In addition, both of these countries have strong support for social blame as well. Estonians are as likely to blame the flaws of the economic system as the individualistic reasons. In Russia, on the other hand, the economic system is blamed even more the individuals themselves (Gorshkov and Tikhonova, 2006; Stephenson, 2000). These results are consistent with Lepianka's (2007: 127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia. Interesting result is that social blame coincides strongly with individual blame. This illustrates the split-consciousness of public perceptions, which means the coexistence of both dominant and potentially challenging beliefs, where the former represents dominant stratification ideology or value structure of a given country and the latter everyday stratification-related experience of an individual (Kluegel et al., 1995).

Based on qualitative poverty studies in Central Asia, Caucasus and Ukraine and Moldova Dudwick et al. (2003: 23) states that, generally, people blamed their own poverty on the failure of the Soviet state and the corruption, indifference, and incompetence of their new leaders. At the same time, however, they often attributed the poverty of others to individual failure, such as laziness, alcoholism, having too many children to support, or having too few children to provide for their old age. They also reported that many of the respondents had become very distrustful of government officials and many suspected that government officials of exploiting the political and economic turmoil for their own gain, at the expense of ordinary citizens.

From the perspective of value structure, the role of "socialist or soviet legacy" in poverty attributions is worth of considering. Murthi and Tiongson (2008) reviewed large body of literature on East-West differences in preferences for redistribution of which vast majority supported the "socialist legacy". By socialist legacy they meant that "owing to the legacy of the past there is a strong preference for income equality and redistributive state spending in post-socialist countries". While testing the findings of previous studies on East-West differences, this study explores the effect of country groups with different socialist and post-socialist past. In order to explore the differences between market economies and post-socialist countries, but also within the "ex-soviet bloc", the study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and Commonwealth of Independent States (CIS) countries. In order to examine the distinctiveness of post-socialist perceptions descriptive analyses include also a selection of Western European countries.

Besides the so called regime-effect, recent comparative studies on attributions for poverty have included country-level explanatory factors in the analysis (e.g. Albrekt Larsen, 2006; Gallie and Paugam, 2002; Lepianka, 2007; Lepianka et al., 2010). For instance, national-level economic conditions have been proved to be associated with welfare attitudes. Increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007). In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual laziness explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved. This result is confirmed by Kallio and Niemelä (2010). They found that the larger the economic growth in a country is, the more likely people are to blame the poor themselves. Thus, we can assume that changes in country’s economic performance are associated with attributions for poverty regardless of country grouping. This study analyses the role of changes in economic performance during the period of global financial crises.

There are also considerable differences in socio-political context between the examined countries. The article analyses whether the level of income inequality and the state of democracy in a given country are associated with attributions for poverty. Regarding the state of democracy we can assume that in those countries where the state of democracy is low, citizens blame more likely society for poverty than in those countries where citizens have political freedom and power to influence socio-political issues. The link between inequality and attributions for poverty is not that easy to predict. On the one hand, the public in countries with greater inequality is more inclined to mistrust "others" (e.g. Uslander, 2002). Therefore, the uneven distribution of wealth should mean greater support for the individualistic explanations for poverty. On the other hand, cross-national studies on attributions for poverty have shown that the support for the individualistic explanation is lower in the countries with lower level of income inequality (e.g. Gallie and Paugam, 2002; Kallio and Niemelä, 2010).

# Risks factors as an individual-level determinants of perceptions

In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalised by the individuals in the society in question (Taylor-Gooby, 1985; Svallfors, 1995). More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice, 2001) and resource/class-based risks (Cusack et al. 2006; Blomberg et al. 2012; Finseraas and Ringdal, 2012) as key to welfare attitudes. Thus, studies have assumed – and to some extent proved – that groups that have a greater risk of facing social problems or being dependent on welfare state benefits might perceive the government’s role in welfare issues in a distinct way compared to those who are not exposed to these risks.

Inspired by the studies which emphasise risks as the key mechanisms, this article tests to what extent factors related to financial hardship are associated – beyond the welfare regime – with perceptions of the causes of poverty. Following this line of reasoning we are interested in whether those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks. Thus, the research design in regard to determinants of poverty attributions enables us to study from a fairly unexplored perspective the different mechanisms through which individuals who have a greater risk to experience material destitution perceive the causes of poverty.

The previous studies on individual-level determinants of poverty perceptions have focused on individual’s socio-demographic characteristics (e.g. Bullock, 1999; Cozzarelli et al., 2001; Furnham, 1982; Morcöl, 1997; Niemelä, 2008). Results have, however, shown that socio-demographic variables explain very little and they do not reveal clear or consistent patterns. On the other hand, there are studies which support the assumption that those who perceive themselves to be poor are more likely to agree with the external cause of poverty than people well above the poverty line (Niemelä, 2008; Saunders, 2002: 155–156). Yet, there are also contradictory findings indicating that personal experience of poverty might lead to individualistic explanations. For instance, claimants of public welfare seem to hold negative views about other recipients (Bullock, 1999; 2004; Golding and Middleton, 1982: 178).

Studies which have explored the importance of individual’s economic situation on attributions for poverty have focused on individual’s cross-sectional situation of financial circumstances. However, as suggested in the literature on welfare attitudes not only the risk patterns but also the way people perceive their future risks might have an effect on attitudes toward the welfare state (Andress and Heien, 2001; Boeri et al., 2001; Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012). In their European comparison Blomberg et al. (2012) found that perceived future risks functioned as an intermediating factor between "objective" risk and welfare state attitudes, i.e. those who perceive their economic future as endangered were more in favour of state responsibility than those who do not expect their future situation to be difficult. Consequently, following the self-interest hypothesis, we can assume that those who perceive their future economic situation to be worse than now endorse more likely external reasons for poverty than those who do not expect difficulties in their future economic situation.

Not only the current situation or the perceived future but also the past experiences might be associated with individual’s evaluations (e.g. Kumlin, 2007). In regard to attributions for poverty, the effects of past experiences are – according to our knowledge – totally unexplored area of research. In this article we will analyse past experiences by two ways. First, we are interested in whether those whose current economic situation is lower than four years ago perceive the causes of poverty distinct way compared to those whose economic situation has improved or remained in the same level. Second, we will focus on the effects of the global financial crisis. We assume that economic crisis have affected especially those who are the most vulnerable groups in the society. Therefore, we will analyse whether those who report that the economic crisis has affected their household at least a fair amount perceive the causes of poverty differently than those who report that it has had no or only a little effect to their household. According to self-interest hypothesis, we assume that those whose economic situation has declined and those whose household has affected by economic crisis are more inclined to support external causes of poverty than those who has not experienced economic problems.

In addition to economic situation, other indicators which measure a greater than average risk of facing more permanent social problems, are "transfer dependency" and low education. Transfer dependency refers to a situation that the respondent’s main source of income relies on private or social transfers (Blomberg et al., 2012; Fridberg and Kangas, 2008). Because of their close relation with welfare system, it can be assumed that welfare recipients perceive more likely structural than individual reasons for poverty compared to the rest of the population.

Low education refers to the lack of capabilities of an individual. Prior findings do not reveal consistent patterns regarding the effects of education. On the one hand, some studies have found the so-called inverted U-relationship which means that people with a middle level of education support the individualistic explanation, whereas people with a low or a high level of education are more likely to endorse structural explanations (Feagin, 1972; cf. Furnham, 1982). Moreover, the relationship between education and fatalistic explanation has been found to be negative (Morcöl, 1997). On the other hand, there are studies which suggest that people with a lower level of education are more likely to favour individualistic explanations. Moreover, when the educational level increases, the likelihood of perceiving causes of poverty in individualistic terms decreases (Niemelä, 2008). It is therefore difficult to form any solid hypotheses as to whether low education as a social risk is associated with perceptions of the causes of poverty in a similar vein than other risk factors discussed above.

# Purpose of the study

Earlier studies have shown that attributions for poverty differ in general between the West and the East of Europe but the variation between East Central European countries is relatively large. There is, however, a shortage of comparative studies which have focused particularly on differences within post-socialist countries. In addition, most of the prior research is based on attitude surveys from the 1990s and most of them have restricted country selection. Therefore, as previous studies have suggested (Kallio and Niemelä, 2010; Lepianka et al., 2010), there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space.

Hence, the general purpose of the study is to analyse public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries. Based on the findings of previous studies, we assume that in general post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations. However, we hypothesise also that findings indicate relatively large variation between countries.

As a country-level determinants, we will analyse, whether this variation can be explained by different regimes within post-socialist countries and the post-socialist legacy. The study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and CIS countries. The survey data enables us also to compare post-socialist countries to a selection of Western European nations. Thus, in order to examine the distinctiveness of post-socialist perception Western European countries are included in the descriptive analyses.  Besides the country groups we will also analyse the role of changes in country’s economic performance during the period of global financial crises, income inequality and the state of democracy.

As an individual-level determinants for the perceptions article test to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty, and whether these factors are associated with perceptions similar way regardless of the  groups of countries studied. Hereby, the research design aims at contributing to the theoretical reasoning regarding the importance of self-interest for attributions for poverty which assumes that those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.

# Data and methods

The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank (WB) in 2010. The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample. More information regarding how households were sampled and people interviewed is available at ERBD, 2011. Second round of the survey had a special focus on how peoples lives were affected by the global economic crisis and it's aftermath. The dataset, documentation and publications are freely available at the survey website.

[^***2]:[ebrd.com/pages/research/publications/special/transitionII.shtml](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml)

We are using LiTS2 data from 21 transition countries with additional 4 Western European countries (France, Germany, Italy and Sweden) (See Figure 1). In total 23513 households from 25 countries are included in the analysis. Transition countries were broken down into two groups. CEE includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania; and CIS[^****] Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia.

[^****]:Georgia withdrawn it's membership from CIS in 2008

## Dependent variables

The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?**

- Because they have been unlucky
- Because of laziness and lack of willpower
- Because of injustice in our society
- It is an inevitable part of modern life
- Other

Thus, the question is a standard forced-choice question which is used also in other comparative surveys such as in European Values Study and in Eurobarometer. According to van Oorschot and Halman (2000) different statements represent a four-tier typology of poverty explanations as follows: While the unlucky refers to individual fate, the laziness and lack of willpower represents individual blame. In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate. The option 'other' was treated as missing data.

As our main interest here is in support for social and blame type of explanation and as we are employing logistic multivariate methods, the answers were recoded in two separate binomial variables. In \emph{social blame} variable we have  value 1 for supporting social blame type of explanation and 0 for supporting other type of explanation. Same logic applies to variable \emph{individual blame}.

## Individual level independent variables

Six individual-level risk predictors are included in the analysis. Variables are recoded as binomial where value 1 is coded for risk and 0 for absence of risk. The items included are described in the table 2.

```{rtablepredictor, results='asis', echo=FALSE}
Variable <- c("Transfer dependency","Low education","Perceived low income level","Income compared to past","Income compared to future","Effect of financial crises")
Description <- c("Social or private transfers \n as main source of income","No or only compulsory level education","Perceived income level below the country median","Perceived change in income compared to situation in four years ago","Expected change in income over the next four years","Whether respondents perceives that their household has suffered the financial crisis great or fair amount")
table <- data.frame(Variable,Description)
library(xtable)
xtbl <- xtable(table, caption="Individual level predictors")
print.xtable(xtbl, type="latex", comment=FALSE, size="scriptsize",include.rownames=FALSE)
```

## Country level independent variables

An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of four variables: 1) change in total GDP in 2007 - 2010, 2) gini coefficient of economic inequality, 3) 2011 democracy index and 4) country grouping. Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), gini coefficients from [UNU-WIDER World Income Inequality Database](http://www.wider.unu.edu/research/Database/en_GB/database/) and Democracy Index from the [Economist Intelligence Unit](http://www.eiu.com/Default.aspx). Gini coefficient for each country is the latest available varying from 2002 to 2006. (see Appendix A)

## Methods

The empirical section starts by presenting the results of descriptive analyses. As we assume that individual level perceptions are dependent both on individual and country-level contextual factors, it is recommended to use multilevel regression analysis for this task (Snijders \& Bosker, 2011).  Multilevel analysis requires hierarchical dataset and allows the estimation of country-level effects on individual level outcomes. The analysis is performed using lmer-function from lme4 package in R [^*****]. All of the models employed are so called multilevel logistic random intercept models at the level of individuals and countries.

[^*****]:R-code for data transformation and analysis is available at [research.muuankarski.org/code/](http://research.muuankarski.org/code/)

Our multilevel modelling focuses first in bivariate connections between two dependent variables and both individual and country level independent variables(See Appendix B). At the second step we include all individual level variables to a single model (models sb1 and ib1). Then we add each of the four country level variables one at the time (models sb2/ib2 - sb5/ib5). With limited number of countries we are obliged to use only few contextual variables. Tables will include both a fixed effects and a random effects part. The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups. Full model summaries currently available at [research.muuankarski.org/code/](http://research.muuankarski.org/code/).

# Results

## Descriptive statistics

Table 3 shows the proportion of the total number of respondents in a country which chose a particular explanation of poverty. We assumed that citizens in post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations. As expected results show relatively large variation within country groups. For the most of the post-socialist countries, social and individual blame types of explanation, i.e. injustice in a society and laziness of the poor, are the most popular poverty attributions. The highest support for social blame is found in Ukraine, Latvia, Lithuania, Hungary, Russia, Slovenia, Armenia, Moldova and Azerbaijan. In these countries majority of population blames the injustice in a society for poverty. Totally, in 19 of 21 examined post-socialist countries injustice in a society receives the highest support.

Also, in most post-socialist countries the individual blame is the second most popular explanation of poverty. Exceptions here are Slovenia, Estonia and Georgia where citizens endorse societal explanations more likely than individual explanations. In Czech Republic social explanations as well as individual blame explanation receive public support quite evenly. Finally, the clearest exception in post-socialist regime is Uzbekistan where poverty is attributed strongly with internal terms.

In regard to differences between the East and the West of Europe, results show that there are some differences. Yet there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al., 2010; Kallio and Niemelä, 2010). Clearest differences between the West and the East is the stronger support for social fate in Western Europe. In Germany and Sweden external reasons for poverty receive the majority of support. On the other hand, attributions for poverty in France and Italy seems to be quite similar with post-socialist countries - blame types of explanations are endorsed more likely than fate types of explanations. Finally, the figures of Great Britain are the example of selective welfare policy dominated liberal regime in which the individual blame type of explanations is strongly supported (see also Albrekt Larsen, 2006).

Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty. Overall, the results indicate that for the most of the post-socialist countries two most popular explanations are individual and social blame types of explanations.Yet, the differences between social fate and individual blame types of explanations are in some countries very small. In general, Western Europeans (except Britons) do not support individual blame type of explanation as much as citizens in the East. Instead they emphasise more social reasons for poverty. Results also clearly emphasise that the idea that poverty is caused by just individual bad luck is the least popular perception of the reason for poverty.

## Determinants of perceptions

The results above indicate that there are cross-national differences between countries and within country groups which means that there are also some other contextual and individual-level factors which should explain the cross-national differences. Therefore, we will also analyse the importance of country-level economic, social and political conditions and individual-level risk factors related to financial hardship as determinants for attributions for poverty. Because the descriptive results showed that the majority of citizens in the post-socialist countries endorsed social and individual blame types of explanations, the subsequent analyses will focus only on these two explanations.

![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)

Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 2, 3 and 4). Correlation coefficients at country group1 level are shown on top of each regression line.Figure 2 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010). In regard to post-socialist countries there is substantial variation between countries in economic change. The general trend is quite clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.35) and more to individual blame (.48). In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries. However, this is greatly due to economic decline in Great Britain. Hence, without Great Britain we can not see any significant association between economic change and poverty perceptions in Western Europe.

![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)

There is also a substantial variation in the level of income inequality among post-socialist countries. Figure 3 shows that the higher level of income inequality is associated with higher level of social blame. As for individual blame type of explanation the picture is more puzzled, showing only weak connection between larger inequality and lower support for individual blame explanation. Again, the Western European countries differ with post-socialist countries. Cross-national differences in income inequality are much smaller in Western Europe than among post-socialist countries. There is a weak negative association between the social blame and income  inequality and a strong positive association between the individual blame and income inequality.

![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)

>Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media. [info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.png)

In regard to the level of democracy we can see great differences between and within groups of countries. Overall association is weak (.19 for social blame and -.44 for individual blame), but in CIS countries the higher level of democracy is clearly indicating higher support for social blame and lower support for individual blame type of explanation. In CEE and in Western Europe the variation in the state of democracy is much smaller than in CIS countries. In these countries the support for blame  types of explanations seems to vanish when the level of democracy increases. Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.

## Multilevel analysis

- click for tables: [code/multilevel_models.pdf](code/multilevel_models.pdf)
- click for RC19 tables: [code/multilevel_modelsRC19.pdf](code/multilevel_modelsRC19.pdf)

>As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables. Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.

Table 4 and 5 present figures for multilevel logistic analysis. Calculating R² or residual standard deviation makes no sense in the case of logistic regression or other binomial models. Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit. Deviance can be treated as similar to residual standard deviation for generalized linear models. Basically, deviance is a measure of error, and lower deviance implicates better fit to data. When a random predictor is added to model, we expect deviance to decrease, on average, by 1. When predictor is more informative deviance should decrease more than 1. (Gelman and Hill, 2007). As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.

Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences. Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation. Inclusion of individual level predictors decreases the variance in both models, but only moderately.

<!-- As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables. Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level. -->

Table 4 and 5 present figures for multilevel logistic analysis. Calculating R² or residual standard deviation makes no sense in the case of logistic regression or other binomial models. Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit. Deviance can be treated as similar to residual standard deviation for generalized linear models. Basically, deviance is a measure of error, and lower deviance implicates better fit to data. When a random predictor is added to model, we expect deviance to decrease, on average, by 1. When predictor is more informative deviance should decrease more than 1. (Gelman and Hill, 2007). As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.

Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences. Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation. Inclusion of individual level predictors decreases the variance in both models, but only moderately.

Looking at the deviance for the empty models both in tables 4 and 5, we can see that there is "less" to be explained for in the model for individual blame than for social blame. When adding the individual level predictors to the models, the deviance decreases by 8512 for social blame and 6636 for individual blame. In both model sb1 and ib1 we have four statistically significant predictors. Low education and worsened household income level do not contribute in explaining neither of the outcomes. Perceived low income level has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation. Coefficient of .368 can be translated so that if respondent perceives belonging to low income group he or she is 9.2\% more likely to blame society for poverty than a non-low income individual. Expected worsening in household income and financial crises experince predict social blame almost in similar manner. All our individual level predictors but low education predict lower probability for supporting individual blame (model ib1). Being greatly affected by financial crisis makes one 11\% less likely to support for individual blame type of explanation.

When adding in the country grouping (CEE/CIS) we can observe decrease close to -1 in deviance in both models sb2 and id2. Country grouping in itself is not significant in our models, whereas gini coefficient has a weak connection to individual blame type of explanation, so that individual blame is supported more in more unequal countries, as we assumed based on studies about the association between generalised trust and equality (Uslander 2002). Inclusion of change in national GDP between 2007 - 2010 substracts the deviance by 14 for social blame and 9 for individual blame model. Change in GDP becomes the most significant predictor of our country level variables for both type of explanation. The better the national economy has performed during the financial crises, the more probably is the support for individual blame type of explanation. Outcome is very much align with the findings of Gallie and Paugam (2002) and Kallio and Niemelä (2010). Our fourth contextual variable, democracy index, is significant in explaining the individual blame in a way that higher level of democracy implies lower support for individual blame type of explanation. Coefficient plots from both full models summarizes the estimates for logistic regression coefficients at 50\% and 90\% intervals in figure 7.

As for explaining the between-country variation the inclusion of country grouping adds a very little. Gini coefficient reduces variation from .24 to .19 for social blame, but it is the inclusion of GDP change that has the greatest effect. Between-country variation  to .12 for social blame and .04 for individual blame. In the final models the inclusion of democracy index results country-level variances at .09 and .02 for social and individual type of explanation, respectively. In other words, only 9\% and 2\% of the variation in the final models are due to country differences at final stage.

# Conclusions

To sum up, the results indicate that in general social blame type of explanation is clearly the most popular explanation in post-socialist countries. Social blame is followed by individual blame, social fate and individual fate types of explanations. Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty in the European post-socialist countries. In regard to differences between the East and the West of Europe, results show that Western Europeans emphasise more social reasons for poverty.

As assumed results also emphasised large between-country differences among post-socialist countries. By employing a multilevel approach we wanted to shed light especially on this issue. To begin with, we found that differences between countries are responsible for almost 30 percent of the variation in the support for social blame and 12 percent for individual blame type of explanation. Our contextual level predictors decreased the variance down to 9 percent for social and 2 percent for individual type of explanation. Yet, findings showed that the membership in the CIS and CEE country group was not significant in this respect.

As we hypothesised changes in country’s economic performance during the period of global financial crises were associated with attributions for poverty. In fact, changes in economic performance was the most significant predictor of our country-level variables for both social and individual blame types of explanations: the better the economy had performed the more support gained the individual blame type of explanation. And on the other hand, increased financial strain is associated with stronger support for the view that poverty is caused injustices of society.

Also, the level of democracy had a substantial impact. Yet, the results were to some extent against our hypothesis. Contrary to our assumption, citizens in those countries where the state of democracy is low do not blame more likely society than citizens in countries where they have political freedom and power to influence socio-political issues. Instead, findings indicate that the lower level of democracy was associated with stronger support for individual blame type of explanation. Social blame type of explanation was, in turn, endorsed more likely in countries with higher level of democracy.

Results of the individual-level determinants of perceptions supported the self-interest hypothesis which states that those who have financial difficulties perceive the causes of poverty differently than those who are not exposed to poverty-related risks. Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crises has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty. Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).

This study focused only on poverty-related risks as individual-level determinants. There are, of course, other variables which might be associated with attributions even more than risk factors. For instance, another line of research have emphasised the role of other beliefs, attitudes and values rather than interest (e.g. Cozzarelli et al., 2001). Treating the values, attitudes and other perceptions as determinants for attributions of poverty is, however, problematic because it is difficult to prove the causal relationship between attributions and other attitudes. This does not mean that they are irrelevant. Conversely, we argue that future research should try to investigate to what extent high levels of social and individual blame are associated with wider welfare attitudes in post-socialist countries. In order, for example, to explore policy implications more thoroughly, there is a need for future research to focus on the relationship between attributions for poverty and the legitimacy of specific types of welfare policies in post-socialist countries.

As a methodological contribution, findings showed interesting differences between the East and the West regarding the effects of contextual variables. Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and income inequality are quite modest compared especially to CIS countries. Consequently our results showed more modest and in some cases opposite associations between dependent and contextual variables among Western European countries than among post-socialist group.

Thus, the finding hold critical implications for future research. On the one hand, in order to examine the effects of contextual factors, comparative social research should try to utilize comparative data encompassing more diverse cases than, for instance, conventional EU17 or OECD countries. On the other hand, the well-known limitation of this kind of variable-oriented approach is that larger number of cases may pose also more obstacles to making interpretive statements about specific cases or even about categories of cases (Ragin, 1987). For instance, in regard to policy implications of research on welfare attitudes, broadening the country selection from Western democracies (in some cases) to more authoritarian societies raises a question whether public attitudes have similar constraining role in policy-making that have been assumed and to some extent proved to be in the policy-making of Western democracies. This kind of research question calls also case-oriented qualitative research methods.

# References

Aidukaite, Jolanta (2004) The Emergence of the Post-Socialist Welfare State - The Case of the Baltic States: Estonia, Latvia and Lithuania, Södertörn University College, Sweden.

Albrekt Larsen, Christian (2006) The Institutional Logic of Welfare Attitudes. How Welfare Regimes Influence Public Support. Hampshire: Ashgate.

Alesina, Alberto and Glaeser, Edward L. (2004) Fighting Poverty in the US and Europe. A World of Difference. Oxford: Oxford University Press.

Andress, Hans-Jürgen and Heien, Thorsten (2001) Four Worlds of Welfare State Attutudes? A Comparison of Germany, Norway, and the United States. European Sociological Review 17 (4), 337–356.

Arts, Wil and Gelissen, John (2001) Welfare States, Solidarity and Justice Principle: Does the Type Really Matter. Acta Sociologica 44 (4), 283–299.

Blomberg, Helena and Kallio, Johanna and Kangas, Olli and Kroll, Christian and Niemelä, Mikko (2012) Attitudes among High-Risk Groups. In Svallfors, Stefan (ed.) Contested Welfare States. Welfare Attitudes in Europe and Beyond. Stanford University Press: Stanford, CA, 58–80.

Boeri, T., Börsch-Supan, A. and Tabellini, G. (2001) Would You Like to Shrink the Welfare State? A Survey of European Citizens. Economic Policy 16, 7–50.

Bullock, Heather E. (1999) Attributions for Poverty: A Comparison of Middle-Class and Welfare Recipient Attitudes. Journal of Applied Social Psychology 29 (10), 2059–2082.

Bullock, Heather E. (2004) From the Front Lines of Welfare Reform: An Analysis of Social Worker and Welfare Recipient Attitudes. The Journal of Social Psychology 144 (6), 571-588.

Bullock, Heather E. and Williams, Wendy R. and Limbert, Wendy M. (2003) Predicting Support for Welfare Policies: The Impact of Attributions and Beliefs about Inequality. Journal of Poverty 7 (3), 35–56.

Campbell, John L. (2004) Institutional Change and Globalization. Princeton and Oxford: Princeton University Press.

Cerami, Aifio and Vanhuysse, Pieter (2009) Post-Communist Welfare Pathways. Theorizing Social Policy Transformations in Central and Eastern Europe. Hampshire: Palgrave Macmillan.

Cook, L. J. (2007) Postcommunist Welfare States: Reform Politics in Russia and Eastern Europe, Cornell University Press.

Cozzarelli, Catherine and Wilkinson, Anna V. and Tagler, Michael J. (2001) Attitudes Toward the Poor and Attributions for Poverty. Journal of Social Issues 57 (2), 207-227.

Cusack, T., Iversen, T. and Rehm, P. (2006) Risks at Work: The Demand and Supply Sides of Government Redistribution. Oxford Review of Economic Policy 22, 365–389.

Dudwick, N., Gomart, E., and Marc, A. (2003) When Things Fall Apart: Qualitative Studies of Poverty in the Former Soviet Union. World Bank Publications.

Economicst Intelligence Unit (2001) Democracy index 2011. Democracy under stress. Downloaded at: \href{http://www.eiu.com/democracyindex2011}{www.eiu.com/democracyindex2011}

Feagin, Joe R. (1972) Poverty: We Still Believe That God Helps Those Who Help Themselves. Psychology Today 6 (2), 101–129.

Feagin, Joe R. (1975) Subordinating the Poor: Welfare and American Beliefs. NJ Prentice Hall: Upper Saddle River.

Feather, Norman T. (1974) Explanations of Poverty in Australian and American Samples: The Person, Society, or Both? Australian Journal of Psychology 26 (3), 199–216.

Fridberg, Torben and Kangas, Olli (2008) The Welfare State, Poverty and Social Exclusion. In Ervasti, H., Fridberg, T., Hjerm, M. and Ringdal, K. (eds.) Nordic Social Attitudes in a European Perspective. Edward Elgar: Cheltenhamn, 22–47.

Furnham, Adrian (1982) Why are the poor always with us? Explanations for poverty in Great Britain. British Journal of Social Psychology 21(3), 311–322.

Gallie, Duncan and Paugam, Serge (2002) Social Precarity and Social Integration. Report for the European Commission Based on Eurobarometer 56.1. Brussels: European Commission Directorate-General Employment.

Giczi, Johanna and Sik, Endre (2009) Trust and Social Capital in Europe. In TARKI European Social Report 2009. TARKI: Budapest, 63-81.

Golding, Peter and Middleton, Sue (1982) Images of Welfare. Press and Public Attitudes to Poverty. Oxford: Basil Blackwell.

Gorshkov, Mikhail K. and Tikhonova, Natalia E. (2006) Wealth and Poverty in the Perceptions of Russians. Sociological Research 45 (1), 27–40.

Hunt, Matthew O. (1996) The Individual, Society, or Both? A Comparison of Black, Latino, and White Beliefs about the Causes of Poverty. Social Forces 75 (1), 293–322.

Inglot, T. (2008). Welfare States in East Central Europe, 1919-2004 (1st ed.). Cambridge University Press.

Iversen, T. and Soskice, D. (2001) An Asset Theory of Social Policy Preferences. American Political Science Review 95, 875–893.

Jaeger, Mads Meier (2006) Welfare Regime and Attitudes Towards Redistribution: The Regime Hypothesis Revisited. European Sociological Review 22 (2), 157–170.

Kallio, Johanna and Niemelä, Mikko (2010) Public Perceptions of the Causes of Poverty: Multilevel Analysis of 26 European Nations. Unpublished manuscript.

Haggard, S. and Kaufman, R. R. (2008) Development, Democracy, and Welfare States: Latin America, East Asia, and Eastern Europe, Princeton Univ Pr.

Kluegel, James R. and Mason, David S. and Wegener, Bernd (eds) (1995) Social Justice and Political Change: Public Opinion in Capitalist and Post-communist States, New York: Aldine De Gruyter.

Kouvo, Antti and Kankainen, Tomi and Niemelä, Mikko (2012) Welfare Benefits and Generalized Trust in Finland and Europe. In Ervasti, Heikki and Goul Andersen Jörgen and Fridberg, Torben and Ringdal, Kristen (eds.) The Future of the Welfare State. Social Policy Attitudes and Social Capital in Europe. Edward Elgar: Cheltenham, 195-213.

Kreidl, Martin (2000) Perceptions of Poverty and Wealth in Western and Post-Communist Countries. Social Justice Research 13 (2), 151–176.

Kumlin, Staffan (2007) Overload or Undermined? European Welfare States in the Face of Performance Dissatisfaction. In Svallfors, Stefan (ed.) The Political Sociology of the Welfare State: Institutions, Social Cleavages, and Orientations. Stanford University Press: Stanford, CA, 80–116.

Lepianka, Dorota (2007) Are the Poor to be Blamed or Pitied? A Comparative Study of Popular Attributions in Europe. Tilburg: Tilburg University.

Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe. Effects of Contextual and Individual Characteristics across 28 European Countries. Acta Sociologica 53 (1), 53–72.

Morçöl, Göktuğ (1997) Lay Explanations for Poverty in Turkey and Their Determinants. Journal of Social Psychology 137 (6) 728–738.

Murthi, M., & Tiongson, E. R. (2008). Attitudes to Equality: The Socialist Legacy Revisited. SSRN eLibrary. Retrieved from http://papers.ssrn.com/sol3/papers.cfm?abstract_id=1098627

Nasser, Ramzi and Abouchedid, Kamal and Khashan, Hilal (2002) Perceptions of the Causes of Poverty Comparing Three National Groups: Lebanon, Portugal, and South Africa. Current Research in Social Psychology 8 (7), 101–119.

Niemelä, Mikko (2008) Perceptions of the Causes of Poverty in Finland. Acta Sociologica, 51 (1), 23–40.

Nilson, Lina B. (1981) Reconsidering Ideological Lines: Beliefs about Poverty in America. The Sociological Quarterly 22 (4), 531-548.

Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of California Press.

Rothstein, Bo and Stolle, Dietlind (2008) The State and Social Capital: An Institutional Theory of Generalized Trust. Comparative Politics 40 (4), 441-459.

Saunders, Peter (2002) The Ends and Means of Welfare. Coping with Economic and Social Change in Australia. Cambridge: Cambridge University Press.

Schumpeter, Joseph A. (1942) Capitalism, Socialism & Democracy. New York: Harper & Brothers.

Szikra, D., & Tomka, B. (2009). Social Policy in East Central Europe: Major Trends in the Twentieht Century. In Post-communist Welfare Pathways (Vol. 2009).

Stephenson, Svetlana (2000) Public Beliefs in the Causes of Wealth and Poverty and Legitimization of Inequalities in Russia and Estonia. Social Justice Research 13 (2), 83–100.

Svallfors, Stefan (1997) Worlds of Welfare and Attitudes to Redistribution: A Comparison of Eight Western Nations. European Sociological Review 13 (3), 283–304.

Tomka, B. (2006). East Central Europe and the European Social Policy Model: A Long-Term View. East European Quarterly, 40(2), 135.

van Oorschot, Wim and Halman, Loek (2000) Blame or Fate, Individual or Social? An International Comparison of Popular Explanations of Poverty. European Societies 2 (1), 1–28.

Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterly 37 (3), 413–428.

### 2013-10-30 code/descriptive_analysis.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/code/")
library(kaRski)
knitpandoc("descriptive_analysis")

```

```{ropts, echo=FALSE}
opts_chunk$set(echo=FALSE, dev='svg', eval=TRUE, cache=TRUE, fig.height=10, message=FALSE, warning=FALSE)
```

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>

# Visualizing the predictors

## Perceived income

### Original variable

```{rdescplot1_1, fig.width=15}
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
library(ggplot2)
library(reshape2)

df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q227),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

#### Transformed variable

```{rdescplot1_2, fig.width=15}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$income),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("Low","High"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Transfer dependency/Income source

### Original variable

```{rdescplot2_1, fig.width=15}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q226m_txt),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot2_2, fig.width=15}

df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$incsour3),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("Dependent","Not.dependent"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Past

### Original variable

```{rdescplot3_1, fig.width=15}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q228),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot3_2, fig.width=15}

tbl <- data.frame(prop.table(table(df$cntry,df$past),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("same or better","worse"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Future

### Original variable

```{rdescplot4_1, fig.width=15}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q229),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot4_2, fig.width=15}

tbl <- data.frame(prop.table(table(df$cntry,df$future),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("same or better","worse"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Education

### Original variable

```{rdescplot5_1, fig.width=15}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q515),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cond <- as.numeric(levels(x$cond))[x$cond]
x <- subset(x, cond >= -1)
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

### Transformed variable

```{rdescplot5_2, fig.width=15}

tbl <- data.frame(prop.table(table(df$cntry,df$edu2),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("no or compulsory","higher"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

## Crises

### Original variable

```{rdescplot6_1, fig.width=15}
df$cntry <- factor(df$cntry)
tbl <- data.frame(prop.table(table(df$cntry,df$q801),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  scale_fill_manual(values=c("red","orange","yellow","green","dim grey","grey"))

```

### Transformed variable

```{rdescplot6_2, fig.width=15}

tbl <- data.frame(prop.table(table(df$cntry,df$crise),1)*100)
reg <- as.data.frame(table(df$cntry,df$group))
reg <- subset(reg, Freq > 0)
reg <- reg[with(reg, order(Var2)), ]
x <- merge(tbl,reg[,1:2],by="Var1")
names(x) <- c("cntry","cond","freq","group")
x$cond <- factor(x$cond, levels=c("Little or not at all","Great or fair amount"))

x <- x[with(x, order(cond)), ]
x$cntry <- factor(x$cntry, levels=reg$Var1)

ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
  geom_bar(stat="identity") +
  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))

```

## Macro level

```{rdescplot7, fig.width=15}
df <- macro[with(macro, order(uw.gini)), ]
df$cntry <- factor(df$cntry, levels=df$cntry)
ggplot(df, aes(x=cntry,y=uw.gini, colour=group)) +
  geom_point(size=5) +
  labs(title="gini coeficient") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{rdescplot8, fig.width=15}
df <- macro[with(macro, order(gdpchange07.10)), ]
df$cntry <- factor(df$cntry, levels=df$cntry)
ggplot(df, aes(x=cntry,y=gdpchange07.10, colour=group)) +
  geom_point(size=5) +
  labs(title="gdp change 2007 - 2010") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{rdescplot9, fig.width=15}
df <- macro[with(macro, order(wbgi.vae)), ]
df$cntry <- factor(df$cntry, levels=df$cntry)

ggplot(df, aes(x=cntry,y=wbgi.vae, colour=group)) +
  geom_point(size=5) +
    labs(title="Voice and accountability") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
```

```{rdescplot10, fig.width=15}
library(GGally)
ggpairs(macro[, c("uw.gini", "gdpchange07.10", "wbgi.vae","social.blame","individual.blame")])

```

# Final plots

![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](../figure/scatter_gdp.png)

![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](../figure/scatter_gini.png)

![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](../figure/scatter_wbgi.png)

### 2013-10-30 code/multilevel_models.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/code/")
library(knitr)
knit("multilevel_models.Rmd")
system("~/.cabal/bin/pandoc multilevel_models.md --number-sections --template=../tex_templates/dcolumn.latex -o multilevel_models.pdf")

system("~/.cabal/bin/pandoc -s multilevel_models.md -o multilevel_models.html")

#system("perl -pi -e 's/footnotesize/tiny/g' multilevel_models.tex")
```

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>

# Final tables

```{rmultilevel_analysis, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, results='hide'}
library(lme4)
#########################################
#########################################
setwd("~/workspace/lits/attrib/attrib_year2013/code")

#########################################
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")

summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.8 <- lmer(pov.log.social ~ uw.gini+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
                        family = binomial, data=df))
summary(bisoc.10 <- lmer(pov.log.social ~ wbgi.vae+(1|cntry),
                         family = binomial, data=df))

summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
                        family = binomial, data=df))
summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
                        family = binomial, data=df))
summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
                        family = binomial, data=df))
summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
                        family = binomial, data=df))
summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
                        family = binomial, data=df))
summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
                        family = binomial, data=df))
summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
                        family = binomial, data=df))
summary(biind.8 <- lmer(pov.log.individual ~ uw.gini+(1|cntry),
                        family = binomial, data=df))
summary(biind.9 <- lmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
                        family = binomial, data=df))
summary(biind.10 <- lmer(pov.log.individual ~ wbgi.vae+(1|cntry),
                         family = binomial, data=df))

summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                      family = binomial, data=df))
summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+(1|cntry),
                      family = binomial, data=df))
summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+(1|cntry),family = binomial, data=df))
summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))

summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
                      family = binomial, data=df))
summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+(1|cntry),
                      family = binomial, data=df))
summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+(1|cntry),family = binomial, data=df))
summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
```

\newpage

```{rmultilevel_social, results='asis', echo=FALSE, eval=FALSE, cache=TRUE}
library(stargazer)
stargazer("Empty model"=soc.0,
          "Model sb1"=soc.1,
          "Model sb2"=soc.2,
          "Model sb3"=soc.3,
          "Model sb4"=soc.4,
          title="Logistic multilevel random intercept model for social blame type of explanation (Using stargazer-package (missing country level random effects)).")
```

```{rmultilevel_individual2_texreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
htmlreg(list(ind.0,
          ind.1,
          ind.2,
          ind.3,
          ind.4),
       use.packages=FALSE,
       include.pvalues=TRUE,
       caption="Logistic multilevel random intercept model for individual blame type of explanation (Using texreg-package (p-value stars missing)).", 
       caption.above=TRUE)
```

```{rmultilevel_individual, results='asis', echo=FALSE, eval=FALSE, cache=TRUE}
stargazer("Empty model"=ind.0,
          "Model ib1"=ind.1,
          "Model ib2"=ind.2,
          "Model ib3"=ind.3,
          "Model ib4"=ind.4,
          title="Logistic multilevel random intercept model for individual blame type of explanation. (Using stargazer-package (missing country level random effects))")
```

```{rmultilevel_social_texreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
htmlreg(list(soc.0,
       soc.1,
       soc.2,
       soc.3,
       soc.4),
       use.package=FALSE,
       include.pvalues=TRUE,
          caption="Logistic multilevel random intercept model for social blame type of explanation( Using texreg-package (p-value stars missing)).", caption.above=TRUE)
```

```{rbivariate_social, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(texreg)
htmlreg(list(bisoc.1,bisoc.2,
          bisoc.3,bisoc.4,
          bisoc.5,bisoc.6,
          bisoc.7,bisoc.8,
          bisoc.9,bisoc.10),
       use.packages=FALSE,
       sideways=TRUE,
       scriptsize=TRUE,
       include.pvalues=TRUE,
       caption="bivatiate social blame", caption.above=TRUE)
```

```{rbivariate_individual, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(texreg)
htmlreg(list(biind.1,biind.2,
          biind.3,biind.4,
          biind.5,biind.6,
          biind.7,biind.8,
          biind.9,biind.10),
       use.packages=FALSE,
       sideways=TRUE,
       scriptsize=TRUE,
       caption="bivatiate individual blame", caption.above=TRUE)
```

### 2013-10-30 index.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

% Attributions for Poverty in Post-Socialist Countries
% Markus Kainu & Mikko Niemelä
% Jan 2013

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013")
library(kaRski)
knitpandoc("index")

cd ~/workspace/lits/attrib/attrib_year2013
lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
```

******

<center>
<table>
<tr>
<td>[Etusivu](index.html)</td>
<td>[Descriptive analysis](code/descriptive_analysis.html)</td>
<td>[Output tables in html](code/multilevel_models.html)</td>
<td>[Output tables in pdf](code/multilevel_models.pdf)</td>
<td>[Article in pdf](article2013.pdf)</td>
<td>[Article in html](article2013.html)</td>
<td>[Presentation slides](slides/slides_attrib.html)</td>
</tr>
</table>
</center>

*******

# Etusivu

### 2013-10-30 multilevel.Rmd
Begin editing for Moscow HSE LCSR seminar in November 2013

% Attributions for Poverty in Post-Socialist Countries
% Markus Kainu & Mikko Niemelä
% Jan 2013

<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>

******

<center>
<table>
<tr>
<td>[Etusivu](index.html)</td>
<td>[Analysis](multilevel.html)</td>
<td>[Article](article2013.html)</td>
</tr>
</table>
</center>

*******

# Attributions for poverty in post-socialist countries - hajontakuviot, **other** vaihtoehdon analyysi sekä monitasomallinnusta eri tavoin

Tässä dokumentissa esitetään vaihtoehtoisua ratkaisuja menetelmäkritiikkiin, jota saimme osaksemme RC19 konffassa. Eka siis vaihtoehtoisia maksomuuttujia - indeksejä

```{rpackagesload, echo=FALSE, warning=FALSE, message=FALSE, results='hide', cache=TRUE}
setwd("~/workspace/lits/attrib/year2013/")

# PACKAGES
# for data manipulation
library(car)
library(reshape2)
library(plyr)
library(gdata)
# for plotting
library(ggplot2)
# for tables
library(xtable)
library(grid)
#library(memisc)
# for survey design
library(survey)
# for multilevel analysis
library(lme4)
#library(coefplot2)

my_coef = function(file, width, height) {pdf(file, width = 5, height = 5, pointsize = 10)}
my_pdf = function(file, width, height) {pdf(file, width = 8, height = 4, pointsize = 10)}
my_pdf_map = function(file, width, height) {pdf(file, width = 8, height = 5, pointsize = 10)}
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

df <- read.csv("data/df.csv")
contextual <- read.table("data/contextual.txt")
df <- join(df,contextual,by="cntry",type="right")

t3 <- read.csv("data/scatter_file.csv", dec=",")
contextual$gdppercapita2010 <- NULL
contextual$Corruption2011 <- NULL

df.cross <- read.csv("~/workspace/general/indexes/1358024_qog_csd_csv_v6apr11.csv", sep=";")
df.cross$group <- recode(df.cross$cname, "c('Armenia','Azerbaijan','Georgia') = 'Caucasus';
                         c('Russia','Ukraine','Moldova','Belarus')= 'CIS North';
                         c('Kyrgyzstan','Kazakhstan','Tajikistan','Turkmenistan','Uzbekistan')='Central Asia';
                         c('Estonia','Latvia','Lithuania')='Baltic States';
                         c('Poland','Czech Republic','Slovakia','Hungary','Slovenia',
                         'Romania','Bulgaria')='CEE EU';
                         c('Croatia','Serbia','Montenegro','Bosnia and Herzegovina')='Ex-Yogoslavia';
                         c('Italy','Germany','United Kingdom','Sweden','France')='West Europe';
                         else='other'")

df.cross.postsoc <- subset(df.cross, group != 'other')

df.cross.postsoc$cname <- recode(df.cross.postsoc$cname, "'United Kingdom'='Great Britain'")

rm(df.cross)

df.qog <- subset(df.cross.postsoc, select=c("cname","group","ffp_fsi","wbgi_vae","undp_hdi","fh_polity2"))

df.qog <- rename.vars(df.qog, from="cname", to="cntry")
# yhdistetään nää aikaisempien datojen kanssa
t3 <- join(t3,df.qog,by="cntry")
df <- join(df,df.qog,by="cntry")

head(df)
#########
d.df <- svydesign(id = ~SerialID, 
                  weights = ~weight, 
                  data = df)
```

## 1. Hajontakuviot makromuuttujien ja *social blame*/*individual blame* selitysten välillä

### Change in GDP between 2007 and 2010

* Tää lienee ihan perusteltu ja kahdenvälinen yhteys on voimakas

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk61, echo=FALSE, results='markup', cache=TRUE#
t3 <- subset(t3, group1 != "Western Europe")
t3$group1 <- car:::recode(t3$group1, "else='group'")

func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$gdpchange07.10)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, gdpchange07.10) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=gdpchange07.10, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
  scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(gdpchange07.10,social.blame, label = COR, group=COR), 
            size=5, color="black") +
  labs(x = "Change in GDP between 2007 and 2010 (%)", 
       y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$gdpchange07.10)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, gdpchange07.10) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=gdpchange07.10, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(gdpchange07.10,individual.blame, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Change in GDP between 2007 and 2010 (%)", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk62, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE#
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))

```

### Gini coefficient

* Tää nyt on vaan hyvä olla. Kertoo myös jotain siitä, että Wilkinson-Picketin eriarvoisuus-teesi ei toimi köyhyysasenteiden osalta.

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk51, echo=FALSE, results='markup', cache=TRUE#
func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$Gini)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, Gini) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=Gini, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(Gini,social.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Gini coefficient", 
                   y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$Gini)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, Gini) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=Gini, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(Gini,individual.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Gini coefficient", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk52, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE#
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

### Democracy index

* Tää ois sitten yksi vaihtoehto näiden seuraavien joukossa.

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk71, echo=FALSE, results='markup', cache=TRUE#
func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$Democracy2011)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, Democracy2011) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=Democracy2011, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(Democracy2011,social.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Democracy Index 2011", 
                   y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$Democracy2011)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, Democracy2011) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=Democracy2011, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(Democracy2011,individual.blame, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Democracy Index 2011", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk72, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Democracy index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

### Democracy (Freedom House/Polity)

* Kahdenvälisissä tarkasteluissa vahvempi korrelaatio verrattuna Economistin Democracy indexiin, jota alunperin käytimme. En osaa sanoa kumpi olisi "akateemisesti" luotettavampi taho

> Scale ranges from 0-10 where 0 is least democratic and 10 most democratic. Average of Freedom House (fh_pr and fh_cl) is transformed to a scale 0-10 and Polity (p_polity2) is transformed to a scale 0-10. These variables are averaged into fh_polity2. The imputed version has imputed values for countries where data on Polity is missing by regressing Polity on the average Freedom House measure. Hadenius & Teorell (2005) show that this average index performs better both in terms of validity and reliability than its constituent parts.

```{rdemo_freedom_load, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk71, echo=FALSE, results='markup', cache=TRUE#
func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$fh_polity2)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, fh_polity2) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=fh_polity2, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(fh_polity2,social.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Democracy (Freedom House/Polity)", 
                   y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$fh_polity2)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, fh_polity2) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=fh_polity2, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(fh_polity2,individual.blame, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Democracy (Freedom House/Polity)", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{rdemo_freedom_plot, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk72, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Democracy index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

### Failed state index - (lower the better)

* ehkä tää menee teoreettisesti taas turhan kauas.. verrattuna noihin muihin indekseihin - tai en tiedä.

>The Failed States Index includes an examination of the pressures on states, their vulnerability to internal conflict and societal deterioration. The country ratings are based on the total scores of 12 indicators: Social Indicators – (1) Mounting Demographic Pressures; (2) Massive Movement of Refugees or Internally Displaced Persons creating Complex Humanitarian Emergencies; (3) Legacy of Vengeance-Seeking Group Grievance or Group Paranoia; and (4) Chronic and Sustained Human Flight. Economic Indicators – (5) Uneven Economic Development along Group Lines; and (6) Sharp and/or Severe Economic Decline. Political Indicators – (7) Criminalization and/or Delegitimization of the State; (8) Progressive Deterioration of Public Services; (9) Suspension or Arbitrary Application of the Rule of Law and Widespread Violation of Human Rights; (10) Security Apparatus Operates as a “State Within a State” (11) Rise of Factionalized Elites; and (12) Intervention of Other States or External Political Actors. For each indicator, the ratings are placed on a scale of 0 to 10, with 0 being the lowest intensity (most stable) and 10 being the highest intensity (least stable). The total score is the sum of the 12 indicators and is on a scale of 0-120

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk71, echo=FALSE, results='markup', cache=TRUE#
func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$ffp_fsi)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, ffp_fsi) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=ffp_fsi, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(ffp_fsi,social.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Failed State Index", 
                   y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$ffp_fsi)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, ffp_fsi) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=ffp_fsi, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(ffp_fsi,individual.blame, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Failed State Index", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk72, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Democracy index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

### Quality of governance index World Bank – Governance Indicators (a.k.a KKZ) - Voice and Accountability

* Tää näyttää teoreettisesti ehkä vahvimmalta ja korrelaatiot on paremman ja huonomman democracy indeksin välissä

>“Voice and Accountability” includes a number of indicators measuring various aspects of the political process, civil liberties and political rights. These indicators measure the extent to which citizens of a country are able to participate in the selection of governments. This category also includes indicators measuring the independence of the media, which serves an important role in monitoring those in authority and holding them accountable for their actions.

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk71, echo=FALSE, results='markup', cache=TRUE#
func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$wbgi_vae)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, wbgi_vae) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=wbgi_vae, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(wbgi_vae,social.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Quality of Governance Index", 
                   y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$wbgi_vae)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, wbgi_vae) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=wbgi_vae, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(wbgi_vae,individual.blame, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Quality of Governance Index", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk72, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Democracy index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

### HDI by UNDP

* tän inhimillisen kehityksen yhteys köyhyysasenteisiin lienee pitkälti tutkimatonta kenttää, joten vahvasta korrelaatiosta huolimatta teoreettisen yhteyden keksiminen on turhan keinotekoista.

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk71, echo=FALSE, results='markup', cache=TRUE#
func <- function(t3)
{
return(data.frame(COR = cor(t3$social.blame, t3$undp_hdi)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(social.blame, undp_hdi) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=undp_hdi, y=social.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(undp_hdi,social.blame, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Human Development Index", 
                   y = "Support for social blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$individual.blame, t3$undp_hdi)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individual.blame, undp_hdi) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=undp_hdi, y=individual.blame, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(undp_hdi,individual.blame, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Human Development Index", 
                   y = "Support for individual blame type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk72, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Democracy index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

## 2. Analysis of answer *other*

Monica peräsi etenkin tietoja siitä, mikä osuus kuuluu tähän meidän skippaamaan "other" kategoriaan ja millaisia tyyppejä ne ovat. Alla ensinnä meidän paperissa ollut neljän vaihtoehdon taulukko ja sen alla alkuperäinen kuuden vaihtoehdon taulukko ja kuvio.

Informatiivisin on ehkä taulukkojen alla olevat kuviot

### The original table from the RC19 paper
Sorted according to **social.blame**
```{rtable_orig, echo=FALSE, results='asis', warning=FALSE, cache=TRUE}
t <- data.frame(prop.table(svytable(~group1+cntry+q308.rec, d.df), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ q308.rec, value.var="Freq")
names(t2) <-  c("group1","country","social.blame","individual.blame","individual.fate","social.fate")
t2 <- t2[order(t2$group,-t2$social.blame),]
########
y <- data.frame(prop.table(svytable(~group1+q308.rec, d.df), 1)*100)
y$Freq <- round(y$Freq, 2)
y$country <- c("CEE Total","CIS Total","Western Europe Total")
y2 <-  dcast(y, group1+country ~ q308.rec, value.var="Freq")
names(y2) <-  c("group1","country","social.blame","individual.blame","individual.fate","social.fate")
#
#
co.var<-function(x)(100*sd(x)/mean(x)) 
#
func.CV <- function(t2)
{
return(data.frame(CV.sb = co.var(t2$social.blame)))
}
df.CV.sb <- ddply(t2, .(group1), func.CV)
#
func.CV <- function(t2)
{
return(data.frame(CV.ib = co.var(t2$individual.blame)))
}
df.CV.ib <- ddply(t2, .(group1), func.CV)
#
func.CV <- function(t2)
{
return(data.frame(CV.sf = co.var(t2$social.fate)))
}
df.CV.sf <- ddply(t2, .(group1), func.CV)
#
func.CV <- function(t2)
{
return(data.frame(CV.if = co.var(t2$individual.fate)))
}
df.CV.if <- ddply(t2, .(group1), func.CV)
##
df.CV <- join(df.CV.sb,df.CV.ib,by="group1")
df.CV <- join(df.CV,df.CV.sf,by="group1")
df.CV <- join(df.CV,df.CV.if,by="group1")
##
df.CV$group1 <- as.character(df.CV$group1)
df.CV$group1 <- recode(df.CV$group1, "'CEE'='CV CEE';
                        'CIS'='CV CIS';
                        'Western Europe'='CV Western Europe'")
##
names(df.CV) <- c("country","social.blame","individual.blame","social.fate","individual.fate")
df.CV$group1 <- c("CV","CV","CV")

t5 <- rbind(t2,y2,df.CV)
#t5 <- arrange(t5, group1,country)
t5$group1 <- NULL
t6 <- xtable(t5)
print.xtable(t6, type="html")
```

### Same table including both *dont know* and *not stated*
Sorted according to **dont.know**
* eli eniten tätä dont knowta on noissa hivenen autoritäärisissä maissa.

```{rtable_other, echo=FALSE, results='asis', warning=FALSE, cache=TRUE}
t <- data.frame(prop.table(svytable(~group1+cntry+q309, d.df), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ q309, value.var="Freq")
names(t2) <-  c("group1","country","social.blame","individual.blame","individual.fate","dont.know","social.fate","not.stated")
t2 <- t2[order(-t2$dont.know),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

### Barplot of population shares of **dont know** and **not stated**

* hankala sanoa mikä olisi yhteistä maille, joissa on korkeampi vastausosuus "other" vaihtoehdossa. Kokeilen vielä alla olisiko yhteys tuon **quality of governance** muuttujan kanssa.

```{rbar_other, echo=FALSE, fig.width=10, fig.height=15, warning=FALSE, cache=TRUE, message=FALSE}
t <- data.frame(prop.table(svytable(~group1+cntry+q309, d.df), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ q309, value.var="Freq")
names(t2) <-  c("group1","country","social.blame","individual.blame","individual.fate","dont.know","social.fate","not.stated")
t2$social.blame <- NULL
t2$individual.blame <- NULL
t2$social.fate <- NULL
t2$individual.fate <- NULL

t2$rank <- rank(t2$dont.know)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("dont.know","not.stated"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

### Onko **dont.know** ja **not.stated** osuudet yhteydessä Quality of governmentiin

* dont know on (-.54)
* not stated **ei ole** (.03)

```{rother_qog, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

t2$group1 <- rep("postsoc", 26)

t2 <- rename.vars(t2, from="country", to="cntry")
t3 <- join(t2,df.qog,by="cntry")

#chunk71, echo=FALSE, results='markup', cache=TRUE#

func <- function(t3)
{
return(data.frame(COR = cor(t3$dont.know, t3$wbgi_vae)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(dont.know, wbgi_vae) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=wbgi_vae, y=dont.know, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(wbgi_vae,dont.know, label = COR, group=COR), 
            size=5, color="black") +
              labs(x = "Quality of Governance Index", 
                   y = "Support for dont know type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))

func <- function(t3)
{
return(data.frame(COR = cor(t3$not.stated, t3$wbgi_vae, use="na.or.complete")))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(not.stated, wbgi_vae) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=wbgi_vae, y=not.stated, color=group1, label=cntry)) +
  geom_text(size=3, hjust=0.5, vjust=1.5) +
  geom_point(size=3) +
    scale_colour_manual(values=cbPalette) +
  geom_smooth(method=lm, se=TRUE) +
  theme(legend.position="top") +
  geom_text(data=cnames, aes(wbgi_vae,not.stated, label = COR, group=COR), 
            size=5, color="black")+
              labs(x = "Quality of Governance Index", 
                   y = "Support for not stated type of explanation (%)") +
   theme(axis.title.x = element_text(size=11),
        axis.title.y  = element_text(angle=90, size=11))
```

```{rother_qog_scatter, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk72, echo=FALSE, results='markup', dev='my_pdf', fig.ext='pdf', fig.cap="Democracy index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty", fig.pos="H", cache=TRUE
grid.newpage()
pushViewport(viewport(layout=grid.layout(1,2)))
print(plot1, vp=viewport(layout.pos.row=1, layout.pos.col=1))
print(plot2, vp=viewport(layout.pos.row=1, layout.pos.col=2))
```

### Who are the **dont know** and **not stated** by country??

Eli tässä kaksi vastausta on yhdistetty. Ensin taulukko, jossa ko. ryhmän väestöosuudet maittain.

```{rother_table, echo=FALSE, cache=TRUE, results='asis'}
t2$not.stated <- recode(t2$not.stated, "NA=0")
t2$other <- t2$dont.know + t2$not.stated

tbl <- subset(t2, select=c("cntry","other"))
tbl <- arrange(df=tbl, -other)

print.xtable(xtable(tbl), type="html")
```

#### In terms of Age groups
* other-tyyppien ikäjakauma vaihtelee helvetisti maittain. länsimaissa näyttää olevan enemmän vanhoja, mut postsos maissa ei selkeitä ryhmiä

```{rother_age, echo=FALSE, cache=TRUE, results='asis'}
df.other <- subset(df, q309 %in% c('Dont know','Not stated'))

d.df.other <- svydesign(id = ~SerialID, 
                  weights = ~weight, 
                  data = df.other)

t <- data.frame(prop.table(svytable(~group1+cntry+ageclass, d.df.other), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ ageclass, value.var="Freq")
names(t2) <-  c("group1","country","a_29_or_younger","b_30_to_44","c_45_to_59","d_60_and_older")
#t2 <- t2[order(-t2$a) 29 or younger),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

```{rbar_other_age, echo=FALSE, fig.width=15, fig.height=10, warning=FALSE, cache=TRUE}
t2$rank <- rank(t2$b_30_to_44)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("a_29_or_younger","b_30_to_44","c_45_to_59","d_60_and_older"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

#### In terms of education
* Liettuaa, Unkaria ja Romaniaa lukuunottamatta korkeammin koulutetut raju enemmistö näissä "other"-tyypeissä. Ruotissakin 96 vs. 4. Tarkoittaako sitä, että ko. kysymys vaatii koulutusta jotta siihen voi vastata????

```{rother_edu2, echo=FALSE, cache=TRUE, results='asis'}
t <- data.frame(prop.table(svytable(~group1+cntry+edu2, d.df.other), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ edu2, value.var="Freq")
names(t2) <-  c("group1","country","higher","no or compulsory")
t2 <- t2[order(-t2$higher),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

```{rbar_other_edu, echo=FALSE, fig.width=15, fig.height=10, warning=FALSE, cache=TRUE}
t2$rank <- rank(t2$higher)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("higher","no or compulsory"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

#### In terms of perceived income
* Tulojen osalta vähemmän selkeä. Tai siis kovin epäselvä.

```{rother_income2, echo=FALSE, cache=TRUE, results='asis'}
t <- data.frame(prop.table(svytable(~group1+cntry+income2, d.df.other), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ income2, value.var="Freq")
names(t2) <-  c("group1","country","High","Low")
t2 <- t2[order(-t2$High),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

```{rbar_other_income, echo=FALSE, fig.width=15, fig.height=10, warning=FALSE, cache=TRUE}
t2$rank <- rank(t2$High)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("High","Low"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

### Yhteenveto **other** tyypeistä
* niiden määrä vaihtelee Tajikistanin 22.9 ja Slovenian 4.12 välillä.
* Ei mielekkäitä maantieteellisia korrelaatioita -> Tatsikistanin (22.9) ja naapurimaa Kirgisia (4.27) ja naapurimaa Uzbekistan (12.46) tai Baltia: Viro (7.16), Latvia(14.50) ja Liettua (8.41) tai länsimaat Ruotti(13.62), Saksa ja Ranska (7), Italia (5)
* länsimaissa enemmän ikääntyneitä vastaajia, postsos ei selkeä
* korkeakoulutetut rajusti enemmistönä valtaosassa maita (pl. liettua, hunkari, romania)
* koetun tulotason osalta tulokset epäselviä

```{rnotstated_age, echo=FALSE, cache=TRUE, results='asis', eval=FALSE}
df.notstated <- subset(df, q309 == 'Not stated')

d.df.notstated <- svydesign(id = ~SerialID, 
                  weights = ~weight, 
                  data = df.notstated)

t <- data.frame(prop.table(svytable(~group1+cntry+ageclass, d.df.notstated), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ ageclass, value.var="Freq")
names(t2) <-  c("group1","country","a_29_or_younger","b_30_to_44","c_45_to_59","d_60_and_older")
#t2 <- t2[order(-t2$a) 29 or younger),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

```{rbar_notstated_age, echo=FALSE, fig.width=15, fig.height=10, warning=FALSE, cache=TRUE, eval=FALSE}
t2$rank <- rank(t2$b_30_to_44)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("a_29_or_younger","b_30_to_44","c_45_to_59","d_60_and_older"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

```{rnotstated_edu2, echo=FALSE, cache=TRUE, results='asis', eval=FALSE}
t <- data.frame(prop.table(svytable(~group1+cntry+edu2, d.df.notstated), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ edu2, value.var="Freq")
names(t2) <-  c("group1","country","higher","no or compulsory")
t2 <- t2[order(-t2$higher),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

```{rbar_notstated_edu, echo=FALSE, fig.width=15, fig.height=10, warning=FALSE, cache=TRUE, eval=FALSE}
t2$rank <- rank(t2$higher)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("higher","no or compulsory"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

```{rnotstated_income2, echo=FALSE, cache=TRUE, results='asis', eval=FALSE}
t <- data.frame(prop.table(svytable(~group1+cntry+income2, d.df.notstated), 2)*100)
t$Freq <- round(t$Freq, 2)
t <- subset(t, Freq > 0)
t2 <-  dcast(t, group1+cntry ~ income2, value.var="Freq")
names(t2) <-  c("group1","country","High","Low")
t2 <- t2[order(-t2$High),]
########
t2$group1 <- NULL
t6 <- xtable(t2)
print.xtable(t6, type="html")
```

```{rbar_notstated_income, echo=FALSE, fig.width=15, fig.height=10, warning=FALSE, cache=TRUE, eval=FALSE}
t2$rank <- rank(t2$High)

t2.l <- melt(data=t2, id.vars=c("country","rank"), measure.vars=c("High","Low"))

ggplot(t2.l, aes(x=rank,y=value,group=variable,fill=variable,label=value)) + 
  geom_bar(stat="identity", position="dodge") + 
  geom_text(position = position_dodge(width=1), size=3, angle=0) +
  theme(legend.position="top") +
  scale_x_discrete(labels=t2.l$country, breaks=t2.l$rank) + 
  #theme(axis.text.x=element_text(angle=90)) +
  coord_flip()

########
```

## 3. Multilevel modelling

Ja sitten monitasomalleihin.

This exercise presents the results of logistic multilevel models with different contextual variable, namely the differences between usage od 1) democracy index by Economist intelligent unit, 2) Failed state index by ?? , 3)  and 4) HDI by UNDP

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk10_1, echo=FALSE, results='hide'#
df <- subset(df, group1 != 'Western Europe')
library(memisc)

#chunk10, echo=FALSE, results='hide', cache=TRUE#
summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.8 <- lmer(pov.log.social ~ Gini+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
                       family = binomial, data=df))
summary(bisoc.10 <- lmer(pov.log.social ~ Democracy2011+(1|cntry),
                       family = binomial, data=df))
```

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

#chunk12, echo=FALSE, results='hide', cache=TRUE#
summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
                       family = binomial, data=df))
summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
                       family = binomial, data=df))
summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
                       family = binomial, data=df))
summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
                       family = binomial, data=df))
summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
                       family = binomial, data=df))
summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
                       family = binomial, data=df))
summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
                       family = binomial, data=df))
summary(biind.8 <- lmer(pov.log.individual ~ Gini+(1|cntry),
                       family = binomial, data=df))
summary(biind.9 <- lmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
                       family = binomial, data=df))
summary(biind.10 <- lmer(pov.log.individual ~ Democracy2011+(1|cntry),
                       family = binomial, data=df))
```

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk19, echo=FALSE, results='hide', cache=TRUE#
# modelling without and with weights
summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                       family = binomial, data=df))

summary(soc.1 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+(1|cntry),
                       family = binomial, data=df))

# koska grouping oli turha se pois. _g merkinnällä kuitenkin mallit
# joissa ryhmittely, jotta saadaan alkuperäinen taulukko myös esille.
summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  (1|cntry), family = binomial, data=df))

summary(soc.2_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+(1|cntry), family = binomial, data=df))

summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+(1|cntry),family = binomial, data=df))

summary(soc.3_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+(1|cntry),family = binomial, data=df))

summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+(1|cntry),family = binomial, data=df))

summary(soc.4_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+gdpchange07.10+(1|cntry),family = binomial, data=df))

summary(soc.5 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+Democracy2011+(1|cntry),family = binomial, data=df))

summary(soc.5_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+gdpchange07.10+Democracy2011+(1|cntry),family = binomial, data=df))
# postsoc+ was left out due to errors...

```

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, eval=FALSE}
#memisc_lme4, echo=FALSE, results='hide', cache=TRUE#
###########################################################################
## ----------------------------------------------------------------------------
## Author: Jason Morgan (borrowing heavily from code contained in Martin Elff's
##         memisc package).
##
## Notes:  Additional methods for mtable formatting of lme4 model
##         objects. Requires that the memisc package be loaded prior to
##         sourcing these functions.
## ----------------------------------------------------------------------------

setSummaryTemplate(mer = c("Log-likelihood" = "($logLik:f#)",
                     "Deviance" = "($deviance:f#)",
                     "AIC" = "($AIC:f#)",
                     "BIC" = "($BIC:f#)",
                     "N" = "($N:d)",
                     "Groups" = "($Groups:d)"))

getSummary.mer <- function (obj, alpha = 0.05, ...) {
  smry <- summary(obj)
  coef <- smry@coefs
  lower <- qnorm(p = alpha/2, mean = coef[, 1], sd = coef[,2])
  upper <- qnorm(p = 1 - alpha/2, mean = coef[, 1], sd = coef[,2])
  if (ncol(smry@coefs) == 3) {
    p <- (1 - pnorm(smry@coefs[, 3])) * 2
    coef <- cbind(coef, p, lower, upper)
  }
  else {
    coef <- cbind(coef, lower, upper)
  }
  RE <- smry@REmat
  ranef <- cbind(as.numeric(RE[,3]), as.numeric(RE[,4]), NA,NA,NA,NA)
  rownames(ranef) <- paste("Ranef", RE[,1], sep = " - ")
  coef <- rbind(coef, ranef)
  colnames(coef) <- c("est", "se", "stat", "p", "lwr", "upr")

  ## Factor levels.
  xlevels <- list()
  Contr <- names(attr(model.matrix(obj), "contrasts"))
  for (c in Contr) xlevels[[c]] <- levels(obj@frame[,c])

  ## Model fit statistics.
  ll <- logLik(obj)[1]
  deviance <- deviance(obj)
  AIC <- AIC(obj)
  BIC <- BIC(obj)
  N <- as.numeric(smry@dims["n"])
  G <- as.numeric(smry@ngrps)
  sumstat <- c(logLik = ll, deviance = deviance, AIC = AIC,
               BIC = BIC, N = N, Groups = G)

  ## Return model summary.
  list(coef = coef, sumstat = sumstat,
       contrasts = attr(model.matrix(obj), "contrasts"),
       xlevels = xlevels, call = obj@call)

```

### Social blame -  Original

```{rmultiorigsb, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk21, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                       "Model sb2"=soc.2,
                       "Model sb3"=soc.3_g,
                       "Model sb4"=soc.4_g,
                       "Model sb5"=soc.5_g)

```

### Social blame - original **without grouping**

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk21, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                      # "Model sb2"=soc.2,
                       "Model sb3"=soc.3,
                       "Model sb4"=soc.4,
                       "Model sb5"=soc.5)

```

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk20, echo=FALSE, results='hide', cache=TRUE#
summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
                       family = binomial, data=df))

summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+(1|cntry),
                       family = binomial, data=df))

summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+(1|cntry), family = binomial, data=df))

summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+(1|cntry),family = binomial, data=df))

summary(ind.3_g <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+(1|cntry),family = binomial, data=df))

summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+(1|cntry),
                       family = binomial, data=df))

summary(ind.4_g <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+gdpchange07.10+(1|cntry),
                       family = binomial, data=df))

summary(ind.5 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+Democracy2011+(1|cntry),
                       family = binomial, data=df))

summary(ind.5_g <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+gdpchange07.10+Democracy2011+(1|cntry),
                       family = binomial, data=df))
# postsoc+ was left out due to errors...

```

### Individual blame - original
```{rmultiorigib, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

  mtable("Empty model"=ind.0,
                       "Model ib1"=ind.1,
                       "Model ib2"=ind.2,
                       "Model ib3"=ind.3_g,
                       "Model ib4"=ind.4_g,
                       "Model ib5"=ind.5_g)

```

### Individual blame - original **without grouping**

```{r, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk22, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=ind.0,
                       "Model ib1"=ind.1,
                      # "Model ib2"=ind.2,
                       "Model ib3"=ind.3,
                       "Model ib4"=ind.4,
                       "Model ib5"=ind.5)
  #,
#                "[(]Intercept[)]"="\\\\emph{Intercept}","[$]"="\\\\$",
 #               Ranef="Random effect",
  #              incsour3="Transfer dep.",
    #            edu2="Low educ.",
     #           income2="Perceived income level",
      #          past.diff="Income compared to past",
       #         future.diff="Income compared to future",
        #        crise="Effect of financial crises",
         #       Gini="Gini coefficient",
          #      gdpchange07.10="Change in GDP 2007-2010",
           #     Democracy2011="Democracy Index 2011",
            #    group1="Country group",
             #   gsub=TRUE), ddigits=3)

```

### Social blame - with Failed State Index

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
summary(soc.5_ffp_fsi <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+ffp_fsi+(1|cntry),family = binomial, data=df))
# postsoc+ was left out due to errors...
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                       "Model sb3"=soc.3,
                       "Model sb4"=soc.4,
                       "Model sb5"=soc.5_ffp_fsi)

```

### Individual blame - with Failed State Index

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
summary(ind.5_ffp_fsi <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+ffp_fsi+(1|cntry),
                       family = binomial, data=df))

```

```{r, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk22, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=ind.0,
                       "Model ib1"=ind.1,
                       "Model ib3"=ind.3,
                       "Model ib4"=ind.4,
                       "Model ib5"=ind.5_ffp_fsi)

```

### Social blame - with Quality of Governance

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
summary(soc.5_wbgi_vae <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+wbgi_vae+(1|cntry),family = binomial, data=df))
# postsoc+ was left out due to errors...
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                       "Model sb3"=soc.3,
                       "Model sb4"=soc.4,
                       "Model sb5"=soc.5_wbgi_vae)

```

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
summary(ind.5_wbgi_vae <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+wbgi_vae+(1|cntry),
                       family = binomial, data=df))

```

### Individual blame - with Quality of Governance

```{r, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk22, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=ind.0,
                       "Model ib1"=ind.1,
                       "Model ib3"=ind.3,
                       "Model ib4"=ind.4,
                       "Model ib5"=ind.5_wbgi_vae)

```

### Social blame - with HDI

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
summary(soc.5_hdi <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+undp_hdi+(1|cntry),family = binomial, data=df))
# postsoc+ was left out due to errors...
```

```{r, fig.width=15, fig.height=10, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}

mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                       "Model sb3"=soc.3,
                       "Model sb4"=soc.4,
                       "Model sb5"=soc.5_hdi)

```

### Individual blame - with HDI

```{r, results='hide', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
summary(ind.5_hdi <- lmer(pov.log.individual ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+undp_hdi+(1|cntry),
                       family = binomial, data=df))

```

```{r, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk22, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=ind.0,
                       "Model ib1"=ind.1,
                       "Model ib3"=ind.3,
                       "Model ib4"=ind.4,
                       "Model ib5"=ind.5_hdi)

```

## 4. Multilevel modelling2

Kirjoitin aluksi näin:

>Edellisissä malleissa selitettävä muuttuja on koodattu siten, että 
>1. vain ko. tavalla vastanneet (social blame tai individual blame)
>0. social fate + individual fate + (social blame tai individual blame)

>Tästä on seurannut mm. se, että tyhjään malliin verrattuna n on pudonnut aina reilulla 6000 kotitaloudella.

>Seuraavissa on käytetty alkuperäistä muuttujaa joilloin mukana ovat myös **dont know** ja **not stated** vaihtoehdon valinneet siis koodattuna arvolle 0. N määrä nousee siten...

mutta pienen kokeilun jälkeen selvisi että ongelma ei ole other-koodatut, he ovat analyysissä mukana, vaan: n ei nouse, sillä n-määrän lasku on peräisin suurelta osin muuttujasta **future.diff** sekä vähemmän muuttujista **past.diff** että **crise**. Se ei liity mitenkään selitettävään köyhyysasennemuuttujaan.

Eli jos n. määrää haluaa nostaa, niin

```{rmulti2_recode, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE}
#chunk10_1, echo=FALSE, results='hide'#
df <- subset(df, group1 != 'Western Europe')
df$cntry <- factor(df$cntry)

###############################
# $pov.log.social
################################

df$pov.log.social <- recode(df$q309, "'Because of injustice in our society'=1;
                               else=0")
################################
# $pov_log_individual
################################
df$pov.log.individual <- recode(df$q309, "'Because of laziness and lack of willpower'=1;
                               else=0")

summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                       family = binomial, data=df))

summary(soc.01 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+crise+(1|cntry),
                       family = binomial, data=df))

summary(soc.1 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+(1|cntry),
                       family = binomial, data=df))
```

### Puuttuvat arvot muuttujissa **future.diff**, **past.diff** että **crise**.

```{rmissing,  echo=FALSE, fig.width=10, fig.height=5, warning=FALSE, cache=TRUE, message=FALSE}
#
library(VIM)
spineMiss(df[, c("cntry", "future.diff")])
spineMiss(df[, c("cntry", "past.diff")])  
spineMiss(df[, c("cntry", "crise")])
```

### Social blame -  Original

```{rmulti2_sb, results='markup', echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, eval=FALSE}
#chunk19, echo=FALSE, results='hide', cache=TRUE#
# modelling without and with weights
summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                       family = binomial, data=df))

summary(soc.1 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+(1|cntry),
                       family = binomial, data=df))

# koska grouping oli turha se pois. _g merkinnällä kuitenkin mallit
# joissa ryhmittely, jotta saadaan alkuperäinen taulukko myös esille.
summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  (1|cntry), family = binomial, data=df))

summary(soc.2_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+(1|cntry), family = binomial, data=df))

summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+(1|cntry),family = binomial, data=df))

summary(soc.3_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+(1|cntry),family = binomial, data=df))

summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+(1|cntry),family = binomial, data=df))

summary(soc.4_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+gdpchange07.10+(1|cntry),family = binomial, data=df))

summary(soc.5 <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  Gini+gdpchange07.10+Democracy2011+(1|cntry),family = binomial, data=df))

summary(soc.5_g <- lmer(pov.log.social ~ incsour3+edu2+income2+
  past.diff+future.diff+crise+
  group1+Gini+gdpchange07.10+Democracy2011+(1|cntry),family = binomial, data=df))
# postsoc+ was left out due to errors...

```

```{rmultiorigsb_2, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, eval=FALSE}
#chunk21, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                       "Model sb2"=soc.2,
                       "Model sb3"=soc.3_g,
                       "Model sb4"=soc.4_g,
                       "Model sb5"=soc.5_g)

```

<!--

### Social blame - original **without grouping**

-->

```{r, echo=FALSE, warning=FALSE, message=FALSE, cache=TRUE, eval=FALSE}
#chunk21, echo=FALSE, results='hide', cache=TRUE#
#toLatex(relabel(
  mtable("Empty model"=soc.0,
                       "Model sb1"=soc.1,
                      # "Model sb2"=soc.2,
                       "Model sb3"=soc.3,
                       "Model sb4"=soc.4,
                       "Model sb5"=soc.5)

```

### 2013-10-30 code/dv_tables.R

@@ -1,0 +1,90 @@
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+library(survey)
+d.df <- svydesign(id = ~SerialID, 
+                  weights = ~weight, 
+                  data = df)
+
+
+
+t <- data.frame(prop.table(svytable(~group_general+cntry+q309, d.df), 2)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+
+### testailua
+maa <- subset(t, q309 == "Not stated")
+maa <- maa[order(maa$group_general),]
+t$cntry <- factor(t$cntry, levels=maa$cntry)
+library(ggplot2)
+ggplot(t, aes(x=cntry,y=Freq,group=cntry)) +
+  geom_bar(stat="identity", position="dodge") +
+  facet_wrap(~q309) + coord_flip()
+####
+library(reshape2)
+t2 <-  dcast(t, group_general+cntry ~ q309, value.var="Freq")
+names(t2) <-  c("group_general","country","social.blame","individual.blame","individual.fate","dont.know","social.fate","not.stated")
+tbl2 <- t2[order(-t2$social.blame),]
+########
+tbl2$group_general <- NULL
+
+#---------------------------------------------------#
+#---------------------------------------------------#
+#---------------------------------------------------#
+#---------------------------------------------------#
+
+t <- data.frame(prop.table(svytable(~group_general+cntry+q309_rec, d.df), 2)*100)
+t$Freq <- round(t$Freq, 2)
+t <- subset(t, Freq > 0)
+t2 <-  dcast(t, group_general+cntry ~ q309_rec, value.var="Freq")
+names(t2) <-  c("group_general","country","social.blame","individual.blame","individual.fate","social.fate")
+t2 <- t2[order(t2$group,-t2$social.blame),]
+########
+y <- data.frame(prop.table(svytable(~group_general+q309_rec, d.df), 1)*100)
+y$Freq <- round(y$Freq, 2)
+y$country <- c("CEE Total","CIS Total","Western Europe Total")
+y2 <-  dcast(y, group_general+country ~ q309_rec, value.var="Freq")
+names(y2) <-  c("group_general","country","social.blame","individual.blame","individual.fate","social.fate")
+#
+#
+library(plyr)
[…]

### 2013-10-30 code/iv_indiv_table.R

@@ -1,0 +1,82 @@
+# text table
+
+Variable <- c("Transfer dependency","Low education","Perceived low income level","Income compared to past","Income compared to future","Effect of financial crises")
+Description <- c("Social or private transfers \n as main source of income","No or only compulsory level education","Perceived income level below the country median","Perceived change in income compared to situation in four years ago","Expected change in income over the next four years","Whether respondents perceives that their household has suffered the financial crisis great or fair amount")
+Varname <- c("q226m","q515","q227","q228","q229","q801")
+table <- data.frame(Variable,Description,Varname)
+
+# numeric table
+
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+library(survey)
+d.df <- svydesign(id = ~SerialID, 
+                  weights = ~weight, 
+                  data = df)
+# Country group
+t <- data.frame(prop.table(svytable(~group_general+q309_rec, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, group_general ~ q309_rec, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+cgroup <- t2
+# PAst difference
+t <- data.frame(prop.table(svytable(~past.diff+q309_rec, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, past.diff ~ q309_rec, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+past.diff <- t2
+# Future difference
+t <- data.frame(prop.table(svytable(~future.diff+q309_rec, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, future.diff ~ q309_rec, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+future.diff <- t2
+# Low education
+t <- data.frame(prop.table(svytable(~edu2+q309_rec, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
[…]

### 2013-10-30 code/iv_macro_table.R

@@ -1,0 +1,31 @@
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+tbl <- macro[, c(1,8:14)]
+tbl[,c(2:7)] <- round(tbl[,c(2:7)], 2)
+tbl <- tbl[order(tbl$undp_hdi), ]
+names(tbl) <- c("country","GDP Change 2007-1010","Failed States Index",
+                "Voice And Accountability (WB)","HDI","Democracy (Freedom House)",
+                "Gini UNU Wider","Country group")
+
+
+Data <- c("Ghange in total GDP 2007 – 2010",
+          "Gini coefficient",
+          "Voice and Accountability")
+Description <- c("Relative change from 2007 to 2010",
+                 "measures the distribution of income or consumption expenditure among individuals or households within an economy deviates from a perfectly equal distribution",
+                 "indicators measuring the political process, civil liberties and political rights. These indicators measure the extent to which citizens of a country are able to participate in the selection of governments. Indicators measuring the independence of the media is also included.")
+Source <- c("World Bank (Quality of Government Data)",
+            "World Bank (Quality of Government Data)",
+            "World Bank (Quality of Government Data)")
+Varname <- c("wdi_gdp",
+             "uw_gini",
+             "wbgi_vae")
+# Year <- c("2011",
+#            "Latest available gini coefficient based either on income or consumption",
+#            "2011")
+macrotable <- data.frame(Data,Description,Source,Varname)
+
+
+# Data & Description & Source & Year
+# & $  $ & \href{http://data.worldbank.org/data-catalog/world-development-indicators}{World Bank World Development Indicators} & 2011 \\
+# &  & \href{http://www.wider.unu.edu/research/Database/en_GB/database/}{UNU-WIDER World Income Inequality Database, Version 2.0c, May 2008} & 2002-2006 \\
+# & . & \href{http://en.wikipedia.org/wiki/Democracy_Index#2011_rankings}{Economist Intelligence Unit} & 2011 \\

### 2013-10-30 code/load_n_clean2013.R

@@ -1,0 +1,395 @@
+##################################################################
+##################################################################
+##################################################################
+# Life in transition survey 2
+# load and clean data
+
+# libraries
+library(car)
+library(stringr)
+library(foreign)
+
+# LITS 2
+# from web
+#temp <- tempfile()
+#download.file("http://www.ebrd.com/downloads/research/surveys/lits2.dta", temp)
+#lits2 <- read.dta(temp)
+# tai 
+#lits2 <- read.dta("~/workspace/lits/lits2.dta")
+#save.image("~/workspace/lits/lits2.RData")
+load("~/workspace/lits/lits2.RData")
+##################################################################
+##################################################################
+## Clean data ##
+##################################################################
+# different country groupings
+
+# remove whitespaces from country
+lits2$cntry <- str_trim(lits2$country_, side="right")
+lits2$cntry <- factor(lits2$cntry)
+
+lits2$socialism <- recode(lits2$cntry,
+                            "c('France','Germany','Great Britain',
+                            'Italy','Sweden')='Market economy';
+                            c('Czech Republic','Estonia','Hungary',
+                          'Bulgaria','Latvia','Lithuania','Poland',
+                          'Slovakia','Slovenia','Romania','Albania',
+                          'Bosnia','Bulgaria','Croatia','Macedonia',
+                          'Montenegro','Kosovo','Serbia','Armenia',
+                          'Azerbaijan','Belarus','Georgia','Kazakhstan',
+                          'Kyrgyzstan','Moldova','Mongolia','Russia',
+                          'Tajikistan','Ukraine','Uzbekistan')='Post-socialist';
+                          else=NA")
+lits2$group_analysis <- recode(lits2$cntry,
+                             "c('France','Germany','Great Britain',
+                             'Italy','Sweden')='Western Europe';
+                             c('Czech Republic','Estonia','Hungary','Bulgaria',
+                             'Latvia','Lithuania','Poland',
+                             'Slovakia','Slovenia','Romania')='CEE';
+                             c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
+                             'Kyrgyzstan','Moldova','Tajikistan',
+                             'Ukraine','Uzbekistan','Russia')='CIS';else=NA")
+# lits2$group_general <- recode(lits2$cntry,
+#                              "c('France','Germany','Great Britain',
+#                              'Italy','Sweden')='Western Europe';
[…]

### 2013-10-30 code/multi_bivariate.R

@@ -1,0 +1,46 @@
+setwd("~/workspace/lits/attrib/attrib_year2013/code")
+
+#########################################
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+library(lme4)
+summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.8 <- lmer(pov.log.social ~ uw_gini+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.10 <- lmer(pov.log.social ~ wbgi_vae+(1|cntry),
+                         family = binomial, data=df))
+
+summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
[…]

### 2013-10-30 code/multi_logit.R

@@ -1,0 +1,38 @@
+setwd("~/workspace/lits/attrib/attrib_year2013/code")
+
+#########################################
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+# malleista länkkärit pois!
+df.multi <- subset(df, group_general %in% c("CEE","CIS"))
+
+library(lme4)
+summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
+                      family = binomial, data=df.multi))
+summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                         past.diff+future.diff+crise+(1|cntry),
+                       family = binomial, data=df.multi))
+summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+(1|cntry),family = binomial, data=df.multi))
+summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+(1|cntry),family = binomial, data=df.multi))
+summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),family = binomial, data=df.multi))
+
+
+summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
+                      family = binomial, data=df.multi))
+summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+(1|cntry),
+                      family = binomial, data=df.multi))
+summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+(1|cntry),family = binomial, data=df.multi))
+summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+(1|cntry),family = binomial, data=df.multi))
+summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),family = binomial, data=df.multi))

### 2013-10-30 code/multilevel_models.R

@@ -1,0 +1,121 @@
+library(xtable)
+library(apsrtable)
+library(survey)
+library(lme4)
+#########################################
+#########################################
+setwd("~/workspace/lits/attrib/attrib_year2013/code")
+#########################################
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+
+#########################################
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+
+summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.8 <- lmer(pov.log.social ~ uw.gini+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
+                        family = binomial, data=df))
+summary(bisoc.10 <- lmer(pov.log.social ~ wbgi.vae+(1|cntry),
+                         family = binomial, data=df))
+
+
+mtable(bisoc.1,bisoc.2,
+       bisoc.3,bisoc.4,
+       bisoc.5,bisoc.6,
+       bisoc.7,bisoc.8,
+       bisoc.9,bisoc.10)
+
+
+summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
+                        family = binomial, data=df))
+summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
[…]

### 2013-10-30 multi_logit.R

@@ -1,0 +1,35 @@
+setwd("~/workspace/lits/attrib/attrib_year2013/code")
+
+#########################################
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+
+summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
+                      family = binomial, data=df))
+summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                         past.diff+future.diff+crise+(1|cntry),
+                       family = binomial, data=df))
+summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw.gini+(1|cntry),family = binomial, data=df))
+summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
+summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
+
+
+summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
+                      family = binomial, data=df))
+summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+(1|cntry),
+                      family = binomial, data=df))
+summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw.gini+(1|cntry),family = binomial, data=df))
+summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
+summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))

### 2013-10-30 index.Rmd
The structure of the document is set

@@
 % Markus Kainu & Mikko Niemelä
-% Jan 2013
+% Nov 2013
 
-<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
-<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>
+<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link> <link href='http://fonts.googleapis.com/css?family=Oswald' rel='stylesheet' type='text/css'>
 
@@
 
-cd ~/workspace/lits/attrib/attrib_year2013
+cd ~/workspace/lits/attrib/attrib_year2013/figures
 lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
@@
 
-******
+# Introduction
 
-<center>
-<table>
-<tr>
-<td>[Etusivu](index.html)</td>
-<td>[Descriptive analysis](code/descriptive_analysis.html)</td>
-<td>[Output tables in html](code/multilevel_models.html)</td>
-<td>[Output tables in pdf](code/multilevel_models.pdf)</td>
-<td>[Article in pdf](article2013.pdf)</td>
-<td>[Article in html](article2013.html)</td>
-<td>[Presentation slides](slides/slides_attrib.html)</td>
-</tr>
-</table>
-</center>
+# Data Analysis
 
-*******
+## Cleaning the data & missing data
 
-# Etusivu
+html-ouput                                                    source code
+-------------------------------------------------------       -----------------------------
+[loading and cleaning](code/descriptive_analysis.html)        [source](code/descriptive_analysis.Rmd)
+[missing data & diagnostics](code/descriptive_analysis.html)  [source](code/descriptive_analysis.Rmd)
+
+## Descriptive analysis
+
+html-ouput                                                source code
+-------------------------------------------------------   -----------------------------
+[Descriptive analysis](code/descriptive_analysis.html)    [source](code/descriptive_analysis.Rmd)
+
+## Modelling the data
+
+html-ouput                                                source code
+-------------------------------------------------------   -----------------------------
+[Output tables in html](code/multilevel_models.html)      [source](code/multilevel_models.Rmd)
+[Output tables in pdf](code/multilevel_models.pdf)        [source](code/multilevel_models.Rmd)
+
+# Paper
+
+html-output                                               source code
+-------------------------------------------------------   -----------------------------
+[Article in html](article2013.html)                       [Article in pdf](article2013.pdf)
+
+# Presentation slides
+
+html-ouput                                                source code
+-------------------------------------------------------   -----------------------------
+[Presentation slides](slides/slides_attrib.html)          [source](code/multilevel_models.Rmd)

### 2013-10-31 descriptive_analysis.Rmd
Converting the code into RMarkdown and cleaning it..

--- 2013-10-30 code/descriptive_analysis.Rmd
@@
+% Descriptive analysis
+
+- [back to index](index.html)
+
+`generated at:`
+```{r, echo=FALSE}
+Sys.time()
+```
+---
+
+<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
+
 ```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013/code/")
+setwd("~/workspace/lits/attrib/attrib_year2013/")
 library(kaRski)
 knitpandoc("descriptive_analysis")
-
 ```
 
-```{ropts, echo=FALSE}
-opts_chunk$set(echo=FALSE, dev='svg', eval=TRUE, cache=TRUE, fig.height=10, message=FALSE, warning=FALSE)
+```{roptsattrib, echo=FALSE}
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
 ```
-
-<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
 
@@
 
-![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](../figure/scatter_gdp.png)
+![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)
 
-![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](../figure/scatter_gini.png)
+![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)
 
-![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](../figure/scatter_wbgi.png)
+![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)

### 2013-10-31 index.Rmd
Converting the code into RMarkdown and cleaning it..

@@
 % Nov 2013
+
+`generated at:`
+```{r, echo=FALSE}
+Sys.time()
+```
+---
 
@@
 
+```{roptsattrib, echo=FALSE}
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
+```
+
 # Introduction
+
+This site gathers together all the data analysis behind the actual paper.
+Data analysis is broken down in three parts and each part is presented in .html format for human eyes and in text source including the R-code required for reproducing the analysis.
+As both the [micro data](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml) and [macro data](http://www.qog.pol.gu.se/data/) are available as open data, anyone can rerun these analysis.
+
+Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.
 
@@
 -------------------------------------------------------   -----------------------------
-[Descriptive analysis](code/descriptive_analysis.html)    [source](code/descriptive_analysis.Rmd)
+[Descriptive analysis](descriptive_analysis.html)         [source](descriptive_analysis.Rmd)
+[Graphics in the paper]()                                 [source](descriptive_analysis.Rmd)
 
@@
 
-html-output                                               source code
--------------------------------------------------------   -----------------------------
-[Article in html](article2013.html)                       [Article in pdf](article2013.pdf)
+- [Article draft in .html](article2013.html)
+- [Article draft in .pdf](article2013.pdf)
+- [Article draft in .epub](article2013.epub)
 
@@
 
-html-ouput                                                source code
--------------------------------------------------------   -----------------------------
-[Presentation slides](slides/slides_attrib.html)          [source](code/multilevel_models.Rmd)
+- HSE Moscow - November 12, 2013
+- Jatkokoulutusseminaari Turku November 7, 2013
+- [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)

### 2013-10-31 loadClean.Rmd
Converting the code into RMarkdown and cleaning it..

% Loading and cleaning the data

- [back to index](index.html)

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(kaRski)
knitpandoc("loadClean")
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
```

# Loading the data from EBRD

```{rLoadClean1}

##################################################################
##################################################################
##################################################################
# Life in transition survey 2
# load and clean data

# libraries
library(car)
library(stringr)
library(foreign)

# LITS 2
# from web
#temp <- tempfile()
#download.file("http://www.ebrd.com/downloads/research/surveys/lits2.dta", temp)
#lits2 <- read.dta(temp)
# tai

load("~/workspace/data/lits/lits2.RData")
##################################################################
##################################################################

```

# Munging the data

## Country names & groupings

```{rLoadClean2}
##################################################################
# different country groupings

# remove whitespaces from country
lits2$cntry <- str_trim(lits2$country_, side="right")
lits2$cntry <- factor(lits2$cntry)

lits2$socialism <- recode(lits2$cntry,
                            "c('France','Germany','Great Britain',
                            'Italy','Sweden')='Market economy';
                            c('Czech Republic','Estonia','Hungary',
                          'Bulgaria','Latvia','Lithuania','Poland',
                          'Slovakia','Slovenia','Romania','Albania',
                          'Bosnia','Bulgaria','Croatia','Macedonia',
                          'Montenegro','Kosovo','Serbia','Armenia',
                          'Azerbaijan','Belarus','Georgia','Kazakhstan',
                          'Kyrgyzstan','Moldova','Mongolia','Russia',
                          'Tajikistan','Ukraine','Uzbekistan')='Post-socialist';
                          else=NA")
lits2$group_analysis <- recode(lits2$cntry,
                             "c('France','Germany','Great Britain',
                             'Italy','Sweden')='Western Europe';
                             c('Czech Republic','Estonia','Hungary','Bulgaria',
                             'Latvia','Lithuania','Poland',
                             'Slovakia','Slovenia','Romania')='CEE';
                             c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
                             'Kyrgyzstan','Moldova','Tajikistan',
                             'Ukraine','Uzbekistan','Russia')='CIS';else=NA")

```

##

```{rLoadClean3}
################################
# reason for poverty
################################

################################
lits2$q309 <- as.factor(str_replace(lits2$q309, "Don't know", "Dont know"))

# All options

lits2$q309 <- recode(lits2$q309, "'Because of injustice in our society'='Social Blame';
                         'Because of laziness and lack of willpower'='Individual Blame';
                         'Because they have been unlucky'='Individual Fate';
                         'It is an inevitable part of modern life'='Social Fate';
                         'Not stated'='Not stated';
                         'Dont know'='Dont know'")

lits2$q309 <- factor(lits2$q309, levels=c("Social Blame","Individual Blame",
                                                  "Individual Fate","Social Fate",
                                          "Not stated","Dont know"))

# not stated and dont know coded NA

lits2$q309_rec <- recode(lits2$q309, "'Because of injustice in our society'='Social Blame';
                         'Because of laziness and lack of willpower'='Individual Blame';
                         'Because they have been unlucky'='Individual Fate';
                         'It is an inevitable part of modern life'='Social Fate';
                         c('Not stated','Dont know')=NA")

lits2$q309_rec <- factor(lits2$q309_rec, levels=c("Social Blame","Individual Blame",
                                                  "Individual Fate","Social Fate"))

################################
# $poverty - muuttuja
################################
#lits2$poverty <- recode(lits2$q309_rec, "'It is an inevitable part of modern life'=NA")
lits2$poverty <- lits2$q309_rec

################################
# $pov.log.social
################################
lits2$pov.log.social <- recode(lits2$q309_rec, "'Social Blame'=1;
                               else=0")
################################
# $pov_log_individual
################################
lits2$pov.log.individual <- recode(lits2$q309_rec, "'Individual Blame'=1;
                               else=0")

```

##

```{rLoadClean4}

################################
# $sex - muuttuja
################################
lits2$sex <- lits2$q102_1

################################
# $income
################################

lits2$income <- factor(recode(lits2$q227, 
                           "1:3='Low income'; 
                          4:7='Middle income'; 
                           8:10='High income'; 
                           else=NA"))

lits2$income <- factor(lits2$income, levels=c("Low income","Middle income","High income"))

lits2$income2 <- factor(recode(lits2$q227, 
                              "1:5='Low'; 
                              6:10='High'; 
                              else=NA"))

lits2$income <- factor(lits2$income2, levels=c("Low","High"))

# median(lits2)
#svyquantile(~q227, d.df, quantile = c(0.25, 0.5, 0.75), ci = TRUE)

lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
################################
# $future
################################
lits2$future <- factor(recode(lits2$q229, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

lits2$q229_rec <- recode(lits2$q229, "-100:0=NA")

lits2$future.diff[lits2$q227_rec > lits2$q229_rec] <- 'worse'
lits2$future.diff[lits2$q227_rec <= lits2$q229_rec] <- 'same or better'
lits2$future.diff <- factor(lits2$future.diff)

################################
# $past
################################
lits2$past <- factor(recode(lits2$q228, 
                              "1:3='a) low income'; 
                          4:7='b) middle income'; 
                              8:10='c) high income'; 
                              else=NA"))

lits2$q228_rec <- recode(lits2$q228, "-111:0=NA")

lits2$past.diff[lits2$q227_rec < lits2$q228_rec] <- 'worse'
lits2$past.diff[lits2$q227_rec >= lits2$q228_rec] <- 'same or better'
lits2$past.diff <- factor(lits2$past.diff)

################################
# $age
################################

lits2$ageclass <- factor(recode(lits2$q104a_1, 
                              "1:29='29 or younger'; 
                          30:44='30 to 44'; 
                              45:59='45 to 59';
                              60:120='60 and older';
                                else=NA"))

lits2$ageclass <- factor(lits2$ageclass, levels=c("29 or younger","30 to 44","45 to 59","60 and older"))

lits2$age <- lits2$q104a_1

################################
# $incsour tulonlähde
################################
lits2$q226m_txt <- recode(lits2$q226m, 
                          "1='Salary or wages in cash or in kind';
                                2='Income from self-employment';
                          3='Sales or bartering of farm products';
                          4='Pensions';5='Benefits from the state';
                          6='Help from relatives or friends';7='Other'")

lits2$incsour3 <- factor(recode(lits2$q226m_txt, 
                                "c('Benefits from the state',
                                'Help from relatives or friends')='Dependent'
                                ;else='Not.dependent'"))

################################
# $edu
################################

lits2$edu2 <- factor(recode(lits2$q515, "c(1,2)='no or compulsory';
                           else='higher'"))

#################
# $crise - kriisikokemus
#################
lits2$q801 <- as.factor(str_replace(lits2$q801, "Don't know", "DONT KNOW"))
summary(lits2$q801)
lits2$q801 <- factor(lits2$q801, levels=c("A GREAT DEAL","A FAIR AMOUNT","JUST A LITTLE","NOT AT ALL","DONT KNOW","Refused"))

library(car)
lits2$crise <- recode(lits2$q801, "c('A GREAT DEAL','A FAIR AMOUNT')='Great or fair amount';
                       c('NOT AT ALL','JUST A LITTLE')='Little or not at all';
                       else=NA")

lits2$crise <- factor(lits2$crise, levels=c("Little or not at all",
                                            "Great or fair amount"))
summary(lits2$crise)
################################################################

```

# Subsetting data for analysis

```{rLoadClean5}

################################################################
## Data for the analysis
df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
                             "q309_rec","sex",
                             "ageclass","age",
                             "future.diff","past.diff",
                             "pov.log.social",
                             "pov.log.individual",
                             "socialism","incsour3",
                             "edu2",
                             "group_analysis",
                             #"group_general","group_fine",
                             "income2",
                             "crise",
                             "q227", # income
                             "q228", # menneisyys
                             "q229", # tulevaisuus
                             "q226m_txt", # tulonlähde
                             "q515", # edu
                             "q801" # crise,
                             ))
# filter out not wanted countries
df <- df[!is.na(df$group_analysis),] # valitaan vaan group1 -muuttujan maat
df$cntry <- factor(df$cntry)
# print data in csv-format

################################################
################################################

```

# Macro level data

## Summarising micro data at the country level

```{rLoadClean6}
################################################

library(reshape2)
library(survey)
d.df <- svydesign(id = ~SerialID, 
                  weights = ~weight, 
                  data = df)

summary309 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
summary309 <- summary309[!is.na(summary309$Freq), ]

summary309w <- dcast(summary309, cntry ~ q309, value.var="Freq")
names(summary309w) <- c("cntry","social.blame","individual.blame","individual.fate","social.fate","not.stated","dont.know")

summary309_rec <- data.frame(prop.table(svytable(~q309_rec + cntry, d.df), 2) * 100)
summary309_rec <- summary309_rec[!is.na(summary309_rec$Freq), ]

summary309_recw <- dcast(summary309_rec, cntry ~ q309_rec, value.var="Freq")
names(summary309_recw) <- c("cntry","social.blame","individual.blame","individual.fate","social.fate")

```

## Data from Quality of Government institute

```{rLoadClean7}
#######
# Quality of governance
library(rustfare)
datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
                   'Kazakhstan','Tajikistan',
                   'Uzbekistan','Estonia','Latvia','Lithuania',
                   'Poland','Czech Republic','Slovakia','Hungary',
                   'Slovenia','Romania','Bulgaria',
                   'Italy','Germany','United Kingdom',
                   'Sweden','France'), 
       indicator = c("ffp_fsi", # Failed States Index
                     "wbgi_vae", # World Bank - Voice and Accountability
                     "undp_hdi", # UNDP Human Development Index
                     "fh_polity2", # Democracy (Freedom House/Polity)
                     "uw_gini"))
# subset the latest value per cntry per indicator
library(plyr)
datQogSum <- ddply(datQog, c("cname","indicator"), subset, year == max(year))

library(ggplot2)
ggplot(datQogSum, aes(x=year,y=value,color=cname)) + 
  geom_point() +
  geom_text(data = datQogSum,
            aes(x=year,y=value,color=cname,label=cname)) +
  facet_wrap(~indicator, scales="free") +
  theme(legend.position="none")

datQogSum.w <- dcast(datQogSum, cname ~ indicator, value.var="value")
datQogSum.w$cname <- recode(datQogSum.w$cname, "'United Kingdom'='Great Britain'")

datQogSum.w$group_fine <- recode(datQogSum.w$cname, "c('Armenia','Azerbaijan','Georgia') = 'Caucasus';
                         c('Russia','Ukraine','Moldova','Belarus')= 'CIS North';
                         c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')='Central Asia';
                         c('Estonia','Latvia','Lithuania')='Baltic States';
                         c('Poland','Czech Republic','Slovakia','Hungary','Slovenia',
                         'Romania','Bulgaria')='CEE EU';
                         c('Italy','Germany','Great Britain','Sweden','France')='West Europe';
                         else='other'")
datQogSum.w$group_general <- recode(datQogSum.w$cname,
                       "c('France','Germany','Great Britain',
                             'Italy','Sweden')='Western Europe';
                       c('Czech Republic','Estonia','Hungary','Bulgaria',
                       'Latvia','Lithuania','Poland',
                       'Slovakia','Slovenia','Romania')='CEE';
                       c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
                       'Kyrgyzstan','Moldova','Tajikistan',
                       'Ukraine','Uzbekistan','Russia')='CIS';else=NA")

################################

df.qog <- datQogSum.w
# underschores into dots for latex...
#names <- names(df.qog)
#library(stringr)
#names <- str_replace_all(names, "_",".")
#names(df.qog) <- names
library(gdata)
df.qog <- rename.vars(df.qog, from="cname", to="cntry")

```

## Data from World Bank

```{rLoadClean8}
## GDP change from WB
### We need iso2c countrycode
library(countrycode)
cntry <- levels(df$cntry) # countrynames
iso2c <- countrycode(cntry, "country.name", "iso2c")
iso2c # copy paste into WDI search
library(WDI)
# WDIsearch(string='gdp', field='name', cache=NULL)
wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), country = c("AL","AM","AZ","BY","BA","BG","HR",
                                                       "CZ","EE","FR","GE","DE","GB","HU",
                                                       "IT","KZ","KG","LV","LT","MK","MD",
                                                       "MN","ME","PL","RO","RU","RS","SK",
                                                       "SI","SE","TJ","TR","UA","UZ"), 
           start = 2007, end = 2010)

library(reshape2)
names(wb) <- c("iso2c","cntry","var","year")
wb.w <- dcast(wb, cntry ~ year, value.var="var")

names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
gdp <- wb.w[,c(1,6)]

gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'

```

## Merging all the data and saving for the analysis

```{rLoadClean9, eval=FALSE}

# yhdistetään nää aikaisempien datojen kanssa
macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
macro <- merge(macro,df.qog,by="cntry")

df <- merge(df,macro,by="cntry")

rm(lits2)
rm(df.qog)
rm(df.cross.postsoc)
rm(gdp)
rm(t3)
rm(summary309)
rm(summary309_rec)
rm(d.df)
rm(wb)
rm(wb.w)
rm(cntry)
rm(iso2c)

save.image("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
#load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
```

### 2013-10-31 index.Rmd
Changing the data munging into vector algebra

@@
 -------------------------------------------------------       -----------------------------
-[loading and cleaning](code/descriptive_analysis.html)        [source](code/descriptive_analysis.Rmd)
+[loading and cleaning](loadClean.html)                        [source](loadClean.Rmd)
 [missing data & diagnostics](code/descriptive_analysis.html)  [source](code/descriptive_analysis.Rmd)

### 2013-10-31 loadClean.Rmd
Changing the data munging into vector algebra

@@
 ```{rLoadClean1}
-
-##################################################################
-##################################################################
-##################################################################
-# Life in transition survey 2
-# load and clean data
-
-# libraries
-library(car)
-library(stringr)
-library(foreign)
-
-# LITS 2
-# from web
+# FROM WEB
 #temp <- tempfile()
 #download.file("http://www.ebrd.com/downloads/research/surveys/lits2.dta", temp)
+#library(foreign)
 #lits2 <- read.dta(temp)
-# tai
 
+# FROM LOCAL DISK
 load("~/workspace/data/lits/lits2.RData")
-##################################################################
-##################################################################
-
 ```
@@
 ```{rLoadClean2}
-##################################################################
 # different country groupings
-
-# remove whitespaces from country
+## remove whitespaces from country
+library(stringr)
 lits2$cntry <- str_trim(lits2$country_, side="right")
@@
 
-lits2$socialism <- recode(lits2$cntry,
-                            "c('France','Germany','Great Britain',
-                            'Italy','Sweden')='Market economy';
-                            c('Czech Republic','Estonia','Hungary',
-                          'Bulgaria','Latvia','Lithuania','Poland',
-                          'Slovakia','Slovenia','Romania','Albania',
-                          'Bosnia','Bulgaria','Croatia','Macedonia',
-                          'Montenegro','Kosovo','Serbia','Armenia',
-                          'Azerbaijan','Belarus','Georgia','Kazakhstan',
-                          'Kyrgyzstan','Moldova','Mongolia','Russia',
-                          'Tajikistan','Ukraine','Uzbekistan')='Post-socialist';
-                          else=NA")
-lits2$group_analysis <- recode(lits2$cntry,
-                             "c('France','Germany','Great Britain',
-                             'Italy','Sweden')='Western Europe';
-                             c('Czech Republic','Estonia','Hungary','Bulgaria',
-                             'Latvia','Lithuania','Poland',
-                             'Slovakia','Slovenia','Romania')='CEE';
-                             c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
-                             'Kyrgyzstan','Moldova','Tajikistan',
-                             'Ukraine','Uzbekistan','Russia')='CIS';else=NA")
+lits2$socialism[lits2$cntry %in% c('France','Germany','Great Britain',
+                                   'Italy','Sweden')]                   <- 'Market economy'
+
+lits2$socialism[lits2$cntry %in% c('Czech Republic','Estonia','Hungary',
+                                   'Bulgaria','Latvia','Lithuania',
+                                   'Poland','Slovakia','Slovenia',
+                                   'Romania','Albania','Bosnia',
+                                   'Bulgaria','Croatia','Macedonia',
+                                   'Montenegro','Kosovo','Serbia',
+                                   'Armenia','Azerbaijan','Belarus',
+                                   'Georgia','Kazakhstan','Kyrgyzstan',
+                                   'Moldova','Mongolia','Russia',
+                                   'Tajikistan','Ukraine','Uzbekistan')] <- 'Post-socialist'
+
+summary(factor(lits2$socialism))
+
+lits2$group_analysis[lits2$cntry %in% c('France','Germany',
+                                        'Great Britain',
+                                        'Italy','Sweden')] <- 'Western Europe'
+
+lits2$group_analysis[lits2$cntry %in% c('Czech Republic','Estonia',
+                                        'Hungary','Bulgaria',
+                                        'Latvia','Lithuania',
+                                        'Poland','Slovakia',
+                                        'Slovenia','Romania')] <- 'CEE'
+
+lits2$group_analysis[lits2$cntry %in% c('Armenia','Azerbaijan',
+                                        'Belarus','Georgia',
+                                        'Kazakhstan','Kyrgyzstan',
+                                        'Moldova','Tajikistan',
+                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'
+
+summary(factor(lits2$group_analysis))
 
@@
 
-##
+## Recoding the dependent variable
 
 ```{rLoadClean3}
-################################
+summary(lits2$q309)
 # reason for poverty
-################################
+library(stringr)
+lits2$q309 <- as.factor(str_replace(lits2$q309, 
+                                    "Don't know", 
+                                    "Dont know"))
+# All options
+lits2$q309 <- as.character(lits2$q309)
+lits2$q309[lits2$q309 %in% 'Because of injustice in our society'] <- 'Social Blame'
+lits2$q309[lits2$q309 %in% 'Because of laziness and lack of willpower'] <- 'Individual Blame'
+lits2$q309[lits2$q309 %in% 'Because they have been unlucky'] <- 'Individual Fate'
+lits2$q309[lits2$q309 %in% 'It is an inevitable part of modern life'] <- 'Social Fate'
+lits2$q309[lits2$q309 %in% 'Not stated'] <- 'Not stated'
+lits2$q309[lits2$q309 %in% 'Dont know'] <- 'Dont know'
 
-################################
-lits2$q309 <- as.factor(str_replace(lits2$q309, "Don't know", "Dont know"))
+lits2$q309 <- factor(lits2$q309, levels=c("Social Blame",
+                                          "Individual Blame",
+                                          "Individual Fate",
+                                          "Social Fate",
+                                          "Not stated",
+                                          "Dont know"))
 
-# All options
-
-lits2$q309 <- recode(lits2$q309, "'Because of injustice in our society'='Social Blame';
-                         'Because of laziness and lack of willpower'='Individual Blame';
-                         'Because they have been unlucky'='Individual Fate';
-                         'It is an inevitable part of modern life'='Social Fate';
-                         'Not stated'='Not stated';
-                         'Dont know'='Dont know'")
-
-lits2$q309 <- factor(lits2$q309, levels=c("Social Blame","Individual Blame",
-                                                  "Individual Fate","Social Fate",
-                                          "Not stated","Dont know"))
-
+summary(lits2$q309)
 # not stated and dont know coded NA
 
-lits2$q309_rec <- recode(lits2$q309, "'Because of injustice in our society'='Social Blame';
-                         'Because of laziness and lack of willpower'='Individual Blame';
-                         'Because they have been unlucky'='Individual Fate';
-                         'It is an inevitable part of modern life'='Social Fate';
-                         c('Not stated','Dont know')=NA")
+lits2$poverty <- lits2$q309
+lits2$poverty[lits2$poverty %in% c('Not stated','Dont know')] <- NA
+lits2$poverty <- factor(lits2$poverty)
+summary(lits2$poverty)
 
-lits2$q309_rec <- factor(lits2$q309_rec, levels=c("Social Blame","Individual Blame",
-                                                  "Individual Fate","Social Fate"))
+## For logistic regressions
+# $pov.log.social
 
-################################
-# $poverty - muuttuja
-################################
-#lits2$poverty <- recode(lits2$q309_rec, "'It is an inevitable part of modern life'=NA")
-lits2$poverty <- lits2$q309_rec
+lits2$pov.log.social[lits2$q309 == 'Social Blame'] <- 1
+lits2$pov.log.social[lits2$q309 != 'Social Blame'] <- 0
 
-################################
-# $pov.log.social
-################################
-lits2$pov.log.social <- recode(lits2$q309_rec, "'Social Blame'=1;
-                               else=0")
-################################
+summary(factor(lits2$pov.log.social))
+
 # $pov_log_individual
-################################
-lits2$pov.log.individual <- recode(lits2$q309_rec, "'Individual Blame'=1;
-                               else=0")
+###
+
+lits2$pov.log.individual[lits2$q309 == 'Individual Blame'] <- 1
+lits2$pov.log.individual[lits2$q309 != 'Individual Blame'] <- 0
+
+summary(factor(lits2$pov.log.individual))
 
@@
 
-##
+## Independent variables
 
 ```{rLoadClean4}
+# $sex
+lits2$sex <- lits2$q102_1
+# $income
 
-################################
-# $sex - muuttuja
-################################
-lits2$sex <- lits2$q102_1
+lits2$income[lits2$q227 == 1:3]  <- 'Low income'
+lits2$income[lits2$q227 == 4:7]  <- 'Middle income'
+lits2$income[lits2$q227 == 8:10] <-'High income'
 
-################################
-# $income
-################################
-
-lits2$income <- factor(recode(lits2$q227, 
-                           "1:3='Low income'; 
-                          4:7='Middle income'; 
-                           8:10='High income'; 
-                           else=NA"))
-
-lits2$income <- factor(lits2$income, levels=c("Low income","Middle income","High income"))
+lits2$income <- factor(lits2$income, 
+                       levels=c("Low income",
+                                "Middle income",
+                                "High income"))
 
@@
 #svyquantile(~q227, d.df, quantile = c(0.25, 0.5, 0.75), ci = TRUE)
+library(car)
+lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
+###
 
-lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
-################################
 # $future
-################################
+###
+
 lits2$future <- factor(recode(lits2$q229, 
@@
 
-################################
+###
+
 # $past
-################################
+###
+
 lits2$past <- factor(recode(lits2$q228, 
@@
 
-################################
+###
+
 # $age
-################################
+###
 
@@
 
-################################
+###
+
 # $incsour tulonlähde
-################################
+###
+
 lits2$q226m_txt <- recode(lits2$q226m, 
@@
 
-################################
+###
+
 # $edu
-################################
+###
 
@@
 summary(lits2$crise)
-################################################################
+###
 
@@
 
-```{rLoadClean5}
+```{rLoadClean5, eval=TRUE}
 
-################################################################
+###
+
 ## Data for the analysis
@@
 
-################################################
-################################################
+summary(df)
 
@@
 
-```{rLoadClean6}
-################################################
+```{rLoadClean6, eval=TRUE}
+###
 
@@
 
-```{rLoadClean7}
+```{rLoadClean7, eval=TRUE}
 #######
@@
 
-################################
+###
 
@@
 
-```{rLoadClean8}
+```{rLoadClean8, eval=TRUE}
 ## GDP change from WB

### 2013-11-01 article2013.Rmd
Changing the data munging into vector algebra
restructuring the whole lot. Creative destruction

@@
 % Markus Kainu[^**];Mikko Niemelä[^***]
-% March 2013
+% October 2013
 
-[^*]:*Work in progress. Please, do not cite without permission of authors*
+[^*]:*Work in progress
 
@@
 
+```{roptsattribArticle, echo=FALSE}
+opts_chunk$set(echo=FALSE,eval=TRUE,cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
+```
+
 ```{r, eval=FALSE, echo=FALSE}
@@
 system("~/.cabal/bin/pandoc -f markdown -t docx article2013.md -o article2013.docx --standalone --toc --number-sections")
+system("~/.cabal/bin/pandoc article2013.md -o article2013.epub --standalone --toc --number-sections")
 
@@
 
-lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions
 ```
+
+**Nootti**
+
+*Paperi on vielä hieman vaiheessa ja suosittelen tutustumaan tutkimuksen analytiikkaan osoitteessa: [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/).
+Yritän parhaillaan saada monitasomallintamista paremmalle tolalle.
+Paperissa olevat mallit ovat alustavia ja pitävät sisällään sekä teoreettista että menetelmällistä epätarkkuutta.*
+
+****
 
@@
 
+![Countries used in the analysis](figure/mapplot.png)
+
 We are using LiTS2 data from 21 transition countries with additional 4 Western European countries (France, Germany, Italy and Sweden) (See Figure 1).
@@
 
-The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?**
+The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?** (Variable `q309` in LiTS-data)
 
@@
 - It is an inevitable part of modern life
-- Other
+- Don't know
+- Not stated
 
@@
 
-```{rtablepredictor, results='asis', echo=FALSE}
-Variable <- c("Transfer dependency","Low education","Perceived low income level","Income compared to past","Income compared to future","Effect of financial crises")
-Description <- c("Social or private transfers \n as main source of income","No or only compulsory level education","Perceived income level below the country median","Perceived change in income compared to situation in four years ago","Expected change in income over the next four years","Whether respondents perceives that their household has suffered the financial crisis great or fair amount")
-table <- data.frame(Variable,Description)
-library(xtable)
-xtbl <- xtable(table, caption="Individual level predictors")
-print.xtable(xtbl, type="latex", comment=FALSE, size="scriptsize",include.rownames=FALSE)
-```
+\newpage
+
+Table: Individual level predictors
+
+-----------------------------------------------------------------------------
+Variable                    Description
+------------------------    -------------------------------------------------
+Transfer dependency         Social or private transfers as main source of income
+
+Low education               No or only compulsory level education
+
+Perceived low income level  Perceived income level below the country median
+
+Income compared to past     Perceived change in income compared to situation in four years ago
+
+Income compared to future   Expected change in income over the next four years
+
+Effect of financial crises  Whether respondents perceives that their household has suffered the financial crisis great or fair amount
+-----------------------------------------------------------------------------
 
@@
 (see Appendix A)
+
+\newpage
+
+Table: Contextual level predictors
+
+-----------------------------------------------------------------------------
+Variable                          Description
+------------------------          -------------------------------------------------
+Ghange in total GDP 2007 – 2010   Relative change from 2007 to 2010
+
+Gini coefficient                  Measures the distribution of income or consumption expenditure among individuals or households
+
+Voice and Accountability          Indicators measuring the political process, civil liberties and political rights.
+These indicators measure the extent to which citizens of a country are able to participate in the selection of governments.
+Indicators measuring the independence of the media is also included.
+Description
+-----------------------------------------------------------------------------
 
@@
 
-[^*****]:R-code for data transformation and analysis is available at [research.muuankarski.org/code/](http://research.muuankarski.org/code/)
+[^*****]:R-code for data transformation and analysis is available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/)
 
@@
 The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups.
-Full model summaries currently available at [research.muuankarski.org/code/](http://research.muuankarski.org/code/).
+Full model summaries currently available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/).
 
@@
 Totally, in 19 of 21 examined post-socialist countries injustice in a society receives the highest support.
+
+\newpage
+
+```{rattribArticleDescTbl1, results='asis'}
+source("code/dv_tables.R")
+print.xtable(xtable(tbl5, 
+                    caption="Weighted population shares of dependent variable by country",),
+              type="latex",
+              comment=FALSE, 
+              include.rownames=FALSE,
+             caption.placement="top")
+
+```
 
@@
 
-![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)
+![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.pdf)
 
@@
 
-![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)
+![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.pdf)
 
@@
 
-![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)
+![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.pdf)
 
@@
 
-- click for tables: [code/multilevel_models.pdf](code/multilevel_models.pdf)
-- click for RC19 tables: [code/multilevel_modelsRC19.pdf](code/multilevel_modelsRC19.pdf)
+*There we some issues in converts regression outputs so the tables are evolving and can be browsed at: [http://research.muuankarski.org/attributions/code/multilevel_models.html](http://research.muuankarski.org/attributions/code/multilevel_models.html)*
 
->As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
+****
+
+```{rattribArticleMultilevel1,results='asis', eval=FALSE}
+source("~/workspace/lits/attrib/attrib_year2013/code/multi_logit.R")
+library(texreg)
+texreg(list(soc.0,
+       soc.1,
+       soc.2,
+       soc.3,
+       soc.4),
+       use.package=FALSE,
+       include.pvalues=TRUE,
+       ci.test=NULL,
+
+          caption="Logistic multilevel random intercept model for social blame type of explanation.", caption.above=TRUE, doctype = FALSE,
+        custom.coef.names=c("Intercept",
+                             "Transfer dependency (Not dependent)",
+                             "Education (No education)",
+                             "Perceived income (Low)",
+                             "Perceived change in income compared to four years ago (Has worsened)",
+                             "Expected change in income over next four years (Will worsen)",
+                             "Effect of financial crisis (Suffered great or fair amount)",
+                             "gini coefficient",
+                             "GDP change 2007-2010",
+                             "Voice And Accountability (WB)"),
+        custom.model.names=c("empty","Model SB1","Model SB2",
+                             "Model SB3","Model SB4"))
+
+```
+
+```{rattribArticleMultilevel2, results='asis', eval=FALSE}
+
+texreg(list(ind.0,
+          ind.1,
+          ind.2,
+          ind.3,
+          ind.4),
+       use.packages=FALSE,
+       include.pvalues=TRUE,
+       ci.test=NULL,
+       caption="Logistic multilevel random intercept model for individual blame type of explanation.", 
+       caption.above=TRUE, 
+        doctype = FALSE,
+                custom.coef.names=c("Intercept",
+                             "Transfer dependency (Not dependent)",
+                             "Education (No education)",
+                             "Perceived income (Low)",
+                             "Perceived change in income compared to four years ago (Has worsened)",
+                             "Expected change in income over next four years (Will worsen)",
+                             "Effect of financial crisis (Suffered great or fair amount)",
+                             "gini coefficient",
+                             "GDP change 2007-2010",
+                             "Voice And Accountability (WB)"),
+        custom.model.names=c("empty","Model IB1","Model IB2",
+                             "Model IB3","Model IB4"))
+```
+
+<!--
+
+As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
 Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
+
+-->
 

### 2013-11-01 descriptive_analysis.Rmd
restructuring the whole lot. Creative destruction

@@
-% Descriptive analysis
+% Descriptive visualizations
 
@@
 library(kaRski)
-knitpandoc("descriptive_analysis")
+knit("descriptive_analysis.Rmd")
+system("pandoc -s descriptive_analysis.md -o descriptive_analysis.html --toc --number-section")
+
+knitpandoc("descriptive_analysis",toc)
 ```
@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
 ```
@@
 
-```{rdescplot1_1, fig.width=15}
+```{rdescplot1_1}
 load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
@@
 tbl <- data.frame(prop.table(table(df$cntry,df$q227),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot1_2, fig.width=15}
+```{rdescplot1_2}
 df$cntry <- factor(df$cntry)
 tbl <- data.frame(prop.table(table(df$cntry,df$income),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot2_1, fig.width=15}
+```{rdescplot2_1}
 df$cntry <- factor(df$cntry)
-tbl <- data.frame(prop.table(table(df$cntry,df$q226m_txt),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+tbl <- data.frame(prop.table(table(df$cntry,df$q226m),1)*100)
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot2_2, fig.width=15}
+```{rdescplot2_2}
 
@@
 tbl <- data.frame(prop.table(table(df$cntry,df$incsour3),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot3_1, fig.width=15}
+```{rdescplot3_1}
 df$cntry <- factor(df$cntry)
 tbl <- data.frame(prop.table(table(df$cntry,df$q228),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot3_2, fig.width=15}
+```{rdescplot3_2}
 
 tbl <- data.frame(prop.table(table(df$cntry,df$past),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot4_1, fig.width=15}
+```{rdescplot4_1}
 df$cntry <- factor(df$cntry)
 tbl <- data.frame(prop.table(table(df$cntry,df$q229),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot4_2, fig.width=15}
+```{rdescplot4_2}
 
 tbl <- data.frame(prop.table(table(df$cntry,df$future),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot5_1, fig.width=15}
+```{rdescplot5_1}
 df$cntry <- factor(df$cntry)
 tbl <- data.frame(prop.table(table(df$cntry,df$q515),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot5_2, fig.width=15}
+```{rdescplot5_2}
 
 tbl <- data.frame(prop.table(table(df$cntry,df$edu2),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot6_1, fig.width=15}
+```{rdescplot6_1}
 df$cntry <- factor(df$cntry)
 tbl <- data.frame(prop.table(table(df$cntry,df$q801),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot6_2, fig.width=15}
+```{rdescplot6_2}
 
 tbl <- data.frame(prop.table(table(df$cntry,df$crise),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group))
+reg <- as.data.frame(table(df$cntry,df$group_general))
 reg <- subset(reg, Freq > 0)
@@
 
-```{rdescplot7, fig.width=15}
-df <- macro[with(macro, order(uw.gini)), ]
+```{rdescplot7}
+df <- macro[with(macro, order(uw_gini)), ]
 df$cntry <- factor(df$cntry, levels=df$cntry)
-ggplot(df, aes(x=cntry,y=uw.gini, colour=group)) +
+ggplot(df, aes(x=cntry,y=uw_gini, colour=group)) +
   geom_point(size=5) +
@@
 
-```{rdescplot8, fig.width=15}
+```{rdescplot8}
 df <- macro[with(macro, order(gdpchange07.10)), ]
@@
 
-```{rdescplot9, fig.width=15}
-df <- macro[with(macro, order(wbgi.vae)), ]
+```{rdescplot9}
+df <- macro[with(macro, order(wbgi_vae)), ]
 df$cntry <- factor(df$cntry, levels=df$cntry)
 
-ggplot(df, aes(x=cntry,y=wbgi.vae, colour=group)) +
+ggplot(df, aes(x=cntry,y=wbgi_vae, colour=group)) +
   geom_point(size=5) +
@@
 
-```{rdescplot10, fig.width=15}
+```{rdescplot10}
 library(GGally)
-ggpairs(macro[, c("uw.gini", "gdpchange07.10", "wbgi.vae","social.blame","individual.blame")])
+ggpairs(macro[, c("uw_gini", "gdpchange07.10", "wbgi_vae","socialBlame","individualBlame")])
 
@@
 
-# Final plots
-
-![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)
-
-![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)
-
-![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)
+# Dependent variable

### 2013-11-01 finalPlots.Rmd
restructuring the whole lot. Creative destruction

% Final graphics for the paper

- [back to index](index.html)

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(kaRski)
knit("finalPlots.Rmd")
system("pandoc -s finalPlots.md -o finalPlots.html --toc --number-section")

```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
```

```{rattribFinalPlotsLoad}
library(ggplot2)
library(gridExtra)
library(plyr)
library(grid)
#########################################
#########################################
setwd("~/workspace/lits/attrib/attrib_year2013")
#########################################
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
#########################################
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
```

# GDP change

```{rattribFinalPlotsGDP}
# valitaan vaan kohdemaat

t3 <- subset(macro, group_fine != 'West Europe')
west <- subset(macro, group_fine == 'West Europe')
#t3 <- macro
#########################################
t3$group1 <- "group1"
func <- function(t3)
{
  return(data.frame(COR = cor(t3$socialBlame, t3$gdpchange07.10)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(socialBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=gdpchange07.10, y=socialBlame, 
                        label=cntry)) +
  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
  geom_text(data=cnames, aes(gdpchange07.10,socialBlame, label = COR, group=COR), 
            size=2.5, color="black") +
  geom_text(size=2, vjust=-0.8, hjust=0.5) +
  geom_point(size=1.5) +
  geom_point(data=t4, aes(x=gdpchange07.10, 
                          y=socialBlame, color=group_general)) +
  # west reference
  geom_point(data=west, aes(x=gdpchange07.10, 
                          y=socialBlame, color=group_general)) +
  geom_text(data=west, aes(x=gdpchange07.10, 
                           y=socialBlame, label=cntry),
            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
  # west reference
  labs(x = "Change in GDP between 2007 and 2010 (%)", 
       y = "Support for social blame type of explanation (%)") + 
  scale_colour_manual(values=cbPalette) +
  coord_cartesian(xlim=c(-25,100)) +
  theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=6)) +
  theme(legend.position="top") +
  theme(axis.title.y = element_text(size=6)) +
  theme(axis.title.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=6)) +
  theme(axis.text.x = element_text(size=6)) +
  #guides(color = guide_legend(nrow = 2)) +
  theme(legend.key.size = unit(3, "mm"))

func <- function(t3)
{
  return(data.frame(COR = cor(t3$individualBlame, t3$gdpchange07.10)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individualBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=gdpchange07.10, 
                        y=individualBlame, label=cntry)) +
  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
  geom_text(size=2, vjust=-0.8, hjust=0.5, color="black") +
  geom_text(data=cnames, aes(gdpchange07.10,individualBlame, label = COR, group=COR), 
            size=2.5, color="black") +
  geom_point(size=1.5) +
  geom_point(data=t4, aes(x=gdpchange07.10, 
             y=individualBlame, color=group_general)) +
  # west reference
  geom_point(data=west, aes(x=gdpchange07.10, 
                            y=individualBlame, color=group_general)) +
  geom_text(data=west, aes(x=gdpchange07.10, 
                           y=individualBlame, label=cntry),
            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
  # west reference
  labs(x = "Change in GDP between 2007 and 2010 (%)", 
       y = "Support for individual blame type of explanation (%)") +
  scale_colour_manual(values=cbPalette) +
  coord_cartesian(xlim=c(-25,100)) +
  theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=6)) +
  theme(legend.position="top") +
  theme(axis.title.y = element_text(size=6)) +
  theme(axis.title.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=6)) +
  theme(axis.text.x = element_text(size=6)) +
  #guides(color = guide_legend(nrow = 2)) +
  theme(legend.key.size = unit(3, "mm"))

pdf("figure/scatter_gdp.pdf", width=14/2.54, height=10/2.54)
grid.arrange(plot1,plot2, ncol=2)
dev.off()

ppi <- 300
png("figure/scatter_gdp.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
grid.arrange(plot1,plot2, ncol=2)
dev.off()
```

![](figure/scatter_gdp.png)

# UNU-Wider Gini

```{rattribFinalPlotsGini}
# UNU-Wider Gini

func <- function(t3)
{
  return(data.frame(COR = cor(t3$socialBlame, t3$uw_gini)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(socialBlame, uw_gini) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=uw_gini, y=socialBlame, label=cntry)) +
  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
  geom_text(size=2, vjust=-0.8, hjust=0.5) +
  geom_text(data=cnames, aes(uw_gini,socialBlame, label = COR, group=COR), 
            size=2.5, color="black") +
  labs(x = "Gini coefficient", 
       y = "Support for social blame type of explanation (%)") +
  geom_point(size=1.5) +
  geom_point(data=t4, aes(x=uw_gini, 
                          y=socialBlame, color=group_general)) +
  # west reference
  geom_point(data=west, aes(x=uw_gini, 
                            y=socialBlame, color=group_general)) +
  geom_text(data=west, aes(x=uw_gini, 
                           y=socialBlame, label=cntry),
            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
  # west reference
  scale_colour_manual(values=cbPalette) +
  coord_cartesian(xlim=c(20,60)) +
  theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=6)) +
  theme(legend.position="top") +
  theme(axis.title.y = element_text(size=6)) +
  theme(axis.title.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=6)) +
  theme(axis.text.x = element_text(size=6)) +
  #guides(color = guide_legend(nrow = 2)) +
  theme(legend.key.size = unit(3, "mm"))

func <- function(t3)
{
  return(data.frame(COR = cor(t3$individualBlame, t3$uw_gini)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individualBlame, uw_gini) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=uw_gini, y=individualBlame, label=cntry)) +
  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
  geom_text(size=2, vjust=-0.8, hjust=0.5) +
  geom_text(data=cnames, aes(uw_gini,individualBlame, label = COR, group=COR), 
            size=2.5, color="black") +
  labs(x = "Gini coefficient", 
       y = "Support for individual blame type of explanation (%)") +
  geom_point(size=1.5) +  
  geom_point(data=t4, aes(x=uw_gini, 
                          y=individualBlame, color=group_general)) +
  # west reference
  geom_point(data=west, aes(x=uw_gini, 
                            y=individualBlame, color=group_general)) +
  geom_text(data=west, aes(x=uw_gini, 
                           y=individualBlame, label=cntry),
            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
  # west reference
  scale_colour_manual(values=cbPalette) +
  coord_cartesian(xlim=c(20,60)) +
  theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=6)) +
  theme(legend.position="top") +
  theme(axis.title.y = element_text(size=6)) +
  theme(axis.title.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=6)) +
  theme(axis.text.x = element_text(size=6)) +
  #guides(color = guide_legend(nrow = 2)) +
  theme(legend.key.size = unit(3, "mm"))

pdf("figure/scatter_gini.pdf", width=14/2.54, height=10/2.54)
grid.arrange(plot1,plot2, ncol=2)
dev.off()

ppi <- 300
png("figure/scatter_gini.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
grid.arrange(plot1,plot2, ncol=2)
dev.off()
```

![](figure/scatter_gini.png)

# World Bank - Voice and Accountability

```{rattribFinalPlotsWB}

# World Bank - Voice and Accountability

func <- function(t3)
{
  return(data.frame(COR = cor(t3$socialBlame, t3$wbgi_vae)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(socialBlame, wbgi_vae) ~ COR, data=t4, mean),2)
plot1 <- ggplot(t4, aes(x=wbgi_vae, y=socialBlame, 
                        label=cntry)) +
  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +
  geom_text(size=2, vjust=-0.8, hjust=0.5) +
  geom_text(data=cnames, aes(wbgi_vae,socialBlame, label = COR, group=COR), 
            size=2.5, color="black") +
  labs(x = "Voice and Accountability Index 2011", 
       y = "Support for social blame type of explanation (%)") +
  geom_point(size=1.5) +
  geom_point(data=t4, aes(x=wbgi_vae, 
                          y=socialBlame, color=group_general)) +
  # west reference
  geom_point(data=west, aes(x=wbgi_vae, 
                            y=socialBlame, color=group_general)) +
  geom_text(data=west, aes(x=wbgi_vae, 
                           y=socialBlame, label=cntry),
            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
  # west reference
  scale_colour_manual(values=cbPalette) +
  coord_cartesian(xlim=c(-2,2)) +
  theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=6)) +
  theme(legend.position="top") +
  theme(axis.title.y = element_text(size=6)) +
  theme(axis.title.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=6)) +
  theme(axis.text.x = element_text(size=6)) +
  #guides(color = guide_legend(nrow = 2)) +
  theme(legend.key.size = unit(3, "mm"))

func <- function(t3)
{
  return(data.frame(COR = cor(t3$individualBlame, t3$wbgi_vae)))
}
df.cor <- ddply(t3, .(group1), func)
t4 <- join(t3,df.cor,by="group1")
cnames <- round(stats:::aggregate.formula(cbind(individualBlame, wbgi_vae) ~ COR, data=t4, mean),2)
plot2 <- ggplot(t4, aes(x=wbgi_vae, 
                        y=individualBlame, label=cntry)) +
geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +
  geom_text(size=2, vjust=-0.8, hjust=0.5) +
  geom_text(data=cnames, aes(wbgi_vae,individualBlame, label = COR, group=COR), 
            size=2.5, color="black") +
  labs(x ="Voice and Accountability Index 2011", 
       y = "Support for individual blame type of explanation (%)") +
  geom_point(size=1.5) +
  geom_point(data=t4, aes(x=wbgi_vae, 
                          y=individualBlame, color=group_general)) +
  # west reference
  geom_point(data=west, aes(x=wbgi_vae, 
                            y=individualBlame, color=group_general)) +
  geom_text(data=west, aes(x=wbgi_vae, 
                           y=individualBlame, label=cntry),
            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
  # west reference
  scale_colour_manual(values=cbPalette) +
  coord_cartesian(xlim=c(-2,2)) +
  theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=6)) +
  theme(legend.position="top") +
  theme(axis.title.y = element_text(size=6)) +
  theme(axis.title.x = element_text(size=6)) +
  theme(axis.text.y = element_text(size=6)) +
  theme(axis.text.x = element_text(size=6)) +
  #guides(color = guide_legend(nrow = 2)) +
  theme(legend.key.size = unit(3, "mm"))

pdf("figure/scatter_wbgi.pdf", width=14/2.54, height=10/2.54)
grid.arrange(plot1,plot2, ncol=2)
dev.off()

ppi <- 300
png("figure/scatter_wbgi.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
grid.arrange(plot1,plot2, ncol=2)
dev.off()

```

![](figure/scatter_wbgi.png)

```{rattribFinalPlotsx, eval=FALSE}

```

### 2013-11-01 finalTables.Rmd
restructuring the whole lot. Creative destruction

--- 2013-11-01 finalPlots.Rmd
@@
-% Final graphics for the paper
+% Final regression ouput tables for the paper
 
@@
 library(kaRski)
-knit("finalPlots.Rmd")
-system("pandoc -s finalPlots.md -o finalPlots.html --toc --number-section")
+knit("finalTables.Rmd")
+system("pandoc -s finalTables.md -o finalTables.html --toc --number-section")
 
@@
 
-```{rattribFinalPlotsLoad}
-library(ggplot2)
-library(gridExtra)
-library(plyr)
-library(grid)
-#########################################
-#########################################
-setwd("~/workspace/lits/attrib/attrib_year2013")
-#########################################
-cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
-#########################################
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
-```
+**In progress..**
 
-# GDP change
-
-```{rattribFinalPlotsGDP}
-# valitaan vaan kohdemaat
-
-t3 <- subset(macro, group_fine != 'West Europe')
-west <- subset(macro, group_fine == 'West Europe')
-#t3 <- macro
-#########################################
-t3$group1 <- "group1"
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$socialBlame, t3$gdpchange07.10)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(socialBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
-plot1 <- ggplot(t4, aes(x=gdpchange07.10, y=socialBlame, 
-                        label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
-  geom_text(data=cnames, aes(gdpchange07.10,socialBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
-  geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=gdpchange07.10, 
-                          y=socialBlame, color=group_general)) +
-  # west reference
-  geom_point(data=west, aes(x=gdpchange07.10, 
-                          y=socialBlame, color=group_general)) +
-  geom_text(data=west, aes(x=gdpchange07.10, 
-                           y=socialBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
-  # west reference
-  labs(x = "Change in GDP between 2007 and 2010 (%)", 
-       y = "Support for social blame type of explanation (%)") + 
-  scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(-25,100)) +
-  theme_minimal() +
-  theme(legend.title=element_blank()) +
-  theme(legend.text=element_text(size=6)) +
-  theme(legend.position="top") +
-  theme(axis.title.y = element_text(size=6)) +
-  theme(axis.title.x = element_text(size=6)) +
-  theme(axis.text.y = element_text(size=6)) +
-  theme(axis.text.x = element_text(size=6)) +
-  #guides(color = guide_legend(nrow = 2)) +
-  theme(legend.key.size = unit(3, "mm"))
-
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$individualBlame, t3$gdpchange07.10)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(individualBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
-plot2 <- ggplot(t4, aes(x=gdpchange07.10, 
-                        y=individualBlame, label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
-  geom_text(size=2, vjust=-0.8, hjust=0.5, color="black") +
-  geom_text(data=cnames, aes(gdpchange07.10,individualBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
-  geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=gdpchange07.10, 
-             y=individualBlame, color=group_general)) +
-  # west reference
-  geom_point(data=west, aes(x=gdpchange07.10, 
-                            y=individualBlame, color=group_general)) +
-  geom_text(data=west, aes(x=gdpchange07.10, 
-                           y=individualBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
-  # west reference
-  labs(x = "Change in GDP between 2007 and 2010 (%)", 
-       y = "Support for individual blame type of explanation (%)") +
-  scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(-25,100)) +
-  theme_minimal() +
-  theme(legend.title=element_blank()) +
-  theme(legend.text=element_text(size=6)) +
-  theme(legend.position="top") +
-  theme(axis.title.y = element_text(size=6)) +
-  theme(axis.title.x = element_text(size=6)) +
-  theme(axis.text.y = element_text(size=6)) +
-  theme(axis.text.x = element_text(size=6)) +
-  #guides(color = guide_legend(nrow = 2)) +
-  theme(legend.key.size = unit(3, "mm"))
-
-pdf("figure/scatter_gdp.pdf", width=14/2.54, height=10/2.54)
-grid.arrange(plot1,plot2, ncol=2)
-dev.off()
-
-ppi <- 300
-png("figure/scatter_gdp.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
-grid.arrange(plot1,plot2, ncol=2)
-dev.off()
-```
-
-![](figure/scatter_gdp.png)
-
-# UNU-Wider Gini
-
-```{rattribFinalPlotsGini}
-# UNU-Wider Gini
-
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$socialBlame, t3$uw_gini)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(socialBlame, uw_gini) ~ COR, data=t4, mean),2)
-plot1 <- ggplot(t4, aes(x=uw_gini, y=socialBlame, label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
-  geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(uw_gini,socialBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
-  labs(x = "Gini coefficient", 
-       y = "Support for social blame type of explanation (%)") +
-  geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=uw_gini, 
-                          y=socialBlame, color=group_general)) +
-  # west reference
-  geom_point(data=west, aes(x=uw_gini, 
-                            y=socialBlame, color=group_general)) +
-  geom_text(data=west, aes(x=uw_gini, 
-                           y=socialBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
-  # west reference
-  scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(20,60)) +
-  theme_minimal() +
-  theme(legend.title=element_blank()) +
-  theme(legend.text=element_text(size=6)) +
-  theme(legend.position="top") +
-  theme(axis.title.y = element_text(size=6)) +
-  theme(axis.title.x = element_text(size=6)) +
-  theme(axis.text.y = element_text(size=6)) +
-  theme(axis.text.x = element_text(size=6)) +
-  #guides(color = guide_legend(nrow = 2)) +
-  theme(legend.key.size = unit(3, "mm"))
-
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$individualBlame, t3$uw_gini)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(individualBlame, uw_gini) ~ COR, data=t4, mean),2)
-plot2 <- ggplot(t4, aes(x=uw_gini, y=individualBlame, label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
-  geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(uw_gini,individualBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
-  labs(x = "Gini coefficient", 
-       y = "Support for individual blame type of explanation (%)") +
-  geom_point(size=1.5) +  
-  geom_point(data=t4, aes(x=uw_gini, 
-                          y=individualBlame, color=group_general)) +
-  # west reference
-  geom_point(data=west, aes(x=uw_gini, 
-                            y=individualBlame, color=group_general)) +
-  geom_text(data=west, aes(x=uw_gini, 
-                           y=individualBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
-  # west reference
-  scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(20,60)) +
-  theme_minimal() +
-  theme(legend.title=element_blank()) +
-  theme(legend.text=element_text(size=6)) +
-  theme(legend.position="top") +
-  theme(axis.title.y = element_text(size=6)) +
-  theme(axis.title.x = element_text(size=6)) +
-  theme(axis.text.y = element_text(size=6)) +
-  theme(axis.text.x = element_text(size=6)) +
-  #guides(color = guide_legend(nrow = 2)) +
-  theme(legend.key.size = unit(3, "mm"))
-
-pdf("figure/scatter_gini.pdf", width=14/2.54, height=10/2.54)
-grid.arrange(plot1,plot2, ncol=2)
-dev.off()
-
-ppi <- 300
-png("figure/scatter_gini.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
-grid.arrange(plot1,plot2, ncol=2)
-dev.off()
-```
-
-![](figure/scatter_gini.png)
-
-# World Bank - Voice and Accountability
-
-```{rattribFinalPlotsWB}
-
-# World Bank - Voice and Accountability
-
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$socialBlame, t3$wbgi_vae)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(socialBlame, wbgi_vae) ~ COR, data=t4, mean),2)
-plot1 <- ggplot(t4, aes(x=wbgi_vae, y=socialBlame, 
-                        label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +
-  geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(wbgi_vae,socialBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
-  labs(x = "Voice and Accountability Index 2011", 
-       y = "Support for social blame type of explanation (%)") +
-  geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=wbgi_vae, 
-                          y=socialBlame, color=group_general)) +
-  # west reference
-  geom_point(data=west, aes(x=wbgi_vae, 
-                            y=socialBlame, color=group_general)) +
-  geom_text(data=west, aes(x=wbgi_vae, 
-                           y=socialBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
-  # west reference
-  scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(-2,2)) +
-  theme_minimal() +
-  theme(legend.title=element_blank()) +
-  theme(legend.text=element_text(size=6)) +
-  theme(legend.position="top") +
-  theme(axis.title.y = element_text(size=6)) +
-  theme(axis.title.x = element_text(size=6)) +
-  theme(axis.text.y = element_text(size=6)) +
-  theme(axis.text.x = element_text(size=6)) +
-  #guides(color = guide_legend(nrow = 2)) +
-  theme(legend.key.size = unit(3, "mm"))
-
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$individualBlame, t3$wbgi_vae)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(individualBlame, wbgi_vae) ~ COR, data=t4, mean),2)
-plot2 <- ggplot(t4, aes(x=wbgi_vae, 
-                        y=individualBlame, label=cntry)) +
-geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +
-  geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(wbgi_vae,individualBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
-  labs(x ="Voice and Accountability Index 2011", 
-       y = "Support for individual blame type of explanation (%)") +
-  geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=wbgi_vae, 
-                          y=individualBlame, color=group_general)) +
-  # west reference
-  geom_point(data=west, aes(x=wbgi_vae, 
-                            y=individualBlame, color=group_general)) +
-  geom_text(data=west, aes(x=wbgi_vae, 
-                           y=individualBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
-  # west reference
-  scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(-2,2)) +
-  theme_minimal() +
-  theme(legend.title=element_blank()) +
-  theme(legend.text=element_text(size=6)) +
-  theme(legend.position="top") +
-  theme(axis.title.y = element_text(size=6)) +
-  theme(axis.title.x = element_text(size=6)) +
-  theme(axis.text.y = element_text(size=6)) +
-  theme(axis.text.x = element_text(size=6)) +
-  #guides(color = guide_legend(nrow = 2)) +
-  theme(legend.key.size = unit(3, "mm"))
-
-pdf("figure/scatter_wbgi.pdf", width=14/2.54, height=10/2.54)
-grid.arrange(plot1,plot2, ncol=2)
-dev.off()
-
-ppi <- 300
-png("figure/scatter_wbgi.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
-grid.arrange(plot1,plot2, ncol=2)
-dev.off()
-
-```
-
-![](figure/scatter_wbgi.png)
-
-```{rattribFinalPlotsx, eval=FALSE}
-
-```
+See old model outputs [here](file:///home/aurelius/workspace/lits/attrib/attrib_year2013/code/multilevel_models.html)

### 2013-11-01 index.Rmd
restructuring the whole lot. Creative destruction

@@
 
-cd ~/workspace/lits/attrib/attrib_year2013/figures
-lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
+cd ~/workspace/lits/attrib/attrib_year2013/
+lftp -u np16003 -e "mput *.html *.Rmd *.pdf *.epub *.docx;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
+
+cd ~/workspace/lits/attrib/attrib_year2013/figure
+lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/figure
 ```
@@
 
-Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.
+**Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.**
 
@@
 
-## Cleaning the data & missing data
+## Cleaning the data & variable summaries
 
+-------------------
 html-ouput                                                    source code
--------------------------------------------------------       -----------------------------
-[loading and cleaning](loadClean.html)                        [source](loadClean.Rmd)
-[missing data & diagnostics](code/descriptive_analysis.html)  [source](code/descriptive_analysis.Rmd)
+-------------------                                           ---------
+[loading, cleaning and summarizing](loadClean.html)           [source](loadClean.Rmd)
+-------------------
 
@@
 
+-------------------
 html-ouput                                                source code
--------------------------------------------------------   -----------------------------
+-----------------------                                   ---------
 [Descriptive analysis](descriptive_analysis.html)         [source](descriptive_analysis.Rmd)
-[Graphics in the paper]()                                 [source](descriptive_analysis.Rmd)
+
+[Graphics in the paper](finalPlots.html)                  [source](finalPlots.Rmd)
+-------------------
 
@@
 
+-------------------
 html-ouput                                                source code
--------------------------------------------------------   -----------------------------
-[Output tables in html](code/multilevel_models.html)      [source](code/multilevel_models.Rmd)
-[Output tables in pdf](code/multilevel_models.pdf)        [source](code/multilevel_models.Rmd)
+-----------------------                                   ---------
+[Regression analysis](code/multilevel_models.html)        [source](code/multilevel_models.Rmd)
+
+[Final tables](finalTables.html)                          [source](finalTables.Rmd)
+-------------------
+
+- [Final tables as A4 .pdf](finalTables.pdf)
 
@@
 - [Article draft in .epub](article2013.epub)
+- [Article draft in .docx](article2013.docx)
 

### 2013-11-01 loadClean.Rmd
restructuring the whole lot. Creative destruction

@@
 
-lits2$income[lits2$q227 == 1:3]  <- 'Low income'
-lits2$income[lits2$q227 == 4:7]  <- 'Middle income'
-lits2$income[lits2$q227 == 8:10] <-'High income'
+lits2$income[lits2$q227 >= 1 & lits2$q227 <= 3]  <- 'Low income'
+lits2$income[lits2$q227 >= 4 & lits2$q227 <= 7]  <- 'Middle income'
+lits2$income[lits2$q227 >= 8] <-'High income'
 
@@
 
-lits2$income2 <- factor(recode(lits2$q227, 
-                              "1:5='Low'; 
-                              6:10='High'; 
-                              else=NA"))
+summary(lits2$income)
 
-lits2$income <- factor(lits2$income2, levels=c("Low","High"))
+lits2$income2[lits2$q227 >= 1 & lits2$q227 <= 5]  <- 'Low'
+lits2$income2[lits2$q227 >= 6 & lits2$q227 <= 10]  <- 'High'
+
+lits2$income <- factor(lits2$income2, 
+                       levels=c("Low",
+                                "High"))
+
+summary(lits2$income2)
 
@@
 library(car)
-lits2$q227_rec <- recode(lits2$q227, "-111:0=NA")
-###
-
+lits2$q227_rec <- lits2$q227
+lits2$q227_rec[lits2$q227_rec <= 0] <- NA
 # $future
@@
 
-lits2$future <- factor(recode(lits2$q229, 
-                              "1:3='a) low income'; 
-                          4:7='b) middle income'; 
-                              8:10='c) high income'; 
-                              else=NA"))
+summary(factor(lits2$q229))
 
-lits2$q229_rec <- recode(lits2$q229, "-100:0=NA")
+lits2$future[lits2$q229 >= 1 & lits2$q229 <= 3] <- "low income"
+lits2$future[lits2$q229 >= 4 & lits2$q229 <= 7] <- "middle income"
+lits2$future[lits2$q229 >= 8 & lits2$q229 <= 10] <- "high income"
+
+summary(factor(lits2$future))
+
+lits2$q229_rec <- lits2$q229
+lits2$q229_rec[lits2$q229_rec <= 0] <- NA
+
+summary(factor(lits2$q229_rec))
 
@@
 lits2$future.diff[lits2$q227_rec <= lits2$q229_rec] <- 'same or better'
-lits2$future.diff <- factor(lits2$future.diff)
+lits2$future.diff <- factor(lits2$future.diff, levels=c("worse",
+                                                        "same or better"))
 
-###
-
+summary(lits2$future.diff)
 # $past
@@
 
-lits2$past <- factor(recode(lits2$q228, 
-                              "1:3='a) low income'; 
-                          4:7='b) middle income'; 
-                              8:10='c) high income'; 
-                              else=NA"))
+summary(factor(lits2$q228))
 
-lits2$q228_rec <- recode(lits2$q228, "-111:0=NA")
+lits2$past[lits2$q228 >= 1 & lits2$q228 <= 3] <- "low income"
+lits2$past[lits2$q228 >= 4 & lits2$q228 <= 7] <- "middle income"
+lits2$past[lits2$q228 >= 8 & lits2$q228 <= 10] <- "high income"
+
+summary(factor(lits2$past))
+
+lits2$q228_rec <- lits2$q228
+lits2$q228_rec[lits2$q228_rec <= 0] <- NA
+
+summary(factor(lits2$q228_rec))
 
@@
 lits2$past.diff[lits2$q227_rec >= lits2$q228_rec] <- 'same or better'
-lits2$past.diff <- factor(lits2$past.diff)
-
-###
-
+lits2$past.diff <- factor(lits2$past.diff, levels=c("worse",
+                                                    "same or better"))
 # $age
@@
 
-lits2$ageclass <- factor(recode(lits2$q104a_1, 
-                              "1:29='29 or younger'; 
-                          30:44='30 to 44'; 
-                              45:59='45 to 59';
-                              60:120='60 and older';
-                                else=NA"))
+lits2$ageclass[lits2$q104a_1 <= 29] <- '29 or younger'
+lits2$ageclass[lits2$q104a_1 >= 30 & lits2$q104a_1 <= 44] <- '30 to 44'
+lits2$ageclass[lits2$q104a_1 >= 45 & lits2$q104a_1 <= 59] <- '45 to 59'
+lits2$ageclass[lits2$q104a_1 >= 60 & lits2$q104a_1 <= 120] <- '60 and older'
 
-lits2$ageclass <- factor(lits2$ageclass, levels=c("29 or younger","30 to 44","45 to 59","60 and older"))
+lits2$ageclass <- factor(lits2$ageclass, 
+                         levels=c("29 or younger",
+                                  "30 to 44",
+                                  "45 to 59",
+                                  "60 and older"))
+
+summary(lits2$ageclass)
 
@@
 ###
+summary(factor(lits2$q226m))
 
-lits2$q226m_txt <- recode(lits2$q226m, 
-                          "1='Salary or wages in cash or in kind';
-                                2='Income from self-employment';
-                          3='Sales or bartering of farm products';
-                          4='Pensions';5='Benefits from the state';
-                          6='Help from relatives or friends';7='Other'")
+lits2$incsour3[lits2$q226m == 5 | lits2$q226m == 6] <- "dependent"
+lits2$incsour3[lits2$q226m > 0 & lits2$q226m < 5 | lits2$q226m > 6 ] <- "notDependent"
 
-lits2$incsour3 <- factor(recode(lits2$q226m_txt, 
-                                "c('Benefits from the state',
-                                'Help from relatives or friends')='Dependent'
-                                ;else='Not.dependent'"))
-
+lits2$incsour3 <- factor(lits2$incsour3, 
+                         levels=c("dependent",
+                                  "notDependent"))
+summary(lits2$incsour3)
 ###
@@
 
-lits2$edu2 <- factor(recode(lits2$q515, "c(1,2)='no or compulsory';
-                           else='higher'"))
+summary(factor(lits2$q515))
 
+lits2$edu2[lits2$q515 == 1 | lits2$q515 == 2]  <- "no or compulsory"
+lits2$edu2[lits2$q515 >= 3]  <- "higher"
+
+lits2$edu2 <- factor(lits2$edu2,
+                     levels=c("no or compulsory",
+                              "higher"))
+
+summary(lits2$edu2)
 #################
@@
 #################
+library(stringr)
 lits2$q801 <- as.factor(str_replace(lits2$q801, "Don't know", "DONT KNOW"))
+
 summary(lits2$q801)
-lits2$q801 <- factor(lits2$q801, levels=c("A GREAT DEAL","A FAIR AMOUNT","JUST A LITTLE","NOT AT ALL","DONT KNOW","Refused"))
 
-library(car)
-lits2$crise <- recode(lits2$q801, "c('A GREAT DEAL','A FAIR AMOUNT')='Great or fair amount';
-                       c('NOT AT ALL','JUST A LITTLE')='Little or not at all';
-                       else=NA")
+lits2$q801 <- factor(lits2$q801, 
+                     levels=c("A GREAT DEAL",
+                              "A FAIR AMOUNT",
+                              "JUST A LITTLE",
+                              "NOT AT ALL",
+                              "DONT KNOW",
+                              "Refused"))
+
+lits2$crise[lits2$q801 %in% c('A GREAT DEAL','A FAIR AMOUNT')] <- 'Great or fair amount'
+lits2$crise[lits2$q801 %in% c('NOT AT ALL','JUST A LITTLE')] <- 'Little or not at all'
 
@@
 summary(lits2$crise)
-###
-
 ```
@@
 ```{rLoadClean5, eval=TRUE}
-
-###
-
 ## Data for the analysis
 df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
-                             "q309_rec","sex",
+                             "sex",
                              "ageclass","age",
@@
                              "q229", # tulevaisuus
-                             "q226m_txt", # tulonlähde
+#                             "q226m_txt", # tulonlähde
                              "q515", # edu
-                             "q801" # crise,
+                             "q801", # crise,
+                             "poverty"
                              ))
@@
 df$cntry <- factor(df$cntry)
-# print data in csv-format
-
 summary(df)
@@
 ###
-
 library(reshape2)
@@
 
-summary309 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
+summary309 <- data.frame(prop.table(svytable(~q309 + 
+                                               cntry, d.df), 
+                                    2) * 100)
+
 summary309 <- summary309[!is.na(summary309$Freq), ]
+summary309w <- dcast(summary309, 
+                     cntry ~ q309, 
+                     value.var="Freq")
+names(summary309w) <- c("cntry","socialBlame",
+                        "individualBlame","individualFate",
+                        "socialFate","not.stated","dont.know")
 
-summary309w <- dcast(summary309, cntry ~ q309, value.var="Freq")
-names(summary309w) <- c("cntry","social.blame","individual.blame","individual.fate","social.fate","not.stated","dont.know")
+summaryPoverty <- data.frame(prop.table(svytable(~poverty + 
+                                                   cntry, d.df), 
+                                        2) * 100)
+summaryPoverty <- summaryPoverty[!is.na(summaryPoverty$Freq), ]
 
-summary309_rec <- data.frame(prop.table(svytable(~q309_rec + cntry, d.df), 2) * 100)
-summary309_rec <- summary309_rec[!is.na(summary309_rec$Freq), ]
-
-summary309_recw <- dcast(summary309_rec, cntry ~ q309_rec, value.var="Freq")
-names(summary309_recw) <- c("cntry","social.blame","individual.blame","individual.fate","social.fate")
+summaryPovertyw <- dcast(summaryPoverty, cntry ~ poverty, value.var="Freq")
+names(summaryPovertyw) <- c("cntry","socialBlame","individualBlame","individualFate","socialFate")
 
@@
 library(plyr)
-datQogSum <- ddply(datQog, c("cname","indicator"), subset, year == max(year))
+datQogSum <- ddply(datQog, c("cname","indicator"), 
+                   subset, year == max(year))
 
@@
 ggplot(datQogSum, aes(x=year,y=value,color=cname)) + 
-  geom_point() +
+  geom_point(size=4) +
   geom_text(data = datQogSum,
-            aes(x=year,y=value,color=cname,label=cname)) +
+            aes(x=year,y=value,
+                color=cname,
+                label=cname),
+            size=3.5) +
   facet_wrap(~indicator, scales="free") +
@@
 
-datQogSum.w <- dcast(datQogSum, cname ~ indicator, value.var="value")
-datQogSum.w$cname <- recode(datQogSum.w$cname, "'United Kingdom'='Great Britain'")
+datQogSum.w <- dcast(datQogSum, 
+                     cname ~ indicator, 
+                     value.var="value")
 
-datQogSum.w$group_fine <- recode(datQogSum.w$cname, "c('Armenia','Azerbaijan','Georgia') = 'Caucasus';
-                         c('Russia','Ukraine','Moldova','Belarus')= 'CIS North';
-                         c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')='Central Asia';
-                         c('Estonia','Latvia','Lithuania')='Baltic States';
-                         c('Poland','Czech Republic','Slovakia','Hungary','Slovenia',
-                         'Romania','Bulgaria')='CEE EU';
-                         c('Italy','Germany','Great Britain','Sweden','France')='West Europe';
-                         else='other'")
-datQogSum.w$group_general <- recode(datQogSum.w$cname,
-                       "c('France','Germany','Great Britain',
-                             'Italy','Sweden')='Western Europe';
-                       c('Czech Republic','Estonia','Hungary','Bulgaria',
+datQogSum.w$cname <-  as.character(datQogSum.w$cname)
+datQogSum.w$cname[datQogSum.w$cname %in% "United Kingdom"] <- "Great Britain"
+
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Georgia')] <- 'Caucasus'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Russia','Ukraine','Moldova','Belarus')] <- 'CIS North'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')] <- 'Central Asia'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Estonia','Latvia','Lithuania')] <- 'Baltic States'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Poland','Czech Republic','Slovakia','Hungary',
+                                                'Slovenia','Romania','Bulgaria')] <- 'CEE EU'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Italy','Germany','Great Britain','Sweden','France')] <- 'West Europe'
+
+datQogSum.w$group_fine <- factor(datQogSum.w$group_fine)
+
+summary(factor(datQogSum.w$group_fine))
+
+datQogSum.w$group_general[datQogSum.w$cname %in% c('France','Germany',
+                                                   'Great Britain','Italy','Sweden')] <- 'Western Europe';
+datQogSum.w$group_general[datQogSum.w$cname %in% c('Czech Republic','Estonia',
+                                                   'Hungary','Bulgaria',
                                                    'Latvia','Lithuania','Poland',
-                       'Slovakia','Slovenia','Romania')='CEE';
-                       c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
-                       'Kyrgyzstan','Moldova','Tajikistan',
-                       'Ukraine','Uzbekistan','Russia')='CIS';else=NA")
+                                                   'Slovakia','Slovenia','Romania')] <- 'CEE'
+datQogSum.w$group_general[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Belarus',
+                                                   'Georgia','Kazakhstan',
+                                                   'Kyrgyzstan','Moldova',
+                                                   'Tajikistan','Ukraine',
+                                                   'Uzbekistan','Russia')] <- 'CIS'
 
-###
+datQogSum.w$group_general <- factor(datQogSum.w$group_general)
+summary(datQogSum.w$group_general)
 
@@
 library(gdata)
-df.qog <- rename.vars(df.qog, from="cname", to="cntry")
-
+names(df.qog)[1] <- "cntry"
 ```
@@
 # WDIsearch(string='gdp', field='name', cache=NULL)
-wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), country = c("AL","AM","AZ","BY","BA","BG","HR",
+wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
+          country = c("AL","AM","AZ","BY","BA","BG","HR",
                       "CZ","EE","FR","GE","DE","GB","HU",
@@
 names(wb) <- c("iso2c","cntry","var","year")
-wb.w <- dcast(wb, cntry ~ year, value.var="var")
+wb.w <- dcast(wb, 
+              cntry ~ year, 
+              value.var="var")
 
@@
 
-```{rLoadClean9, eval=FALSE}
+```{rLoadClean9}
 

### 2013-11-01 regressionModelling.Rmd
restructuring the whole lot. Creative destruction

--- 2013-10-30 code/multilevel_models.Rmd
@@
 
-summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.8 <- lmer(pov.log.social ~ uw.gini+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.10 <- lmer(pov.log.social ~ wbgi.vae+(1|cntry),
-                         family = binomial, data=df))
+summary(bisoc.1 <- glmer(pov.log.social ~ incsour3+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.2 <- glmer(pov.log.social ~ edu2+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.3 <- glmer(pov.log.social ~ income2+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.4 <- glmer(pov.log.social ~ past.diff+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.5 <- glmer(pov.log.social ~ future.diff+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.6 <- glmer(pov.log.social ~ crise+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.7 <- glmer(pov.log.social ~ group1+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.8 <- glmer(pov.log.social ~ uw.gini+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.9 <- glmer(pov.log.social ~ gdpchange07.10+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(bisoc.10 <- glmer(pov.log.social ~ wbgi.vae+(1|cntry),
+                         family = binomial(logit), data=df))
 
-summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.8 <- lmer(pov.log.individual ~ uw.gini+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.9 <- lmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.10 <- lmer(pov.log.individual ~ wbgi.vae+(1|cntry),
-                         family = binomial, data=df))
+summary(biind.1 <- glmer(pov.log.individual ~ incsour3+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.2 <- glmer(pov.log.individual ~ edu2+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.3 <- glmer(pov.log.individual ~ income2+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.4 <- glmer(pov.log.individual ~ past.diff+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.5 <- glmer(pov.log.individual ~ future.diff+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.6 <- glmer(pov.log.individual ~ crise+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.7 <- glmer(pov.log.individual ~ group1+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.8 <- glmer(pov.log.individual ~ uw.gini+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.9 <- glmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.10 <- glmer(pov.log.individual ~ wbgi.vae+(1|cntry),
+                         family = binomial(logit), data=df))
 
-summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
-                      family = binomial, data=df))
+summary(soc.0 <- glmer(pov.log.social ~ (1|cntry),
+                      family = binomial(logit)(logit),
+                       data=df.multi))
 summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                          past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.2 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.3 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
 
-summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
-                      family = binomial, data=df))
-summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+summary(ind.0 <- glmer(pov.log.individual ~ (1|cntry),
+                      family = binomial(logit),
+                       data=df.multi))
+summary(ind.1 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                      family = binomial(logit),
+                       data=df.multi))
+summary(ind.2 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(ind.3 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(ind.4 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
 ```

### 2013-11-01 code/dv_tables.R

@@ -31,18 +31,18 @@ tbl2$group_general <- NULL
 #---------------------------------------------------#
 #---------------------------------------------------#
 
-t <- data.frame(prop.table(svytable(~group_general+cntry+q309_rec, d.df), 2)*100)
+t <- data.frame(prop.table(svytable(~group_general+cntry+poverty, d.df), 2)*100)
 t$Freq <- round(t$Freq, 2)
 t <- subset(t, Freq > 0)
-t2 <-  dcast(t, group_general+cntry ~ q309_rec, value.var="Freq")
-names(t2) <-  c("group_general","country","social.blame","individual.blame","individual.fate","social.fate")
-t2 <- t2[order(t2$group,-t2$social.blame),]
+t2 <-  dcast(t, group_general+cntry ~ poverty, value.var="Freq")
+names(t2) <-  c("group_general","country","socialBlame","individualBlame","individualFate","socialFate")
+t2 <- t2[order(t2$group,-t2$socialBlame),]
 ########
-y <- data.frame(prop.table(svytable(~group_general+q309_rec, d.df), 1)*100)
+y <- data.frame(prop.table(svytable(~group_general+poverty, d.df), 1)*100)
 y$Freq <- round(y$Freq, 2)
 y$country <- c("CEE Total","CIS Total","Western Europe Total")
-y2 <-  dcast(y, group_general+country ~ q309_rec, value.var="Freq")
-names(y2) <-  c("group_general","country","social.blame","individual.blame","individual.fate","social.fate")
+y2 <-  dcast(y, group_general+country ~ poverty, value.var="Freq")
+names(y2) <-  c("group_general","country","socialBlame","individualBlame","individualFate","socialFate")
 #
 #
 library(plyr)
@@ -50,25 +50,25 @@ co.var<-function(x)(100*sd(x)/mean(x))
 #
 func.CV <- function(t2)
 {
-  return(data.frame(CV.sb = co.var(t2$social.blame)))
+  return(data.frame(CV.sb = co.var(t2$socialBlame)))
 }
 df.CV.sb <- ddply(t2, .(group_general), func.CV)
 #
 func.CV <- function(t2)
 {
-  return(data.frame(CV.ib = co.var(t2$individual.blame)))
+  return(data.frame(CV.ib = co.var(t2$individualBlame)))
 }
 df.CV.ib <- ddply(t2, .(group_general), func.CV)
 #
[…]

### 2013-11-01 code/iv_indiv_table.R

@@ -1,9 +1,3 @@
-# text table
-
-Variable <- c("Transfer dependency","Low education","Perceived low income level","Income compared to past","Income compared to future","Effect of financial crises")
-Description <- c("Social or private transfers \n as main source of income","No or only compulsory level education","Perceived income level below the country median","Perceived change in income compared to situation in four years ago","Expected change in income over the next four years","Whether respondents perceives that their household has suffered the financial crisis great or fair amount")
-Varname <- c("q226m","q515","q227","q228","q229","q801")
-table <- data.frame(Variable,Description,Varname)
 
 # numeric table
 
@@ -13,59 +7,59 @@ d.df <- svydesign(id = ~SerialID,
                   weights = ~weight, 
                   data = df)
 # Country group
-t <- data.frame(prop.table(svytable(~group_general+q309_rec, d.df), 1)*100)
+t <- data.frame(prop.table(svytable(~group_general+poverty, d.df), 1)*100)
 t$Freq <- round(t$Freq, 1)
 t <- subset(t, Freq > 0)
 library(reshape2)
-t2 <-  dcast(t, group_general ~ q309_rec, value.var="Freq")
+t2 <-  dcast(t, group_general ~ poverty, value.var="Freq")
 names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
 cgroup <- t2
 # PAst difference
-t <- data.frame(prop.table(svytable(~past.diff+q309_rec, d.df), 1)*100)
+t <- data.frame(prop.table(svytable(~past.diff+poverty, d.df), 1)*100)
 t$Freq <- round(t$Freq, 1)
 t <- subset(t, Freq > 0)
 library(reshape2)
-t2 <-  dcast(t, past.diff ~ q309_rec, value.var="Freq")
+t2 <-  dcast(t, past.diff ~ poverty, value.var="Freq")
 names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
 past.diff <- t2
 # Future difference
-t <- data.frame(prop.table(svytable(~future.diff+q309_rec, d.df), 1)*100)
[…]

### 2013-11-01 code/iv_macro_table.R

@@ -5,27 +5,3 @@ tbl <- tbl[order(tbl$undp_hdi), ]
 names(tbl) <- c("country","GDP Change 2007-1010","Failed States Index",
                 "Voice And Accountability (WB)","HDI","Democracy (Freedom House)",
                 "Gini UNU Wider","Country group")
-
-
-Data <- c("Ghange in total GDP 2007 – 2010",
-          "Gini coefficient",
-          "Voice and Accountability")
-Description <- c("Relative change from 2007 to 2010",
-                 "measures the distribution of income or consumption expenditure among individuals or households within an economy deviates from a perfectly equal distribution",
-                 "indicators measuring the political process, civil liberties and political rights. These indicators measure the extent to which citizens of a country are able to participate in the selection of governments. Indicators measuring the independence of the media is also included.")
-Source <- c("World Bank (Quality of Government Data)",
-            "World Bank (Quality of Government Data)",
-            "World Bank (Quality of Government Data)")
-Varname <- c("wdi_gdp",
-             "uw_gini",
-             "wbgi_vae")
-# Year <- c("2011",
-#            "Latest available gini coefficient based either on income or consumption",
-#            "2011")
-macrotable <- data.frame(Data,Description,Source,Varname)
-
-
-# Data & Description & Source & Year
-# & $  $ & \href{http://data.worldbank.org/data-catalog/world-development-indicators}{World Bank World Development Indicators} & 2011 \\
-# &  & \href{http://www.wider.unu.edu/research/Database/en_GB/database/}{UNU-WIDER World Income Inequality Database, Version 2.0c, May 2008} & 2002-2006 \\
-# & . & \href{http://en.wikipedia.org/wiki/Democracy_Index#2011_rankings}{Economist Intelligence Unit} & 2011 \\

### 2013-11-01 code/multi_logit.R

@@ -6,33 +6,49 @@ load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
 df.multi <- subset(df, group_general %in% c("CEE","CIS"))
 
 library(lme4)
-summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
-                      family = binomial, data=df.multi))
+summary(soc.0 <- glmer(pov.log.social ~ (1|cntry),
+                      family = binomial(logit)(logit),
+                       data=df.multi))
 summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                          past.diff+future.diff+crise+(1|cntry),
-                       family = binomial, data=df.multi))
-summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.2 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw_gini+(1|cntry),family = binomial, data=df.multi))
-summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.3 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw_gini+gdpchange07.10+(1|cntry),family = binomial, data=df.multi))
-summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),family = binomial, data=df.multi))
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
 
 
-summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
-                      family = binomial, data=df.multi))
-summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+summary(ind.0 <- glmer(pov.log.individual ~ (1|cntry),
+                      family = binomial(logit),
+                       data=df.multi))
[…]

### 2013-11-05 article2013.Rmd
regression modelling. Just before going into multinomial regression

@@
 
-<link href="http://markuskainu.fi/material/css/article.css" rel="stylesheet" type="text/css" title="compact"></link>
-<link href='http://fonts.googleapis.com/css?family=Alegreya+SC|Dosis:400,200' rel='stylesheet' type='text/css'>
+<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
 
@@
 source("code/dv_tables.R")
+library(xtable)
 print.xtable(xtable(tbl5, 
                     caption="Weighted population shares of dependent variable by country",),
-              type="latex",
+              type="html",
               comment=FALSE, 
@@
 
-![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.pdf)
+![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)
 
@@
 
-![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.pdf)
+![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)
 
@@
 
-![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.pdf)
+![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)
 
@@
 ****
-
-```{rattribArticleMultilevel1,results='asis', eval=FALSE}
-source("~/workspace/lits/attrib/attrib_year2013/code/multi_logit.R")
-library(texreg)
-texreg(list(soc.0,
-       soc.1,
-       soc.2,
-       soc.3,
-       soc.4),
-       use.package=FALSE,
-       include.pvalues=TRUE,
-       ci.test=NULL,
-
-          caption="Logistic multilevel random intercept model for social blame type of explanation.", caption.above=TRUE, doctype = FALSE,
-        custom.coef.names=c("Intercept",
-                             "Transfer dependency (Not dependent)",
-                             "Education (No education)",
-                             "Perceived income (Low)",
-                             "Perceived change in income compared to four years ago (Has worsened)",
-                             "Expected change in income over next four years (Will worsen)",
-                             "Effect of financial crisis (Suffered great or fair amount)",
-                             "gini coefficient",
-                             "GDP change 2007-2010",
-                             "Voice And Accountability (WB)"),
-        custom.model.names=c("empty","Model SB1","Model SB2",
-                             "Model SB3","Model SB4"))
-
-```
-
-```{rattribArticleMultilevel2, results='asis', eval=FALSE}
-
-texreg(list(ind.0,
-          ind.1,
-          ind.2,
-          ind.3,
-          ind.4),
-       use.packages=FALSE,
-       include.pvalues=TRUE,
-       ci.test=NULL,
-       caption="Logistic multilevel random intercept model for individual blame type of explanation.", 
-       caption.above=TRUE, 
-        doctype = FALSE,
-                custom.coef.names=c("Intercept",
-                             "Transfer dependency (Not dependent)",
-                             "Education (No education)",
-                             "Perceived income (Low)",
-                             "Perceived change in income compared to four years ago (Has worsened)",
-                             "Expected change in income over next four years (Will worsen)",
-                             "Effect of financial crisis (Suffered great or fair amount)",
-                             "gini coefficient",
-                             "GDP change 2007-2010",
-                             "Voice And Accountability (WB)"),
-        custom.model.names=c("empty","Model IB1","Model IB2",
-                             "Model IB3","Model IB4"))
-```
 

### 2013-11-05 index.Rmd
regression modelling. Just before going into multinomial regression

@@
 
+`log of changes`
+
+date      comment
+--------  -------------------
+20131103  Regression modelling in process. Final models will be published on Tuesday afternoon
+
 # Introduction
@@
 html-ouput                                          source code
--------------------                                           ---------
+------------                                        --------------
 [loading, cleaning and summarizing](loadClean.html) [source](loadClean.Rmd)
@@
 html-ouput                                        source code
------------------------                                   ---------
+------------                                      --------------
 [Descriptive analysis](descriptive_analysis.html) [source](descriptive_analysis.Rmd)
+
+[Spatial visualization](mapPlots.html)            [source]((mapPlots.Rmd))
 
@@
 html-ouput                                                source code
------------------------                                   ---------
-[Regression analysis](code/multilevel_models.html)        [source](code/multilevel_models.Rmd)
+------------                                              --------------
+[Regression analysis](regressionModelling.html)           [source](regressionModelling.Rmd)
 

### 2013-11-05 loadClean.Rmd
regression modelling. Just before going into multinomial regression

@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
 ```
@@
 
-lits2$income <- factor(lits2$income2, 
+lits2$income2 <- factor(lits2$income2, 
                        levels=c("Low",
@@
 
-lits2$crise <- factor(lits2$crise, levels=c("Little or not at all",
-                                            "Great or fair amount"))
+lits2$crise <- factor(lits2$crise, levels=c("Great or fair amount",
+                                            "Little or not at all"))
 summary(lits2$crise)
@@
 
-library(ggplot2)
-ggplot(datQogSum, aes(x=year,y=value,color=cname)) + 
-  geom_point(size=4) +
-  geom_text(data = datQogSum,
-            aes(x=year,y=value,
-                color=cname,
-                label=cname),
-            size=3.5) +
-  facet_wrap(~indicator, scales="free") +
-  theme(legend.position="none")
+# library(ggplot2)
+# ggplot(datQogSum, aes(x=year,y=value,color=cname)) + 
+#   geom_point(size=4) +
+#   geom_text(data = datQogSum,
+#             aes(x=year,y=value,
+#                 color=cname,
+#                 label=cname),
+#             size=3.5) +
+#   facet_wrap(~indicator, scales="free") +
+#   theme(legend.position="none")
 
@@
 datQogSum.w$group_general[datQogSum.w$cname %in% c('France','Germany',
-                                                   'Great Britain','Italy','Sweden')] <- 'Western Europe';
+                                                   'Great Britain','Italy','Sweden')] <- 'Western Europe'
 datQogSum.w$group_general[datQogSum.w$cname %in% c('Czech Republic','Estonia',
@@
 datQogSum.w$group_general <- factor(datQogSum.w$group_general)
-summary(datQogSum.w$group_general)
 
@@
 #names(df.qog) <- names
-library(gdata)
 names(df.qog)[1] <- "cntry"

### 2013-11-05 mapPlots.Rmd
regression modelling. Just before going into multinomial regression

% Spatial visualization

- [back to index](index.html)

`generated at:`
```{r, echo=FALSE}
Sys.time()
```
---

<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/")
library(knitr)
knit("mapPlots.Rmd")
system("pandoc -s mapPlots.md -o mapPlots.html --toc --number-section")
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=15,fig.height=11)
```

# Country grouping

```{rmapPlotsLoadData}
library(RColorBrewer)
library(mapproj)
library(maptools)
library(rgdal)
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
                                layer="lits_all_simple"))

map$NAME <- as.character(map$NAME)
map$NAME[map$NAME == "United Kingdom"] <- 'Great Britain'
map$group_analysis[map$NAME %in% c('France','Germany',
                                        'Great Britain',
                                        'Italy','Sweden')] <- 'Western Europe'

map$group_analysis[map$NAME %in% c('Czech Republic','Estonia',
                                        'Hungary','Bulgaria',
                                        'Latvia','Lithuania',
                                        'Poland','Slovakia',
                                        'Slovenia','Romania')] <- 'CEE'

map$group_analysis[map$NAME %in% c('Armenia','Azerbaijan',
                                        'Belarus','Georgia',
                                        'Kazakhstan','Kyrgyzstan',
                                        'Moldova','Tajikistan',
                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'
map$NAME <- as.factor(map$NAME)

# shapefile into data.frame
map@data$id <- rownames(map@data)
library(ggplot2)
map.points <- fortify(map, region="id")
map.df <- merge(map.points, map@data, by="id")
# only east of longitude -25
map.df <- subset(map.df, long >= -15 & long <= 80)
# names for countries in analysis (excluding others)
namedata <- map.df[!is.na(map.df$group_analysis),]
cnames <- stats:::aggregate.formula(cbind(long, lat) ~ NAME, data=namedata, mean)
cbPalette <- c("#999999", "#E69F00",
               "#56B4E9", "#009E73",
               "#F0E442", "#0072B2",
               "gray90", "#CC79A7")
```

```{rmapPlotsGroup1}
ggplot(map.df, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
  scale_fill_manual(values=cbPalette) +
  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
  coord_map(project="orthographic") +
  theme(legend.position="top")
```

# Dependent variables

**The darker the higher the support**

## Support for Social blame

```{rmapPlotsDV1}
### Faceted showing population shares of explanations
library(scales)
plotMap <- merge(map.df,macro,by.x="NAME",by.y="cntry")
plotMap <- plotMap[order(plotMap$order),]

ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = socialBlame)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top")
```

## Individual blame

```{rmapPlotsDV2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = individualBlame)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top")
```

## Social fate

```{rmapPlotsDV3}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = socialFate)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top")
```

## Individual fate

```{rmapPlotsDV4}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = individualFate)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top")
```

## Not stated

```{rmapPlotsDV5}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = not.stated)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top")
```

## Dont know

```{rmapPlotsDV6}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = dont.know)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low="white", high=muted("blue")) +
  theme(legend.position="top")
```

# Independent variables

## Individual level

## Macro level

**Blue is "positive", red is "negative"**

### GDP change

```{rmapPlotsIvMacro1}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = gdpchange07.10)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("red"), high=muted("blue")) +
  theme(legend.position="top")
```

### Gini coefficient

```{rmapPlotsIvMacro2}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = uw_gini)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("blue"), high=muted("red"), midpoint=33.55) +
  theme(legend.position="top")
```

### Voice and Accountability

```{rmapPlotsIvMacro3}
ggplot(plotMap, aes(long,lat,group=group)) +
  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
  geom_polygon(aes(fill = wbgi_vae)) +
  geom_polygon(data = plotMap, aes(long,lat), 
               fill=NA, color = "white", size=0.1) +
  coord_map(project="orthographic") +
  scale_fill_gradient2(low=muted("red"), high=muted("blue")) +
  theme(legend.position="top")
```

### 2013-11-05 regressionModelling.Rmd
regression modelling. Just before going into multinomial regression

--- 2013-10-30 code/multilevel_models.Rmd
@@
+% Regression modellining
+
+- [back to index](index.html)
+
+`generated at:`
+```{r, echo=FALSE}
+Sys.time()
+```
+---
+
+<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
+
 ```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013/code/")
+setwd("~/workspace/lits/attrib/attrib_year2013/")
 library(knitr)
-knit("multilevel_models.Rmd")
-system("~/.cabal/bin/pandoc multilevel_models.md --number-sections --template=../tex_templates/dcolumn.latex -o multilevel_models.pdf")
+knit("regressionModelling.Rmd")
+system("pandoc -s regressionModelling.md -o regressionModelling.html --toc --number-section")
 
-system("~/.cabal/bin/pandoc -s multilevel_models.md -o multilevel_models.html")
-
-#system("perl -pi -e 's/footnotesize/tiny/g' multilevel_models.tex")
 ```
 
-<link href="http://markuskainu.fi/material/css/article_wide.css" rel="stylesheet" type="text/css" title="compact"></link>
+```{roptsattrib, echo=FALSE}
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
+```
 
-# Final tables
-
-```{rmultilevel_analysis, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, results='hide'}
+```{rmultilevelLoadData}
 library(lme4)
@@
 #########################################
-setwd("~/workspace/lits/attrib/attrib_year2013/code")
+setwd("~/workspace/lits/attrib/attrib_year2013/")
 
@@
 load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+df.model <- df[df$group_general != "Western Europe",]
 
-summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.8 <- lmer(pov.log.social ~ uw.gini+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.10 <- lmer(pov.log.social ~ wbgi.vae+(1|cntry),
-                         family = binomial, data=df))
-
-summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.8 <- lmer(pov.log.individual ~ uw.gini+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.9 <- lmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.10 <- lmer(pov.log.individual ~ wbgi.vae+(1|cntry),
-                         family = binomial, data=df))
-
-summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
-                      family = binomial, data=df))
-summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
-
-summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
-                      family = binomial, data=df))
-summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
 ```
 
-\newpage
+`log of changes`
 
-```{rmultilevel_social, results='asis', echo=FALSE, eval=FALSE, cache=TRUE}
-library(stargazer)
-stargazer("Empty model"=soc.0,
-          "Model sb1"=soc.1,
-          "Model sb2"=soc.2,
-          "Model sb3"=soc.3,
-          "Model sb4"=soc.4,
-          title="Logistic multilevel random intercept model for social blame type of explanation (Using stargazer-package (missing country level random effects)).")
+date      comment
+--------  -------------------
+20131103  Non multi-level logistic regression done but not interpret yet. Bivariate multilevel models the same. Multivariate multilevel models running over night.
+
+# Description
+
+This page includes everything related to modelling of the data. [Click here for the final tables](finalTables.html).
+
+In this paper we have two depentend variables (DV) that are both binomial
+
+1. support for social blame type of explanation
+2. support for individual blame type of explanation
+
+# Unilevel regression analysis
+
+Unilevel is used here for making distinction between multilevel analysis later. *In the analysis of this chapter **no** contextual level variables are used*.
+
+I'm using the University of California's resource [R Data Analysis Examples: Logit Regression](http://www.ats.ucla.edu/stat/r/dae/logit.htm) as a reference here.
+
+## Bivariate logistic regression
+
+### Explaining social blame
+
+```{rUnilevelBivarSoc}
+unbisoc.1 <- glm(pov.log.social ~ incsour3, data = df.model, family = "binomial")
+unbisoc.2 <- glm(pov.log.social ~ edu2, data = df.model, family = "binomial")
+unbisoc.3 <- glm(pov.log.social ~ income2, data = df.model, family = "binomial")
+unbisoc.4 <- glm(pov.log.social ~ past.diff, data = df.model, family = "binomial")
+unbisoc.5 <- glm(pov.log.social ~ future.diff, data = df.model, family = "binomial")
+unbisoc.6 <- glm(pov.log.social ~ crise, data = df.model, family = "binomial")
+
 ```
 
-```{rmultilevel_individual2_texreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
-htmlreg(list(ind.0,
-          ind.1,
-          ind.2,
-          ind.3,
-          ind.4),
+```{rUnilevelBivarSocOutput, results='asis'}
+library(texreg)
+
+model.names = c("Transfer Dependency", "Education level",
+                "Perceived income","Income compared to past",
+                "Income compared to future","Effect of financial crises")
+
+coef.names = c("(intercept)", 
+               "Dependent vs. non-dependent",
+               "Low education vs. non-low education",
+               "Low income vs. high income",
+               "Income has worsened vs. has not",
+               "Income will worsen vs. will not",
+               "Has affected great or fair amount vs. has not")
+
+htmlreg(list(unbisoc.1,unbisoc.2,
+             unbisoc.3,unbisoc.4,
+             unbisoc.5,unbisoc.6),
+        custom.model.names = model.names,
+        custom.coef.names = coef.names,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+       caption="Simple logistic regression on social blame", caption.above=TRUE)
+
+```
+
+#### Interpretation
+
+- all the predictors are statistically significant which means that the *null hypothesis* of 
+    - The probability of supporting the social blame type of explanation decreases by `-0.16` if we compare individual who is dependent on tranfers with individual who is not dependent
+    - The probability of supporting the social blame type of explanation decreases by `-0.08` if we compare individual with low education with individual with no-low education
+    - The probability of supporting the social blame type of explanation decreases by `-0.59` if we compare individual with low perceived income with individual with high perceived income
+    - The probability of supporting the social blame type of explanation decreases by `-0.39` if we compare individual whose income has worsened during the financial crisis with individual whose income has not worsened
+    - The probability of supporting the social blame type of explanation decreases by `-0.49` if we compare individual who is expecting his income to worsen during the next four years with individual expects his income to remain at the same level or to increase
+    - The probability of supporting the social blame type of explanation decreases by `-0.46` if we compare individual who has suffered *Great or fair amount* from the financial crisis with individual who has suffered *Little or not at all*
+- As you would expect the model fit is the best in the case of *expected change in income*
+
+### Explaining individual blame
+
+```{rUnilevelBivarInd}
+unbiind.1 <- glm(pov.log.individual ~ incsour3, data = df.model, family = "binomial")
+unbiind.2 <- glm(pov.log.individual ~ edu2, data = df.model, family = "binomial")
+unbiind.3 <- glm(pov.log.individual ~ income2, data = df.model, family = "binomial")
+unbiind.4 <- glm(pov.log.individual ~ past.diff, data = df.model, family = "binomial")
+unbiind.5 <- glm(pov.log.individual ~ future.diff, data = df.model, family = "binomial")
+unbiind.6 <- glm(pov.log.individual ~ crise, data = df.model, family = "binomial")
+```
+
+```{rUnilevelBivarIndOutput, results='asis'}
+library(texreg)
+htmlreg(list(unbiind.1,unbiind.2,
+             unbiind.3,unbiind.4,
+             unbiind.5,unbiind.6),
+        custom.model.names = model.names,
+        custom.coef.names = coef.names,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+       caption="Simple logistic regression on individual blame", caption.above=TRUE)
+```
+
+#### Interpretation
+
+- all the predictors, but *education level* are statistically significant
+    - The probability of supporting the individual blame type of explanation increases by `0.49` if we compare individual who is dependent on tranfers to individual who is not dependent
+    - The probability of supporting the individual blame type of explanation increases by `0.03` if we compare individual with low education with individual with no-low education.
+*This is not statistically significant*
+    - The probability of supporting the individual blame type of explanation increases by `0.38` if we compare individual with low perceived income with individual with high perceived income
+    - The probability of supporting the individual blame type of explanation increases by `0.32` if we compare individual whose income has worsened during the financial crisis with individual whose income has not worsened
+    - The probability of supporting the individual blame type of explanation increases by `0.40` if we compare individual who is expecting his income to worsen during the next four years with individual expects his income to remain at the same level or to increase
+    - The probability of supporting the individual blame type of explanation increases by `-0.53` if we compare individual who has suffered *Great or fair amount* from the financial crisis with individual who has suffered *Little or not at all*
+- As you would expect the model fit is the best in the case of *expected change in income*
+
+## Multivariate logistic regression
+
+### Social blame
+
+```{rUnilevelMulvarSoc}
+mulbisoc.1 <- glm(pov.log.social ~  incsour3, data = df.model, family = "binomial")
+mulbisoc.2 <- glm(pov.log.social ~  incsour3 +
+                                    edu2, data = df.model, family = "binomial")
+mulbisoc.3 <- glm(pov.log.social ~  incsour3 +
+                                    edu2 +
+                                    income2, data = df.model, family = "binomial")
+mulbisoc.4 <- glm(pov.log.social ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff, data = df.model, family = "binomial")
+mulbisoc.4 <- glm(pov.log.social ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff, data = df.model, family = "binomial")
+mulbisoc.5 <- glm(pov.log.social ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff + 
+                                    future.diff, data = df.model, family = "binomial")
+mulbisoc.6 <- glm(pov.log.social ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff + 
+                                    future.diff +
+                                    crise, data = df.model, family = "binomial")
+
+```
+
+```{rUnilevelMulvarSocOutput, results='asis'}
+library(texreg)
+htmlreg(list(mulbisoc.1,mulbisoc.2,
+             mulbisoc.3,mulbisoc.4,
+             mulbisoc.5,mulbisoc.6),
+        custom.coef.names = coef.names,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+       caption="Multiple logistic regression on social blame", caption.above=TRUE)
+
+```
+
+#### Diagnostics of the final model
+
+```{r}
+confint(mulbisoc.6)
+```
+
+```{r}
+exp(cbind(OR = coef(mulbisoc.6), confint(mulbisoc.6)))
+```
+
+```{r}
+newdata2 <- with(df.model, data.frame(gre = rep(seq(from = 200, to = 800, length.out = 100),
+    4), gpa = mean(gpa), rank = factor(rep(1:4, each = 100))))
+
+newdata3 <- cbind(newdata2, predict(mylogit, newdata = newdata2, type = "link",
+    se = TRUE))
+newdata3 <- within(newdata3, {
+    PredictedProb <- plogis(fit)
+    LL <- plogis(fit - (1.96 * se.fit))
+    UL <- plogis(fit + (1.96 * se.fit))
+})
+
+ggplot(newdata3, aes(x = gre, y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
+    ymax = UL, fill = rank), alpha = 0.2) + geom_line(aes(colour = rank),
+    size = 1)
+
+```
+
+#### Interpretation
+
+- all the predictors, but *education level* are statistically significant
+    - The probability of supporting the individual blame type of explanation increases by `0.49` if we compare individual who is dependent on tranfers to individual who is not dependent
+    - The probability of supporting the individual blame type of explanation increases by `0.03` if we compare individual with low education with individual with no-low education.
+*This is not statistically significant*
+    - The probability of supporting the individual blame type of explanation increases by XXX` if we compare individual with low perceived income with individual with high perceived income
+    - The probability of supporting the individual blame type of explanation increases by `0.32` if we compare individual whose income has worsened during the financial crisis with individual whose income has not worsened
+    - The probability of supporting the individual blame type of explanation increases by `0.40` if we compare individual who is expecting his income to worsen during the next four years with individual expects his income to remain at the same level or to increase
+    - The probability of supporting the individual blame type of explanation increases by `-XXX` if we compare individual who has suffered *Great or fair amount* from the financial crisis with individual who has suffered *Little or not at all*
+- As you would expect the model fit is the best in the case of *expected change in income*
+
+### Individual blame
+
+```{rUnilevelMulvarInd}
+mulbiind.1 <- glm(pov.log.individual ~  incsour3, data = df.model, family = "binomial")
+mulbiind.2 <- glm(pov.log.individual ~  incsour3 +
+                                    edu2, data = df.model, family = "binomial")
+mulbiind.3 <- glm(pov.log.individual ~  incsour3 +
+                                    edu2 +
+                                    income2, data = df.model, family = "binomial")
+mulbiind.4 <- glm(pov.log.individual ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff, data = df.model, family = "binomial")
+mulbiind.4 <- glm(pov.log.individual ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff, data = df.model, family = "binomial")
+mulbiind.5 <- glm(pov.log.individual ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff + 
+                                    future.diff, data = df.model, family = "binomial")
+mulbiind.6 <- glm(pov.log.individual ~  incsour3 +
+                                    edu2 +
+                                    income2 +
+                                    past.diff + 
+                                    future.diff +
+                                    crise, data = df.model, family = "binomial")
+```
+
+```{rUnilevelMulvarIndOutput, results='asis'}
+library(texreg)
+htmlreg(list(mulbiind.1,mulbiind.2,
+             mulbiind.3,mulbiind.4,
+             mulbiind.5,mulbiind.6),
+        custom.coef.names = coef.names,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+       caption="Multiple logistic regression on individual blame", caption.above=TRUE)
+```
+
+#### Interpretation
+
+# Multinomial logistic regression
+
+In multinomial setting we use variable `poverty`as DV. `Dont know` and `not stated` values are exluded from the analysis and we are left with following four values:
+
+- Social Blame 
+- Individual Blame
+- Individual Fate
+- Social Fate
+
+Of the DV values we use `Social Blame` as a reference category as it is distinctive category and allows me to make clear interpretations of the significance of the predictors.
+([source](http://stats.stackexchange.com/questions/34574/is-there-any-criterion-about-choosing-reference-factor-in-multinomial-logistic-r))
+
+## Simple multinomial logistic regression
+
+```{rmultiNomSimple}
+
+```
+
+```{rmultiNomSimpleOutput, results='asis'}
+
+```
+
+### Interpretation
+
+## Multivariate multinomial logistic regression
+
+```{rmultiNomMulti}
+
+```
+
+```{rmultiNomMultiOutput, results='asis'}
+
+```
+
+### Interpretation
+
+# Multilevel regression
+
+## Bivariate analysis
+
+### Social blame explained
+
+```{rmultilevelBivarSoc}
+bisoc.1 <- glmer(pov.log.social ~ incsour3+(1|cntry),
+                        family = binomial(logit), data = df.model)
+bisoc.2 <- glmer(pov.log.social ~ edu2+(1|cntry),
+                        family = binomial(logit), data = df.model)
+bisoc.3 <- glmer(pov.log.social ~ income2+(1|cntry),
+                        family = binomial(logit), data = df.model)
+bisoc.4 <- glmer(pov.log.social ~ past.diff+(1|cntry),
+                        family = binomial(logit), data = df.model)
+bisoc.5 <- glmer(pov.log.social ~ future.diff+(1|cntry),
+                        family = binomial(logit), data = df.model)
+ bisoc.6 <- glmer(pov.log.social ~ crise+(1|cntry),
+                         family = binomial(logit), data = df.model)
+#  bisoc.7 <- glmer(pov.log.social ~ group_general+(1|cntry),
+#                          family = binomial(logit), data = df.model)
+ bisoc.8 <- glmer(pov.log.social ~ uw_gini+(1|cntry),
+                         family = binomial(logit), data = df.model)
+ bisoc.9 <- glmer(pov.log.social ~ gdpchange07.10+(1|cntry),
+                         family = binomial(logit), data = df.model)
+ bisoc.10 <- glmer(pov.log.social ~ wbgi_vae+(1|cntry),
+                          family = binomial(logit), data = df.model)
+
+```
+
+```{rregressionBivarSocOutput, results='asis'}
+
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      #"Country grouping (CIS vs. CEE)",
+                      "Gini coefficient",
+                      "Change in GDP between 2007 to 2010",
+                      "Good governance")
+
+library(texreg)
+htmlreg(list(bisoc.1,bisoc.2,
+          bisoc.3,bisoc.4,
+          bisoc.5,bisoc.6,
+          #bisoc.7,
+          bisoc.8,
+          bisoc.9,bisoc.10),
+        custom.coef.names = coef.names.multi,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+        include.pvalues=TRUE,
+        caption="Bivariate social blame", caption.above=TRUE)
+
+```
+
+#### Interpretation
+
+### Individual blame explained
+
+```{rmultilevelBivarInd}
+biind.1 <- glmer(pov.log.individual ~ incsour3+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.2 <- glmer(pov.log.individual ~ edu2+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.3 <- glmer(pov.log.individual ~ income2+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.4 <- glmer(pov.log.individual ~ past.diff+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.5 <- glmer(pov.log.individual ~ future.diff+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.6 <- glmer(pov.log.individual ~ crise+(1|cntry),
+                        family = binomial(logit), data = df.model)
+# biind.7 <- glmer(pov.log.individual ~ group_general+(1|cntry),
+#                         family = binomial(logit), data = df.model)
+biind.8 <- glmer(pov.log.individual ~ uw_gini+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.9 <- glmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
+                        family = binomial(logit), data = df.model)
+biind.10 <- glmer(pov.log.individual ~ wbgi_vae+(1|cntry),
+                         family = binomial(logit), data = df.model)
+```
+
+```{rmultilevelBivarIndOuput, results='asis'}
+library(texreg)
+htmlreg(list(biind.1,biind.2,
+          biind.3,biind.4,
+          biind.5,biind.6,
+          #biind.7,
+          biind.8,
+          biind.9,biind.10),
+        custom.coef.names = coef.names.multi,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
         use.packages=FALSE,
-       include.pvalues=TRUE,
-       caption="Logistic multilevel random intercept model for individual blame type of explanation (Using texreg-package (p-value stars missing)).", 
+        caption="Bivariate individual blame", 
         caption.above=TRUE)
@@
 
-```{rmultilevel_individual, results='asis', echo=FALSE, eval=FALSE, cache=TRUE}
-stargazer("Empty model"=ind.0,
-          "Model ib1"=ind.1,
-          "Model ib2"=ind.2,
-          "Model ib3"=ind.3,
-          "Model ib4"=ind.4,
-          title="Logistic multilevel random intercept model for individual blame type of explanation. (Using stargazer-package (missing country level random effects))")
+#### Interpretation
+
+## Multivariate multilevel regressions
+
+### Social blame type of explanation
+
+```{rmultilevelMultiSoc}
+library(lme4)
+soc.0 <- glmer(pov.log.social ~ (1|cntry),
+                      family = binomial(logit),
+                       data = df.model)
+soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+(1|cntry),
+                      family = binomial(logit),
+                      data = df.model)
+soc.2 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+soc.3 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+
 ```
 
-```{rmultilevel_social_texreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
+```{rmultilevelMultiSocOutput, results='asis'}
+model.names.multi = c("Empty",
+                      "Only individual",
+                      "Gini added",
+                      "GDP change added",
+                      "Good governance added")
+
 htmlreg(list(soc.0,
@@
         soc.4),
-       use.package=FALSE,
+        custom.model.names = model.names.multi,
+        custom.coef.names = coef.names.multi,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
         include.pvalues=TRUE,
-          caption="Logistic multilevel random intercept model for social blame type of explanation( Using texreg-package (p-value stars missing)).", caption.above=TRUE)
+          caption="Logistic multilevel random intercept model for social blame type of explanation", 
+        caption.above=TRUE)
 ```
 
-```{rbivariate_social, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
-library(texreg)
-htmlreg(list(bisoc.1,bisoc.2,
-          bisoc.3,bisoc.4,
-          bisoc.5,bisoc.6,
-          bisoc.7,bisoc.8,
-          bisoc.9,bisoc.10),
-       use.packages=FALSE,
-       sideways=TRUE,
-       scriptsize=TRUE,
-       include.pvalues=TRUE,
-       caption="bivatiate social blame", caption.above=TRUE)
+#### Interpretation
+
+### Individual blame type of explanation
+
+```{rmultilevelMultiInd, eval=FALSE}
+ind.0 <- glmer(pov.log.individual ~ (1|cntry),
+                      family = binomial(logit),
+                       data = df.model)
+ind.1 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+(1|cntry),
+                      family = binomial(logit),
+                       data = df.model)
+ind.2 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+ind.3 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+ind.4 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
 ```
 
-```{rbivariate_individual, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
-library(texreg)
-htmlreg(list(biind.1,biind.2,
-          biind.3,biind.4,
-          biind.5,biind.6,
-          biind.7,biind.8,
-          biind.9,biind.10),
-       use.packages=FALSE,
-       sideways=TRUE,
-       scriptsize=TRUE,
-       caption="bivatiate individual blame", caption.above=TRUE)
+```{rmultilevelMultiIndOutput, results='asis', eval=FALSE}
+htmlreg(list(soc.0,
+       ind.1,
+       ind.2,
+       ind.3,
+       ind.4),
+        custom.model.names = model.names.multi,
+        custom.coef.names = coef.names.multi,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+        include.pvalues=TRUE,
+          caption="Logistic multilevel random intercept model for social blame type of explanation", 
+        caption.above=TRUE)
 ```
+
+#### Interpretation

### 2013-11-06 code/multilevel_models.Rmd
All ready for Turku seminar. Now back in using the latest version of texreg 1.30.1. Thanks Philip!

@@
+% Multilevel modelling
+
 ```{r, eval=FALSE, echo=FALSE}
@@
 
-summary(bisoc.1 <- lmer(pov.log.social ~ incsour3+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.2 <- lmer(pov.log.social ~ edu2+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.3 <- lmer(pov.log.social ~ income2+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.4 <- lmer(pov.log.social ~ past.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.5 <- lmer(pov.log.social ~ future.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.6 <- lmer(pov.log.social ~ crise+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.7 <- lmer(pov.log.social ~ group1+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.8 <- lmer(pov.log.social ~ uw.gini+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.9 <- lmer(pov.log.social ~ gdpchange07.10+(1|cntry),
-                        family = binomial, data=df))
-summary(bisoc.10 <- lmer(pov.log.social ~ wbgi.vae+(1|cntry),
-                         family = binomial, data=df))
+summary(biind.1 <- glmer(pov.log.individual ~ incsour3+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.2 <- glmer(pov.log.individual ~ edu2+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.3 <- glmer(pov.log.individual ~ income2+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.4 <- glmer(pov.log.individual ~ past.diff+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.5 <- glmer(pov.log.individual ~ future.diff+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.6 <- glmer(pov.log.individual ~ crise+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.7 <- glmer(pov.log.individual ~ group1+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.8 <- glmer(pov.log.individual ~ uw.gini+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.9 <- glmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
+                        family = binomial(logit), data=df))
+summary(biind.10 <- glmer(pov.log.individual ~ wbgi.vae+(1|cntry),
+                         family = binomial(logit), data=df))
 
-summary(biind.1 <- lmer(pov.log.individual ~ incsour3+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.2 <- lmer(pov.log.individual ~ edu2+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.3 <- lmer(pov.log.individual ~ income2+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.4 <- lmer(pov.log.individual ~ past.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.5 <- lmer(pov.log.individual ~ future.diff+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.6 <- lmer(pov.log.individual ~ crise+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.7 <- lmer(pov.log.individual ~ group1+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.8 <- lmer(pov.log.individual ~ uw.gini+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.9 <- lmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
-                        family = binomial, data=df))
-summary(biind.10 <- lmer(pov.log.individual ~ wbgi.vae+(1|cntry),
-                         family = binomial, data=df))
-
-summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
-                      family = binomial, data=df))
+summary(soc.0 <- glmer(pov.log.social ~ (1|cntry),
+                      family = binomial(logit)(logit),
+                       data=df.multi))
 summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                          past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.2 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.3 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
 
-summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
-                      family = binomial, data=df))
-summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+summary(ind.0 <- glmer(pov.log.individual ~ (1|cntry),
+                      family = binomial(logit),
+                       data=df.multi))
+summary(ind.1 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                      family = binomial(logit),
+                       data=df.multi))
+summary(ind.2 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(ind.3 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(ind.4 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
                         past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
 ```

### 2013-11-06 index.Rmd
All ready for Turku seminar. Now back in using the latest version of texreg 1.30.1. Thanks Philip!

@@
 lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/figure
+
+cd ~/workspace/lits/attrib/attrib_year2013/turku2013
+lftp -u np16003 -e "mput *.html;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/turku2013
 ```
@@
 20131103  Regression modelling in process. Final models will be published on Tuesday afternoon
+20131105  Regression modelling almost finished. Bug in a software caused a delay. Will be fixed by Wed morning. Now the tables are there, but coef and model names are not customized.
+20131106  Regression modelling somewhat done for Turku seminar.
 
@@
 
-- [Final tables as A4 .pdf](finalTables.pdf)
-
 # Paper
@@
 - HSE Moscow - November 12, 2013
-- Jatkokoulutusseminaari Turku November 7, 2013
+- [Jatkokoulutusseminaari Turku November 7, 2013](turku2013/slidesTkuJatko20131107.html)
 - [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)

### 2013-11-06 regressionModelling.Rmd
All ready for Turku seminar. Now back in using the latest version of texreg 1.30.1. Thanks Philip!

@@
 ```{rmultilevelLoadData}
-library(lme4)
 #########################################
@@
 20131103  Non multi-level logistic regression done but not interpret yet. Bivariate multilevel models the same. Multivariate multilevel models running over night.
+20131105  Regression modelling almost finished. Bug in a software caused a delay. Will be fixed by Wed morning. Now the tables are there, but coef and model names are not customized.
+20131106  Model names fixed. Multinomial models coming by midday
 
@@
 
-#### Diagnostics of the final model
-
-```{r}
-confint(mulbisoc.6)
-```
-
-```{r}
-exp(cbind(OR = coef(mulbisoc.6), confint(mulbisoc.6)))
-```
-
-```{r}
-newdata2 <- with(df.model, data.frame(gre = rep(seq(from = 200, to = 800, length.out = 100),
-    4), gpa = mean(gpa), rank = factor(rep(1:4, each = 100))))
-
-newdata3 <- cbind(newdata2, predict(mylogit, newdata = newdata2, type = "link",
-    se = TRUE))
-newdata3 <- within(newdata3, {
-    PredictedProb <- plogis(fit)
-    LL <- plogis(fit - (1.96 * se.fit))
-    UL <- plogis(fit + (1.96 * se.fit))
-})
-
-ggplot(newdata3, aes(x = gre, y = PredictedProb)) + geom_ribbon(aes(ymin = LL,
-    ymax = UL, fill = rank), alpha = 0.2) + geom_line(aes(colour = rank),
-    size = 1)
-
-```
-
 #### Interpretation
@@
 
-```{rmultiNomSimple}
+<!--
+
+### using nnet-package
+
+This model follows tutorial by UCLA [Multinomial regression](http://www.ats.ucla.edu/stat/r/dae/mlogit.htm) and it uses [nnet](http://cran.r-project.org/web/packages/nnet/index.html)-package.
+
+-->
+
+```{rmultiNomSimplennet, eval=FALSE, echo=FALSE}
+library(nnet)
+df.model$poverty2 <- relevel(df.model$poverty, ref = "Social Blame")
+test <- multinom(poverty2 ~ incsour3 + edu2 + income2 +
+                   past.diff + future.diff + crise, data = df.model)
+#library(texreg)
+#screenreg(list(test))
+summary(test)$coefficients/summary(test)$standard.errors
+
+# The ratio of the probability of choosing one outcome category over the
+# probability of choosing the baseline category is often referred as relative
+# risk (and it is also sometimes referred as odds as we have just used to
+# described the regression parameters above). The relative risk is the right-hand
+# side linear equation exponentiated, leading to the fact that the exponentiated
+# regression coefficients are relative risk ratios for a unit change in the
+# predictor variable. We can exponentiate the coefficients from our model to see
+# these risk ratios.
+
+exp(coef(test))
 
@@
 
-```{rmultiNomSimpleOutput, results='asis'}
+### using mlogit-package
 
+This follows the example from [University of Toronto](http://www.utstat.toronto.edu/~brunner/oldclass/312f12/lectures/MultinomialLogitWithR.pdf) and [Discrete-Choice Logit Models with R by Philip A.
+Viton (October 26, 2013)](http://facweb.knowlton.ohio-state.edu/pviton/courses2/crp5700/5700-mlogit.pdf)
+
+```{rmultiNomSimpleMlogit, results='markup'}
+library(mlogit)
+dat.mlogit <- df.model[,c("poverty","edu2","incsour3","income2",
+                          "past.diff","future.diff","crise")]
+ dat.mlogit <- mlogit.data(dat.mlogit, shape="wide", choice="poverty")
+# 
+# 
+# model.mlogit1 <- mlogit(poverty ~ 0 | incsour3,
+#                                data = dat.mlogit, 
+#                                reflevel ="Social Blame")
+# model.mlogit2 <- mlogit(poverty ~ 0 | incsour3 + edu2,
+#                                data = dat.mlogit, 
+#                                reflevel ="Social Blame")
+# model.mlogit3 <- mlogit(poverty ~ 0 | incsour3 + edu2 + 
+#                                       past.diff,
+#                                data = dat.mlogit, 
+#                                reflevel ="Social Blame")
+# model.mlogit4 <- mlogit(poverty ~ 0 | incsour3 + edu2 + 
+#                                       past.diff + future.diff,
+#                                data = dat.mlogit, 
+#                                reflevel ="Social Blame")
+model.mlogit5 <- mlogit(poverty ~ 0 | incsour3 + edu2 + 
+                                      past.diff + future.diff +
+                                      crise, 
+                               data = dat.mlogit, 
+                               reflevel ="Social Blame")
+
+summary(model.mlogit5)
 ```
@@
 
-## Multivariate multinomial logistic regression
+```{r, eval=FALSE, echo=FALSE}
+
+screenreg(list(model.mlogit1,model.mlogit2))
+
+f <- coeftest(model.mlogit)
+xtable(as.data.frame(f))model1 <- multinom(formula = reason ~ supu+income, data = ml.06)
+
+summary(m)
+coeftest(m)
+
+X <- mlogit.data(ml.10, shape="wide", choice="poverty")
+coeftest(mlogit(poverty ~ 0 | success,
+               data = X, reflevel ="Because of injustice in our society"))
+summary(mlogit(poverty ~ 0 | success,
+                data = X, reflevel ="Because of injustice in our society"))
+
+coef(mlogit(poverty ~ 0 | success,
+               data = X, reflevel ="Because of injustice in our society"))
+
+```
+
+## Multilevel multinomial logistic regression
+
+Can be modelled in R in Bayesian manner using either [ChoiceModelR](http://cran.r-project.org/web/packages/ChoiceModelR/index.html) and  [Zelig](http://cran.r-project.org/web/packages/Zelig/).
+
+Marco Steenbergen discusses the stat theory and imlementation in Stata in this video
+
+<iframe width="420" height="315" src="//www.youtube.com/embed/q_t9VerQ2ko" frameborder="0" allowfullscreen></iframe>
 
@@
 
-### Interpretation
-
 # Multilevel regression
@@
 ```{rmultilevelBivarSoc}
+library(lme4)
 bisoc.1 <- glmer(pov.log.social ~ incsour3+(1|cntry),
@@
                          family = binomial(logit), data = df.model)
-#  bisoc.7 <- glmer(pov.log.social ~ group_general+(1|cntry),
-#                          family = binomial(logit), data = df.model)
  bisoc.8 <- glmer(pov.log.social ~ uw_gini+(1|cntry),
@@
                           family = binomial(logit), data = df.model)
-
 ```
 
-```{rregressionBivarSocOutput, results='asis'}
+```{rmultilevelBivarSocOutputNOCI, results='asis'}
 
@@
                       "Income will worsen","Has affected great or fair amount",
-                      #"Country grouping (CIS vs. CEE)",
                       "Gini coefficient",
@@
           bisoc.5,bisoc.6,
-          #bisoc.7,
           bisoc.8,
@@
         custom.coef.names = coef.names.multi,
+        naive = TRUE,
         inline.css = FALSE,
@@
                         family = binomial(logit), data = df.model)
-# biind.7 <- glmer(pov.log.individual ~ group_general+(1|cntry),
-#                         family = binomial(logit), data = df.model)
 biind.8 <- glmer(pov.log.individual ~ uw_gini+(1|cntry),
@@
 
-```{rmultilevelBivarIndOuput, results='asis'}
+```{rmultilevelBivarIndOuputNOCI, results='asis'}
 library(texreg)
@@
           biind.5,biind.6,
-          #biind.7,
           biind.8,
@@
         use.packages=FALSE,
+        include.pvalues=TRUE,
         caption="Bivariate individual blame", 
@@
 library(lme4)
+
 soc.0 <- glmer(pov.log.social ~ (1|cntry),
@@
 
-```{rmultilevelMultiSocOutput, results='asis'}
+```{rmultilevelMultiSocOutputNOCI, results='asis'}
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      "Gini coefficient",
+                      "Change in GDP between 2007 to 2010",
+                      "Good governance")
+
 model.names.multi = c("Empty",
@@
         caption.above=TRUE)
+
 ```
@@
 
-```{rmultilevelMultiInd, eval=FALSE}
+```{rmultilevelMultiInd}
+
 ind.0 <- glmer(pov.log.individual ~ (1|cntry),
@@
                        data = df.model)
+
 ```
 
-```{rmultilevelMultiIndOutput, results='asis', eval=FALSE}
-htmlreg(list(soc.0,
+```{rmultilevelMultiIndOutputNOCI, results='asis'}
+
+htmlreg(list(ind.0,
        ind.1,
@@
         custom.coef.names = coef.names.multi,
+        naive = TRUE,
+        ci.test = NULL,
         inline.css = FALSE,
@@
         caption.above=TRUE)
+
 ```
 
-#### Interpretation
+### Comparison of full models with our last model
+
+```{rmultilevelMultivarFull}
+
+soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+
+ind.4 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+
+df.model$pov.log.socVsInc[df.model$poverty == "Social Blame"] <- 1
+df.model$pov.log.socVsInc[df.model$poverty == "Individual Blame"] <- 0
+
+socVsInc <- glmer(pov.log.socVsInc ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+
+```
+
+```{rrmultilevelMultivarFullOutput, results='asis', eval=TRUE}
+model.names.multi.2 = c("Social blame vs. not",
+                      "Individual blame vs. not",
+                        "Social blame vs. Individual blame")
+
+library(texreg)
+htmlreg(list(soc.4,ind.4,socVsInc),
+        naive = TRUE,
+        custom.model.names = model.names.multi.2,
+        custom.coef.names = coef.names.multi,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+        include.pvalues=TRUE,
+        caption="Logistic multilevel random intercept model for social and individual blame type of explanation", 
+        caption.above=TRUE)
+
+```

### 2013-11-08 article2013.Rmd
ennen kuin aloitan semman korjaukset

@@
 ```
-
-**Nootti**
-
-*Paperi on vielä hieman vaiheessa ja suosittelen tutustumaan tutkimuksen analytiikkaan osoitteessa: [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/).
-Yritän parhaillaan saada monitasomallintamista paremmalle tolalle.
-Paperissa olevat mallit ovat alustavia ja pitävät sisällään sekä teoreettista että menetelmällistä epätarkkuutta.*
-
-****
 
@@
 
-![Dimensions and types of explanations of poverty](figure/nelikentta.png)
+![Dimensions and types of explanations of poverty](figure/nelikentta.pdf)
 
@@
 
-![Countries used in the analysis](figure/mapplot.png)
+![Countries used in the analysis](figure/mapplot.pdf)
 
@@
 
-Perceived low income level  Perceived income level below the country median
+Perceived low income level  Perceived income level below the country median (median calculated from the 10 step scale)
 
@@
                     caption="Weighted population shares of dependent variable by country",),
-              type="html",
+              type="latex",
               comment=FALSE, 
@@
 
-![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)
+![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.pdf)
 
 >Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media.
-[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.png)
+[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf)
 
@@
 
-*There we some issues in converts regression outputs so the tables are evolving and can be browsed at: [http://research.muuankarski.org/attributions/code/multilevel_models.html](http://research.muuankarski.org/attributions/code/multilevel_models.html)*
+```{rrmultilevelMultivarFullOutput, results='asis', eval=TRUE}
+load("data/finalModels.RData")
 
-****
+model.names.multi.2 = c("Social blame vs. not",
+                      "Individual blame vs. not",
+                        "Social blame vs. Individual blame")
+
+library(texreg)
+texreg(list(soc.4,ind.4,socVsInc),
+        naive = TRUE,
+        custom.model.names = model.names.multi.2,
+        custom.coef.names = coef.names.multi,
+        inline.css = FALSE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+        include.pvalues=TRUE,
+        caption="Logistic multilevel random intercept model for social and individual blame type of explanation", 
+        caption.above=TRUE)
+
+```
 

### 2013-11-08 article2013.Rmd
Regression tables added and pandoc works ok. Booktabs and dcolumn would be nice

@@
 knit("article2013.Rmd")
+system("~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --latex-engine=xelatex -o article2013.pdf")
+
 system("~/.cabal/bin/pandoc article2013.md  --number-sections -o article2013.pdf")
+
 system("~/.cabal/bin/pandoc -s article2013.md --toc --number-sections -o article2013.html")
@@
 
-![Countries used in the analysis](figure/mapplot.pdf)
-
 We are using LiTS2 data from 21 transition countries with additional 4 Western European countries (France, Germany, Italy and Sweden) (See Figure 1).
@@
 print.xtable(xtable(tbl5, 
-                    caption="Weighted population shares of dependent variable by country",),
+                    caption="Weighted population shares of dependent variable by country",),table.placement="H",
               type="latex",
@@
               include.rownames=FALSE,
-             caption.placement="top")
+             caption.placement="top",)
 
@@
 
+\newpage
+
 ## Multilevel analysis
 
-```{rrmultilevelMultivarFullOutput, results='asis', eval=TRUE}
+*As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
+Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.*
+
+```{rarticleModels, results='asis', eval=TRUE}
 load("data/finalModels.RData")
 
-model.names.multi.2 = c("Social blame vs. not",
-                      "Individual blame vs. not",
-                        "Social blame vs. Individual blame")
+model.names.multi.2 = c("SB vs. not",
+                      "IB vs. not",
+                        "SB vs. IB")
+
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      "Gini coefficient",
+                      "Change in GDP between 2007 to 2010",
+                      "Good governance")
 
 library(texreg)
-texreg(list(soc.4,ind.4,socVsInc),
+texreg(list(soc.4,ind.4,socVsInc),float.pos="H",
+        use.packages=TRUE,
+        #booktabs=TRUE,
+        #dcolumn=TRUE,
+        #sideways=TRUE, 
         naive = TRUE, 
+        caption.above=TRUE,
+        doctype=FALSE,
         custom.model.names = model.names.multi.2,
         custom.coef.names = coef.names.multi,
-        inline.css = FALSE,
-        doctype = FALSE,
-        html.tag = FALSE,
-        head.tag = FALSE,
-        body.tag = FALSE,
         include.pvalues=TRUE,
-        caption="Logistic multilevel random intercept model for social and individual blame type of explanation", 
-        caption.above=TRUE)
+        caption="Logistic multilevel random intercept model for social and individual blame type of explanation")
 
 ```
-
-<!--
-
-As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
-Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
-
--->
 

### 2013-11-08 finalPlots.Rmd
Regression tables added and pandoc works ok. Booktabs and dcolumn would be nice

@@
 
-# GDP change
+# Scatterplots
+
+## GDP change
 
@@
   geom_point(data=west, aes(x=gdpchange07.10, 
-                          y=socialBlame, color=group_general)) +
+                          y=socialBlame, 
+                            color=group_general), alpha=.5) +
   geom_text(data=west, aes(x=gdpchange07.10, 
                            y=socialBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+            alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
   # west reference
@@
   geom_point(data=west, aes(x=gdpchange07.10, 
-                            y=individualBlame, color=group_general)) +
+                            y=individualBlame, 
+                            color=group_general), alpha=.5) +
   geom_text(data=west, aes(x=gdpchange07.10, 
                            y=individualBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+            alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
   # west reference
@@
 
-pdf("figure/scatter_gdp.pdf", width=14/2.54, height=10/2.54)
-grid.arrange(plot1,plot2, ncol=2)
+pdf("figure/scatter_gdp.pdf", width=14/2.54, height=22/2.54)
+grid.arrange(plot1,plot2, ncol=1)
 dev.off()
@@
 ppi <- 300
-png("figure/scatter_gdp.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
-grid.arrange(plot1,plot2, ncol=2)
+png("figure/scatter_gdp.png", width=14/2.54*ppi, height=22/2.54*ppi, res=ppi)
+grid.arrange(plot1,plot2, ncol=1)
 dev.off()
@@
 
-# UNU-Wider Gini
+## UNU-Wider Gini
 
@@
   geom_point(data=west, aes(x=uw_gini, 
-                            y=socialBlame, color=group_general)) +
+                            y=socialBlame, 
+                            color=group_general), alpha=.5) +
   geom_text(data=west, aes(x=uw_gini, 
                            y=socialBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+            alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
   # west reference
@@
   geom_point(data=west, aes(x=uw_gini, 
-                            y=individualBlame, color=group_general)) +
+                            y=individualBlame, 
+                            color=group_general), alpha=.5) +
   geom_text(data=west, aes(x=uw_gini, 
                            y=individualBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+            alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
   # west reference
@@
 
-pdf("figure/scatter_gini.pdf", width=14/2.54, height=10/2.54)
-grid.arrange(plot1,plot2, ncol=2)
+pdf("figure/scatter_gini.pdf", width=14/2.54, height=22/2.54)
+grid.arrange(plot1,plot2, ncol=1)
 dev.off()
@@
 ppi <- 300
-png("figure/scatter_gini.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
-grid.arrange(plot1,plot2, ncol=2)
+png("figure/scatter_gini.png", width=14/2.54*ppi, height=22/2.54*ppi, res=ppi)
+grid.arrange(plot1,plot2, ncol=1)
 dev.off()
@@
 
-# World Bank - Voice and Accountability
+## World Bank - Voice and Accountability
 
@@
   geom_point(data=west, aes(x=wbgi_vae, 
-                            y=socialBlame, color=group_general)) +
+                            y=socialBlame,
+                            color=group_general), alpha=.5) +
   geom_text(data=west, aes(x=wbgi_vae, 
                            y=socialBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+            alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
   # west reference
@@
   geom_point(data=west, aes(x=wbgi_vae, 
-                            y=individualBlame, color=group_general)) +
+                            y=individualBlame, 
+                            color=group_general), alpha=.5) +
   geom_text(data=west, aes(x=wbgi_vae, 
                            y=individualBlame, label=cntry),
-            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+            alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
   # west reference
@@
 
-pdf("figure/scatter_wbgi.pdf", width=14/2.54, height=10/2.54)
-grid.arrange(plot1,plot2, ncol=2)
+pdf("figure/scatter_wbgi.pdf", width=14/2.54, height=22/2.54)
+grid.arrange(plot1,plot2, ncol=1)
 dev.off()
@@
 ppi <- 300
-png("figure/scatter_wbgi.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
-grid.arrange(plot1,plot2, ncol=2)
+png("figure/scatter_wbgi.png", width=14/2.54*ppi, height=22/2.54*ppi, res=ppi)
+grid.arrange(plot1,plot2, ncol=1)
 dev.off()
@@
 ![](figure/scatter_wbgi.png)
-
-```{rattribFinalPlotsx, eval=FALSE}
-
-```

### 2013-11-09 article2013.Rmd
hieman epätoivoinen commit enne mikon tekstin lisäämistä

--- 2013-11-05 article2013.Rmd
@@
 
-As Joseph Schumpeter (1942: 12) has once argued, "attitudes are coins that do not readily melt".
+As Joseph Schumpeter (1942: 12) once argued, "attitudes are coins that do not readily melt".
 This means that normative expectations, justice beliefs or welfare perceptions are very hard to change.
 In this way, welfare attitudes can be associated with policy changes by highlighting public sentiments, which are normative background assumptions or ideas that constrain decision-making and institutional change by limiting the range of programme that decision-making elites are likely to perceive as acceptable and legitimate to both their constituents and themselves (Campbell, 2004: 96-100).
-On the other hand, "the worlds on welfare" literature on welfare attitudes has in turn emphasised that type of welfare policy affects public attitudes (Andress and Heien, 2001; Arts and Gelissen, 2001; Jaeger, 2006; Svallfors, 1997).
-In regard to public perceptions of the causes of poverty, in the selective welfare-policy dominated liberal regimes explanations of poverty are more often based on individual morality that they are in the Continental and Nordic welfare states – where external and structural explanations prevail (Albrekt Larsen, 2006: 47-56; Alesina and Gleaser, 2004).
+On the other hand, "the worlds on welfare" literature on welfare attitudes has in turn emphasized that type of welfare policy affects public attitudes (Andress and Heien, 2001; Arts and Gelissen, 2001; Jaeger, 2006; Svallfors, 1997).
+In regard to public perceptions of the causes of poverty, in the marginal welfare-policy dominated liberal regime, the poor will be asked to fulfil much harder deservingness criteria than in a universal social democratic regime (van Oorschot 2006).
+In countries, in which welfare programmes are more marginal and less generous, explanations of poverty are more often based on individual morality, reflecting the strength of the dominant individualistic ideology, than they are in the Continental and Nordic welfare states – where external and structural explanations prevail (Albrekt Larsen, 2006: 47-56; Alesina and Gleaser, 2004).
 
 The focus of this study is on the public perceptions of the causes of poverty in post-socialist countries.
-The development of post-socialist welfare state has received abundant scholarly attention over last years (e.g.
+Since the early 1990s, the development of post-socialist welfare state has received abundant scholarly attention (e.g.
 Cerami and Vanhuysse, 2009, Inglot, 2008, Haggard and Kauffman, 2008; Cook, 2007; Aidukaite, 2004; Deacon 1992).
-As the decades under socialist regime have had an undeniable influence on societies, these studies have emphasised the role of institutional differences within region as well as pre-socialist historical origins of the welfare states.
+As the decades under socialist regime have had an undeniable influence on societies, these studies have emphasized the role of institutional differences within region as well as pre-socialist historical origins of the welfare states.
 Now after 20 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka, 2009, Tomka, 2006, Cerami and Vanhuysse, 2009, Inglot, 2008).
 
-Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasised clear differences between Western and Eastern European countries (Gallie and Paugam, 2002; Kallio and Niemelä, 2010; Lepianka, 2007; van Oorschot and Halman, 2000).
-East Europeans are more inclined to blame either the individual behaviour or the flaws of the economic system of country than their Western counterparts.
-However, there is a shortage of comparative studies on attributions for poverty in particular focusing on the East Central European countries or on the whole post-socialist space.
-There is therefore a need for a detailed and more up-to-date analysis of the public perceptions of the causes of poverty in post-socialist countries.
+The purpose of this article is to examine cross-country differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries.
+Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and Eastern European countries (Gallie and Paugam, 2002; Kallio and Niemelä, 2013; Lepianka, 2007; van Oorschot and Halman, 2000).
+East Europeans are more inclined to blame either the individual behavior or the flaws of the economic system of country than their Western counterparts.
+Hence, it seems that unlike in Western Europe traditionally opposite attitudinal dimensions, i.e.
+structural and individualistic explanations of poverty coincide with each other in East Europeans perceptions.
+However, there is shortage of comparative studies on attributions for poverty in particular focusing on the East Central European countries or on the whole post-socialist space.
+There is therefore a need for a more detailed and up-to-date analysis of the public perceptions of the causes of poverty in post-socialist countries.
 
-Hence, utilising The Life in Transition Survey (LiTS) conducted jointly by the European Bank for Reconstruction and Development (EBRD) and the World Bank in 2010, the purpose of the article is to examine cross-country differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries.
-The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions.
+The novelty of this study is that it covers wide range of countries from Europe and Western Asia.
+Thus, compared to previous studies, it provides more valid evidence on the differences in perceptions between the East and the West and within post-socialist regime as well as on the importance of the post-socialist legacy.
+In addition, the study applies multilevel techniques in order to analyze the role of individual and contextual level explanatory factors as determinants of poverty attributions.
+The article explores to what extent specific risks related to household financial circumstances at the individual level and the state of democracy, income inequality and changes in national economy at the contextual level are associated with different types of lay explanations of poverty.
+Finally, extensive country-selection of our data provides a novel opportunity to examine the effects of the financial crises both at the individual and contextual level and their association with attributions for poverty.
+At the individual level, the study focuses not only on the individuals’ current financial situation.
+Instead, the article provides an empirical example of how past experiences, current financial position as well as future expectations concerning individual’s economic situation are associated with perceptions of the causes of poverty.
+
+The article is structured as follows: A brief overview of theoretical discourse on explanations of poverty and prior findings on cross-national differences in and contextual level determinants of attributions for poverty is presented in Section 2.
+Section 3 focuses on the self-interest approach as an individual level explanatory factor of the perceptions with particular interest of risk factors related to individual’s financial situation.
+Research design, survey data and methods are described in detail in Section 4.
+The results of the study are presented in Section 5.
+Finally, the main conclusions, implications and limitations of the study are discussed in Section 6.
 
@@
 Pioneering empirical studies on lay explanations of poverty emerged at the early 1970s when Joe R.
-Feagin (1972; 1975) examined Americans perceptions of the causes of poverty.
-He categorised reasons for poverty into three basic categories: individualistic, structural and fatalistic.
-Later, many factor analytic studies have given empirical support to Feagin’s categorisation (e.g.
+Feagin (1972; 1975) examined Americans' perceptions of the causes of poverty.
+He categorized reasons for poverty into three basic categories: individualistic, structural and fatalistic.
+Later, many factor analytic studies have given empirical support to Feagin’s categorization (e.g.
 Feather, 1974; Furnham, 1982; Hunt, 1996; Morçöl, 1997; Nasser et al., 2002; Niemelä, 2008).
-Yet some of the recent studies have expanded our understanding by incorporating more contemporary beliefs into the attributional scales.
-These recent works emphasise especially the relevance of cultural attributions for poverty (Bullock et al., 2003; Cozzarelli et al., 2001).
+Yet some studies have expanded our understanding by incorporating more contemporary beliefs into the attributional scales.
+These works emphasize especially the relevance of cultural attributions for poverty (Bullock et al., 2003; Cozzarelli et al., 2001).
 
-van Oorschot and Halman (2000) have suggested a four-tier typology (Table 1) in which individual blame emphasises internal factors, such as undesirable behaviour of the poor, moral looseness and laziness.
-Individual fate type of explanation emphasises that poverty can be caused by uncontrollable and inescapable internal factors.
+van Oorschot and Halman (2000) have suggested a four-tier typology (Table 1) in which individual blame emphasizes internal factors, such as undesirable behaviour of the poor, moral looseness and laziness.
+Individual fate type of explanation emphasizes that poverty can be caused by uncontrollable and inescapable internal factors.
 These are factors such as personal misfortune, illness or just bad luck.
@@
 Prior empirical findings on attributions for poverty in Europe have shown that there are some differences between the West and the East of Europe.
-Western Europeans are more inclined to support structural explanations for poverty than their Eastern counterparts.
-On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä, 2010; Kreidl, 2000; Lepianka, 2007; Lepianka et al., 2010; van Oorschot and Halman, 2000).
+Western Europeans are more inclined to support social explanations for poverty than their Eastern counterparts.
+On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä, 2013; Kreidl, 2000; Lepianka, 2007; Lepianka et al., 2010; van Oorschot and Halman, 2000).
 Earlier results suggest, however, that there are variations both within and across groups of countries.
-Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä, 2010).
+Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä, 2013).
 Also Lepianka et al.
@@
 Estonians are as likely to blame the flaws of the economic system as the individualistic reasons.
-In Russia, on the other hand, the economic system is blamed even more the individuals themselves (Gorshkov and Tikhonova, 2006; Stephenson, 2000).
+In Russia, on the other hand, the economic system is blamed even more than the individuals themselves (Gorshkov and Tikhonova, 2006; Stephenson, 2000).
 These results are consistent with Lepianka's (2007: 127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia.
@@
 At the same time, however, they often attributed the poverty of others to individual failure, such as laziness, alcoholism, having too many children to support, or having too few children to provide for their old age.
-They also reported that many of the respondents had become very distrustful of government officials and many suspected that government officials of exploiting the political and economic turmoil for their own gain, at the expense of ordinary citizens.
+Emphazising the positive association between horizontal and vertical trust (see e.g.
+Rothstein, 2000), they also reported that many of the respondents had become very distrustful of government officials and many suspected that government officials of exploiting the political and economic turmoil for their own gain, at the expense of ordinary citizens.
+With extensive country selection, the data utilized in this study provides a marvellous possibility to examine the association between attributions for poverty and the context regarding political process, civil liberties and political rights.
+We can assume that in those countries where the state of democracy is low, citizens blame more likely society for poverty than in those countries where citizens have political freedom and power to influence socio-political issues.
+
+There are also considerable differences in other aspects of socio-political context between examined countries.
+One of these aspects is the extent of income inequality which varies extensively between examined countries.
+Thus, this study examines the role of income inequality as a contextual-level explanatory factor of public perceptions of the causes of poverty.
+The link between inequality and attributions for poverty is not that easy to predict.
+One the one hand, the public in countries with greater inequality is more inclined to mistrust “others” (e.g.
+Uslander, 2002).
+Therefore, the uneven distribution of wealth should mean stronger support for the individualistic explanation for poverty.
+On the other hand, cross-national studies on attributions for poverty have shown that the support for the individual blame is lower in countries with lower level of income inequality (e.g.
+Gallie and Paugam, 2002; Kallio and Niemelä, 2013).
 
@@
 
-Besides the so called regime-effect, recent comparative studies on attributions for poverty have included country-level explanatory factors in the analysis (e.g.
-Albrekt Larsen, 2006; Gallie and Paugam, 2002; Lepianka, 2007; Lepianka et al., 2010).
-For instance, national-level economic conditions have been proved to be associated with welfare attitudes.
-Increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007).
-In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual laziness explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved.
-This result is confirmed by Kallio and Niemelä (2010).
-They found that the larger the economic growth in a country is, the more likely people are to blame the poor themselves.
+Besides the so called regime-effect, national-level economic conditions have been proved to be associated with welfare attitudes.
+This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence.
+In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crises in distinct ways.
+While many Western and Eastern European countries have witnessed economic decline, the economic performance in Western Asia has been strong.
+Hence, by analyzing the role of changes in economic performance during the period of global financial crises, the research design of this article provides empirical evidence on the effects of crises to public perceptions of the causes of poverty.
+Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007).
+In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved.
+This result is confirmed by Kallio and Niemelä (2013).
+They found that the larger the economic growth in a country is, more likely people are to blame the poor themselves.
 Thus, we can assume that changes in country’s economic performance are associated with attributions for poverty regardless of country grouping.
-This study analyses the role of changes in economic performance during the period of global financial crises.
-
-There are also considerable differences in socio-political context between the examined countries.
-The article analyses whether the level of income inequality and the state of democracy in a given country are associated with attributions for poverty.
-Regarding the state of democracy we can assume that in those countries where the state of democracy is low, citizens blame more likely society for poverty than in those countries where citizens have political freedom and power to influence socio-political issues.
-The link between inequality and attributions for poverty is not that easy to predict.
-On the one hand, the public in countries with greater inequality is more inclined to mistrust "others" (e.g.
-Uslander, 2002).
-Therefore, the uneven distribution of wealth should mean greater support for the individualistic explanations for poverty.
-On the other hand, cross-national studies on attributions for poverty have shown that the support for the individualistic explanation is lower in the countries with lower level of income inequality (e.g.
-Gallie and Paugam, 2002; Kallio and Niemelä, 2010).
 
@@
 
-In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalised by the individuals in the society in question (Taylor-Gooby, 1985; Svallfors, 1995).
+In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalized by the individuals in the society in question (Taylor-Gooby, 1985; Svallfors, 1995).
 More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice, 2001) and resource/class-based risks (Cusack et al.
@@
 Inspired by the studies which emphasise risks as the key mechanisms, this article tests to what extent factors related to financial hardship are associated – beyond the welfare regime – with perceptions of the causes of poverty.
-Following this line of reasoning we are interested in whether those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.
+Following this line of reasoning, we are interested in whether those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.
 Thus, the research design in regard to determinants of poverty attributions enables us to study from a fairly unexplored perspective the different mechanisms through which individuals who have a greater risk to experience material destitution perceive the causes of poverty.
@@
 Results have, however, shown that socio-demographic variables explain very little and they do not reveal clear or consistent patterns.
-On the other hand, there are studies which support the assumption that those who perceive themselves to be poor are more likely to agree with the external cause of poverty than people well above the poverty line (Niemelä, 2008; Saunders, 2002: 155–156).
+On the other hand, there are studies which support the self-interest hypothesis that those who perceive themselves to be poor are more likely to blame society and social structures for poverty than people well above the poverty line (Niemelä, 2008; Saunders, 2002: 155–156).
 Yet, there are also contradictory findings indicating that personal experience of poverty might lead to individualistic explanations.
@@
 Studies which have explored the importance of individual’s economic situation on attributions for poverty have focused on individual’s cross-sectional situation of financial circumstances.
-However, as suggested in the literature on welfare attitudes not only the risk patterns but also the way people perceive their future risks might have an effect on attitudes toward the welfare state (Andress and Heien, 2001; Boeri et al., 2001; Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).
+However, as suggested in the literature on welfare attitudes, not only the risk patterns but also the way people perceive their future risks might have an effect on attitudes toward the welfare state (Andress and Heien, 2001; Boeri et al., 2001; Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).
 In their European comparison Blomberg et al.
@@
 those who perceive their economic future as endangered were more in favour of state responsibility than those who do not expect their future situation to be difficult.
-Consequently, following the self-interest hypothesis, we can assume that those who perceive their future economic situation to be worse than now endorse more likely external reasons for poverty than those who do not expect difficulties in their future economic situation.
+Consequently, following the self-interest hypothesis, we can assume that those who perceive their future economic situation to be worse than now endorse more likely social blame type of explanation of poverty than those who do not expect difficulties in their future economic situation.
 
@@
 Therefore, we will analyse whether those who report that the economic crisis has affected their household at least a fair amount perceive the causes of poverty differently than those who report that it has had no or only a little effect to their household.
-According to self-interest hypothesis, we assume that those whose economic situation has declined and those whose household has affected by economic crisis are more inclined to support external causes of poverty than those who has not experienced economic problems.
+According to self-interest hypothesis, we assume that those whose economic situation has declined and those whose household has affected by economic crisis are more inclined to blame society for poverty than those who has not experienced economic problems.
 
@@
 Transfer dependency refers to a situation that the respondent’s main source of income relies on private or social transfers (Blomberg et al., 2012; Fridberg and Kangas, 2008).
-Because of their close relation with welfare system, it can be assumed that welfare recipients perceive more likely structural than individual reasons for poverty compared to the rest of the population.
+Because of their close relation with welfare system, it can be assumed that welfare recipients perceive more likely social blame than indivual blame type of explanation of poverty compared to the rest of the population.
 
@@
 Prior findings do not reveal consistent patterns regarding the effects of education.
-On the one hand, some studies have found the so-called inverted U-relationship which means that people with a middle level of education support the individualistic explanation, whereas people with a low or a high level of education are more likely to endorse structural explanations (Feagin, 1972; cf.
+On the one hand, some studies have found the so-called inverted U-relationship which means that people with a middle level of education support the individualistic explanation, whereas people with a low or a high level of education are more likely to endorse structural explanations emphazising social inequalities and the flaws of the economic system of a given country (Feagin, 1972; cf.
 Furnham, 1982).
-Moreover, the relationship between education and fatalistic explanation has been found to be negative (Morcöl, 1997).
 On the other hand, there are studies which suggest that people with a lower level of education are more likely to favour individualistic explanations.
@@
 
-# Purpose of the study
+# Research design
 
@@
 In addition, most of the prior research is based on attitude surveys from the 1990s and most of them have restricted country selection.
-Therefore, as previous studies have suggested (Kallio and Niemelä, 2010; Lepianka et al., 2010), there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space.
-
-Hence, the general purpose of the study is to analyse public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
-Based on the findings of previous studies, we assume that in general post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations.
-However, we hypothesise also that findings indicate relatively large variation between countries.
-
-As a country-level determinants, we will analyse, whether this variation can be explained by different regimes within post-socialist countries and the post-socialist legacy.
-The study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and CIS countries.
-The survey data enables us also to compare post-socialist countries to a selection of Western European nations.
-Thus, in order to examine the distinctiveness of post-socialist perception Western European countries are included in the descriptive analyses.
-Besides the country groups we will also analyse the role of changes in country’s economic performance during the period of global financial crises, income inequality and the state of democracy.
-
-As an individual-level determinants for the perceptions article test to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty, and whether these factors are associated with perceptions similar way regardless of the  groups of countries studied.
-Hereby, the research design aims at contributing to the theoretical reasoning regarding the importance of self-interest for attributions for poverty which assumes that those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.
-
-# Data and methods
+Therefore, as previous studies have suggested (Kallio and Niemelä, 2013; Lepianka et al., 2010), there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space.
+Hence, the general purpose of the study is to analyze public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
 
 The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank (WB) in 2010.
-The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample.
-More information regarding how households were sampled and people interviewed is available at ERBD, 2011.
-Second round of the survey had a special focus on how peoples lives were affected by the global economic crisis and it's aftermath.
-The dataset, documentation and publications are freely available at the survey website.
+The analyses cover 22 post-socialist countries and 4 Western European countries.
+The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see ERBD, 2011).
+The survey had a special focus on how people’s lives were affected by the global economic crises and its aftermath.
 
@@
 
-![Countries used in the analysis](figure/mapplot.png)
-
-We are using LiTS2 data from 21 transition countries with additional 4 Western European countries (France, Germany, Italy and Sweden) (See Figure 1).
-In total 23513 households from 25 countries are included in the analysis.
-Transition countries were broken down into two groups.
-CEE includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania; and CIS[^****] Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia.
-
-[^****]:Georgia withdrawn it's membership from CIS in 2008
-
-## Dependent variables
-
-The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?** (Variable `q309` in LiTS-data)
+The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?**
 
@@
 
-Thus, the question is a standard forced-choice question which is used also in other comparative surveys such as in European Values Study and in Eurobarometer.
+Thus, the question is a standard forced-choice question which is used also in other comparative surveys such as in World Values Survey and in Eurobarometer.
 According to van Oorschot and Halman (2000) different statements represent a four-tier typology of poverty explanations as follows: While the unlucky refers to individual fate, the laziness and lack of willpower represents individual blame.
-In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate.
-The option 'other' was treated as missing data.
+In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate (see Table 1).
+Based on the findings of previous studies, we assume that in general post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations.
+However, we hypothesize also that findings indicate relatively large variation between countries.
 
-As our main interest here is in support for social and blame type of explanation and as we are employing logistic multivariate methods, the answers were recoded in two separate binomial variables.
-In \emph{social blame} variable we have  value 1 for supporting social blame type of explanation and 0 for supporting other type of explanation.
-Same logic applies to variable \emph{individual blame}.
+As country-level determinants of perceptions, we will analyze, whether this variation can be explained by different regimes within post-socialist countries and the post-socialist legacy.
+The study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and Commonwealth of Independent States (CIS).
+CEE includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania; and CIS[^****] Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia.
+The survey data enables us also to compare post-socialist countries to a selection of Western European nations (France, Germany, Italy and Sweden).
+Thus, in order to examine the distinctiveness of post-socialist perception, Western European countries are included in the descriptive analyses.
+(See Figure 1.)
 
-## Individual level independent variables
+![Countries used in the analysis](figure/mapplot.png)
 
-Six individual-level risk predictors are included in the analysis.
+[^****]:Georgia withdrawn its membership from CIS in 2008
+
+Besides the country groups we will also analyze the role of changes in country’s economic performance during the period of global financial crises, income inequality and the state of democracy.
+As described in previous sections, we assume that in those countries where the economic crises has meant declining economic conditions, public are more likely to endorse social explanations for poverty than in countries where national economy has continued to grow despite the global economic turbulence.
+Based on the theoretical discussion on civil liberties, political rights and institutional trust, we assume that the public support to social blame type of explanation is high in those countries where the state of democracy is low.
+Due to contradictory findings on the direction of the effect of income inequality, it is more difficult to provide any straightforward hypothesis on the importance of income inequality.
+
+As an individual-level determinants for the perceptions article test to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty, and whether these factors are associated with perceptions similar way regardless of the  groups of countries studied.
+Hereby, the research design aims at contributing to the theoretical reasoning regarding the importance of self-interest for attributions for poverty which assumes that those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.
+Analyses include six individual-level predictors, namely transfer dependency, low education, perceived low income level, income-level compared to the past and expected change in income in the future as well as perceived effect of the financial crises.
 Variables are recoded as binomial where value 1 is coded for risk and 0 for absence of risk.
-The items included are described in the table 2.
+Both contextual and individual-level variables are described in the Table 2.
 
@@
 
-Table: Individual level predictors
+Table: Description of independent variables in the analyses.
 
@@
 ------------------------    -------------------------------------------------
+Individual-level
+
 Transfer dependency         Social or private transfers as main source of income
@@
 Effect of financial crises  Whether respondents perceives that their household has suffered the financial crisis great or fair amount
------------------------------------------------------------------------------
 
-## Country level independent variables
+Country-level
 
-An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of four variables: 1) change in total GDP in 2007 - 2010, 2) gini coefficient of economic inequality, 3) 2011 democracy index and 4) country grouping.
-Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), gini coefficients from [UNU-WIDER World Income Inequality Database](http://www.wider.unu.edu/research/Database/en_GB/database/) and Democracy Index from the [Economist Intelligence Unit](http://www.eiu.com/Default.aspx).
-Gini coefficient for each country is the latest available varying from 2002 to 2006.
-(see Appendix A)
-
-\newpage
-
-Table: Contextual level predictors
-
------------------------------------------------------------------------------
-Variable                          Description
-------------------------          -------------------------------------------------
 Ghange in total GDP 2007 – 2010   Relative change from 2007 to 2010
@@
 
-## Methods
+An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of four variables: 1) change in total GDP in 2007 - 2010, 2) gini coefficient of economic inequality, 3) 2011 democracy index and 4) country grouping.
+Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), gini coefficients from [UNU-WIDER World Income Inequality Database](http://www.wider.unu.edu/research/Database/en_GB/database/) and Democracy Index from the [Economist Intelligence Unit](http://www.eiu.com/Default.aspx).
+Gini coefficient for each country is the latest available varying from 2002 to 2006.
+(see Appendix A)
+
+TÄMÄ YLLÄ MAINITTU TEKSTI VOIDAAN KUITATA SÄRMÄSTI SILLÄ ETTÄ LISÄTÄÄN TUOHON TABLE 2:EN OIKEALLE PUOLELLE YKSI SARAKE OTSIKOLLA "DATA SOURCE" SIIHEN SITTEN INDIVIDUAL-LEVEL MUUTTUJIEN KOHDALLE JOKA RIVILLE LIFE IN TRANSITION SURVEY JA MAKROMUUTTUJIEN KOHDALLE TYYLIIN WORLD BANK 2013 JA SITTEN LÄHTEISIIN LÄHDEVIITEMERKINNÄT NÄILLE ELI WORLD BANK (2013) JA ONLINEVIITTAUS LEHDEN SÄÄTÖJEN MUKAISESTI.
 
@@
 [^*****]:R-code for data transformation and analysis is available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/)
+
+TÄTÄ YLLÄ MAINITTUA ALAVIITETTÄ EI VARMAANKAAN SAA LAITTAA SUBMITTED-VERSIOON. SE VOIDAAN HALUTTAESSA LISÄTÄ SITTEN KUN PAPERI ON HYVÄKSYTTY LEHTEEN.
 
@@
 
+Rothstein, Bo (2000) Trust, Social Dilemmas and Collective Memories. Journal of Theoretical Politics, 12 (4), 477–501.
+
 Rothstein, Bo and Stolle, Dietlind (2008) The State and Social Capital: An Institutional Theory of Generalized Trust. Comparative Politics 40 (4), 441-459.

### 2013-11-09 article2013.Rmd
Tein takautuvasti vielä muutokset artikkelin taulukkokoodeihin ja poistin jotain mitä olin jo poistanut. (siis olisi pitänyt committaa uusimmat muutokset ennen mikko branchin yhdistämistä)

@@
 knit("article2013.Rmd")
+system("~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --latex-engine=xelatex -o article2013.pdf")
 system("~/.cabal/bin/pandoc article2013.md  --number-sections -o article2013.pdf")
+
 system("~/.cabal/bin/pandoc -s article2013.md --toc --number-sections -o article2013.html")
@@
 ------------------------    -------------------------------------------------
-Individual-level
+**Individual-level**
 
@@
 
-Country-level
+**Country-level**
 
@@
 print.xtable(xtable(tbl5, 
-                    caption="Weighted population shares of dependent variable by country",),
-              type="html",
+                    caption="Weighted population shares of dependent variable by country",),table.placement="H",
+             type="latex",
              comment=FALSE, 
@@
 
-![Change in total GDP in 2007 - 2010 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gdp.png)
+![Change in total GDP in 2007 - 2010 and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_gdp.pdf)
 
@@
 
-![Gini coefficient and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_gini.png)
+![Gini coefficient and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_gini.pdf)
 
@@
 
-![World Bank's Voice and Accountability index in 2011 and support for social (left) and individual (right) blame types of attributions for poverty](figure/scatter_wbgi.png)
+![World Bank's Voice and Accountability index in 2011 and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_wbgi.pdf)
 
 >Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media.
-[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.png)
+[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf)
 
@@
 
-*There we some issues in converts regression outputs so the tables are evolving and can be browsed at: [http://research.muuankarski.org/attributions/code/multilevel_models.html](http://research.muuankarski.org/attributions/code/multilevel_models.html)*
+*As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
+Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.*
 
-****
+```{rarticleModels, results='asis', eval=TRUE}
+load("data/finalModels.RData")
 
-<!--
+model.names.multi.2 = c("SB vs. not",
+                        "IB vs. not",
+                        "SB vs. IB")
 
-As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
-Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
+coef.names.multi = c("(intercept)",
+                     "Dependent","Low education",
+                     "Low income","Income has worsened",
+                     "Income will worsen","Has affected great or fair amount",
+                     "Gini coefficient",
+                     "Change in GDP between 2007 to 2010",
+                     "Good governance")
 
--->
+library(texreg)
+texreg(list(soc.4,ind.4,socVsInc),float.pos="H",
+       use.packages=TRUE,
+       #booktabs=TRUE,
+       #dcolumn=TRUE,
+       #sideways=TRUE,
+       naive = TRUE,
+       caption.above=TRUE,
+       doctype=FALSE,
+       custom.model.names = model.names.multi.2,
+       custom.coef.names = coef.names.multi,
+       include.pvalues=TRUE,
+       caption="Logistic multilevel random intercept model for social and individual blame type of explanation")
+
+```
 

### 2013-11-09 finalPlots.Rmd
hieman epätoivoinen commit enne mikon tekstin lisäämistä
Tein takautuvasti vielä muutokset artikkelin taulukkokoodeihin ja poistin jotain mitä olin jo poistanut. (siis olisi pitänyt committaa uusimmat muutokset ennen mikko branchin yhdistämistä)

@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
+opts_chunk$set(echo=FALSE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
 ```
@@
 #########################################
-cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2","#D55E00", "#CC79A7")
+
 #########################################
@@
                         label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
+  geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
   geom_text(data=cnames, aes(gdpchange07.10,socialBlame, label = COR, group=COR), 
@@
                         y=individualBlame, label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
+  geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
   geom_text(size=2, vjust=-0.8, hjust=0.5, color="black") +
@@
 plot1 <- ggplot(t4, aes(x=uw_gini, y=socialBlame, label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
+  geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
@@
 plot2 <- ggplot(t4, aes(x=uw_gini, y=individualBlame, label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
+  geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
@@
                         label=cntry)) +
-  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +
+  geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
@@
                         y=individualBlame, label=cntry)) +
-geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +
+geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
@@
 ![](figure/scatter_wbgi.png)
+
+```{rattribFinalPlotsMap}
+library(RColorBrewer)
+library(mapproj)
+library(maptools)
+library(rgdal)
+map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
+                                layer="lits_all_simple"))
+
+map$NAME <- as.character(map$NAME)
+map$NAME[map$NAME == "United Kingdom"] <- 'Great Britain'
+map$group_analysis[map$NAME %in% c('France','Germany',
+                                        'Great Britain',
+                                        'Italy','Sweden')] <- 'Western Europe'
+
+map$group_analysis[map$NAME %in% c('Czech Republic','Estonia',
+                                        'Hungary','Bulgaria',
+                                        'Latvia','Lithuania',
+                                        'Poland','Slovakia',
+                                        'Slovenia','Romania')] <- 'CEE'
+
+map$group_analysis[map$NAME %in% c('Armenia','Azerbaijan',
+                                        'Belarus','Georgia',
+                                        'Kazakhstan','Kyrgyzstan',
+                                        'Moldova','Tajikistan',
+                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'
+map$NAME <- as.factor(map$NAME)
+
+map$group_analysis <- factor(map$group_analysis, levels=c("Western Europe",
+                                                             "CEE",
+                                                             "CIS",
+                                                             NA))
+
+# shapefile into data.frame
+map@data$id <- rownames(map@data)
+library(ggplot2)
+map.points <- fortify(map, region="id")
+map.df <- merge(map.points, map@data, by="id")
+# only east of longitude -25
+map.df <- subset(map.df, long >= -15 & long <= 80)
+# names for countries in analysis (excluding others)
+namedata <- map.df[!is.na(map.df$group_analysis),]
+cnames <- stats:::aggregate.formula(cbind(long, lat) ~ NAME, data=namedata, mean)
+
+mapplot <- ggplot(map.df, aes(long,lat,group=group)) +
+  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
+  scale_fill_manual(values=cbPalette) +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white", size=.3) +
+  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
+  coord_map(project="orthographic") +
+  theme_minimal() + 
+  theme(legend.position="top") +
+  theme(legend.title=element_blank()) +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) +
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank()) +
+  theme(axis.ticks.x = element_blank(),
+        axis.ticks.y = element_blank()) +
+  theme(legend.key.size = unit(3, "mm"))
+
+pdf("figure/mapplot.pdf", width=14/2.54, height=10/2.54)
+mapplot
+dev.off()
+
+ppi <- 300
+png("figure/mapplot.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
+mapplot
+dev.off()
+
+```
+
+![](figure/mapplot.png)

### 2013-11-09 article2013.Rmd
Now it works. Mikko text included and my corrections with the plots. Now using KOMA script and everything looks good!

@@
 % Markus Kainu[^**];Mikko Niemelä[^***]
-% October 2013
+% November 2013
 
-[^*]:*Work in progress
+[^*]:*Work in progress. See the computation analysis at [research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/)
 
@@
 knit("article2013.Rmd")
-system("~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --latex-engine=xelatex -o article2013.pdf")
-system("~/.cabal/bin/pandoc article2013.md  --number-sections -o article2013.pdf")
+system("~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=en -V papersize:a4paper -V documentclass=scrreprt -o article2013.pdf")
+# Kuvat kohdalleen - avaa päätte
+~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=en -V papersize:a4paper -V documentclass=scrreprt -o article2013.tex
+sed -i 's/begin{figure}\[htbp\]/begin{figure}\[H\]/g' article2013.tex
+xelatex article2013.tex
 
@@
 
-# jos multilevelit mukana
-#system("~/.cabal/bin/pandoc article2013.md --number-sections --template=tex_templates/dcolumn.latex -o article2013.tex")
-#system("perl -pi -e 's/footnotesize/tiny/g' article2013.tex")
-#system("pdflatex article2013.tex")
-
 ```
 
-**Abstract**
-
-*Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
+\begin{abstract}
+Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
 The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions.
 As country-level determinants, paper analyses the changes in country's economic performance during the period of global financial crises, income inequality and the state of democracy.
-As individual-level determinants, the paper tests to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.*
-
-\newpage
+As individual-level determinants, the paper tests to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.
+\end{abstract}
 
@@
 
-```{rarticleModels, results='asis', eval=TRUE}
+```{rarticleModelsSoc, results='asis', eval=TRUE}
+load("data/finalModels.RData")
+
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      "Gini coefficient",
+                      "Change in GDP between 2007 to 2010",
+                      "Good governance")
+
+model.names.multi = c("Empty",
+                      "Only individual",
+                      "Gini added",
+                      "GDP change added",
+                      "Good governance added")
+
+texreg(list(soc.0,
+        soc.1,
+        soc.2,
+        soc.3,
+        soc.4),
+        custom.model.names = model.names.multi,
+        custom.coef.names = coef.names.multi,
+        naive = TRUE,
+        use.packages=TRUE,
+        #sideways=TRUE,
+        scriptsize=TRUE,
+        include.pvalues=TRUE,
+        caption="Logistic multilevel random intercept model for social blame type of explanation", 
+        caption.above=TRUE)
+
+```
+
+```{rarticleModelsInd, results='asis', eval=TRUE}
+texreg(list(ind.0,
+        ind.1,
+        ind.2,
+        ind.3,
+        ind.4),
+        custom.model.names = model.names.multi,
+        custom.coef.names = coef.names.multi,
+        naive = TRUE,
+        use.packages=TRUE,
+        include.pvalues=TRUE,
+        #sideways=TRUE,
+        scriptsize=TRUE,
+        caption="Logistic multilevel random intercept model for individual blame type of explanation", 
+        caption.above=TRUE)
+
+```
+
+```{rarticleModelsFull, results='asis', eval=TRUE}
 load("data/finalModels.RData")
@@
        #dcolumn=TRUE,
-       #sideways=TRUE,
+       scriptsize=TRUE,
        naive = TRUE,

### 2013-11-09 finalPlots.Rmd
Now it works. Mikko text included and my corrections with the plots. Now using KOMA script and everything looks good!

@@
 #########################################
-cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2","#D55E00", "#CC79A7")
+cbPalette <- c("#E69F00","#56B4E9","#999999","#009E73", "#F0E442", "#0072B2","#D55E00", "#CC79A7")
 
@@
 ```{rattribFinalPlotsGDP}
-# valitaan vaan kohdemaat
-
+# valitaan vaan kohdemaat #
 t3 <- subset(macro, group_fine != 'West Europe')
@@
 cnames <- round(stats:::aggregate.formula(cbind(socialBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
+
 plot1 <- ggplot(t4, aes(x=gdpchange07.10, y=socialBlame, 
@@
 
-pdf("figure/scatter_gdp.pdf", width=14/2.54, height=22/2.54)
+pdf("figure/scatter_gdp.pdf", width=12/2.54, height=18/2.54)
 grid.arrange(plot1,plot2, ncol=1)
@@
 ppi <- 300
-png("figure/scatter_gdp.png", width=14/2.54*ppi, height=22/2.54*ppi, res=ppi)
+png("figure/scatter_gdp.png", width=12/2.54*ppi, height=18/2.54*ppi, res=ppi)
 grid.arrange(plot1,plot2, ncol=1)
@@
 
-pdf("figure/scatter_gini.pdf", width=14/2.54, height=22/2.54)
+pdf("figure/scatter_gini.pdf", width=12/2.54, height=18/2.54)
 grid.arrange(plot1,plot2, ncol=1)
@@
 ppi <- 300
-png("figure/scatter_gini.png", width=14/2.54*ppi, height=22/2.54*ppi, res=ppi)
+png("figure/scatter_gini.png", width=12/2.54*ppi, height=18/2.54*ppi, res=ppi)
 grid.arrange(plot1,plot2, ncol=1)
@@
 
-pdf("figure/scatter_wbgi.pdf", width=14/2.54, height=22/2.54)
+pdf("figure/scatter_wbgi.pdf", width=12/2.54, height=18/2.54)
 grid.arrange(plot1,plot2, ncol=1)
@@
 ppi <- 300
-png("figure/scatter_wbgi.png", width=14/2.54*ppi, height=22/2.54*ppi, res=ppi)
+png("figure/scatter_wbgi.png", width=12/2.54*ppi, height=18/2.54*ppi, res=ppi)
 grid.arrange(plot1,plot2, ncol=1)
@@
 ```{rattribFinalPlotsMap}
+
 library(RColorBrewer)
@@
 map$NAME <- as.factor(map$NAME)
-
-map$group_analysis <- factor(map$group_analysis, levels=c("Western Europe",
-                                                             "CEE",
-                                                             "CIS",
-                                                             NA))
 
@@
 
-pdf("figure/mapplot.pdf", width=14/2.54, height=10/2.54)
+pdf("figure/mapplot.pdf", width=14/2.54, height=12/2.54)
 mapplot
@@
 ppi <- 300
-png("figure/mapplot.png", width=14/2.54*ppi, height=10/2.54*ppi, res=ppi)
+png("figure/mapplot.png", width=14/2.54*ppi, height=12/2.54*ppi, res=ppi)
 mapplot

### 2013-11-09 index.Rmd
Now it works. Mikko text included and my corrections with the plots. Now using KOMA script and everything looks good!

@@
 20131106  Regression modelling somewhat done for Turku seminar.
+20131109  Corrections from Turku and the text restuctured. Methods needs to be opened
 
@@
 
-# Paper
+# Article draft
 
-- [Article draft in .html](article2013.html)
-- [Article draft in .pdf](article2013.pdf)
-- [Article draft in .epub](article2013.epub)
-- [Article draft in .docx](article2013.docx)
+-------------------
+pdf-output                                                source code
+------------                                              --------------
+[Article draft in .pdf](article2013.pdf)                  [source](article2013.Rmd)
+-------------------
 

### 2013-11-09 regressionModelling.Rmd
ennen kuin aloitan semman korjaukset
Regression tables added and pandoc works ok. Booktabs and dcolumn would be nice
Now it works. Mikko text included and my corrections with the plots. Now using KOMA script and everything looks good!

@@
         include.pvalues=TRUE,
-          caption="Logistic multilevel random intercept model for social blame type of explanation", 
+          caption="Logistic multilevel random intercept model for individual blame type of explanation", 
         caption.above=TRUE)
@@
 
+save(ind.0,ind.1,ind.2,ind.3,ind.4,
+     soc.0,soc.1,soc.2,soc.3,soc.4,
+     socVsInc, file="data/finalModels.RData")
+
 ```

### 2013-11-11 article2013.Rmd
Tried out the country grouping. I will leave it out from the presentation and paper, too. Edited the tables, but many issues still remains, mainly explaining the analysis and later intrepretations

@@
 # Kuvat kohdalleen - avaa päätte
-~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=en -V papersize:a4paper -V documentclass=scrreprt -o article2013.tex
+~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=english -V papersize:a4paper -V documentclass=scrreprt -o article2013.tex
 sed -i 's/begin{figure}\[htbp\]/begin{figure}\[H\]/g' article2013.tex
@@
 
------------------------------------------------------------------------------
-Variable                          Description
-------------------------          -------------------------------------------------
+--------------------------------------------------------------------------------------------
+Variable                          Description                                                                                                   Source
+------------------------          ------------------------                                                                                      -----------
 **Individual-level**
 
-Transfer dependency               Social or private transfers as main source of income
+Transfer dependency               Social or private transfers as main source of income                                                          lits
 
-Low education                     No or only compulsory level education
+Low education                     No or only compulsory level education                                                                         lits
 
-Perceived low income level        Perceived income level below the country median
+Perceived low income level        Perceived income level below the country median                                                               lits
 
-Income compared to past           Perceived change in income compared to situation in four years ago
+Income compared to past           Perceived change in income compared to situation in four years ago                                            lits
 
-Income compared to future         Expected change in income over the next four years
+Income compared to future         Expected change in income over the next four years                                                            lits
 
-Effect of financial crises        Whether respondents perceives that their household has suffered the financial crisis great or fair amount
+Effect of financial crises        Whether respondents perceives that their household has suffered the financial crisis great or fair amount     list
 
@@
 
-Ghange in total GDP 2007 – 2010   Relative change from 2007 to 2010
+Ghange in total GDP 2007 – 2010   Relative change from 2007 to 2010                                                                             World Bank
 
-Gini coefficient                  Measures the distribution of income or consumption expenditure among individuals or households
+Gini coefficient                  Measures the distribution of income or consumption expenditure among individuals or households                World Bank
 
-Voice and Accountability          Indicators measuring the political process, civil liberties and political rights.
-These indicators measure the extent to which citizens of a country are able to participate in the selection of governments.
-Indicators measuring the independence of the media is also included.
-Description
------------------------------------------------------------------------------
+Voice and Accountability          Indicators measuring the political process, civil liberties and political rights.                             World Bank
+--------------------------------------------------------------------------------------------
 
@@
 
-TÄMÄ YLLÄ MAINITTU TEKSTI VOIDAAN KUITATA SÄRMÄSTI SILLÄ ETTÄ LISÄTÄÄN TUOHON TABLE 2:EN OIKEALLE PUOLELLE YKSI SARAKE OTSIKOLLA "DATA SOURCE" SIIHEN SITTEN INDIVIDUAL-LEVEL MUUTTUJIEN KOHDALLE JOKA RIVILLE LIFE IN TRANSITION SURVEY JA MAKROMUUTTUJIEN KOHDALLE TYYLIIN WORLD BANK 2013 JA SITTEN LÄHTEISIIN LÄHDEVIITEMERKINNÄT NÄILLE ELI WORLD BANK (2013) JA ONLINEVIITTAUS LEHDEN SÄÄTÖJEN MUKAISESTI.
-
 The empirical section starts by presenting the results of descriptive analyses.
@@
 Multilevel analysis requires hierarchical dataset and allows the estimation of country-level effects on individual level outcomes.
-The analysis is performed using lmer-function from lme4 package in R [^*****].
+The analysis is performed using lmer-function from lme4 package in R.
 All of the models employed are so called multilevel logistic random intercept models at the level of individuals and countries.
-
-[^*****]:R-code for data transformation and analysis is available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/)
-
-TÄTÄ YLLÄ MAINITTUA ALAVIITETTÄ EI VARMAANKAAN SAA LAITTAA SUBMITTED-VERSIOON. SE VOIDAAN HALUTTAESSA LISÄTÄ SITTEN KUN PAPERI ON HYVÄKSYTTY LEHTEEN.
 
@@
 The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups.
-Full model summaries currently available at [http://research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/).
+
+*A more throughout description of modelling will be added as the final details in modelling will be decided*
 
 # Results
+
+*Results and conclusions are not yet reshaped to match the recent changes in the analysis...*
 
@@
 
-texreg(list(soc.0,
-        soc.1,
-        soc.2,
-        soc.3,
-        soc.4),
-        custom.model.names = model.names.multi,
+model.names.multi.A = c("SB empty",
+                      "SB individual",
+                      "SB all",
+                      "IB empty",
+                      "IB individual",
+                        "IB all")
+
+texreg(list(soc.0,soc.1,soc.4,
+            ind.0,ind.1,ind.4),
+        custom.model.names = model.names.multi.A,
         custom.coef.names = coef.names.multi,
@@
         caption="Logistic multilevel random intercept model for social blame type of explanation", 
-        caption.above=TRUE)
-
-```
-
-```{rarticleModelsInd, results='asis', eval=TRUE}
-texreg(list(ind.0,
-        ind.1,
-        ind.2,
-        ind.3,
-        ind.4),
-        custom.model.names = model.names.multi,
-        custom.coef.names = coef.names.multi,
-        naive = TRUE,
-        use.packages=TRUE,
-        include.pvalues=TRUE,
-        #sideways=TRUE,
-        scriptsize=TRUE,
-        caption="Logistic multilevel random intercept model for individual blame type of explanation", 
         caption.above=TRUE)

### 2013-11-11 finalPlots.Rmd
Tried out the country grouping. I will leave it out from the presentation and paper, too. Edited the tables, but many issues still remains, mainly explaining the analysis and later intrepretations

@@
 dev.off()
+
+png("figure/scatter_gdpSB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot1
+dev.off()
+png("figure/scatter_gdpIB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot2
+dev.off()
+
 ```
@@
 grid.arrange(plot1,plot2, ncol=1)
+dev.off()
+
+png("figure/scatter_giniSB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot1
+dev.off()
+png("figure/scatter_giniIB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot2
 dev.off()
@@
 
+png("figure/scatter_wbgiSB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot1
+dev.off()
+png("figure/scatter_wbgiIB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot2
+dev.off()
+
 ```

### 2013-11-11 index.Rmd
Tried out the country grouping. I will leave it out from the presentation and paper, too. Edited the tables, but many issues still remains, mainly explaining the analysis and later intrepretations

@@
 cd ~/workspace/lits/attrib/attrib_year2013/
-lftp -u np16003 -e "mput *.html *.Rmd *.pdf *.epub *.docx;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
+lftp -u np16003 -e "mput *.html *.Rmd *.pdf;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
 
@@
 lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/figure
+
+cd ~/workspace/lits/attrib/attrib_year2013/moscow2013
+lftp -u np16003 -e "mput *.html;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/moscow2013
 
@@
 20131109  Corrections from Turku and the text restuctured. Methods needs to be opened
+20131111  Presentation for HSE conference and improving the text a bit
 
@@
 
-[Final tables](finalTables.html)                          [source](finalTables.Rmd)
 -------------------
@@
 
-- HSE Moscow - November 12, 2013
+- [HSE Moscow - November 12, 2013](moscow2013/lcsr20131112.html)
 - [Jatkokoulutusseminaari Turku November 7, 2013](turku2013/slidesTkuJatko20131107.html)

### 2013-11-11 mapPlots.Rmd
Tried out the country grouping. I will leave it out from the presentation and paper, too. Edited the tables, but many issues still remains, mainly explaining the analysis and later intrepretations

@@
   scale_fill_gradient2(low="white", high=muted("blue")) +
-  theme(legend.position="top")
+  theme(legend.position="top") +
+  theme(legend.text = element_text(size=14))
 ```
@@
   scale_fill_gradient2(low=muted("red"), high=muted("blue")) +
-  theme(legend.position="top")
+  theme(legend.position="top") +
+  theme(legend.text = element_text(size=14))
 ```
@@
   scale_fill_gradient2(low=muted("blue"), high=muted("red"), midpoint=33.55) +
-  theme(legend.position="top")
+  theme(legend.position="top") +
+  theme(legend.text = element_text(size=14))
 ```

### 2013-11-11 regressionModelling.Rmd
Tried out the country grouping. I will leave it out from the presentation and paper, too. Edited the tables, but many issues still remains, mainly explaining the analysis and later intrepretations

@@
  dat.mlogit <- mlogit.data(dat.mlogit, shape="wide", choice="poverty")
-# 
-# 
-# model.mlogit1 <- mlogit(poverty ~ 0 | incsour3,
-#                                data = dat.mlogit, 
-#                                reflevel ="Social Blame")
-# model.mlogit2 <- mlogit(poverty ~ 0 | incsour3 + edu2,
-#                                data = dat.mlogit, 
-#                                reflevel ="Social Blame")
-# model.mlogit3 <- mlogit(poverty ~ 0 | incsour3 + edu2 + 
-#                                       past.diff,
-#                                data = dat.mlogit, 
-#                                reflevel ="Social Blame")
-# model.mlogit4 <- mlogit(poverty ~ 0 | incsour3 + edu2 + 
-#                                       past.diff + future.diff,
-#                                data = dat.mlogit, 
-#                                reflevel ="Social Blame")
+# ------- #
 model.mlogit5 <- mlogit(poverty ~ 0 | incsour3 + edu2 + 
@@
 
-```{r, eval=FALSE, echo=FALSE}
-
-screenreg(list(model.mlogit1,model.mlogit2))
-
-f <- coeftest(model.mlogit)
-xtable(as.data.frame(f))model1 <- multinom(formula = reason ~ supu+income, data = ml.06)
-
-summary(m)
-coeftest(m)
-
-X <- mlogit.data(ml.10, shape="wide", choice="poverty")
-coeftest(mlogit(poverty ~ 0 | success,
-               data = X, reflevel ="Because of injustice in our society"))
-summary(mlogit(poverty ~ 0 | success,
-                data = X, reflevel ="Because of injustice in our society"))
-
-coef(mlogit(poverty ~ 0 | success,
-               data = X, reflevel ="Because of injustice in our society"))
-
-```
-
 ## Multilevel multinomial logistic regression
@@
 <iframe width="420" height="315" src="//www.youtube.com/embed/q_t9VerQ2ko" frameborder="0" allowfullscreen></iframe>
-
-```{rmultiNomMulti}
-
-```
-
-```{rmultiNomMultiOutput, results='asis'}
-
-```
 
@@
 library(lme4)
+df.model$group_balinCis[df.model$cntry %in% c('Czech Republic',
+                                        'Hungary','Bulgaria',
+                                        'Poland','Slovakia',
+                                        'Slovenia','Romania')] <- 'CEE'
+
+df.model$group_balinCis[df.model$cntry %in% c('Armenia','Azerbaijan',
+                                        'Belarus','Georgia',
+                                        'Kazakhstan','Kyrgyzstan',
+                                        'Moldova','Tajikistan',
+                                        'Ukraine','Uzbekistan','Russia',
+                                        'Estonia','Latvia','Lithuania')] <- 'CIS'
+df.model$group_balinCis <- factor(df.model$group_balinCis, levels=c("CEE","CIS"))
+
 bisoc.1 <- glmer(pov.log.social ~ incsour3+(1|cntry),
@@
                           family = binomial(logit), data = df.model)
+bisoc.11 <- glmer(pov.log.social ~ group_analysis+(1|cntry),
+                          family = binomial(logit), data = df.model)
+bisoc.12 <- glmer(pov.log.social ~ group_balinCis+(1|cntry),
+                          family = binomial(logit), data = df.model)
+
 ```
@@
 ```{rmultilevelBivarSocOutputNOCI, results='asis'}
-
 coef.names.multi = c("(intercept)", 
@@
                       "Change in GDP between 2007 to 2010",
-                      "Good governance")
+                      "Good governance","Cgroup","CgroupMod")
 
@@
           bisoc.8,
-          bisoc.9,bisoc.10),
+          bisoc.9,bisoc.10,
+             bisoc.11,bisoc.12),
         custom.coef.names = coef.names.multi,
@@
                          family = binomial(logit), data = df.model)
+biind.11 <- glmer(pov.log.individual ~ group_analysis+(1|cntry),
+                          family = binomial(logit), data = df.model)
+bisoc.12 <- glmer(pov.log.individual ~ group_balinCis+(1|cntry),
+                          family = binomial(logit), data = df.model)
+
 ```
@@
           biind.8,
-          biind.9,biind.10),
+          biind.9,biind.10,
+          biind.11,biind.12),
         custom.coef.names = coef.names.multi,
@@
 ```{rmultilevelMultiIndOutputNOCI, results='asis'}
-
 htmlreg(list(ind.0,
@@
 
+# Tables for presentation
+
+## Partial
+
+```{rmultilevelMultivarSlideOutput, results='asis', eval=TRUE}
+library(texreg)
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      "Gini coefficient",
+                      "Change in GDP between 2007 to 2010",
+                      "Good governance")
+
+model.names.multi = c("Empty",
+                      "Only individual",
+                      "Gini added",
+                      "GDP change added",
+                      "Good governance added")
+
+model.names.multi.A = c("SB empty",
+                      "SB individual",
+                      "SB all",
+                      "IB empty",
+                      "IB individual",
+                        "IB all")
+
+htmlreg(list(soc.0,soc.1,soc.4,
+            ind.0,ind.1,ind.4),
+        custom.model.names = model.names.multi.A,
+        custom.coef.names = coef.names.multi,
+        naive = TRUE,
+        inline.css = TRUE,
+        doctype = FALSE,
+        html.tag = FALSE,
+        head.tag = FALSE,
+        body.tag = FALSE,
+        include.pvalues=TRUE,
+        caption="Logistic multilevel random intercept model for social and individual blame type of explanation", 
+        caption.above=TRUE)
+```
+
+## Comparison of full models
+
 ```{rrmultilevelMultivarFullOutput, results='asis', eval=TRUE}
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      "Gini coefficient",
+                      "Change in GDP between 2007 to 2010",
+                      "Good governance")
+
 model.names.multi.2 = c("Social blame vs. not",
@@
         caption.above=TRUE)
-
 ```

### 2013-11-13 index.Rmd
Final slides for LCSR conference. Major failure due to outdated browser.

@@
 - [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)
+
+# Misc links
+
+- [Article code in github](https://github.com/muuankarski/attributions)
+- [rustfare-package](http://markuskainu.fi/rustfare/) *- toolbox for R language for open welfare data from Russian Federation*
+- [rOpenGov](http://ropengov.github.io/) *-  R Ecosystem for Open Government Data and Computational Social Science*
+- [markuskainu.fi](http://markuskainu.fi) *- personal website*
+
+-----

### 2013-11-19 finalPlots.Rmd
Some basic fixing since Moscow conference. DV table fixed, analysis cleared. Baseline for trying out some more modelling after Moscow discussions.

@@
 ```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013/")
+setwd("~/workspace/lits/attributions/")
 library(kaRski)
@@
 #########################################
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+load("~/workspace/lits/attributions/data/lits.RData")
+
 ```

### 2013-11-19 mapPlots.Rmd
Some basic fixing since Moscow conference. DV table fixed, analysis cleared. Baseline for trying out some more modelling after Moscow discussions.

@@
 ```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013/")
+setwd("~/workspace/lits/attributions/")
 library(knitr)
@@
 library(rgdal)
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+load("~/workspace/lits/attributions/data/lits.RData")
 map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 

### 2013-11-19 code/dv_tables.R

@@ -1,4 +1,4 @@
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+load("~/workspace/lits/attributions/data/lits.RData")
 library(survey)
 d.df <- svydesign(id = ~SerialID, 
                   weights = ~weight, 
@@ -11,38 +11,42 @@ t$Freq <- round(t$Freq, 1)
 t <- subset(t, Freq > 0)
 
 ### testailua
-maa <- subset(t, q309 == "Not stated")
-maa <- maa[order(maa$group_general),]
-t$cntry <- factor(t$cntry, levels=maa$cntry)
-library(ggplot2)
-ggplot(t, aes(x=cntry,y=Freq,group=cntry)) +
-  geom_bar(stat="identity", position="dodge") +
-  facet_wrap(~q309) + coord_flip()
+# maa <- subset(t, q309 == "Not stated")
+# maa <- maa[order(maa$group_general),]
+# t$cntry <- factor(t$cntry, levels=maa$cntry)
+# library(ggplot2)
+# ggplot(t, aes(x=cntry,y=Freq,group=cntry)) +
+#   geom_bar(stat="identity", position="dodge") +
+#   facet_wrap(~q309) + coord_flip()
 ####
 library(reshape2)
 t2 <-  dcast(t, group_general+cntry ~ q309, value.var="Freq")
-names(t2) <-  c("group_general","country","social.blame","individual.blame","individual.fate","dont.know","social.fate","not.stated")
-tbl2 <- t2[order(-t2$social.blame),]
+names(t2) <-  c("group_general","country","socialBlame","individualBlame",
+                "individualFate","socialFate","notStated","dontKnow")
+t2 <- t2[,c(1,2,3,4,6,5,7,8)]
+t2 <- t2[order(-t2$socialBlame),]
 ########
-tbl2$group_general <- NULL
+#t2$group_general <- NULL
 
 #---------------------------------------------------#
 #---------------------------------------------------#
 #---------------------------------------------------#
 #---------------------------------------------------#
 
-t <- data.frame(prop.table(svytable(~group_general+cntry+poverty, d.df), 2)*100)
-t$Freq <- round(t$Freq, 2)
-t <- subset(t, Freq > 0)
-t2 <-  dcast(t, group_general+cntry ~ poverty, value.var="Freq")
[…]

### 2013-11-19 article2013.Rmd
Final slides for LCSR conference. Major failure due to outdated browser.
Some basic fixing since Moscow conference. DV table fixed, analysis cleared. Baseline for trying out some more modelling after Moscow discussions.

@@
 % Markus Kainu[^**];Mikko Niemelä[^***]
-% November 2013
+% November 19, 2013
 
@@
 ```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013/")
+setwd("~/workspace/lits/attributions/")
 library(knitr)
 knit("article2013.Rmd")
-system("~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=en -V papersize:a4paper -V documentclass=scrreprt -o article2013.pdf")
-# Kuvat kohdalleen - avaa päätte
-~/.cabal/bin/pandoc article2013.md --template=../../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=english -V papersize:a4paper -V documentclass=scrreprt -o article2013.tex
+
+~/.cabal/bin/pandoc article2013.md --template=../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=english -V papersize:a4paper -V documentclass=scrartcl -o article2013.tex
 sed -i 's/begin{figure}\[htbp\]/begin{figure}\[H\]/g' article2013.tex
@@
 ```
+
+\newpage
 
@@
 
+--------
+
 *A more throughout description of modelling will be added as the final details in modelling will be decided*
@@
 library(xtable)
-print.xtable(xtable(tbl5, 
+print.xtable(xtable(tbl5,digits=1,
                     caption="Weighted population shares of dependent variable by country",),table.placement="H",
@@
              include.rownames=FALSE,
-             caption.placement="top")
+             caption.placement="top",
+             size="scriptsize")
 

### 2013-11-19 descriptive_analysis.Rmd
Some basic fixing since Moscow conference. DV table fixed, analysis cleared. Baseline for trying out some more modelling after Moscow discussions.

@@
 ```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013/")
+setwd("~/workspace/lits/attributions")
 library(kaRski)
@@
 ```{rdescplot1_1}
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+load("~/workspace/lits/attributions/data/lits.RData")
 library(ggplot2)

### 2013-11-19 loadClean.Rmd
Now it works. Mikko text included and my corrections with the plots. Now using KOMA script and everything looks good!
Some basic fixing since Moscow conference. DV table fixed, analysis cleared. Baseline for trying out some more modelling after Moscow discussions.

@@
 
-summary(factor(lits2$group_analysis))
+lits2$group_analysis <- factor(lits2$group_analysis, levels=c("Western Europe",
+                                                             "CEE",
+                                                             "CIS",
+                                                             NA))
 
+summary(lits2$group_analysis)
 ```
@@
 
-datQogSum.w$group_general <- factor(datQogSum.w$group_general)
+datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
+                                    levels=c("Western Europe",
+                                             "CEE",
+                                             "CIS"))
 
@@
 
-save.image("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
-#load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+save.image("~/workspace/lits/attributions/data/lits.RData")
+#load("~/workspace/lits/attributions/data/lits.RData")
 ```

### 2013-11-19 regressionModelling.Rmd
Some basic fixing since Moscow conference. DV table fixed, analysis cleared. Baseline for trying out some more modelling after Moscow discussions.

@@
 #########################################
-setwd("~/workspace/lits/attrib/attrib_year2013/")
+setwd("~/workspace/lits/attributions/")
 
 #########################################
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+load("~/workspace/lits/attributions/data/lits.RData")
 df.model <- df[df$group_general != "Western Europe",]

### 2013-11-20 loadClean.Rmd
Intermediate commit.. Adding perceived gini and 2006 data

@@
 # FROM WEB
+# temp <- tempfile()
+# download.file("http://www.ebrd.com/downloads/research/surveys/litscsv.zip", temp)
+# unzip(temp)
+# lits1 <- read.csv("LITS2006_1.csv")
+# lits2 <- read.csv("LITS2006_2.csv")
+# lits3 <- read.csv("LITS2006_3.csv")
+# lits4 <- read.csv("LITS2006_4.csv")
+# lits5 <- read.csv("LITS2006_5.csv")
+# 
+# lits <- merge(lits1,lits2, by="id")
+# lits <- merge(lits,lits3, by="id")
+# lits <- merge(lits,lits4, by="id")
+# lits1 <- merge(lits,lits5, by="id")
+# save(lits1, 
+#       file="~/workspace/data/lits/lits1.RData")
+
+load("~/workspace/data/lits/lits1.RData")
+# FROM WEB
 #temp <- tempfile()
@@
 ```{rLoadClean2}
+### ------ BEGIN 2006 ---------- ####
+
+# different country groupings
+# Function the uppercase the first characters
+.simpleCap <- function(x) {
+    s <- strsplit(x, " ")[[1]]
+    paste(toupper(substring(s, 1, 1)), substring(s, 2),
+          sep = "", collapse = " ")
+}
+lits1$countryname <- as.character(lits1$countryname)
+lits1$cntry <- sapply(lits1$countryname, .simpleCap)
+# recongin to match lits2 cntry names
+lits1$cntry[lits1$cntry %in% "Czechrep"] <- "Czech Republic"
+
+### ------ END 2006 ---------- ####
 # different country groupings
@@
 lits2$cntry <- factor(lits2$cntry)
+
+### ------ BEGIN 2006 ---------- ####
+
+lits1$socialism[lits1$cntry %in% c('Czech Republic','Estonia','Hungary',
+                                   'Bulgaria','Latvia','Lithuania',
+                                   'Poland','Slovakia','Slovenia',
+                                   'Romania','Albania','Bosnia',
+                                   'Bulgaria','Croatia','Macedonia',
+                                   'Montenegro','Kosovo','Serbia',
+                                   'Armenia','Azerbaijan','Belarus',
+                                   'Georgia','Kazakhstan','Kyrgyzstan',
+                                   'Moldova','Mongolia','Russia',
+                                   'Tajikistan','Ukraine','Uzbekistan')] <- 'Post-socialist'
+summary(factor(lits1$socialism))
+
+lits1$group_analysis[lits1$cntry %in% c('Czech Republic','Estonia',
+                                        'Hungary','Bulgaria',
+                                        'Latvia','Lithuania',
+                                        'Poland','Slovakia',
+                                        'Slovenia','Romania')] <- 'CEE'
+
+lits1$group_analysis[lits1$cntry %in% c('Armenia','Azerbaijan',
+                                        'Belarus','Georgia',
+                                        'Kazakhstan','Kyrgyzstan',
+                                        'Moldova','Tajikistan',
+                                        'Ukraine','Uzbekistan','Russia')] <- 'CIS'
+
+lits1$group_analysis <- factor(lits1$group_analysis, levels=c("CEE",
+                                                             "CIS",
+                                                             NA))
+
+summary(lits1$group_analysis)
+
+### ------ END 2006 ---------- ####
 
@@
 summary(lits2$group_analysis)
+
+tbl <- merge(as.data.frame(table(lits2$cntry)),as.data.frame(table(lits1$cntry)),by="Var1",all.x=TRUE)
+names(tbl) <- c("cntry","n2010","n2006")
+tbl
 ```
@@
 ```{rLoadClean3}
+
+### ------ BEGIN 2006 ---------- ####
+summary(lits1$q308)
+# All options
+lits1$q309 <- as.character(lits1$q308)
+lits1$q309[lits1$q309 %in% 'injustice in society'] <- 'Social Blame'
+lits1$q309[lits1$q309 %in% 'laziness and lack of willpower'] <- 'Individual Blame'
+lits1$q309[lits1$q309 %in% 'unlucky'] <- 'Individual Fate'
+lits1$q309[lits1$q309 %in% 'inevitable part of modern life'] <- 'Social Fate'
+lits1$q309[lits1$q309 %in% 'other'] <- 'Not stated'
+lits1$q309[!(lits1$q309 %in% c("Social Blame",
+                              "Individual Blame",
+                              "Individual Fate",
+                              "Social Fate",
+                              "Not stated"))] <- NA
+
+lits1$q309 <- factor(lits1$q309, levels=c("Social Blame",
+                                          "Individual Blame",
+                                          "Individual Fate",
+                                          "Social Fate",
+                                          "Not stated"))
+summary(factor(lits1$q309))
+# not stated and dont know coded NA
+
+lits1$poverty <- lits1$q309
+lits1$poverty[lits1$poverty %in% 'Not stated'] <- NA
+lits1$poverty <- factor(lits1$poverty)
+summary(lits1$poverty)
+
+## For logistic regressions
+# $pov.log.social
+
+lits1$pov.log.social[lits1$q309 == 'Social Blame'] <- 1
+lits1$pov.log.social[lits1$q309 != 'Social Blame'] <- 0
+
+summary(factor(lits1$pov.log.social))
+
+# $pov_log_individual
+###
+
+lits1$pov.log.individual[lits1$q309 == 'Individual Blame'] <- 1
+lits1$pov.log.individual[lits1$q309 != 'Individual Blame'] <- 0
+
+summary(factor(lits1$pov.log.individual))
+
+### ------ END 2006 ---------- ####
+
 summary(lits2$q309)
@@
 
+#
+tbl <- merge(as.data.frame(table(lits2$cntry,lits2$pov.log.social)),
+      as.data.frame(table(lits1$cntry,lits1$pov.log.social)),
+      by="Var1",all.x=TRUE)
+names(tbl) <- c("cntry","Var2010","N2010","Var2006","N2006")
+tbl
+#
+tbl <- merge(as.data.frame(table(lits2$cntry,lits2$pov.log.individual)),
+      as.data.frame(table(lits1$cntry,lits1$pov.log.individual)),
+      by="Var1",all.x=TRUE)
+names(tbl) <- c("cntry","Var2010","N2010","Var2006","N2006")
+tbl
+#
+
 ```
@@
 ```{rLoadClean4}
+
+### ------ BEGIN 2006 ---------- ####
+# $sex
+lits1$sex <- lits1$q102_1
+# $income
+library(plyr)
+medianData <- ddply(lits1,~cntry,summarise,incMedian=median(q211, na.rm=TRUE))
+lits1 <- merge(lits1,medianData,by="cntry",all.x=TRUE)
+
+lits1$income2[lits1$q211 < lits1$incMedian]  <- 'Low'
+lits1$income2[lits1$q211 >= lits1$incMedian]  <- 'High'
+
+lits1$income2 <- factor(lits1$income2, 
+                       levels=c("Low",
+                                "High"))
+
+summary(lits1$income2)
+### ------ END 2006 ---------- ####
+
 # $sex
@@
 # $income
-
-lits2$income[lits2$q227 >= 1 & lits2$q227 <= 3]  <- 'Low income'
-lits2$income[lits2$q227 >= 4 & lits2$q227 <= 7]  <- 'Middle income'
-lits2$income[lits2$q227 >= 8] <-'High income'
-
-lits2$income <- factor(lits2$income, 
-                       levels=c("Low income",
-                                "Middle income",
-                                "High income"))
-
-summary(lits2$income)
 
@@
 
-summary(lits2$income2)
+lits2$q227_rec <- lits2$q227
+lits2$q227_rec[lits2$q227 <= 0] <- NA
 
-# median(lits2)
-#svyquantile(~q227, d.df, quantile = c(0.25, 0.5, 0.75), ci = TRUE)
-library(car)
-lits2$q227_rec <- lits2$q227
-lits2$q227_rec[lits2$q227_rec <= 0] <- NA
-# $future
+library(plyr)
+medianData <- ddply(lits2,~cntry,summarise,incMedian=median(q227_rec, na.rm=TRUE))
+lits2 <- merge(lits2,medianData,by="cntry",all.x=TRUE)
+
+lits2$income2[lits2$q227 < lits2$incMedian]  <- 'Low'
+lits2$income2[lits2$q227 >= lits2$incMedian]  <- 'High'
+
+lits2$income2 <- factor(lits2$income2, 
+                       levels=c("Low",
+                                "High"))
+
+### ------ BEGIN 2006 ---------- ####
+## perGini
+library(laeken)
+perGini2006 <- as.data.frame(gini(inc="q211",
+                      weights = "weight_1",
+                      data = lits1,
+                      na.rm = TRUE,
+                      breakdown="cntry")[2])
+
+names(perGini2006) <- c("cntry","perGini")
+perGini2006
+
+### ------ END 2006 ---------- ####
+
+library(laeken)
+perGini2010 <- as.data.frame(gini(inc="q227_rec",
+                      weights = "weight",
+                      data = lits2,
+                      na.rm = TRUE,
+                      breakdown="cntry")[2])
+
+names(perGini2010) <- c("cntry","perGini")
+perGini2010
+
+lits1$age <- lits1$age1
+lits2$age <- lits2$q104a_1
+
+### ------ BEGIN 2006 ---------- ####
+
+# $incsour tulonlähde
+###
+summary(factor(lits1$q210b))
+
+lits1$incsour3[lits1$q210b %in% c('state provided social benefits',
+                                  'unemployment benefits',
+                                  'pensions',
+                                  'community/privately provided social benefits',
+                                  'help from charities and NGOs',
+                                  'help from relatives/friend in the country',
+                                  'help from relatives/friends abroad')] <- "dependent"
+lits1$incsour3[!(lits1$q210b %in% c('state provided social benefits',
+                                  'unemployment benefits',
+                                  'pensions',
+                                  'community/privately provided social benefits',
+                                  'help from charities and NGOs',
+                                  'help from relatives/friend in the country',
+                                  'help from relatives/friends abroad'))] <- "notDependent"
+
+lits1$incsour3 <- factor(lits1$incsour3, 
+                         levels=c("dependent",
+                                  "notDependent"))
+summary(lits1$incsour3)
 ###
 
-summary(factor(lits2$q229))
-
-lits2$future[lits2$q229 >= 1 & lits2$q229 <= 3] <- "low income"
-lits2$future[lits2$q229 >= 4 & lits2$q229 <= 7] <- "middle income"
-lits2$future[lits2$q229 >= 8 & lits2$q229 <= 10] <- "high income"
-
-summary(factor(lits2$future))
-
-lits2$q229_rec <- lits2$q229
-lits2$q229_rec[lits2$q229_rec <= 0] <- NA
-
-summary(factor(lits2$q229_rec))
-
-lits2$future.diff[lits2$q227_rec > lits2$q229_rec] <- 'worse'
-lits2$future.diff[lits2$q227_rec <= lits2$q229_rec] <- 'same or better'
-lits2$future.diff <- factor(lits2$future.diff, levels=c("worse",
-                                                        "same or better"))
-
-summary(lits2$future.diff)
-# $past
+# $edu
 ###
 
-summary(factor(lits2$q228))
+lits1$edu2[lits1$q501 %in% c('compulsory school education',
+                         'no degree/education')] <- 'no or compulsory'
+lits1$edu2[!(lits1$q501 %in% c('compulsory school education',
+                         'no degree/education'))] <- 'higher'
 
-lits2$past[lits2$q228 >= 1 & lits2$q228 <= 3] <- "low income"
-lits2$past[lits2$q228 >= 4 & lits2$q228 <= 7] <- "middle income"
-lits2$past[lits2$q228 >= 8 & lits2$q228 <= 10] <- "high income"
+lits1$edu2 <- factor(lits1$edu2,
+                     levels=c("no or compulsory",
+                              "higher"))
 
-summary(factor(lits2$past))
-
-lits2$q228_rec <- lits2$q228
-lits2$q228_rec[lits2$q228_rec <= 0] <- NA
-
-summary(factor(lits2$q228_rec))
-
-lits2$past.diff[lits2$q227_rec < lits2$q228_rec] <- 'worse'
-lits2$past.diff[lits2$q227_rec >= lits2$q228_rec] <- 'same or better'
-lits2$past.diff <- factor(lits2$past.diff, levels=c("worse",
-                                                    "same or better"))
-# $age
-###
-
-lits2$ageclass[lits2$q104a_1 <= 29] <- '29 or younger'
-lits2$ageclass[lits2$q104a_1 >= 30 & lits2$q104a_1 <= 44] <- '30 to 44'
-lits2$ageclass[lits2$q104a_1 >= 45 & lits2$q104a_1 <= 59] <- '45 to 59'
-lits2$ageclass[lits2$q104a_1 >= 60 & lits2$q104a_1 <= 120] <- '60 and older'
-
-lits2$ageclass <- factor(lits2$ageclass, 
-                         levels=c("29 or younger",
-                                  "30 to 44",
-                                  "45 to 59",
-                                  "60 and older"))
-
-summary(lits2$ageclass)
-
-lits2$age <- lits2$q104a_1
-
-###
+summary(lits1$edu2)
+### ------ END 2006 ---------- ####
 
@@
 #################
+
+# $future
+###
+lits2$q229_rec <- lits2$q229
+lits2$q229_rec[lits2$q229 <= 0] <- NA
+
+lits2$future.diff[lits2$q227_rec > lits2$q229_rec] <- 'worse'
+lits2$future.diff[lits2$q227_rec <= lits2$q229_rec] <- 'same or better'
+lits2$future.diff <- factor(lits2$future.diff, levels=c("worse",
+                                                        "same or better"))
+# $past
+lits2$q228_rec <- lits2$q228
+lits2$q228_rec[lits2$q228 <= 0] <- NA
+
+lits2$past.diff[lits2$q227_rec < lits2$q228_rec] <- 'worse'
+lits2$past.diff[lits2$q227_rec >= lits2$q228_rec] <- 'same or better'
+lits2$past.diff <- factor(lits2$past.diff, levels=c("worse",
+                                                    "same or better"))
 # $crise - kriisikokemus
@@
 df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
-                             "sex",
-                             "ageclass","age",
+                             "sex","age",
                              "future.diff","past.diff",
@@
                              "group_analysis",
-                             #"group_general","group_fine",
                              "income2",
                              "crise",
-                             "q227", # income
-                             "q228", # menneisyys
-                             "q229", # tulevaisuus
-#                             "q226m_txt", # tulonlähde
-                             "q515", # edu
-                             "q801", # crise,
                              "poverty"
@@
 # Quality of governance
-library(rustfare)
-datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
-                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
-                   'Kazakhstan','Tajikistan',
-                   'Uzbekistan','Estonia','Latvia','Lithuania',
-                   'Poland','Czech Republic','Slovakia','Hungary',
-                   'Slovenia','Romania','Bulgaria',
-                   'Italy','Germany','United Kingdom',
-                   'Sweden','France'), 
-       indicator = c("ffp_fsi", # Failed States Index
-                     "wbgi_vae", # World Bank - Voice and Accountability
-                     "undp_hdi", # UNDP Human Development Index
-                     "fh_polity2", # Democracy (Freedom House/Polity)
-                     "uw_gini"))
-# subset the latest value per cntry per indicator
-library(plyr)
-datQogSum <- ddply(datQog, c("cname","indicator"), 
-                   subset, year == max(year))
+# library(rustfare)
+# datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
+#                    'Ukraine','Moldova','Belarus','Kyrgyzstan',
+#                    'Kazakhstan','Tajikistan',
+#                    'Uzbekistan','Estonia','Latvia','Lithuania',
+#                    'Poland','Czech Republic','Slovakia','Hungary',
+#                    'Slovenia','Romania','Bulgaria',
+#                    'Italy','Germany','United Kingdom',
+#                    'Sweden','France'), 
+#        indicator = c("ffp_fsi", # Failed States Index
+#                      "wbgi_vae", # World Bank - Voice and Accountability
+#                      "undp_hdi", # UNDP Human Development Index
+#                      "fh_polity2", # Democracy (Freedom House/Polity)
+#                      "uw_gini"))
+# # subset the latest value per cntry per indicator
+# library(plyr)
+# datQogSum <- ddply(datQog, c("cname","indicator"), 
+#                    subset, year == max(year))
 
-# library(ggplot2)
-# ggplot(datQogSum, aes(x=year,y=value,color=cname)) + 
-#   geom_point(size=4) +
-#   geom_text(data = datQogSum,
-#             aes(x=year,y=value,
-#                 color=cname,
-#                 label=cname),
-#             size=3.5) +
-#   facet_wrap(~indicator, scales="free") +
-#   theme(legend.position="none")
+# datQogSum.w <- dcast(datQogSum, 
+#                      cname ~ indicator, 
+#                      value.var="value")
+# 
+# datQogSum.w$cname <-  as.character(datQogSum.w$cname)
+# datQogSum.w$cname[datQogSum.w$cname %in% "United Kingdom"] <- "Great Britain"
+# 
+# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Georgia')] <- 'Caucasus'
+# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Russia','Ukraine','Moldova','Belarus')] <- 'CIS North'
+# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')] <- 'Central Asia'
+# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Estonia','Latvia','Lithuania')] <- 'Baltic States'
+# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Poland','Czech Republic','Slovakia','Hungary',
+#                                                 'Slovenia','Romania','Bulgaria')] <- 'CEE EU'
+# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Italy','Germany','Great Britain','Sweden','France')] <- 'West Europe'
+#                          
+# datQogSum.w$group_fine <- factor(datQogSum.w$group_fine)
+# 
+# summary(factor(datQogSum.w$group_fine))
+# 
+# datQogSum.w$group_general[datQogSum.w$cname %in% c('France','Germany',
+#                                                    'Great Britain','Italy','Sweden')] <- 'Western Europe'
+# datQogSum.w$group_general[datQogSum.w$cname %in% c('Czech Republic','Estonia',
+#                                                    'Hungary','Bulgaria',
+#                                                    'Latvia','Lithuania','Poland',
+#                                                    'Slovakia','Slovenia','Romania')] <- 'CEE'
+# datQogSum.w$group_general[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Belarus',
+#                                                    'Georgia','Kazakhstan',
+#                                                    'Kyrgyzstan','Moldova',
+#                                                    'Tajikistan','Ukraine',
+#                                                    'Uzbekistan','Russia')] <- 'CIS'
+# 
+# datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
+#                                     levels=c("Western Europe",
+#                                              "CEE",
+#                                              "CIS"))
+# 
+#df.qog <- datQogSum.w
+# 
+# names(df.qog)[1] <- "cntry"
 
-datQogSum.w <- dcast(datQogSum, 
-                     cname ~ indicator, 
-                     value.var="value")
-
-datQogSum.w$cname <-  as.character(datQogSum.w$cname)
-datQogSum.w$cname[datQogSum.w$cname %in% "United Kingdom"] <- "Great Britain"
-
-datQogSum.w$group_fine[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Georgia')] <- 'Caucasus'
-datQogSum.w$group_fine[datQogSum.w$cname %in% c('Russia','Ukraine','Moldova','Belarus')] <- 'CIS North'
-datQogSum.w$group_fine[datQogSum.w$cname %in% c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')] <- 'Central Asia'
-datQogSum.w$group_fine[datQogSum.w$cname %in% c('Estonia','Latvia','Lithuania')] <- 'Baltic States'
-datQogSum.w$group_fine[datQogSum.w$cname %in% c('Poland','Czech Republic','Slovakia','Hungary',
-                                                'Slovenia','Romania','Bulgaria')] <- 'CEE EU'
-datQogSum.w$group_fine[datQogSum.w$cname %in% c('Italy','Germany','Great Britain','Sweden','France')] <- 'West Europe'
-
-datQogSum.w$group_fine <- factor(datQogSum.w$group_fine)
-
-summary(factor(datQogSum.w$group_fine))
-
-datQogSum.w$group_general[datQogSum.w$cname %in% c('France','Germany',
-                                                   'Great Britain','Italy','Sweden')] <- 'Western Europe'
-datQogSum.w$group_general[datQogSum.w$cname %in% c('Czech Republic','Estonia',
-                                                   'Hungary','Bulgaria',
-                                                   'Latvia','Lithuania','Poland',
-                                                   'Slovakia','Slovenia','Romania')] <- 'CEE'
-datQogSum.w$group_general[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Belarus',
-                                                   'Georgia','Kazakhstan',
-                                                   'Kyrgyzstan','Moldova',
-                                                   'Tajikistan','Ukraine',
-                                                   'Uzbekistan','Russia')] <- 'CIS'
-
-datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
-                                    levels=c("Western Europe",
-                                             "CEE",
-                                             "CIS"))
-
-df.qog <- datQogSum.w
-# underschores into dots for latex...
-#names <- names(df.qog)
-#library(stringr)
-#names <- str_replace_all(names, "_",".")
-#names(df.qog) <- names
-names(df.qog)[1] <- "cntry"
 ```
@@
 ```{rLoadClean8, eval=TRUE}
-## GDP change from WB
-### We need iso2c countrycode
-library(countrycode)
-cntry <- levels(df$cntry) # countrynames
-iso2c <- countrycode(cntry, "country.name", "iso2c")
-iso2c # copy paste into WDI search
-library(WDI)
-# WDIsearch(string='gdp', field='name', cache=NULL)
-wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
-          country = c("AL","AM","AZ","BY","BA","BG","HR",
-                      "CZ","EE","FR","GE","DE","GB","HU",
-                      "IT","KZ","KG","LV","LT","MK","MD",
-                      "MN","ME","PL","RO","RU","RS","SK",
-                      "SI","SE","TJ","TR","UA","UZ"), 
-           start = 2007, end = 2010)
-
-library(reshape2)
-names(wb) <- c("iso2c","cntry","var","year")
-wb.w <- dcast(wb, 
-              cntry ~ year, 
-              value.var="var")
-
-names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
-wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
-gdp <- wb.w[,c(1,6)]
-
-gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
-gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
-gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
-gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'
+# ## GDP change from WB
+# ### We need iso2c countrycode
+# library(countrycode)
+# cntry <- levels(df$cntry) # countrynames
+# iso2c <- countrycode(cntry, "country.name", "iso2c")
+# iso2c # copy paste into WDI search
+# library(WDI)
+# # WDIsearch(string='gdp', field='name', cache=NULL)
+# wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
+#           country = c("AL","AM","AZ","BY","BA","BG","HR",
+#                       "CZ","EE","FR","GE","DE","GB","HU",
+#                       "IT","KZ","KG","LV","LT","MK","MD",
+#                       "MN","ME","PL","RO","RU","RS","SK",
+#                       "SI","SE","TJ","TR","UA","UZ"), 
+#            start = 2007, end = 2010)
+# 
+# library(reshape2)
+# names(wb) <- c("iso2c","cntry","var","year")
+# wb.w <- dcast(wb, 
+#               cntry ~ year, 
+#               value.var="var")
+# 
+# names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
+# wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
+# gdp <- wb.w[,c(1,6)]
+# 
+# gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
+# gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
+# gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
+# gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'
 
@@
 # yhdistetään nää aikaisempien datojen kanssa
-macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
-macro <- merge(macro,df.qog,by="cntry")
+# macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
+# macro <- merge(macro,df.qog,by="cntry")
+#save(macro, file="data/macro.RData")
+load("data/macro.RData")
+macro <- merge(macro,perGini2010,by="cntry")
 
@@
 rm(iso2c)
+rm(lits1)
+rm(medianData)
+rm(perGini2006)
+rm(perGini2010)
 

### 2013-11-20 regressionModelling.Rmd
Intermediate commit.. Adding perceived gini and 2006 data

@@
                           family = binomial(logit), data = df.model)
+bisoc.13 <- glmer(pov.log.social ~ perGini+(1|cntry),
+                          family = binomial(logit), data = df.model)
 
@@
 ```{rmultilevelBivarSocOutputNOCI, results='asis'}
+
 coef.names.multi = c("(intercept)", 
@@
                       "Change in GDP between 2007 to 2010",
-                      "Good governance","Cgroup","CgroupMod")
+                      "Good governance","Cgroup","CgroupMod",
+                     "Perceived Gini")
 
@@
           bisoc.9,bisoc.10,
-             bisoc.11,bisoc.12),
+             bisoc.11,bisoc.12,
+             bisoc.13),
         custom.coef.names = coef.names.multi,
@@
                           family = binomial(logit), data = df.model)
+bisoc.13 <- glmer(pov.log.individual ~ perGini+(1|cntry),
+                          family = binomial(logit), data = df.model)
 
@@
 ```{rmultilevelBivarIndOuputNOCI, results='asis'}
+
 library(texreg)
@@
           biind.9,biind.10,
-          biind.11,biind.12),
+          biind.11,biind.12,
+             biind.13),
         custom.coef.names = coef.names.multi,

### 2013-11-20 regressionModelling.Rmd
Just another test commit whether rawgithub can handle with .css ver2

@@
-% Regression modellining
+# Regression modellining
 
@@
         custom.coef.names = coef.names,
-        inline.css = FALSE,
+        inline.css = TRUE,
         doctype = FALSE,
@@
         naive = TRUE,
-        inline.css = FALSE,
+        inline.css = TRUE,
         doctype = FALSE,
@@
         ci.test = NULL,
-        inline.css = FALSE,
+        inline.css = TRUE,
         doctype = FALSE,
@@
         custom.coef.names = coef.names.multi,
-        inline.css = FALSE,
+        inline.css = TRUE,
         doctype = FALSE,

### 2013-11-20 regressionModelling.Rmd
Just another test commit whether rawgithub can handle with .css
toc should work now

@@
 # Regression modellining
-
-- [back to index](index.html)
-
-`generated at:`
-```{r, echo=FALSE}
-Sys.time()
-```
----
 
@@
 library(knitr)
-knit("regressionModelling.Rmd")
-system("pandoc -s regressionModelling.md -o regressionModelling.html --toc --number-section")
+knit2html('regressionModelling.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 

### 2013-11-20 regressionModelling.Rmd
toc should work now

--- 2013-11-20 regressionModelling.Rmd
@@
 library(knitr)
-knit("regressionModelling.Rmd")
+knit2html('regressionModelling.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 system("pandoc -s regressionModelling.md -o regressionModelling.html --toc --number-section")

### 2013-11-20 finalPlots.Rmd

@@
-% Final graphics for the paper
-
-- [back to index](index.html)
-
-`generated at:`
-```{r, echo=FALSE}
-Sys.time()
-```
----
+# Final graphics for the paper
 
@@
 setwd("~/workspace/lits/attributions/")
-library(kaRski)
-knit("finalPlots.Rmd")
-system("pandoc -s finalPlots.md -o finalPlots.html --toc --number-section")
+library(knitr)
+knit2html('finalPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 
@@
 #########################################
-setwd("~/workspace/lits/attrib/attrib_year2013")
+setwd("~/workspace/lits/attributions")
 #########################################
@@
 ```{rattribFinalPlotsGDP}
+
 # valitaan vaan kohdemaat #
@@
 
+## Perceived inequality from data
+
+```{rattribFinalPlotsPG}
+
+# Perceived inequality from data
+
+func <- function(t3)
+{
+  return(data.frame(COR = cor(t3$socialBlame, t3$perGini)))
+}
+df.cor <- ddply(t3, .(group1), func)
+t4 <- join(t3,df.cor,by="group1")
+cnames <- round(stats:::aggregate.formula(cbind(socialBlame, perGini) ~ COR, data=t4, mean),2)
+plot1 <- ggplot(t4, aes(x=perGini, y=socialBlame, 
+                        label=cntry)) +
+  geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +
+  geom_text(size=2, vjust=-0.8, hjust=0.5) +
+  geom_text(data=cnames, aes(perGini,socialBlame, label = COR, group=COR), 
+            size=2.5, color="black") +
+  labs(x = "Perceived inequality from data", 
+       y = "Support for social blame type of explanation (%)") +
+  geom_point(size=1.5) +
+  geom_point(data=t4, aes(x=perGini, 
+                          y=socialBlame, color=group_general)) +
+  # west reference
+  geom_point(data=west, aes(x=perGini, 
+                            y=socialBlame, 
+                            color=group_general), alpha=.5) +
+  geom_text(data=west, aes(x=perGini, 
+                           y=socialBlame, label=cntry),
+             alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
+  # west reference
+  scale_colour_manual(values=cbPalette) +
+  #coord_cartesian(xlim=c(-2,2)) +
+  theme_minimal() +
+  theme(legend.title=element_blank()) +
+  theme(legend.text=element_text(size=6)) +
+  theme(legend.position="top") +
+  theme(axis.title.y = element_text(size=6)) +
+  theme(axis.title.x = element_text(size=6)) +
+  theme(axis.text.y = element_text(size=6)) +
+  theme(axis.text.x = element_text(size=6)) +
+  #guides(color = guide_legend(nrow = 2)) +
+  theme(legend.key.size = unit(3, "mm"))
+
+func <- function(t3)
+{
+  return(data.frame(COR = cor(t3$individualBlame, t3$perGini)))
+}
+df.cor <- ddply(t3, .(group1), func)
+t4 <- join(t3,df.cor,by="group1")
+cnames <- round(stats:::aggregate.formula(cbind(individualBlame, perGini) ~ COR, data=t4, mean),2)
+plot2 <- ggplot(t4, aes(x=perGini, 
+                        y=individualBlame, label=cntry)) +
+geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +
+  geom_text(size=2, vjust=-0.8, hjust=0.5) +
+  geom_text(data=cnames, aes(perGini,individualBlame, label = COR, group=COR), 
+            size=2.5, color="black") +
+  labs(x ="Perceived inequality from data", 
+       y = "Support for individual blame type of explanation (%)") +
+  geom_point(size=1.5) +
+  geom_point(data=t4, aes(x=perGini, 
+                          y=individualBlame, color=group_general)) +
+  # west reference
+  geom_point(data=west, aes(x=perGini, 
+                            y=individualBlame, 
+                            color=group_general), alpha=.5) +
+  geom_text(data=west, aes(x=perGini, 
+                           y=individualBlame, label=cntry),
+             alpha=.5, size=2, vjust=-0.8, hjust=0.5) +
+  # west reference
+  scale_colour_manual(values=cbPalette) +
+  #coord_cartesian(xlim=c(-2,2)) +
+  theme_minimal() +
+  theme(legend.title=element_blank()) +
+  theme(legend.text=element_text(size=6)) +
+  theme(legend.position="top") +
+  theme(axis.title.y = element_text(size=6)) +
+  theme(axis.title.x = element_text(size=6)) +
+  theme(axis.text.y = element_text(size=6)) +
+  theme(axis.text.x = element_text(size=6)) +
+  #guides(color = guide_legend(nrow = 2)) +
+  theme(legend.key.size = unit(3, "mm"))
+
+pdf("figure/scatter_perGini.pdf", width=12/2.54, height=18/2.54)
+grid.arrange(plot1,plot2, ncol=1)
+dev.off()
+
+ppi <- 300
+png("figure/scatter_perGini.png", width=12/2.54*ppi, height=18/2.54*ppi, res=ppi)
+grid.arrange(plot1,plot2, ncol=1)
+dev.off()
+
+png("figure/scatter_perGiniSB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot1
+dev.off()
+png("figure/scatter_perGiniIB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+plot2
+dev.off()
+
+```
+
+![](figure/scatter_perGini.png)
+
 ```{rattribFinalPlotsMap}

### 2013-11-20 regressionModelling.Rmd

--- 2013-11-20 regressionModelling.Rmd
@@
 knit2html('regressionModelling.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-
 ```

### 2013-11-20 loadClean.Rmd
Adding the plots

@@
-% Loading and cleaning the data
-
-- [back to index](index.html)
-
-`generated at:`
-```{r, echo=FALSE}
-Sys.time()
-```
----
+# Loading and cleaning the data
 
@@
 setwd("~/workspace/lits/attributions/")
-library(kaRski)
-knitpandoc("loadClean")
+library(knitr)
+knit2html('loadClean.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+
 ```
@@
 # Quality of governance
-# library(rustfare)
-# datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
-#                    'Ukraine','Moldova','Belarus','Kyrgyzstan',
-#                    'Kazakhstan','Tajikistan',
-#                    'Uzbekistan','Estonia','Latvia','Lithuania',
-#                    'Poland','Czech Republic','Slovakia','Hungary',
-#                    'Slovenia','Romania','Bulgaria',
-#                    'Italy','Germany','United Kingdom',
-#                    'Sweden','France'), 
-#        indicator = c("ffp_fsi", # Failed States Index
-#                      "wbgi_vae", # World Bank - Voice and Accountability
-#                      "undp_hdi", # UNDP Human Development Index
-#                      "fh_polity2", # Democracy (Freedom House/Polity)
-#                      "uw_gini"))
-# # subset the latest value per cntry per indicator
-# library(plyr)
-# datQogSum <- ddply(datQog, c("cname","indicator"), 
-#                    subset, year == max(year))
+library(rustfare)
+datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
+                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
+                   'Kazakhstan','Tajikistan',
+                   'Uzbekistan','Estonia','Latvia','Lithuania',
+                   'Poland','Czech Republic','Slovakia','Hungary',
+                   'Slovenia','Romania','Bulgaria',
+                   'Italy','Germany','United Kingdom',
+                   'Sweden','France'), 
+       indicator = c("ffp_fsi", # Failed States Index
+                     "wbgi_vae", # World Bank - Voice and Accountability
+                     "undp_hdi", # UNDP Human Development Index
+                     "fh_polity2", # Democracy (Freedom House/Polity)
+                     "uw_gini"))
+# subset the latest value per cntry per indicator
+library(plyr)
+datQogSum <- ddply(datQog, c("cname","indicator"), 
+                   subset, year == max(year))
 
-# datQogSum.w <- dcast(datQogSum, 
-#                      cname ~ indicator, 
-#                      value.var="value")
-# 
-# datQogSum.w$cname <-  as.character(datQogSum.w$cname)
-# datQogSum.w$cname[datQogSum.w$cname %in% "United Kingdom"] <- "Great Britain"
-# 
-# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Georgia')] <- 'Caucasus'
-# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Russia','Ukraine','Moldova','Belarus')] <- 'CIS North'
-# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')] <- 'Central Asia'
-# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Estonia','Latvia','Lithuania')] <- 'Baltic States'
-# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Poland','Czech Republic','Slovakia','Hungary',
-#                                                 'Slovenia','Romania','Bulgaria')] <- 'CEE EU'
-# datQogSum.w$group_fine[datQogSum.w$cname %in% c('Italy','Germany','Great Britain','Sweden','France')] <- 'West Europe'
-#                          
-# datQogSum.w$group_fine <- factor(datQogSum.w$group_fine)
-# 
-# summary(factor(datQogSum.w$group_fine))
-# 
-# datQogSum.w$group_general[datQogSum.w$cname %in% c('France','Germany',
-#                                                    'Great Britain','Italy','Sweden')] <- 'Western Europe'
-# datQogSum.w$group_general[datQogSum.w$cname %in% c('Czech Republic','Estonia',
-#                                                    'Hungary','Bulgaria',
-#                                                    'Latvia','Lithuania','Poland',
-#                                                    'Slovakia','Slovenia','Romania')] <- 'CEE'
-# datQogSum.w$group_general[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Belarus',
-#                                                    'Georgia','Kazakhstan',
-#                                                    'Kyrgyzstan','Moldova',
-#                                                    'Tajikistan','Ukraine',
-#                                                    'Uzbekistan','Russia')] <- 'CIS'
-# 
-# datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
-#                                     levels=c("Western Europe",
-#                                              "CEE",
-#                                              "CIS"))
-# 
-#df.qog <- datQogSum.w
-# 
-# names(df.qog)[1] <- "cntry"
+datQogSum.w <- dcast(datQogSum, 
+                     cname ~ indicator, 
+                     value.var="value")
+
+datQogSum.w$cname <-  as.character(datQogSum.w$cname)
+datQogSum.w$cname[datQogSum.w$cname %in% "United Kingdom"] <- "Great Britain"
+
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Georgia')] <- 'Caucasus'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Russia','Ukraine','Moldova','Belarus')] <- 'CIS North'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')] <- 'Central Asia'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Estonia','Latvia','Lithuania')] <- 'Baltic States'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Poland','Czech Republic','Slovakia','Hungary',
+                                                'Slovenia','Romania','Bulgaria')] <- 'CEE EU'
+datQogSum.w$group_fine[datQogSum.w$cname %in% c('Italy','Germany','Great Britain','Sweden','France')] <- 'West Europe'
+
+datQogSum.w$group_fine <- factor(datQogSum.w$group_fine)
+
+summary(factor(datQogSum.w$group_fine))
+
+datQogSum.w$group_general[datQogSum.w$cname %in% c('France','Germany',
+                                                   'Great Britain','Italy','Sweden')] <- 'Western Europe'
+datQogSum.w$group_general[datQogSum.w$cname %in% c('Czech Republic','Estonia',
+                                                   'Hungary','Bulgaria',
+                                                   'Latvia','Lithuania','Poland',
+                                                   'Slovakia','Slovenia','Romania')] <- 'CEE'
+datQogSum.w$group_general[datQogSum.w$cname %in% c('Armenia','Azerbaijan','Belarus',
+                                                   'Georgia','Kazakhstan',
+                                                   'Kyrgyzstan','Moldova',
+                                                   'Tajikistan','Ukraine',
+                                                   'Uzbekistan','Russia')] <- 'CIS'
+
+datQogSum.w$group_general <- factor(datQogSum.w$group_general, 
+                                    levels=c("Western Europe",
+                                             "CEE",
+                                             "CIS"))
+
+df.qog <- datQogSum.w
+
+names(df.qog)[1] <- "cntry"
 
@@
 ```{rLoadClean8, eval=TRUE}
-# ## GDP change from WB
-# ### We need iso2c countrycode
-# library(countrycode)
-# cntry <- levels(df$cntry) # countrynames
-# iso2c <- countrycode(cntry, "country.name", "iso2c")
-# iso2c # copy paste into WDI search
-# library(WDI)
-# # WDIsearch(string='gdp', field='name', cache=NULL)
-# wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
-#           country = c("AL","AM","AZ","BY","BA","BG","HR",
-#                       "CZ","EE","FR","GE","DE","GB","HU",
-#                       "IT","KZ","KG","LV","LT","MK","MD",
-#                       "MN","ME","PL","RO","RU","RS","SK",
-#                       "SI","SE","TJ","TR","UA","UZ"), 
-#            start = 2007, end = 2010)
-# 
-# library(reshape2)
-# names(wb) <- c("iso2c","cntry","var","year")
-# wb.w <- dcast(wb, 
-#               cntry ~ year, 
-#               value.var="var")
-# 
-# names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
-# wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
-# gdp <- wb.w[,c(1,6)]
-# 
-# gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
-# gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
-# gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
-# gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'
+## GDP change from WB
+### We need iso2c countrycode
+library(countrycode)
+cntry <- levels(df$cntry) # countrynames
+iso2c <- countrycode(cntry, "country.name", "iso2c")
+iso2c # copy paste into WDI search
+library(WDI)
+# WDIsearch(string='gdp', field='name', cache=NULL)
+wb <- WDI(indicator = c("NY.GDP.MKTP.CD"), 
+          country = c("AL","AM","AZ","BY","BA","BG","HR",
+                      "CZ","EE","FR","GE","DE","GB","HU",
+                      "IT","KZ","KG","LV","LT","MK","MD",
+                      "MN","ME","PL","RO","RU","RS","SK",
+                      "SI","SE","TJ","TR","UA","UZ"), 
+           start = 2007, end = 2010)
+
+library(reshape2)
+names(wb) <- c("iso2c","cntry","var","year")
+wb.w <- dcast(wb, 
+              cntry ~ year, 
+              value.var="var")
+
+names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
+wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
+gdp <- wb.w[,c(1,6)]
+
+gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
+gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
+gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
+gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'
 
@@
 # yhdistetään nää aikaisempien datojen kanssa
-# macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
-# macro <- merge(macro,df.qog,by="cntry")
-#save(macro, file="data/macro.RData")
-load("data/macro.RData")
+macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
+macro <- merge(macro,df.qog,by="cntry")
 macro <- merge(macro,perGini2010,by="cntry")

### 2013-11-20 descriptive_analysis.Rmd
All the files and figures moved to GH

@@
-% Descriptive visualizations
-
-- [back to index](index.html)
-
-`generated at:`
-```{r, echo=FALSE}
-Sys.time()
-```
----
+# Descriptive visualizations
 
@@
 setwd("~/workspace/lits/attributions")
-library(kaRski)
-knit("descriptive_analysis.Rmd")
-system("pandoc -s descriptive_analysis.md -o descriptive_analysis.html --toc --number-section")
+library(knitr)
+knit2html('descriptive_analysis.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 

### 2013-11-20 mapPlots.Rmd
All the files and figures moved to GH

@@
-% Spatial visualization
-
-- [back to index](index.html)
-
-`generated at:`
-```{r, echo=FALSE}
-Sys.time()
-```
----
+# Spatial visualization
 
@@
 library(knitr)
-knit("mapPlots.Rmd")
-system("pandoc -s mapPlots.md -o mapPlots.html --toc --number-section")
+knit2html('mapPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+
 ```

### 2013-11-20 multi_logit.R

@@ -1,35 +1,0 @@
-setwd("~/workspace/lits/attrib/attrib_year2013/code")
-
-#########################################
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
-
-summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
-                      family = binomial, data=df))
-summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
-                         past.diff+future.diff+crise+(1|cntry),
-                       family = binomial, data=df))
-summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
-
-
-summary(ind.0 <- lmer(pov.log.individual ~ (1|cntry),
-                      family = binomial, data=df))
-summary(ind.1 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+(1|cntry),
-                      family = binomial, data=df))
-summary(ind.2 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+(1|cntry),family = binomial, data=df))
-summary(ind.3 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
-summary(ind.4 <- lmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))

### 2013-11-20 regressionModelling.Rmd
perGini in the regression

@@
                       "IB individual",
-                        "IB all")
+                        "IB all",
+                        "SB perGini",
+                        "IB perGini")
+
+SBperGini <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        perGini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+IBperGini <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        perGini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
 
 htmlreg(list(soc.0,soc.1,soc.4,
-            ind.0,ind.1,ind.4),
+            ind.0,ind.1,ind.4,
+             SBperGini,IBperGini),
         custom.model.names = model.names.multi.A,
-        custom.coef.names = coef.names.multi,
+        #custom.coef.names = coef.names.multi,
         naive = TRUE,

### 2013-11-21 descriptive_analysis.Rmd
A new css with left side navigation

@@
 library(knitr)
-knit2html('descriptive_analysis.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+library(markdown)
+knit2html('descriptive_analysis.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 
-knitpandoc("descriptive_analysis",toc)
 ```

### 2013-11-21 finalPlots.Rmd
A new css with left side navigation

@@
-# Final graphics for the paper
+% Final graphics for the paper
+
+- [back to index](index.html)
+
+`generated at:`
+```{r, echo=FALSE}
+Sys.time()
+```
+---
 
@@
 setwd("~/workspace/lits/attributions/")
-library(knitr)
-knit2html('finalPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+library(kaRski)
+knit("finalPlots.Rmd")
+system("pandoc -s finalPlots.md -o finalPlots.html --toc --number-section")
 

### 2013-11-21 loadClean.Rmd
A new css with left side navigation

@@
 library(knitr)
-knit2html('loadClean.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+library(markdown)
+knit2html('loadClean.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 

### 2013-11-21 regressionModelling.Rmd
A new css with left side navigation

--- 2013-11-20 regressionModelling.Rmd
@@
 
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
+```{r, eval=FALSE, echo=FALSE}
+library(knitr)
+library(markdown)
+knit2html('regressionModelling.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attributions/")
-library(knitr)
-knit2html('regressionModelling.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 ```
@@
         custom.coef.names = coef.names,
-        inline.css = TRUE,
+        inline.css = FALSE,
         doctype = FALSE,
@@
         naive = TRUE,
-        inline.css = TRUE,
+        inline.css = FALSE,
         doctype = FALSE,
@@
         ci.test = NULL,
-        inline.css = TRUE,
+        inline.css = FALSE,
         doctype = FALSE,
@@
         custom.coef.names = coef.names.multi,
-        inline.css = TRUE,
+        inline.css = FALSE,
         doctype = FALSE,

### 2013-11-21 descriptive_analysis.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.

--- 2013-11-20 descriptive_analysis.Rmd
@@
 # Descriptive visualizations
-
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
 
@@
 library(knitr)
-knit2html('descriptive_analysis.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-
-knitpandoc("descriptive_analysis",toc)
+library(markdown)
+knit2html('descriptive_analysis.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 ```

### 2013-11-21 finalPlots.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.

--- 2013-11-20 finalPlots.Rmd
@@
 # Final graphics for the paper
-
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
 
@@
 library(knitr)
-knit2html('finalPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-
+library(markdown)
+knit2html('finalPlots.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
 ```

### 2013-11-21 loadClean.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.

@@
 # Loading and cleaning the data
-
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
 
@@
 library(plyr)
-medianData <- ddply(lits1,~cntry,summarise,incMedian=median(q211, na.rm=TRUE))
+lits1$q211_rec <- lits1$q211
+lits1$q211_rec[lits1$q211_rec >= 11] <- NA
+
+medianData <- ddply(lits1,~cntry,summarise,incMedian=median(q211_rec, na.rm=TRUE))
 lits1 <- merge(lits1,medianData,by="cntry",all.x=TRUE)
 
-lits1$income2[lits1$q211 < lits1$incMedian]  <- 'Low'
-lits1$income2[lits1$q211 >= lits1$incMedian]  <- 'High'
+lits1$income2[lits1$q211_rec < lits1$incMedian]  <- 'Low'
+lits1$income2[lits1$q211_rec >= lits1$incMedian]  <- 'High'
 
@@
 # $income
-
-lits2$income2[lits2$q227 >= 1 & lits2$q227 <= 5]  <- 'Low'
-lits2$income2[lits2$q227 >= 6 & lits2$q227 <= 10]  <- 'High'
-
-lits2$income2 <- factor(lits2$income2, 
-                       levels=c("Low",
-                                "High"))
 
@@
 
-### ------ BEGIN 2006 ---------- ####
-## perGini
-library(laeken)
-perGini2006 <- as.data.frame(gini(inc="q211",
-                      weights = "weight_1",
-                      data = lits1,
-                      na.rm = TRUE,
-                      breakdown="cntry")[2])
-
-names(perGini2006) <- c("cntry","perGini")
-perGini2006
-
-### ------ END 2006 ---------- ####
-
-library(laeken)
-perGini2010 <- as.data.frame(gini(inc="q227_rec",
-                      weights = "weight",
-                      data = lits2,
-                      na.rm = TRUE,
-                      breakdown="cntry")[2])
-
-names(perGini2010) <- c("cntry","perGini")
-perGini2010
-
 lits1$age <- lits1$age1
@@
                                   'unemployment benefits',
-                                  'pensions',
                                   'community/privately provided social benefits',
@@
 ###
-
-# $edu
-###
-
-lits1$edu2[lits1$q501 %in% c('compulsory school education',
-                         'no degree/education')] <- 'no or compulsory'
-lits1$edu2[!(lits1$q501 %in% c('compulsory school education',
-                         'no degree/education'))] <- 'higher'
-
-lits1$edu2 <- factor(lits1$edu2,
-                     levels=c("no or compulsory",
-                              "higher"))
-
-summary(lits1$edu2)
-### ------ END 2006 ---------- ####
 
@@
 summary(lits2$incsour3)
+
+# $edu
 ###
+summary(lits1$q501)
+
+lits1$edu2[lits1$q501 %in% c(#'compulsory school education',
+                         'no degree/education')] <- 'no or compulsory'
+lits1$edu2[!(lits1$q501 %in% c(#'compulsory school education',
+                         'no degree/education'))] <- 'higher'
+
+lits1$edu2 <- factor(lits1$edu2,
+                     levels=c("no or compulsory",
+                              "higher"))
+
+summary(lits1$edu2)
+### ------ END 2006 ---------- ####
 
@@
 
-lits2$edu2[lits2$q515 == 1 | lits2$q515 == 2]  <- "no or compulsory"
-lits2$edu2[lits2$q515 >= 3]  <- "higher"
+lits2$edu2[lits2$q515 == 1]  <- "no or compulsory"
+lits2$edu2[lits2$q515 >= 2]  <- "higher"
 
@@
 ## Data for the analysis
-df <- subset(lits2, select=c("SerialID","weight","cntry","country0","q309",
+
+df <- subset(lits2, select=c("SerialID","weight",
+                             "cntry","q309",
                              "sex","age",
-                             "future.diff","past.diff",
                              "pov.log.social",
@@
                              "income2",
-                             "crise",
+                             "poverty",
+                             "future.diff","past.diff"
+                             ))
+names(lits1)[2] <- "SerialID"
+names(lits1)[1165] <- "weight"
+
+df2 <- lits1[,c("SerialID","weight",
+                             "cntry","q309",
+                             "sex","age",
+                             "pov.log.social",
+                             "pov.log.individual",
+                             "socialism","incsour3",
+                             "edu2",
+                             "group_analysis",
+                             "income2",
                              "poverty"
-                             ))
+                             )]
+
 # filter out not wanted countries
@@
 ```{rLoadClean6, eval=TRUE}
-###
-library(reshape2)
+### Dependent var country level values
+## 2010
 library(survey)
-d.df <- svydesign(id = ~SerialID, 
-                  weights = ~weight, 
-                  data = df)
+d.df <- svydesign(id = ~SerialID, weights = ~weight, data = df)
+q309_10 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
+q309_10 <- q309_10[!is.na(q309_10$Freq), ]
+q309_10w <- dcast(q309_10, cntry ~ q309, value.var="Freq")
+names(q309_10w) <- c("cntry","socialBlame",
+                        "individualBlame","individualFate",
+                        "socialFate","notStated","dontKnow")
 
-summary309 <- data.frame(prop.table(svytable(~q309 + 
-                                               cntry, d.df), 
+poverty_10 <- data.frame(prop.table(svytable(~poverty + cntry, d.df), 
                                         2) * 100)
+poverty_10 <- poverty_10[!is.na(poverty_10$Freq), ]
+poverty_10w <- dcast(poverty_10, cntry ~ poverty, value.var="Freq")
+names(poverty_10w) <- c("cntry","socialBlameRec","individualBlameRec",
+                        "individualFateRec","socialFateRec")
+library(laeken)
+perGini2010 <- as.data.frame(gini(inc="q227_rec",
+                      weights = "weight",
+                      data = lits2,
+                      na.rm = TRUE,
+                      breakdown="cntry")[2])
+names(perGini2010) <- c("cntry","perGini")
 
-summary309 <- summary309[!is.na(summary309$Freq), ]
-summary309w <- dcast(summary309, 
-                     cntry ~ q309, 
-                     value.var="Freq")
-names(summary309w) <- c("cntry","socialBlame",
+macro2010 <- merge(q309_10w,poverty_10w,by="cntry",all.x=TRUE)
+macro2010 <- merge(macro2010,perGini2010,by="cntry",all.x=TRUE)
+macro2010$year <- 2010
+### ------ BEGIN 2006 ---------- ####
+library(survey)
+df2 <- df2[!is.na(df2$weight),]
+d.df <- svydesign(id = ~SerialID, weights = ~weight, data = df2)
+q309_06 <- data.frame(prop.table(svytable(~q309 + cntry, d.df), 2) * 100)
+q309_06 <- q309_06[!is.na(q309_06$Freq), ]
+q309_06w <- dcast(q309_06, cntry ~ q309, value.var="Freq")
+names(q309_06w) <- c("cntry","socialBlame",
                         "individualBlame","individualFate",
-                        "socialFate","not.stated","dont.know")
+                        "socialFate","notStated")
+q309_06w$dontKnow <- NA
 
-summaryPoverty <- data.frame(prop.table(svytable(~poverty + 
-                                                   cntry, d.df), 
+poverty_06 <- data.frame(prop.table(svytable(~poverty + cntry, d.df), 
                                         2) * 100)
-summaryPoverty <- summaryPoverty[!is.na(summaryPoverty$Freq), ]
+poverty_06 <- poverty_06[!is.na(poverty_06$Freq), ]
+poverty_06w <- dcast(poverty_06, cntry ~ poverty, value.var="Freq")
+names(poverty_06w) <- c("cntry","socialBlameRec","individualBlameRec",
+                        "individualFateRec","socialFateRec")
+## perGini
+library(laeken)
+perGini2006 <- as.data.frame(gini(inc="q211_rec",
+                      weights = "weight_1",
+                      data = lits1,
+                      na.rm = TRUE,
+                      breakdown="cntry")[2])
 
-summaryPovertyw <- dcast(summaryPoverty, cntry ~ poverty, value.var="Freq")
-names(summaryPovertyw) <- c("cntry","socialBlame","individualBlame","individualFate","socialFate")
+names(perGini2006) <- c("cntry","perGini")
+macro2006 <- merge(q309_06w,poverty_06w,by="cntry",all.x=TRUE)
+macro2006 <- merge(macro2006,perGini2006,by="cntry",all.x=TRUE)
+macro2006$year <- 2006
 
+#macro <- rbind(macro2006,macro2010)
+### ------ END 2006 ---------- ####
 ```
@@
                    'Sweden','France'), 
-       indicator = c("ffp_fsi", # Failed States Index
-                     "wbgi_vae", # World Bank - Voice and Accountability
-                     "undp_hdi", # UNDP Human Development Index
-                     "fh_polity2", # Democracy (Freedom House/Polity)
-                     "uw_gini"))
+       indicator = c("wbgi_vae"#, # World Bank - Voice and Accountability
+                     #"uw_gini"
+                     ))
+
+# datQog <- datQog[datQog$year >= 2000, ]
+# table(factor(datQog$cname),factor(datQog$year),factor(datQog$indicator))
+
 # subset the latest value per cntry per indicator
-library(plyr)
-datQogSum <- ddply(datQog, c("cname","indicator"), 
-                   subset, year == max(year))
+#library(plyr)
+#datQogSum <- ddply(datQog, c("cname","indicator"), 
+#                   subset, year == max(year))
+datQogSum <- datQog[datQog$year %in% c(2006,2009),]
+datQogSum$year[datQogSum$year == 2009] <- 2010
+names(datQogSum)[3] <- "wbgi_vae"
+datQogSum$indicator <- NULL
 
-datQogSum.w <- dcast(datQogSum, 
-                     cname ~ indicator, 
-                     value.var="value")
+# datQogSum.w <- dcast(datQogSum, 
+#                      cname+indicator ~ year, 
+#                      value.var="value")
+
+datQogSum.w <- datQogSum
 
@@
                                              "CIS"))
+df.qog <- datQogSum.w
+names(df.qog)[1] <- "cntry"
 
-df.qog <- datQogSum.w
-
-names(df.qog)[1] <- "cntry"
+macro2006 <- merge(macro2006,df.qog[df.qog$year == 2006,],by=c("cntry","year"))
+macro2010 <- merge(macro2010,df.qog[df.qog$year == 2010,],by=c("cntry","year"))
 
@@
                       "SI","SE","TJ","TR","UA","UZ"), 
-           start = 2007, end = 2010)
-
+           start = 2003, end = 2010)
 library(reshape2)
 names(wb) <- c("iso2c","cntry","var","year")
+wb$year <- paste("x",wb$year,sep="")
 wb.w <- dcast(wb, 
@@
 
-names(wb.w) <- c("cntry","x2007","x2008","x2009","x2010")
 wb.w$gdpchange07.10 <- wb.w$x2010/wb.w$x2007 * 100 - 100
-gdp <- wb.w[,c(1,6)]
+wb.w$gdpchange03.06 <- wb.w$x2006/wb.w$x2003 * 100 - 100
 
-gdp$cntry[gdp$cntry == 'Russian Federation'] <- 'Russia'
-gdp$cntry[gdp$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
-gdp$cntry[gdp$cntry == 'United Kingdom'] <- 'Great Britain'
-gdp$cntry[gdp$cntry == 'Slovak Republic'] <- 'Slovakia'
+wb.w$cntry[wb.w$cntry == 'Russian Federation'] <- 'Russia'
+wb.w$cntry[wb.w$cntry == 'Kyrgyz Republic'] <- 'Kyrgyzstan'
+wb.w$cntry[wb.w$cntry == 'United Kingdom'] <- 'Great Britain'
+wb.w$cntry[wb.w$cntry == 'Slovak Republic'] <- 'Slovakia'
 
+gdp10 <- wb.w[,c(1,10)]
+names(gdp10) <- c("cntry","gdpChange")
+gdp06 <- wb.w[,c(1,11)]
+names(gdp06) <- c("cntry","gdpChange")
+
+macro2006 <- merge(macro2006,gdp06,by="cntry")
+macro2010 <- merge(macro2010,gdp10,by="cntry")
 ```
@@
 ```{rLoadClean9}
+# yhdistetään nää aikaisempien datojen kanssa
+macro <- rbind(macro2006,macro2010)
 
-# yhdistetään nää aikaisempien datojen kanssa
-macro <- merge(summary309w,gdp,by="cntry", all.x=TRUE)
-macro <- merge(macro,df.qog,by="cntry")
-macro <- merge(macro,perGini2010,by="cntry")
+df <- merge(df,macro[macro$year == 2010,],by="cntry")
+df2 <- merge(df2,macro[macro$year == 2006,],by="cntry")
 
-df <- merge(df,macro,by="cntry")
-
-rm(lits2)
-rm(df.qog)
-rm(df.cross.postsoc)
-rm(gdp)
-rm(t3)
-rm(summary309)
-rm(summary309_rec)
-rm(d.df)
-rm(wb)
-rm(wb.w)
-rm(cntry)
-rm(iso2c)
-rm(lits1)
-rm(medianData)
-rm(perGini2006)
-rm(perGini2010)
+rm(list=setdiff(ls(), c("macro","df","df2")))
 

### 2013-11-21 compareWaves.Rmd
Changed the toc font size in css

# Comparing waves 2006 and 2010

```{r, eval=FALSE, echo=FALSE}
library(knitr)
library(markdown)
knit2html('compareWaves.Rmd', 
          stylesheet="~/workspace/web/css/rmarkdown.css",
          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
```

```{roptsattrib, echo=FALSE}
opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
```

```{rcompload}
load("~/workspace/lits/attributions/data/lits.RData")
```

# Dependent variable

## Original coding

### By country

```{rcompareDvCntryOrig, fig.height=16}
dfPlot <- macro[,1:8]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("cntry","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=cntry,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

### By country group

```{rcompareDvCgOrig, fig.height=16}
dfPlot <- macro[,c(2:8,16)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("group_general","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=group_general,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

## Only the four values

### By country

```{rcompareDvCntryRec, fig.height=16}
dfPlot <- macro[,c(1:2,9:12)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("cntry","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=cntry,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

### By country group

```{rcompareDvCgRec, fig.height=16}
dfPlot <- macro[,c(2,9:12,16)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("group_general","year"))
library(ggplot2)
ggplot(dfPlotW, aes(x=group_general,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free_x") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))
```

# Independent variables

## Individual level

### Transfer dependency (share of dependent hh's)

```{rcompareIvTransfer}
library(survey)
d.df10 <- svydesign(id = ~SerialID, weights = ~weight, data = df)
d.df06 <- svydesign(id = ~SerialID, weights = ~weight, data = df2)

dat10 <- data.frame(prop.table(svytable(~incsour3 + cntry, d.df10), 2) * 100)
names(dat10) <- c("Var","cntry","Freq")
dat10$year <- 2010
dat06 <- data.frame(prop.table(svytable(~incsour3 + cntry, d.df06), 2) * 100)
names(dat06) <- c("Var","cntry","Freq")
dat06$year <- 2006
datPlot <- rbind(dat06,dat10)

datPlot2 <- datPlot[datPlot$Var == "dependent",]

ggplot(datPlot2, aes(x=cntry,y=Freq,fill=factor(year))) +
        geom_bar(stat="identity",position="dodge") +
        theme(legend.position="top") +
        theme(axis.text.x = element_text(angle = 90))

```

### Education level (share of low education hh's)

```{rcompareIvEdu}
dat10 <- data.frame(prop.table(svytable(~edu2 + cntry, d.df10), 2) * 100)
names(dat10) <- c("Var","cntry","Freq")
dat10$year <- 2010
dat06 <- data.frame(prop.table(svytable(~edu2 + cntry, d.df06), 2) * 100)
names(dat06) <- c("Var","cntry","Freq")
dat06$year <- 2006
datPlot <- rbind(dat06,dat10)

datPlot2 <- datPlot[datPlot$Var == "no or compulsory",]

ggplot(datPlot2, aes(x=cntry,y=Freq,fill=factor(year))) +
        geom_bar(stat="identity",position="dodge") +
        theme(legend.position="top") +
        theme(axis.text.x = element_text(angle = 90))

```

### Income level (share of low income hh's)

```{rcompareIvInc}
dat10 <- data.frame(prop.table(svytable(~income2 + cntry, d.df10), 2) * 100)
names(dat10) <- c("Var","cntry","Freq")
dat10$year <- 2010
dat06 <- data.frame(prop.table(svytable(~income2 + cntry, d.df06), 2) * 100)
names(dat06) <- c("Var","cntry","Freq")
dat06$year <- 2006
datPlot <- rbind(dat06,dat10)

datPlot2 <- datPlot[datPlot$Var == "Low",]

ggplot(datPlot2, aes(x=cntry,y=Freq,fill=factor(year))) +
        geom_bar(stat="identity",position="dodge") +
        theme(legend.position="top") +
        theme(axis.text.x = element_text(angle = 90))

```

## Macro level

```{rcompareIvMacro, fig.height=16}
dfPlot <- macro[,c(1:2,13,14,17)]
library(reshape2)
dfPlotW <- melt(dfPlot,id.vars=c("cntry","year"),
                value.vars=c("perGini",
                             "wbgi_vae",
                             "gdpChange"))
library(ggplot2)
ggplot(dfPlotW, aes(x=cntry,y=value,fill=factor(year))) +
    geom_bar(stat="identity",position="dodge") +
    facet_wrap(~variable, ncol=1, scales="free") +
    theme(legend.position="top") +
    theme(axis.text.x = element_text(angle = 90))

```

# Associations

*Name is in the 2010 end of the line*

## Good governanance

### vs. Social blame

```{rcompareAssoGoodSB}
cbPalette <- c("#E69F00","#56B4E9","#999999","#009E73", "#F0E442", "#0072B2","#D55E00", "#CC79A7")
dfPlot <- macro[,c(1:4,13,14,16,17)]
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=wbgi_vae,y=socialBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=socialBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### vs. Individual blame

```{rcompareAssoGoodIB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=wbgi_vae,y=individualBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=wbgi_vae,y=individualBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

## Change in GDP

### vs. Social blame

```{rcompareAssoGdpSB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=gdpChange,y=socialBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=socialBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### vs. Individual blame

```{rcompareAssoGdbIB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=gdpChange,y=individualBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=gdpChange,y=individualBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

## Perceived Gini

### vs. Social blame

```{rcompareAssoPerginiSB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=socialBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=socialBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=socialBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=perGini,y=socialBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=socialBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### vs. Individual blame

```{rcompareAssoPerginiIB}
library(ggplot2)
ggplot(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=individualBlame,group=cntry,color=group_general)) +
    geom_point() +
    geom_text(data=dfPlot[dfPlot$year == 2006,], 
       aes(x=perGini,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_point(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=individualBlame,group=cntry,color=group_general)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=individualBlame,group=cntry,label=year),
               vjust=-1,size=4) +
    geom_line(data=dfPlot,
              aes(x=perGini,y=individualBlame,group=cntry)) +
    geom_text(data=dfPlot[dfPlot$year == 2010,], 
       aes(x=perGini,y=individualBlame,group=cntry,label=cntry),
               vjust=1,size=4) +
  scale_colour_manual(values=cbPalette) +
   theme_minimal() +
  theme(legend.title=element_blank()) +
  theme(legend.text=element_text(size=12)) +
  theme(legend.position="top")

```

### 2013-11-21 loadClean.Rmd
Changed the toc font size in css

@@
 macro2006$year <- 2006
-
-#macro <- rbind(macro2006,macro2010)
-### ------ END 2006 ---------- ####
 ```

### 2013-11-22 loadClean.Rmd
Regression models updated with perceived gini

@@
                              "poverty",
+                             "crise",
                              "future.diff","past.diff"

### 2013-11-22 regressionModelling.Rmd
Wave 2006 also included. Numerous minor corrections to data munging also for 2010 data. Nothing major I think. Well, low education is currently covering only uneducated individuals.
Changed the toc font size in css
Regression models updated with perceived gini

@@
 ```
-
-`log of changes`
-
-date      comment
---------  -------------------
-20131103  Non multi-level logistic regression done but not interpret yet. Bivariate multilevel models the same. Multivariate multilevel models running over night.
-20131105  Regression modelling almost finished. Bug in a software caused a delay. Will be fixed by Wed morning. Now the tables are there, but coef and model names are not customized.
-20131106  Model names fixed. Multinomial models coming by midday
 
@@
                          family = binomial(logit), data = df.model)
- bisoc.8 <- glmer(pov.log.social ~ uw_gini+(1|cntry),
-                         family = binomial(logit), data = df.model)
- bisoc.9 <- glmer(pov.log.social ~ gdpchange07.10+(1|cntry),
+# bisoc.8 <- glmer(pov.log.social ~ perGini+(1|cntry),
+#                         family = binomial(logit), data = df.model)
+bisoc.9 <- glmer(pov.log.social ~ gdpChange+(1|cntry),
                          family = binomial(logit), data = df.model)
@@
                       "Income will worsen","Has affected great or fair amount",
-                      "Gini coefficient",
+                      #"Gini coefficient",
                       "Change in GDP between 2007 to 2010",
@@
           bisoc.5,bisoc.6,
-          bisoc.8,
+          #bisoc.8,
           bisoc.9,bisoc.10,
@@
           bisoc.13),
-        custom.coef.names = coef.names.multi,
+        #custom.coef.names = coef.names.multi,
         naive = TRUE,
@@
                         family = binomial(logit), data = df.model)
-biind.8 <- glmer(pov.log.individual ~ uw_gini+(1|cntry),
-                        family = binomial(logit), data = df.model)
-biind.9 <- glmer(pov.log.individual ~ gdpchange07.10+(1|cntry),
+#biind.8 <- glmer(pov.log.individual ~ perGini+(1|cntry),
+#                        family = binomial(logit), data = df.model)
+biind.9 <- glmer(pov.log.individual ~ gdpChange+(1|cntry),
                         family = binomial(logit), data = df.model)
@@
                           family = binomial(logit), data = df.model)
-bisoc.12 <- glmer(pov.log.individual ~ group_balinCis+(1|cntry),
+biind.12 <- glmer(pov.log.individual ~ group_balinCis+(1|cntry),
                           family = binomial(logit), data = df.model)
-bisoc.13 <- glmer(pov.log.individual ~ perGini+(1|cntry),
+biind.13 <- glmer(pov.log.individual ~ perGini+(1|cntry),
                           family = binomial(logit), data = df.model)
@@
           biind.5,biind.6,
-          biind.8,
+          #biind.8,
           biind.9,biind.10,
@@
              biind.13),
-        custom.coef.names = coef.names.multi,
+        #custom.coef.names = coef.names.multi,
         naive = TRUE,
@@
                         past.diff+future.diff+crise+
-                        uw_gini+(1|cntry),
+                        perGini+(1|cntry),
                        family = binomial(logit),
@@
                         past.diff+future.diff+crise+
-                        uw_gini+gdpchange07.10+(1|cntry),
+                        perGini+gdpChange+(1|cntry),
                        family = binomial(logit),
@@
                         past.diff+future.diff+crise+
-                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                        perGini+gdpChange+wbgi_vae+(1|cntry),
                        family = binomial(logit),
@@
                       "Income will worsen","Has affected great or fair amount",
-                      "Gini coefficient",
+                      "Perceived Gini",
                       "Change in GDP between 2007 to 2010",
@@
 ```{rmultilevelMultiInd}
-
 ind.0 <- glmer(pov.log.individual ~ (1|cntry),

### 2013-12-01 index.Rmd
Some Sunday evening edits..

@@
-% Attributions for Poverty in Post-Socialist Countries
-% Markus Kainu & Mikko Niemelä
-% Nov 2013
+# Attributions for Poverty in Post-Socialist Countries - Markus Kainu & Mikko Niemelä
 
-`generated at:`
-```{r, echo=FALSE}
-Sys.time()
-```
----
-
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link> <link href='http://fonts.googleapis.com/css?family=Oswald' rel='stylesheet' type='text/css'>
-
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attrib/attrib_year2013")
-library(kaRski)
-knitpandoc("index")
-
-cd ~/workspace/lits/attrib/attrib_year2013/
-lftp -u np16003 -e "mput *.html *.Rmd *.pdf;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/
-
-cd ~/workspace/lits/attrib/attrib_year2013/figure
-lftp -u np16003 -e "mirror -R;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/figure
-
-cd ~/workspace/lits/attrib/attrib_year2013/moscow2013
-lftp -u np16003 -e "mput *.html;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/moscow2013
-
-cd ~/workspace/lits/attrib/attrib_year2013/turku2013
-lftp -u np16003 -e "mput *.html;quit" ftp://np16003@www.muuankarski.org/domains/muuankarski.org/public_html/research/attributions/turku2013
-```
+****
 
@@
 ```
-
-`log of changes`
-
-date      comment
---------  -------------------
-20131103  Regression modelling in process. Final models will be published on Tuesday afternoon
-20131105  Regression modelling almost finished. Bug in a software caused a delay. Will be fixed by Wed morning. Now the tables are there, but coef and model names are not customized.
-20131106  Regression modelling somewhat done for Turku seminar.
-20131109  Corrections from Turku and the text restuctured. Methods needs to be opened
-20131111  Presentation for HSE conference and improving the text a bit
 

### 2013-12-01 loadClean.Rmd
Some Sunday evening edits..

@@
 
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attributions/")
-library(knitr)
-library(markdown)
-knit2html('loadClean.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-
-```
-
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
 ```
@@
 ## 2010
+library(reshape2)
 library(survey)
@@
 perGini2006 <- as.data.frame(gini(inc="q211_rec",
-                      weights = "weight_1",
+                      weights = "weight",
                       data = lits1,

### 2013-12-01 komennot.R

@@ -1,0 +1,35 @@
+
+setwd("~/workspace/lits/attributions")
+library(knitr)
+library(markdown)
+#
+knit('index.Rmd','README.md')
+knit2html('index.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+#
+knit2html('loadClean.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+#
+knit2html('finalPlots.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+#
+knit2html('descriptive_analysis.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+#
+library(knitr)
+library(markdown)
+knit2html('compareWaves.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+#
+knit2html('mapPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
+#
+library(knitr)
+library(markdown)
+knit2html('regressionModelling.Rmd', 
+          stylesheet="~/workspace/web/css/rmarkdown.css",
+          options = c('toc', markdown::markdownHTMLOptions(TRUE)))

### 2013-12-01 index.Rmd
a quick table edit

@@
 
--------------------
-html-ouput                                          source code
-------------                                        --------------
-[loading, cleaning and summarizing](loadClean.html) [source](loadClean.Rmd)
--------------------
+| html-ouput | source code |
+| ------------ | -------------- |
+| [loading, cleaning and summarizing](loadClean.html) | [source](loadClean.Rmd) |
 
@@
 
--------------------
-html-ouput                                        source code
-------------                                      --------------
-[Descriptive analysis](descriptive_analysis.html) [source](descriptive_analysis.Rmd)
-
-[Spatial visualization](mapPlots.html)            [source]((mapPlots.Rmd))
-
-[Graphics in the paper](finalPlots.html)          [source](finalPlots.Rmd)
--------------------
+| html-ouput | source code |
+| ------------ | -------------- |
+| [Descriptive analysis](descriptive_analysis.html) | [source](descriptive_analysis.Rmd) |
+| [Spatial visualization](mapPlots.html) | [source](mapPlots.Rmd) |
+| [Graphics in the paper](finalPlots.html) | [source](finalPlots.Rmd) |
 
@@
 
--------------------
-html-ouput                                                source code
-------------                                              --------------
-[Regression analysis](regressionModelling.html)           [source](regressionModelling.Rmd)
-
--------------------
+| html-ouput | source code |
+| ------------ | -------------- |
+| [Regression analysis](regressionModelling.html) | [source](regressionModelling.Rmd) |
 
@@
 
--------------------
-pdf-output                                                source code
-------------                                              --------------
-[Article draft in .pdf](article2013.pdf)                  [source](article2013.Rmd)
--------------------
+| pdf-ouput | source code |
+| ------------ | -------------- |
+| [Article draft in .pdf](article2013.pdf) | [source](article2013.Rmd) |
 

### 2013-12-09 komennot.R
removed .R from repo

@@ -1,35 +1,0 @@
-
-setwd("~/workspace/lits/attributions")
-library(knitr)
-library(markdown)
-#
-knit('index.Rmd','README.md')
-knit2html('index.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-#
-knit2html('loadClean.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-#
-knit2html('finalPlots.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-#
-knit2html('descriptive_analysis.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-#
-library(knitr)
-library(markdown)
-knit2html('compareWaves.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-#
-knit2html('mapPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-#
-library(knitr)
-library(markdown)
-knit2html('regressionModelling.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))

### 2013-12-11 article2013.Rmd
Commented on the country grouping part of the article
more commenting
another row added
Esimerkkiviesti Anulle
Tähän saakka tehtyä:
- koska gini-muuttujassa oli vähän ongelmia (1. tiedot vähän miltä vuosilta sattuu, 2. EI TILASTOLLISESTI MERKITSEVÄ, 3. teoreettisesti hieman epävarma) niin kokeilin laskea koetun eriarvoisuuden ginin aineistosta. Se on nyt mukana niin paperin kuvioissa kun myös ainakin osassa regressioita.
- Maaryhmittely on vielä mukana teksteissä, mutta sen voisi jättää kokonaan pois. Sitä ei ole analyyseissä ollutkaan hetkeen, mutta sen voisi jättää pois teoriasta, abstraktis

@@
 % Markus Kainu[^**];Mikko Niemelä[^***]
-% November 19, 2013
+% December 11, 2013
 
-[^*]:*Work in progress. See the computation analysis at [research.muuankarski.org/attributions/](http://research.muuankarski.org/attributions/)
+[^*]:*Work in progress. See the analysis outputs at [http://muuankarski.github.io/attributions/](http://muuankarski.github.io/attributions/)
 
@@
 [^***]:Senior Researcher. Social Insurance Institution of Finland, KELA a<a href="mailto:mikko.niemela@kela.fi">mikko.niemela@kela.fi</a>
-
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
 
@@
 
-From the perspective of value structure, the role of "socialist or soviet legacy" in poverty attributions is worth of considering.
+>From the perspective of value structure, the role of "socialist or soviet legacy" in poverty attributions is worth of considering.
 Murthi and Tiongson (2008) reviewed large body of literature on East-West differences in preferences for redistribution of which vast majority supported the "socialist legacy".
@@
 
+<!--
+
 ![Gini coefficient and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_gini.pdf)
 
-There is also a substantial variation in the level of income inequality among post-socialist countries.
+-->
+
+![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
+
+>There is also a substantial variation in the level of income inequality among post-socialist countries.
 Figure 3 shows that the higher level of income inequality is associated with higher level of social blame.
@@
 
-Table 4 and 5 present figures for multilevel logistic analysis.
-Calculating R² or residual standard deviation makes no sense in the case of logistic regression or other binomial models.
-Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit.
-Deviance can be treated as similar to residual standard deviation for generalized linear models.
-Basically, deviance is a measure of error, and lower deviance implicates better fit to data.
-When a random predictor is added to model, we expect deviance to decrease, on average, by 1.
-When predictor is more informative deviance should decrease more than 1.
-(Gelman and Hill, 2007).
-As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.
-
-Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences.
-Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation.
-Inclusion of individual level predictors decreases the variance in both models, but only moderately.
-
 <!-- As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
@@
 Table 4 and 5 present figures for multilevel logistic analysis.
-Calculating R² or residual standard deviation makes no sense in the case of logistic regression or other binomial models.
+Calculating R squared or residual standard deviation makes no sense in the case of logistic regression or other binomial models.
 Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit.

### 2013-12-11 finalPlots.Rmd
Changed the toc font size in css
Some Sunday evening edits..
Tähän saakka tehtyä:
- koska gini-muuttujassa oli vähän ongelmia (1. tiedot vähän miltä vuosilta sattuu, 2. EI TILASTOLLISESTI MERKITSEVÄ, 3. teoreettisesti hieman epävarma) niin kokeilin laskea koetun eriarvoisuuden ginin aineistosta. Se on nyt mukana niin paperin kuvioissa kun myös ainakin osassa regressioita.
- Maaryhmittely on vielä mukana teksteissä, mutta sen voisi jättää kokonaan pois. Sitä ei ole analyyseissä ollutkaan hetkeen, mutta sen voisi jättää pois teoriasta, abstraktista ja todeta jotain siitä, miksi emme alaryhmittel

@@
 # Final graphics for the paper
-
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attributions/")
-library(knitr)
-library(markdown)
-knit2html('finalPlots.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-```
 
@@
 {
-  return(data.frame(COR = cor(t3$socialBlame, t3$gdpchange07.10)))
+  return(data.frame(COR = cor(t3$socialBlame, t3$gdpChange)))
 }
@@
 t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(socialBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
+cnames <- round(stats:::aggregate.formula(cbind(socialBlame, gdpChange) ~ COR, data=t4, mean),2)
 
-plot1 <- ggplot(t4, aes(x=gdpchange07.10, y=socialBlame, 
+plot1 <- ggplot(t4, aes(x=gdpChange, y=socialBlame, 
                         label=cntry)) +
   geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
-  geom_text(data=cnames, aes(gdpchange07.10,socialBlame, label = COR, group=COR), 
+  geom_text(data=cnames, aes(gdpChange,socialBlame, label = COR, group=COR), 
             size=2.5, color="black") +
@@
   geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=gdpchange07.10, 
+  geom_point(data=t4, aes(x=gdpChange, 
                           y=socialBlame, color=group_general)) +
   # west reference
-  geom_point(data=west, aes(x=gdpchange07.10, 
+  geom_point(data=west, aes(x=gdpChange, 
                           y=socialBlame, 
                             color=group_general), alpha=.5) +
-  geom_text(data=west, aes(x=gdpchange07.10, 
+  geom_text(data=west, aes(x=gdpChange, 
                            y=socialBlame, label=cntry),
@@
 {
-  return(data.frame(COR = cor(t3$individualBlame, t3$gdpchange07.10)))
+  return(data.frame(COR = cor(t3$individualBlame, t3$gdpChange)))
 }
@@
 t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(individualBlame, gdpchange07.10) ~ COR, data=t4, mean),2)
-plot2 <- ggplot(t4, aes(x=gdpchange07.10, 
+cnames <- round(stats:::aggregate.formula(cbind(individualBlame, gdpChange) ~ COR, data=t4, mean),2)
+plot2 <- ggplot(t4, aes(x=gdpChange, 
                         y=individualBlame, label=cntry)) +
@@
   geom_text(size=2, vjust=-0.8, hjust=0.5, color="black") +
-  geom_text(data=cnames, aes(gdpchange07.10,individualBlame, label = COR, group=COR), 
+  geom_text(data=cnames, aes(gdpChange,individualBlame, label = COR, group=COR), 
             size=2.5, color="black") +
   geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=gdpchange07.10, 
+  geom_point(data=t4, aes(x=gdpChange, 
              y=individualBlame, color=group_general)) +
   # west reference
-  geom_point(data=west, aes(x=gdpchange07.10, 
+  geom_point(data=west, aes(x=gdpChange, 
                             y=individualBlame, 
                             color=group_general), alpha=.5) +
-  geom_text(data=west, aes(x=gdpchange07.10, 
+  geom_text(data=west, aes(x=gdpChange, 
                            y=individualBlame, label=cntry),
@@
 
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$socialBlame, t3$perGini)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(socialBlame, perGini) ~ COR, data=t4, mean),2)
+# func <- function(t3)
+# {
+#   return(data.frame(COR = cor(t3$socialBlame, t3$perGini)))
+# }
+# df.cor <- ddply(t3, .(group1), func)
+# t4 <- join(t3,df.cor,by="group1")
+# cnames <- round(stats:::aggregate.formula(cbind(socialBlame, perGini) ~ COR, data=t4, mean),2)
 plot1 <- ggplot(t4, aes(x=perGini, y=socialBlame, 
@@
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(perGini,socialBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
+#   geom_text(data=cnames, aes(perGini,socialBlame, label = COR, group=COR), 
+#             size=2.5, color="black") +
   labs(x = "Perceived inequality from data", 
@@
 
-func <- function(t3)
-{
-  return(data.frame(COR = cor(t3$individualBlame, t3$perGini)))
-}
-df.cor <- ddply(t3, .(group1), func)
-t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(individualBlame, perGini) ~ COR, data=t4, mean),2)
+# func <- function(t3)
+# {
+#   return(data.frame(COR = cor(t3$individualBlame, t3$perGini)))
+# }
+# df.cor <- ddply(t3, .(group1), func)
+# t4 <- join(t3,df.cor,by="group1")
+# cnames <- round(stats:::aggregate.formula(cbind(individualBlame, perGini) ~ COR, data=t4, mean),2)
 plot2 <- ggplot(t4, aes(x=perGini, 
@@
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(perGini,individualBlame, label = COR, group=COR), 
-            size=2.5, color="black") +
+#   geom_text(data=cnames, aes(perGini,individualBlame, label = COR, group=COR), 
+#             size=2.5, color="black") +
   labs(x ="Perceived inequality from data", 
@@
 library(rgdal)
+library(grid)
 map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 

### 2013-12-11 loadClean.Rmd
Tähän saakka tehtyä:
- koska gini-muuttujassa oli vähän ongelmia (1. tiedot vähän miltä vuosilta sattuu, 2. EI TILASTOLLISESTI MERKITSEVÄ, 3. teoreettisesti hieman epävarma) niin kokeilin laskea koetun eriarvoisuuden ginin aineistosta. Se on nyt mukana niin paperin kuvioissa kun myös ainakin osassa regressioita.
- Maaryhmittely on vielä mukana teksteissä, mutta sen voisi jättää kokonaan pois. Sitä ei ole analyyseissä ollutkaan hetkeen, mutta sen voisi jättää pois teoriasta, abstraktista ja todeta jotain siitä, miksi emme alaryhmittele post-sos maita.

Seuraavaksi:
- mun pitäisi kirjoittaa vähä

@@
                    'Sweden','France'), 
-       indicator = c("wbgi_vae"#, # World Bank - Voice and Accountability
-                     #"uw_gini"
+       indicator = c("wbgi_vae", # World Bank - Voice and Accountability
+                     "uw_gini"
                      ))
@@
 datQogSum$year[datQogSum$year == 2009] <- 2010
-names(datQogSum)[3] <- "wbgi_vae"
-datQogSum$indicator <- NULL
+#names(datQogSum)[3] <- "wbgi_vae"
+#datQogSum$indicator <- NULL
 
-# datQogSum.w <- dcast(datQogSum, 
-#                      cname+indicator ~ year, 
-#                      value.var="value")
+datQogSum.w <- dcast(datQogSum, 
+                   cname+year ~ indicator, 
+                       value.var="value")
 

### 2013-12-13 article2013.Rmd
Muutin päivämäärän
Koodin korjailua, korrelaatiokokeet ja maaryhmäkappaleen poisto
Korjasin koodia niin että hajontakuviot ovat oikein paperissa.

Katsoin läpi shinylla tehdyn korrelaatiokoneen ja kokosin selkeimmät yhteydet dokkariin correlations.

Otin tekstistä pois maaryhmiä koskevan teorialuvun.

@@
 % Markus Kainu[^**];Mikko Niemelä[^***]
-% December 11, 2013
+% December 12, 2013
 
@@
 Gallie and Paugam, 2002; Kallio and Niemelä, 2013).
-
->From the perspective of value structure, the role of "socialist or soviet legacy" in poverty attributions is worth of considering.
-Murthi and Tiongson (2008) reviewed large body of literature on East-West differences in preferences for redistribution of which vast majority supported the "socialist legacy".
-By socialist legacy they meant that "owing to the legacy of the past there is a strong preference for income equality and redistributive state spending in post-socialist countries".
-While testing the findings of previous studies on East-West differences, this study explores the effect of country groups with different socialist and post-socialist past.
-In order to explore the differences between market economies and post-socialist countries, but also within the "ex-soviet bloc", the study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and Commonwealth of Independent States (CIS) countries.
-In order to examine the distinctiveness of post-socialist perceptions descriptive analyses include also a selection of Western European countries.
 

### 2013-12-13 code/descriptive_analysis.Rmd
First pages commit

@@
+`generated at:`
+```{r, echo=FALSE}
+Sys.time()
+```
+---
+
+<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link> <link href='http://fonts.googleapis.com/css?family=Oswald' rel='stylesheet' type='text/css'>
+
 ```{r, eval=FALSE, echo=FALSE}
@@
 
-```{ropts, echo=FALSE}
-opts_chunk$set(echo=FALSE, dev='svg', eval=TRUE, cache=TRUE, fig.height=10, message=FALSE, warning=FALSE)
+```{roptsattrib, echo=FALSE}
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
 ```
@@
 ```{rdescplot1_1, fig.width=15}
-load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+load("~/workspace/data/lits2.RData")
 library(ggplot2)

### 2013-12-13 code/multilevel_models_test.Rmd
First pages commit

% Multilevel modelling

```{r, eval=FALSE, echo=FALSE}
setwd("~/workspace/lits/attrib/attrib_year2013/code/")
library(knitr)
knit("multilevel_models_test.Rmd")
system("~/.cabal/bin/pandoc multilevel_models_test.md --number-sections --template=../tex_templates/dcolumn.latex -o multilevel_models_test.tex")
#system("perl -pi -e 's/footnotesize/tiny/g' multilevel_models_test.tex")
system("pdflatex multilevel_models_test.tex")
```

```{rmultilevel_analysis, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE, results='hide', eval=TRUE}
library(lme4.0)
#########################################
load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")

library(lme4)
summary(soc.0 <- lmer(pov.log.social ~ (1|cntry),
                      family = binomial, data=df))
summary(soc.1 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+(1|cntry),
                      family = binomial, data=df))
summary(soc.2 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+(1|cntry),family = binomial, data=df))
summary(soc.3 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+(1|cntry),family = binomial, data=df))
summary(soc.4 <- lmer(pov.log.social ~ incsour3+edu2+income2+
                        past.diff+future.diff+crise+
                        uw.gini+gdpchange07.10+wbgi.vae+(1|cntry),family = binomial, data=df))
```

```{rstargazer, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(stargazer)
stargazer(soc.1,soc.2,soc.3,soc.4,
          title="testailua", ord.intercepts = TRUE)
```

```{rmemisc, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(memisc)
toLatex(mtable(soc.1,soc.2,soc.3,soc.4))
```

```{rtexreg, results='asis', echo=FALSE, eval=TRUE, cache=TRUE}
library(texreg)
texreg(list(soc.1,soc.2,soc.3,soc.4), use.packages=FALSE, bold=0.5)
#extract.mer(soc.1, include.pvalues=TRUE)
#summary(soc.1)
```

### 2013-12-13 finalPlots.Rmd
First pages commit

@@
 # valitaan vaan kohdemaat #
-t3 <- subset(macro, group_fine != 'West Europe')
-west <- subset(macro, group_fine == 'West Europe')
+t3 <- subset(macro, group_fine != 'West Europe' & year == 2010)
+west <- subset(macro, group_fine == 'West Europe' & year == 2010)
 #t3 <- macro
@@
 
-```{rattribFinalPlotsGini}
+```{rattribFinalPlotsGini, eval=FALSE}
 # UNU-Wider Gini

### 2013-12-13 code/dv_tables.R

@@ -1,0 +1,112 @@
+load("~/workspace/lits/attributions/data/lits.RData")
+library(survey)
+d.df <- svydesign(id = ~SerialID, 
+                  weights = ~weight, 
+                  data = df)
+
+
+
+t <- data.frame(prop.table(svytable(~group_general+cntry+q309, d.df), 2)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+
+### testailua
+# maa <- subset(t, q309 == "Not stated")
+# maa <- maa[order(maa$group_general),]
+# t$cntry <- factor(t$cntry, levels=maa$cntry)
+# library(ggplot2)
+# ggplot(t, aes(x=cntry,y=Freq,group=cntry)) +
+#   geom_bar(stat="identity", position="dodge") +
+#   facet_wrap(~q309) + coord_flip()
+####
+library(reshape2)
+t2 <-  dcast(t, group_general+cntry ~ q309, value.var="Freq")
+names(t2) <-  c("group_general","country","socialBlame","individualBlame",
+                "individualFate","socialFate","notStated","dontKnow")
+t2 <- t2[,c(1,2,3,4,6,5,7,8)]
+t2 <- t2[order(-t2$socialBlame),]
+########
+#t2$group_general <- NULL
+
+#---------------------------------------------------#
+#---------------------------------------------------#
+#---------------------------------------------------#
+#---------------------------------------------------#
+
+# t <- data.frame(prop.table(svytable(~group_general+cntry+poverty, d.df), 2)*100)
+# t$Freq <- round(t$Freq, 2)
+# t <- subset(t, Freq > 0)
+# t2 <-  dcast(t, group_general+cntry ~ poverty, value.var="Freq")
+# names(t2) <-  c("group_general","country","socialBlame","individualBlame","individualFate","socialFate")
+# t2 <- t2[order(t2$group,-t2$socialBlame),]
+########
+y <- data.frame(prop.table(svytable(~group_general+q309, d.df), 1)*100)
+y$Freq <- round(y$Freq, 2)
+y$country <- c("CEE Total","CIS Total","Western Europe Total")
+y2 <-  dcast(y, group_general+country ~ q309, value.var="Freq")
[…]

### 2013-12-13 code/figures.R

@@ -1,0 +1,287 @@
+library(ggplot2)
+library(gridExtra)
+library(plyr)
+library(grid)
+#########################################
+#########################################
+setwd("~/workspace/lits/attrib/attrib_year2013/code")
+#########################################
+cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
+#########################################
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+#########################################
+#########################################
+#########################################
+# valitaan vaan kohdemaat
+
+t3 <- subset(macro, group_fine != 'West Europe')
+west <- subset(macro, group_fine == 'West Europe')
+#t3 <- macro
+#########################################
+t3$group1 <- "group1"
+func <- function(t3)
+{
+  return(data.frame(COR = cor(t3$social.blame, t3$gdpchange07.10)))
+}
+df.cor <- ddply(t3, .(group1), func)
+t4 <- join(t3,df.cor,by="group1")
+cnames <- round(stats:::aggregate.formula(cbind(social.blame, gdpchange07.10) ~ COR, data=t4, mean),2)
+plot1 <- ggplot(t4, aes(x=gdpchange07.10, y=social.blame, 
+                        label=cntry)) +
+  geom_smooth(method=lm, se=FALSE, color="Dim Grey", linetype="dashed", size=0.5) +  
+  geom_text(data=cnames, aes(gdpchange07.10,social.blame, label = COR, group=COR), 
+            size=2.5, color="black") +
+  geom_text(size=2, vjust=-0.8, hjust=0.5) +
+  geom_point(size=1.5) +
+  geom_point(data=t4, aes(x=gdpchange07.10, 
+                          y=social.blame, color=group_general)) +
+  # west reference
+  geom_point(data=west, aes(x=gdpchange07.10, 
+                          y=social.blame, color=group_general)) +
+  geom_text(data=west, aes(x=gdpchange07.10, 
+                           y=social.blame, label=cntry),
+            color="Dim Grey", size=2, vjust=-0.8, hjust=0.5) +
+  # west reference
+  labs(x = "Change in GDP between 2007 and 2010 (%)", 
[…]

### 2013-12-13 code/iv_indiv_table.R

@@ -1,0 +1,76 @@
+
+# numeric table
+
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+library(survey)
+d.df <- svydesign(id = ~SerialID, 
+                  weights = ~weight, 
+                  data = df)
+# Country group
+t <- data.frame(prop.table(svytable(~group_general+poverty, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, group_general ~ poverty, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+cgroup <- t2
+# PAst difference
+t <- data.frame(prop.table(svytable(~past.diff+poverty, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, past.diff ~ poverty, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+past.diff <- t2
+# Future difference
+t <- data.frame(prop.table(svytable(~future.diff+poverty, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, future.diff ~ poverty, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+future.diff <- t2
+# Low education
+t <- data.frame(prop.table(svytable(~edu2+poverty, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, edu2 ~ poverty, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+edu2 <- t2
+# Crisis
+t <- data.frame(prop.table(svytable(~crise+poverty, d.df), 1)*100)
+t$Freq <- round(t$Freq, 1)
+t <- subset(t, Freq > 0)
+library(reshape2)
+t2 <-  dcast(t, crise ~ poverty, value.var="Freq")
+names(t2) <- c("value","Social Blame","Individual Blame","Individual Fate","Social Fate")
+crise <- t2
+# Transfer dependency
[…]

### 2013-12-13 code/iv_macro_table.R

@@ -1,0 +1,7 @@
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+tbl <- macro[, c(1,8:14)]
+tbl[,c(2:7)] <- round(tbl[,c(2:7)], 2)
+tbl <- tbl[order(tbl$undp_hdi), ]
+names(tbl) <- c("country","GDP Change 2007-1010","Failed States Index",
+                "Voice And Accountability (WB)","HDI","Democracy (Freedom House)",
+                "Gini UNU Wider","Country group")

### 2013-12-13 code/multi_logit.R

@@ -1,0 +1,54 @@
+setwd("~/workspace/lits/attrib/attrib_year2013/code")
+
+#########################################
+load("~/workspace/lits/attrib/attrib_year2013/data/lits.RData")
+# malleista länkkärit pois!
+df.multi <- subset(df, group_general %in% c("CEE","CIS"))
+
+library(lme4)
+summary(soc.0 <- glmer(pov.log.social ~ (1|cntry),
+                      family = binomial(logit)(logit),
+                       data=df.multi))
+summary(soc.1 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                         past.diff+future.diff+crise+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.2 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.3 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+
+
+summary(ind.0 <- glmer(pov.log.individual ~ (1|cntry),
+                      family = binomial(logit),
+                       data=df.multi))
+summary(ind.1 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+(1|cntry),
+                      family = binomial(logit),
+                       data=df.multi))
+summary(ind.2 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
+summary(ind.3 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
+                        past.diff+future.diff+crise+
+                        uw_gini+gdpchange07.10+(1|cntry),
+                       family = binomial(logit),
+                       data=df.multi))
[…]

### 2013-12-13 code/plot_map.R

@@ -1,0 +1,119 @@
+library(gpclib)
+library(RColorBrewer)
+library(mapproj)
+library(maptools)
+library(rgdal)
+map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/lits/shape/countries_shp", 
+               layer="lits_all"))
+library(car)
+map$NAME <- recode(map$NAME, "'United Kingdom'='Great Britain'")
+map$group_general <- recode(map$NAME,
+                       "c('France','Germany','Great Britain',
+                             'Italy','Sweden')='Western Europe';
+                             c('Czech Republic','Estonia','Hungary','Bulgaria',
+                             'Latvia','Lithuania','Poland',
+                             'Slovakia','Slovenia','Romania')='CEE';
+                             c('Armenia','Azerbaijan','Belarus','Georgia','Kazakhstan',
+                             'Kyrgyzstan','Moldova','Tajikistan',
+                             'Ukraine','Uzbekistan','Russia')='CIS';else=NA")
+
+map$group_fine <- recode(map$NAME,
+                    "c('Armenia','Azerbaijan','Georgia') = 'Caucasus';
+                         c('Russia','Ukraine','Moldova','Belarus')= 'CIS North';
+                         c('Kyrgyzstan','Kazakhstan','Tajikistan','Uzbekistan')='Central Asia';
+                         c('Estonia','Latvia','Lithuania')='Baltic States';
+                         c('Poland','Czech Republic','Slovakia','Hungary','Slovenia',
+                         'Romania','Bulgaria')='CEE EU';
+                         c('Italy','Germany','Great Britain','Sweden','France')='West Europe';
+                         else='other'")
+map$group_fine <- factor(map$group_fine, levels=c('West Europe','CEE EU',
+                                                  'Baltic States','CIS North',
+                                                  'Caucasus','Central Asia',
+                                                  'other'))
+
+# shapefile into data.frame
+gpclibPermit() 
+map@data$id <- rownames(map@data)
+library(ggplot2)
+map.points <- fortify(map, region="id")
+map.df <- merge(map.points, map@data, by="id")
+# only east of longitude -25
[…]

### 2013-12-13 article2013.Rmd
Makromuuttujat lukittu gdoChange,hdi,voice&Account
Makromuuttujat nyt lukittu
1) bkt-muutos 2007-2010
2) HDI
3) Voice and accountability (WB)

perusteet:
1) makrotalouden muutos kriisin aikana
2) elinolojen indikaattori
3) demokratiaindikaattori

Lisäsin ko, muuttujat paperin hajontakuvioihin ja lisäsin ekaan mallitaulukkoon molempiin malleihin sarakkeen jossa vaan makrot.

@@
 xelatex article2013.tex
+
+knitpandoc_pdf("article2013")
 
@@
 
+![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
+
 -->
 
-![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
+![Human Development Index and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_hdi.pdf)
+
+AND SOMETHING ABOUT THE NEW FIGURE ON HUMAN DEVELPMENT INDEX AND LIVING CONDITIONS OF THE PEOPLE AND SKIPPING THE NEXT PARAGRAPH ON INEQUALITY
 
@@
                       "Income will worsen","Has affected great or fair amount",
-                      "Gini coefficient",
+                      "Human Deveplopment Index",
                       "Change in GDP between 2007 to 2010",
-                      "Good governance")
+                      "Voice and Accountability")
 
@@
                       "Only individual",
-                      "Gini added",
+                      "HDI added",
                       "GDP change added",
-                      "Good governance added")
+                      "Voice and Accountability added")
 
@@
                       "SB individual",
+                      "SB contextual",
                       "SB all",
@@
                       "IB individual",
+                      "IB contextual",
                       "IB all")
 
-texreg(list(soc.0,soc.1,soc.4,
-            ind.0,ind.1,ind.4),
+texreg(list(soc.0,soc.1,soc.Cont,soc.4,
+            ind.0,ind.1,ind.Cont,ind.4),
         custom.model.names = model.names.multi.A,
@@
         use.packages=TRUE,
-        #sideways=TRUE,
+        sideways=TRUE,
         scriptsize=TRUE,
@@
 coef.names.multi = c("(intercept)",
-                     "Dependent","Low education",
-                     "Low income","Income has worsened",
-                     "Income will worsen","Has affected great or fair amount",
-                     "Gini coefficient",
+                     "Dependent",
+                     "Low education",
+                     "Low income",
+                     "Income has worsened",
+                     "Income will worsen",
+                     "Has affected great or fair amount",
+                     "Human Developmet Index",
                      "Change in GDP between 2007 to 2010",
-                     "Good governance")
+                     "Voice and Accountability")
 

### 2013-12-13 finalPlots.Rmd
Makromuuttujat lukittu gdoChange,hdi,voice&Account
Makromuuttujat nyt lukittu
1) bkt-muutos 2007-2010
2) HDI
3) Voice and accountability (WB)

perusteet:
1) makrotalouden muutos kriisin aikana
2) elinolojen indikaattori
3) demokratiaindikaattori

Lisäsin ko, muuttujat paperin hajontakuvioihin ja lisäsin ekaan mallitaulukkoon molempiin malleihin sarakkeen jossa vaan makrot.

@@
 
-## UNU-Wider Gini
+## UNDP HDI
 
-```{rattribFinalPlotsGini, eval=FALSE}
+```{rattribFinalPlotsHDI, eval=FALSE}
 # UNU-Wider Gini
@@
 {
-  return(data.frame(COR = cor(t3$socialBlame, t3$uw_gini)))
+  return(data.frame(COR = cor(t3$socialBlame, t3$undp_hdi)))
 }
@@
 t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(socialBlame, uw_gini) ~ COR, data=t4, mean),2)
-plot1 <- ggplot(t4, aes(x=uw_gini, y=socialBlame, label=cntry)) +
+cnames <- round(stats:::aggregate.formula(cbind(socialBlame, undp_hdi) ~ COR, data=t4, mean),2)
+plot1 <- ggplot(t4, aes(x=undp_hdi, y=socialBlame, label=cntry)) +
   geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(uw_gini,socialBlame, label = COR, group=COR), 
+  geom_text(data=cnames, aes(undp_hdi,socialBlame, label = COR, group=COR), 
             size=2.5, color="black") +
-  labs(x = "Gini coefficient", 
+  labs(x = "Human Development Index", 
        y = "Support for social blame type of explanation (%)") +
   geom_point(size=1.5) +
-  geom_point(data=t4, aes(x=uw_gini, 
+  geom_point(data=t4, aes(x=undp_hdi, 
                           y=socialBlame, color=group_general)) +
   # west reference
-  geom_point(data=west, aes(x=uw_gini, 
+  geom_point(data=west, aes(x=undp_hdi, 
                             y=socialBlame, 
                             color=group_general), alpha=.5) +
-  geom_text(data=west, aes(x=uw_gini, 
+  geom_text(data=west, aes(x=undp_hdi, 
                            y=socialBlame, label=cntry),
@@
   scale_colour_manual(values=cbPalette) +
-  coord_cartesian(xlim=c(20,60)) +
+  #coord_cartesian(xlim=c(20,60)) +
   theme_minimal() +
@@
 {
-  return(data.frame(COR = cor(t3$individualBlame, t3$uw_gini)))
+  return(data.frame(COR = cor(t3$individualBlame, t3$undp_hdi)))
 }
@@
 t4 <- join(t3,df.cor,by="group1")
-cnames <- round(stats:::aggregate.formula(cbind(individualBlame, uw_gini) ~ COR, data=t4, mean),2)
-plot2 <- ggplot(t4, aes(x=uw_gini, y=individualBlame, label=cntry)) +
+cnames <- round(stats:::aggregate.formula(cbind(individualBlame, undp_hdi) ~ COR, data=t4, mean),2)
+plot2 <- ggplot(t4, aes(x=undp_hdi, y=individualBlame, label=cntry)) +
   geom_smooth(method=lm, se=FALSE,  alpha=.5, linetype="dashed", size=0.5) +  
   geom_text(size=2, vjust=-0.8, hjust=0.5) +
-  geom_text(data=cnames, aes(uw_gini,individualBlame, label = COR, group=COR), 
+  geom_text(data=cnames, aes(undp_hdi,individualBlame, label = COR, group=COR), 
             size=2.5, color="black") +
-  labs(x = "Gini coefficient", 
+  labs(x = "Human Development Index", 
        y = "Support for individual blame type of explanation (%)") +
   geom_point(size=1.5) +  
-  geom_point(data=t4, aes(x=uw_gini, 
+  geom_point(data=t4, aes(x=undp_hdi, 
                           y=individualBlame, color=group_general)) +
   # west reference
-  geom_point(data=west, aes(x=uw_gini, 
+  geom_point(data=west, aes(x=undp_hdi, 
                             y=individualBlame, 
                             color=group_general), alpha=.5) +
-  geom_text(data=west, aes(x=uw_gini, 
+  geom_text(data=west, aes(x=undp_hdi, 
                            y=individualBlame, label=cntry),
@@
 
-pdf("figure/scatter_gini.pdf", width=12/2.54, height=18/2.54)
+pdf("figure/scatter_hdi.pdf", width=12/2.54, height=18/2.54)
 grid.arrange(plot1,plot2, ncol=1)
@@
 ppi <- 300
-png("figure/scatter_gini.png", width=12/2.54*ppi, height=18/2.54*ppi, res=ppi)
+png("figure/scatter_hdi.png", width=12/2.54*ppi, height=18/2.54*ppi, res=ppi)
 grid.arrange(plot1,plot2, ncol=1)
@@
 
-png("figure/scatter_giniSB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+png("figure/scatter_hdiSB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
 plot1
 dev.off()
-png("figure/scatter_giniIB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
+png("figure/scatter_hdiIB.png", width=12/2.54*ppi, height=8/2.54*ppi, res=ppi)
 plot2
@@
 
-![](figure/scatter_gini.png)
+![](figure/scatter_hdi.png)
 

### 2013-12-13 loadClean.Rmd
Koodin korjailua, korrelaatiokokeet ja maaryhmäkappaleen poisto
Korjasin koodia niin että hajontakuviot ovat oikein paperissa.

Katsoin läpi shinylla tehdyn korrelaatiokoneen ja kokosin selkeimmät yhteydet dokkariin correlations.

Otin tekstistä pois maaryhmiä koskevan teorialuvun.
Makromuuttujat lukittu gdoChange,hdi,voice&Account
Makromuuttujat nyt lukittu
1) bkt-muutos 2007-2010
2) HDI
3) Voice and accountability (WB)

perusteet:
1) makrotalouden muutos kriisin aikana
2) elinolojen indikaattori
3) demokratiaindikaattori

Lisäsin ko, muuttujat paperin hajontakuvioihin ja lisäsin ekaan mallit

@@
 # Quality of governance
-library(rustfare)
-datQog <- GetQog(country = c('Armenia','Azerbaijan','Georgia','Russia',
+
+getQog <- function(data = "basic",
+                   country = "all", 
+                   indicator = "undp_hdi",
+                   year = "all") {
+# URL's for different data's
+  basicURL <- "http://www.qogdata.pol.gu.se/data/qog_bas_ts_30aug13.csv"
+  standardURL <- "http://www.qogdata.pol.gu.se/data/qog_std_ts_15may13.csv"
+  socialPolicyURL <- "http://www.qogdata.pol.gu.se/data/qog_soc_tsl_4apr12.csv"
+
+# Select the data to download
+  if (data == "basic")         dat <- read.csv(basicURL,  sep = ";")
+  if (data == "standard")      dat <- read.csv(standardURL,  sep = ";")
+  if (data == "socialPolicy")  dat <- read.csv(socialPolicyURL,  sep = ";")
+# subset the countries
+  if (country != "all") df <- dat[dat$cname %in% c(country), ]
+  if (country == "all") df <- dat
+# subset the variables
+  indicatorx <- c("cname","year",indicator)
+  df <- df[, c(indicatorx)]
+# reshape the data into long format  
+  library(reshape2)
+  df.l <- melt(df, id.vars=c("cname","year"))
+  df.l <- df.l[!is.na(df.l$value), ]
+  df.l$indicator <- df.l$variable
+  df.l$variable <- NULL
+# subset the years
+  if (year != "all") df.l <- df.l[df.l$year == year,]
+  if (year == "all") df.l <- df.l
+  df.l
+}
+
+datQog <- getQog(data="basic",
+                 country = c('Armenia','Azerbaijan','Georgia','Russia',
                    'Ukraine','Moldova','Belarus','Kyrgyzstan',
@@
        indicator = c("wbgi_vae", # World Bank - Voice and Accountability
-                     "uw_gini"
-                     ))
+                     "undp_hdi"))
+
+datQogSum <- datQog[datQog$year %in% c(2006,2010),]
 
@@
 #datQogSum <- ddply(datQog, c("cname","indicator"), 
-#                   subset, year == max(year))
-datQogSum <- datQog[datQog$year %in% c(2006,2009),]
-datQogSum$year[datQogSum$year == 2009] <- 2010
+#                   subset, year == 2010)
+#datQogSum <- datQog[datQog$year %in% c(2006,2009),]
+#datQog$year[datQog$year == 2009] <- 2010
 #names(datQogSum)[3] <- "wbgi_vae"
@@
 
-datQogSum.w <- datQogSum
+#datQogSum.w <- datQogSum
 

### 2013-12-13 regressionModelling.Rmd
Some Sunday evening edits..
Makromuuttujat lukittu gdoChange,hdi,voice&Account
Makromuuttujat nyt lukittu
1) bkt-muutos 2007-2010
2) HDI
3) Voice and accountability (WB)

perusteet:
1) makrotalouden muutos kriisin aikana
2) elinolojen indikaattori
3) demokratiaindikaattori

Lisäsin ko, muuttujat paperin hajontakuvioihin ja lisäsin ekaan mallitaulukkoon molempiin malleihin sarakkeen jossa vaan makrot.

@@
 
-```{r, eval=FALSE, echo=FALSE}
-library(knitr)
-library(markdown)
-knit2html('regressionModelling.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-
+```{roptsattrib, echo=FALSE}
+opts_chunk$set(echo=TRUE,eval=FALSE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
 ```
 
-```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
-```
-
-```{rmultilevelLoadData}
+```{rmultilevelLoadData, eval=TRUE}
 #########################################
@@
 
-```{rmultilevelBivarSoc}
+```{rmultilevelBivarSoc, eval=TRUE}
 library(lme4)
@@
                          family = binomial(logit), data = df.model)
-# bisoc.8 <- glmer(pov.log.social ~ perGini+(1|cntry),
-#                         family = binomial(logit), data = df.model)
+bisoc.8 <- glmer(pov.log.social ~ undp_hdi+(1|cntry),
+                         family = binomial(logit), data = df.model)
 bisoc.9 <- glmer(pov.log.social ~ gdpChange+(1|cntry),
@@
 
-```{rmultilevelBivarSocOutputNOCI, results='asis'}
+```{rmultilevelBivarSocOutputNOCI, results='asis', eval=TRUE}
 
@@
                       "Income will worsen","Has affected great or fair amount",
-                      #"Gini coefficient",
+                      "Human Development Index",
                       "Change in GDP between 2007 to 2010",
@@
           bisoc.5,bisoc.6,
-          #bisoc.8,
+          bisoc.8,
           bisoc.9,bisoc.10,
@@
 
-```{rmultilevelBivarInd}
+```{rmultilevelBivarInd, eval=TRUE}
 biind.1 <- glmer(pov.log.individual ~ incsour3+(1|cntry),
@@
                         family = binomial(logit), data = df.model)
-#biind.8 <- glmer(pov.log.individual ~ perGini+(1|cntry),
-#                        family = binomial(logit), data = df.model)
+biind.8 <- glmer(pov.log.individual ~ undp_hdi+(1|cntry),
+                        family = binomial(logit), data = df.model)
 biind.9 <- glmer(pov.log.individual ~ gdpChange+(1|cntry),
@@
                           family = binomial(logit), data = df.model)
-
 ```
 
-```{rmultilevelBivarIndOuputNOCI, results='asis'}
-
+```{rmultilevelBivarIndOuputNOCI, results='asis', eval=TRUE}
 library(texreg)
@@
           biind.5,biind.6,
-          #biind.8,
+          biind.8,
           biind.9,biind.10,
@@
 
-```{rmultilevelMultiSoc}
+```{rmultilevelMultiSoc, eval=TRUE}
 library(lme4)
-
 soc.0 <- glmer(pov.log.social ~ (1|cntry),
@@
                         past.diff+future.diff+crise+
-                        perGini+(1|cntry),
+                        undp_hdi+(1|cntry),
                        family = binomial(logit),
@@
                         past.diff+future.diff+crise+
-                        perGini+gdpChange+(1|cntry),
+                        undp_hdi+gdpChange+(1|cntry),
                        family = binomial(logit),
@@
                         past.diff+future.diff+crise+
-                        perGini+gdpChange+wbgi_vae+(1|cntry),
+                        undp_hdi+gdpChange+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+soc.Cont <- glmer(pov.log.social ~ undp_hdi+gdpChange+
+                      wbgi_vae+(1|cntry),
                        family = binomial(logit),
@@
 
-```{rmultilevelMultiSocOutputNOCI, results='asis'}
+```{rmultilevelMultiSocOutputNOCI, results='asis', eval=TRUE}
 coef.names.multi = c("(intercept)", 
@@
                       "Income will worsen","Has affected great or fair amount",
-                      "Perceived Gini",
+                      "Human Deveplopment Index",
                       "Change in GDP between 2007 to 2010",
@@
                       "Only individual",
-                      "Gini added",
+                      "HDI added",
                       "GDP change added",
@@
 
-```{rmultilevelMultiInd}
+```{rmultilevelMultiInd, eval=TRUE}
 ind.0 <- glmer(pov.log.individual ~ (1|cntry),
@@
                         past.diff+future.diff+crise+
-                        perGini+gdpChange+wbgi_vae+(1|cntry),
+                        undp_hdi+gdpChange+wbgi_vae+(1|cntry),
+                       family = binomial(logit),
+                       data = df.model)
+ind.Cont <- glmer(pov.log.individual ~ undp_hdi+gdpChange+
+                      wbgi_vae+(1|cntry),
                        family = binomial(logit),
@@
 
-```{rmultilevelMultiIndOutputNOCI, results='asis'}
+```{rmultilevelMultiIndOutputNOCI, results='asis', eval=TRUE}
 htmlreg(list(ind.0,
@@
 
-```{rmultilevelMultivarFull}
+```{rmultilevelMultivarFull, eval=TRUE}
 
@@
                         past.diff+future.diff+crise+
-                        perGini+gdpChange+wbgi_vae+(1|cntry),
+                        undp_hdi+gdpChange+wbgi_vae+(1|cntry),
                        family = binomial(logit),
@@
 
-save(ind.0,ind.1,ind.2,ind.3,ind.4,
-     soc.0,soc.1,soc.2,soc.3,soc.4,
+save(ind.0,ind.1,ind.2,ind.3,ind.4,ind.Cont,
+     soc.0,soc.1,soc.2,soc.3,soc.4,soc.Cont,
      socVsInc, file="data/finalModels.RData")
@@
 
-```{rmultilevelMultivarSlideOutput, results='asis', eval=TRUE}
+```{rmultilevelMultivarSlideOutput, results='asis', eval=TRUE, eval=TRUE}
 library(texreg)
@@
                       "Income will worsen","Has affected great or fair amount",
-                      "Gini coefficient",
+                      "Human Development Index",
                       "Change in GDP between 2007 to 2010",

### 2013-12-16 article2013.Rmd
Puristin kaikki mallit yhteen taulukkoon

@@
                       "IB contextual",
-                      "IB all")
+                      "IB all",
+                      "SB vs. IB")
 
 texreg(list(soc.0,soc.1,soc.Cont,soc.4,
-            ind.0,ind.1,ind.Cont,ind.4),
+            ind.0,ind.1,ind.Cont,ind.4,socVsInc),
         custom.model.names = model.names.multi.A,
@@
 
-```{rarticleModelsFull, results='asis', eval=TRUE}
+```{rarticleModelsFull, results='asis', eval=FALSE}
 load("data/finalModels.RData")

### 2013-12-17 article2013.Rmd
Mallit ajettu luottamusväleillä, kesti 22 tuntia.
Pdf päivittynyt, mutta luottamusvälien merkinnästä johtuen niin että SB vs. IB malli ei mahtunut sivulle. No, ne löytyy sitten versionhallinnasta jos pitää verrata. Teen iltapäivällä tulkintaa..

--- 2013-12-13 article2013.Rmd
@@
 
-Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe.
-Effects of Contextual and Individual Characteristics across 28 European Countries.
-Acta Sociologica 53 (1), 53–72.
-
-Morçöl, Göktuğ (1997) Lay Explanations for Poverty in Turkey and Their Determinants. Journal of Social Psychology 137 (6) 728–738.
-
-Murthi, M., & Tiongson, E. R. (2008). Attitudes to Equality: The Socialist Legacy Revisited. SSRN eLibrary. Retrieved from http://papers.ssrn.com/sol3/papers.cfm?abstract_id=1098627
-
-Nasser, Ramzi and Abouchedid, Kamal and Khashan, Hilal (2002) Perceptions of the Causes of Poverty Comparing Three National Groups: Lebanon, Portugal, and South Africa.
-Current Research in Social Psychology 8 (7), 101–119.
-
-Niemelä, Mikko (2008) Perceptions of the Causes of Poverty in Finland. Acta Sociologica, 51 (1), 23–40.
-
-Nilson, Lina B. (1981) Reconsidering Ideological Lines: Beliefs about Poverty in America. The Sociological Quarterly 22 (4), 531-548.
-
-Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of California Press.
-
-Rothstein, Bo (2000) Trust, Social Dilemmas and Collective Memories. Journal of Theoretical Politics, 12 (4), 477–501.
-
-Rothstein, Bo and Stolle, Dietlind (2008) The State and Social Capital: An Institutional Theory of Generalized Trust. Comparative Politics 40 (4), 441-459.
-
-Saunders, Peter (2002) The Ends and Means of Welfare. Coping with Economic and Social Change in Australia. Cambridge: Cambridge University Press.
-
-Schumpeter, Joseph A. (1942) Capitalism, Socialism & Democracy. New York: Harper & Brothers.
-
-Szikra, D., & Tomka, B. (2009). Social Policy in East Central Europe: Major Trends in the Twentieht Century. In Post-communist Welfare Pathways (Vol. 2009).
-
-Stephenson, Svetlana (2000) Public Beliefs in the Causes of Wealth and Poverty and Legitimization of Inequalities in Russia and Estonia. Social Justice Research 13 (2), 83–100.
-
-Svallfors, Stefan (1997) Worlds of Welfare and Attitudes to Redistribution: A Comparison of Eight Western Nations. European Sociological Review 13 (3), 283–304.
-
-Tomka, B. (2006). East Central Europe and the European Social Policy Model: A Long-Term View. East European Quarterly, 40(2), 135.
-
-van Oorschot, Wim and Halman, Loek (2000) Blame or Fate, Individual or Social? An International Comparison of Popular Explanations of Poverty. European Societies 2 (1), 1–28.
-
-Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterly 37 (3), 413–428.
+Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe. E

### 2013-12-17 regressionModelling.Rmd
Mallit ajettu luottamusväleillä, kesti 22 tuntia.
Pdf päivittynyt, mutta luottamusvälien merkinnästä johtuen niin että SB vs. IB malli ei mahtunut sivulle. No, ne löytyy sitten versionhallinnasta jos pitää verrata. Teen iltapäivällä tulkintaa..

@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=FALSE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
+opts_chunk$set(echo=FALSE,eval=FALSE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
 ```
@@
 
-```{rmultilevelBivarSoc, eval=TRUE}
+```{rmultilevelBivarSoc}
 library(lme4)
@@
 
-```{rmultilevelBivarSocOutputNOCI, results='asis', eval=TRUE}
+```{rmultilevelBivarSocOutputNOCI, results='asis'}
 
@@
 
-```{rmultilevelBivarInd, eval=TRUE}
+```{rmultilevelBivarInd}
 biind.1 <- glmer(pov.log.individual ~ incsour3+(1|cntry),
@@
 
-```{rmultilevelBivarIndOuputNOCI, results='asis', eval=TRUE}
+```{rmultilevelBivarIndOuputNOCI, results='asis'}
 library(texreg)
@@
 
-```{rmultilevelMultiSoc, eval=TRUE}
+```{rmultilevelMultiSoc, eval=TRUE, echo=TRUE}
 library(lme4)
@@
 
-```{rmultilevelMultiSocOutputNOCI, results='asis', eval=TRUE}
+```{rmultilevelMultiSocOutputNOCI, results='asis', eval=FALSE}
 coef.names.multi = c("(intercept)", 
@@
 
-```{rmultilevelMultiInd, eval=TRUE}
+```{rmultilevelMultiInd, eval=TRUE, echo=TRUE}
 ind.0 <- glmer(pov.log.individual ~ (1|cntry),
@@
 
-```{rmultilevelMultiIndOutputNOCI, results='asis', eval=TRUE}
+```{rmultilevelMultiIndOutputNOCI, results='asis', eval=FALSE}
 htmlreg(list(ind.0,
@@
 
-```{rmultilevelMultivarFull, eval=TRUE}
-
-soc.4 <- glmer(pov.log.social ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        undp_hdi+gdpChange+wbgi_vae+(1|cntry),
-                       family = binomial(logit),
-                       data = df.model)
-
-ind.4 <- glmer(pov.log.individual ~ incsour3+edu2+income2+
-                        past.diff+future.diff+crise+
-                        undp_hdi+gdpChange+wbgi_vae+(1|cntry),
-                       family = binomial(logit),
-                       data = df.model)
+```{rmultilevelMultivarFull, eval=TRUE, echo=TRUE}
 
@@
 
-# Tables for presentation
+# FINAL TABLES SAME AS IN PAPER
 
-## Partial
+```{rmultilevelMultivarSlideOutput, results='asis', eval=TRUE, echo=TRUE}
+library(texreg)
 
-```{rmultilevelMultivarSlideOutput, results='asis', eval=TRUE, eval=TRUE}
-library(texreg)
+#load("data/finalModels.RData")
+
 coef.names.multi = c("(intercept)", 
@@
                       "Income will worsen","Has affected great or fair amount",
-                      "Human Development Index",
+                      "Human Deveplopment Index",
                       "Change in GDP between 2007 to 2010",
-                      "Good governance")
+                      "Voice and Accountability")
 
@@
                       "GDP change added",
-                      "Good governance added")
+                      "Voice and Accountability added")
 
@@
                       "IB individual",
-                        "IB all")
+                      "IB contextual",
+                      "IB all",
+                      "SB vs. IB")
 
-htmlreg(list(soc.0,soc.1,soc.4,
-            ind.0,ind.1,ind.4),
+htmlreg(list(soc.0,soc.1,soc.Cont,soc.4,
+            ind.0,ind.1,ind.Cont,ind.4,socVsInc),
         custom.model.names = model.names.multi.A,
         custom.coef.names = coef.names.multi,
-        naive = TRUE,
+        naive=TRUE,
         inline.css = TRUE,
@@
 
+```
+
 ## Comparison of full models
 
-```{rrmultilevelMultivarFullOutput, results='asis', eval=TRUE}
+```{rrmultilevelMultivarFullOutput, results='asis', eval=FALSE}
 coef.names.multi = c("(intercept)", 

### 2014-01-04 mapPlots.Rmd
First pages commit
pientä testailua vaan

@@
 
-<link href="http://markuskainu.fi/material/css/rmarkdown.css" rel="stylesheet" type="text/css" title="compact"></link>
-
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attributions/")
-library(knitr)
-knit2html('mapPlots.Rmd', options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-
-```
-
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=15,fig.height=11)
+opts_chunk$set(echo=FALSE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=15,fig.height=11)
 ```
@@
   theme(legend.position="top")
+
 ```
@@
   theme(legend.text = element_text(size=14))
+
 ```

### 2014-02-11 article2013.Rmd
Siirryin Rmarkdowniin

@@
-% Attributions for Poverty in Post-Socialist Countries[^*]
-% Markus Kainu[^**];Mikko Niemelä[^***]
-% December 12, 2013
-
-[^*]:*Work in progress. See the analysis outputs at [http://muuankarski.github.io/attributions/](http://muuankarski.github.io/attributions/)
-
-[^**]:PhD Student. Aleksanteri institute, University of Helsinki & Department of Social Research, University of Turku <a href="mailto:markuskainu@gmail.com">markuskainu@gmail.com</a>
-
-[^***]:Senior Researcher. Social Insurance Institution of Finland, KELA a<a href="mailto:mikko.niemela@kela.fi">mikko.niemela@kela.fi</a>
-
-```{roptsattribArticle, echo=FALSE}
-opts_chunk$set(echo=FALSE,eval=TRUE,cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
-```
-
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attributions/")
-library(knitr)
-knit("article2013.Rmd")
-
-~/.cabal/bin/pandoc article2013.md --template=../../template/pandoc_article/tex_templates/pandoc.latex.template --number-sections --toc --latex-engine=xelatex -V lang=english -V papersize:a4paper -V documentclass=scrartcl -o article2013.tex
-sed -i 's/begin{figure}\[htbp\]/begin{figure}\[H\]/g' article2013.tex
-xelatex article2013.tex
-
-knitpandoc_pdf("article2013")
-
-system("~/.cabal/bin/pandoc -s article2013.md --toc --number-sections -o article2013.html")
-system("~/.cabal/bin/pandoc -f markdown -t docx article2013.md -o article2013.docx --standalone --toc --number-sections")
-system("~/.cabal/bin/pandoc article2013.md -o article2013.epub --standalone --toc --number-sections")
-
-```
-
-\newpage
-
-\begin{abstract}
-Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
+---
+title: "Attributions for Poverty in Post-Socialist Countries"
+author: [Markus Kainu & Mikko Niemelä]
+date: February 11, 2014
+abstract: Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
 The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions.
@@
 As individual-level determinants, the paper tests to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.
-\end{abstract}
+keywords: poverty
+output:
+  pdf_document:
+    toc: true
+    highlight: zenburn
+  html_document:
+    toc: true
+    theme: flatly
+---
+
+\newpage
 
@@
 
+<!--
 ![Countries used in the analysis](figure/mapplot.pdf)
+-->
 
@@
 
---------------------------------------------------------------------------------------------
+----------------------------
 Variable                          Description                                                                                                   Source
-------------------------          ------------------------                                                                                      -----------
+---------                         ------------                                                                                                  -----------
 **Individual-level**
@@
 Voice and Accountability          Indicators measuring the political process, civil liberties and political rights.                             World Bank
---------------------------------------------------------------------------------------------
+----------------------------
 
@@
 
-```{rattribArticleDescTbl1, results='asis'}
+```{rattribArticleDescTbl1, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
 source("code/dv_tables.R")
@@
 print.xtable(xtable(tbl5,digits=1,
-                    caption="Weighted population shares of dependent variable by country",),table.placement="H",
+                    caption="Weighted population shares of dependent variable by country",), #table.placement="H",
              type="latex",
@@
              size="scriptsize")
-
 ```
@@
 
-```{rarticleModelsSoc, results='asis', eval=TRUE}
+```{rarticleModelsSoc, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
 load("data/finalModels.RData")
-
+library(texreg)
 coef.names.multi = c("(intercept)", 
@@
                       "SB contextual",
-                      "SB all",
-                      "IB empty",
+                      "SB all")
+
+texreg(list(soc.0,soc.1,soc.Cont,soc.4),
+        custom.model.names = model.names.multi.A,
+        custom.coef.names = coef.names.multi,
+        naive = TRUE,
+        #use.packages=TRUE,
+        #sideways=TRUE,
+        scriptsize=TRUE,
+        include.pvalues=TRUE,
+        caption="Logistic multilevel random intercept model for social blame type of explanation", 
+        caption.above=TRUE)
+```
+
+```{rarticleModelsInd, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
+load("data/finalModels.RData")
+library(texreg)
+coef.names.multi = c("(intercept)", 
+                      "Dependent","Low education",
+                      "Low income","Income has worsened",
+                      "Income will worsen","Has affected great or fair amount",
+                      "Human Deveplopment Index",
+                      "Change in GDP between 2007 to 2010",
+                      "Voice and Accountability")
+
+model.names.multi = c("Empty",
+                      "Only individual",
+                      "HDI added",
+                      "GDP change added",
+                      "Voice and Accountability added")
+
+model.names.multi.A = c("IB empty",
                       "IB individual",
@@
 
-texreg(list(soc.0,soc.1,soc.Cont,soc.4,
-            ind.0,ind.1,ind.Cont,ind.4),
+texreg(list(ind.0,ind.1,ind.Cont,ind.4),
         custom.model.names = model.names.multi.A,
@@
         naive = TRUE,
-        use.packages=TRUE,
-        sideways=TRUE,
+        #use.packages=TRUE,
+        #sideways=TRUE,
         scriptsize=TRUE,
@@
         caption.above=TRUE)
-
 ```
 
-```{rarticleModelsFull, results='asis', eval=TRUE}
+```{rarticleModelsFull, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
 load("data/finalModels.RData")
@@
 library(texreg)
-texreg(list(soc.4,ind.4,socVsInc),float.pos="H",
-       use.packages=TRUE,
+texreg(list(soc.4,ind.4,socVsInc),
+       #float.pos="H",
+       #use.packages=TRUE,
        #booktabs=TRUE,

### 2014-02-11 article2013.Rmd

@@
              size="scriptsize")
+
 ```
@@
 
+-->
+
+![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
+
+>There is also a substantial variation in the level of income inequality among post-socialist countries.
+Figure 3 shows that the higher level of income inequality is associated with higher level of social blame.
+As for individual blame type of explanation the picture is more puzzled, showing only weak connection between larger inequality and lower support for individual blame explanation.
+Again, the Western European countries differ with post-socialist countries.
+Cross-national differences in income inequality are much smaller in Western Europe than among post-socialist countries.
+There is a weak negative association between the social blame and income  inequality and a strong positive association between the individual blame and income inequality.
+
+=======
+
 ![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
@@
 There is a weak negative association between the social blame and income  inequality and a strong positive association between the individual blame and income inequality.
+
+>>>>>>> gh-pages
 
@@
 
-Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe. E
+Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe.
+Effects of Contextual and Individual Characteristics across 28 European Countries.
+Acta Sociologica 53 (1), 53–72.
+
+Morçöl, Göktuğ (1997) Lay Explanations for Poverty in Turkey and Their Determinants. Journal of Social Psychology 137 (6) 728–738.
+
+Murthi, M., & Tiongson, E. R. (2008). Attitudes to Equality: The Socialist Legacy Revisited. SSRN eLibrary. Retrieved from http://papers.ssrn.com/sol3/papers.cfm?abstract_id=1098627
+
+Nasser, Ramzi and Abouchedid, Kamal and Khashan, Hilal (2002) Perceptions of the Causes of Poverty Comparing Three National Groups: Lebanon, Portugal, and South Africa.
+Current Research in Social Psychology 8 (7), 101–119.
+
+Niemelä, Mikko (2008) Perceptions of the Causes of Poverty in Finland. Acta Sociologica, 51 (1), 23–40.
+
+Nilson, Lina B. (1981) Reconsidering Ideological Lines: Beliefs about Poverty in America. The Sociological Quarterly 22 (4), 531-548.
+
+Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of California Press.
+
+Rothstein, Bo (2000) Trust, Social Dilemmas and Collective Memories. Journal of Theoretical Politics, 12 (4), 477–501.
+
+Rothstein, Bo and Stolle, Dietlind (2008) The State and Social Capital: An Institutional Theory of Generalized Trust. Comparative Politics 40 (4), 441-459.
+
+Saunders, Peter (2002) The Ends and Means of Welfare. Coping with Economic and Social Change in Australia. Cambridge: Cambridge University Press.
+
+Schumpeter, Joseph A. (1942) Capitalism, Socialism & Democracy. New York: Harper & Brothers.
+
+Szikra, D., & Tomka, B. (2009). Social Policy in East Central Europe: Major Trends in the Twentieht Century. In Post-communist Welfare Pathways (Vol. 2009).
+
+Stephenson, Svetlana (2000) Public Beliefs in the Causes of Wealth and Poverty and Legitimization of Inequalities in Russia and Estonia. Social Justice Research 13 (2), 83–100.
+
+Svallfors, Stefan (1997) Worlds of Welfare and Attitudes to Redistribution: A Comparison of Eight Western Nations. European Sociological Review 13 (3), 283–304.
+
+Tomka, B. (2006). East Central Europe and the European Social Policy Model: A Long-Term View. East European Quarterly, 40(2), 135.
+
+van Oorschot, Wim and Halman, Loek (2000) Blame or Fate, Individual or Social? An International Comparison of Popular Explanations of Poverty. European Societies 2 (1), 1–28.
+
+Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterly 37 (3), 413–428.

### 2014-02-11 descriptive_analysis.Rmd
Changed the toc font size in css
First pages commit
Getting back to business. Nothing changed, but rerunning the script.
Pientä päivtystä. Huomenna iso remmi päälle

@@
-# Descriptive visualizations
-
-```{r, eval=FALSE, echo=FALSE}
-setwd("~/workspace/lits/attributions")
-library(knitr)
-library(markdown)
-knit2html('descriptive_analysis.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-```
-
-```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
-```
+---
+title: Descriptive visualizations
+output:
+  html_document:
+    toc: true
+    theme: flatly
+---
 

### 2014-02-11 loadClean.Rmd
Pientä päivtystä. Huomenna iso remmi päälle

@@
+---
+title: Descriptive visualizations
+output:
+  html_document:
+    toc: true
+    theme: flatly
+---
+
 # Loading and cleaning the data
-
-```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
-```
 
@@
 # Quality of governance
-
 getQog <- function(data = "basic",

### 2014-02-11 article2013.Rmd
Pientä päivtystä. Huomenna iso remmi päälle

@@
 ---
-title: "Attributions for Poverty in Post-Socialist Countries"
+title: Attributions for Poverty in Post-Socialist Countries
 author: [Markus Kainu & Mikko Niemelä]
@@
     toc: true
+    number_section: true
     highlight: zenburn
@@
 
-# Introduction
+Introduction
+==============================================
 
@@
 
-# Prior findings on cross-national differences
+Prior findings on cross-national differences
+==============================================
 
@@
 
-# Risks factors as an individual-level determinants of perceptions
+Risks factors as an individual-level determinants of perceptions
+==============================================
 
@@
 
-# Research design
+Research design
+==============================================
 
@@
 
-<!--
 ![Countries used in the analysis](figure/mapplot.pdf)
--->
 
@@
 
-# Results
+Results
+==============================================
 
@@
 
-## Descriptive statistics
+Descriptive statistics
+----------------------------------------------
 
@@
 
-## Determinants of perceptions
+Determinants of perceptions
+----------------------------------------------
 
@@
 
-=======
-
 ![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
-
--->
 
@@
 
->>>>>>> gh-pages
-
 ![World Bank's Voice and Accountability index in 2011 and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_wbgi.pdf)
@@
 
-## Multilevel analysis
+Multilevel analysis
+-------------------------------------------------
 
-*As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
-Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.*
+As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
+Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
 
@@
 
-<!-- As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
-Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
--->
-
 Table 4 and 5 present figures for multilevel logistic analysis.
@@
 
-# Conclusions
+Conclusions
+==============================================
 
@@
 
-# References
+References
+==============================================
 

### 2014-02-11 compareWaves.Rmd
Some Sunday evening edits..
Pientä päivtystä. Huomenna iso remmi päälle

@@
-# Comparing waves 2006 and 2010
-
-```{r, eval=FALSE, echo=FALSE}
-library(knitr)
-library(markdown)
-knit2html('compareWaves.Rmd', 
-          stylesheet="~/workspace/web/css/rmarkdown.css",
-          options = c('toc', markdown::markdownHTMLOptions(TRUE)))
-```
-
-```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
-```
+---
+title: Comparing waves 2006 and 2010 
+output:
+  html_document:
+    toc: true
+    theme: flatly
+---
 

### 2014-02-11 mapPlots.Rmd
Pientä päivtystä. Huomenna iso remmi päälle

@@
-# Spatial visualization
+---
+title: Spatial visualization
+output:
+  html_document:
+    toc: true
+    theme: flatly
+---
 
@@
   geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
-  geom_polygon(aes(fill = not.stated)) +
+  geom_polygon(aes(fill = notStated)) +
   geom_polygon(data = plotMap, aes(long,lat), 
@@
   geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
-  geom_polygon(aes(fill = dont.know)) +
+  geom_polygon(aes(fill = dontKnow)) +
   geom_polygon(data = plotMap, aes(long,lat), 
@@
   geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
-  geom_polygon(aes(fill = gdpchange07.10)) +
+  geom_polygon(aes(fill = gdpChange)) +
   geom_polygon(data = plotMap, aes(long,lat), 
@@
 
-### Gini coefficient
+### Perceived Gini coefficient
 
@@
   geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
-  geom_polygon(aes(fill = uw_gini)) +
+  geom_polygon(aes(fill = perGini)) +
   geom_polygon(data = plotMap, aes(long,lat), 

### 2014-02-11 regressionModelling.Rmd
Getting back to business. Nothing changed, but rerunning the script.
Pientä päivtystä. Huomenna iso remmi päälle

@@
-# Regression modellining
-
-```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=FALSE,eval=FALSE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
-```
+---
+title: Regression modellining
+output:
+  html_document:
+    toc: true
+    theme: flatly
+---
 
@@
 
+# Tables for presentation
+
+## Partial
+
 # FINAL TABLES SAME AS IN PAPER
@@
 
-```
-
 ## Comparison of full models

### 2014-02-12 article2013.Rmd
Poistin eriarvoisuus+maaryhmaläpän ja lisäsin melko lopulliset kuviot&taulukot. Esitän tän Mikolle2

@@
 author: [Markus Kainu & Mikko Niemelä]
-date: February 11, 2014
+date: February 12, 2014
 abstract: Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
@@
     number_section: true
-    highlight: zenburn
-  html_document:
-    toc: true
-    theme: flatly
 ---
@@
 
-There are also considerable differences in other aspects of socio-political context between examined countries.
-One of these aspects is the extent of income inequality which varies extensively between examined countries.
-Thus, this study examines the role of income inequality as a contextual-level explanatory factor of public perceptions of the causes of poverty.
-The link between inequality and attributions for poverty is not that easy to predict.
-One the one hand, the public in countries with greater inequality is more inclined to mistrust “others” (e.g.
-Uslander, 2002).
-Therefore, the uneven distribution of wealth should mean stronger support for the individualistic explanation for poverty.
-On the other hand, cross-national studies on attributions for poverty have shown that the support for the individual blame is lower in countries with lower level of income inequality (e.g.
-Gallie and Paugam, 2002; Kallio and Niemelä, 2013).
+----
+
+*Eriarvoisuus lähti tästä*
+
+*Maaryhmittely on tästä pois*
+
+**Tähän tulisi HDI:n/elinolojen perustelut**
+
+----
 
@@
 
-Gini coefficient                  Measures the distribution of income or consumption expenditure among individuals or households                World Bank
+Human Development Index           Measures life expectancy, education, and income indices as a composite statistic
+UNDP
 
-Voice and Accountability          Indicators measuring the political process, civil liberties and political rights.                             World Bank
+Voice and Accountability          Indicators measuring the political process, civil liberties and political rights                             World Bank
 ----------------------------
@@
 
+\newpage
+
 Determinants of perceptions
@@
 
-<!--
-
-![Gini coefficient and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_gini.pdf)
-
--->
-
-![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
-
->There is also a substantial variation in the level of income inequality among post-socialist countries.
-Figure 3 shows that the higher level of income inequality is associated with higher level of social blame.
-As for individual blame type of explanation the picture is more puzzled, showing only weak connection between larger inequality and lower support for individual blame explanation.
-Again, the Western European countries differ with post-socialist countries.
-Cross-national differences in income inequality are much smaller in Western Europe than among post-socialist countries.
-There is a weak negative association between the social blame and income  inequality and a strong positive association between the individual blame and income inequality.
-
-![**Perceived inequality from data** and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_perGini.pdf)
-
 ![Human Development Index and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_hdi.pdf)
 
-AND SOMETHING ABOUT THE NEW FIGURE ON HUMAN DEVELPMENT INDEX AND LIVING CONDITIONS OF THE PEOPLE AND SKIPPING THE NEXT PARAGRAPH ON INEQUALITY
-
->There is also a substantial variation in the level of income inequality among post-socialist countries.
-Figure 3 shows that the higher level of income inequality is associated with higher level of social blame.
-As for individual blame type of explanation the picture is more puzzled, showing only weak connection between larger inequality and lower support for individual blame explanation.
-Again, the Western European countries differ with post-socialist countries.
-Cross-national differences in income inequality are much smaller in Western Europe than among post-socialist countries.
-There is a weak negative association between the social blame and income  inequality and a strong positive association between the individual blame and income inequality.
+HERE WE NEED AN INTREPETATIIONS OF HDI-FIGURE. BUT THE RESOURCES FIRST
 
@@
 Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.
+
+\newpage
 

### 2014-02-12 index.Rmd
Etusivu päivitetty..
Getting back to business. Nothing changed, but rerunning the script.
Poistin eriarvoisuus+maaryhmaläpän ja lisäsin melko lopulliset kuviot&taulukot. Esitän tän Mikolle2

@@
 | ------------ | -------------- |
-| [loading, cleaning and summarizing](loadClean.html) | [source](loadClean.Rmd) |
+| [loading, cleaning and summarizing](http://muuankarski.github.io/attributions/loadClean.html) | [source](http://muuankarski.github.io/attributions/loadClean.Rmd) |
 
@@
 | ------------ | -------------- |
-| [Descriptive analysis](descriptive_analysis.html) | [source](descriptive_analysis.Rmd) |
-| [Spatial visualization](mapPlots.html) | [source](mapPlots.Rmd) |
-| [Graphics in the paper](finalPlots.html) | [source](finalPlots.Rmd) |
+| [Descriptive analysis](http://muuankarski.github.io/attributions/descriptive_analysis.html) | [source](http://muuankarski.github.io/attributions/descriptive_analysis.Rmd) |
+| [Spatial visualization](http://muuankarski.github.io/attributions/mapPlots.html) | [source](http://muuankarski.github.io/attributions/mapPlots.Rmd) |
+| [Graphics in the paper](http://muuankarski.github.io/attributions/finalPlots.html) | [source](http://muuankarski.github.io/attributions/finalPlots.Rmd) |
+
+### Misc analysis
+
+| html-ouput | source code |
+| ------------ | -------------- |
+| [Comparing waves 2006 and 2011](http://muuankarski.github.io/attributions/compareWaves.html) | [source](http://muuankarski.github.io/attributions/compareWaves.Rmd) |
+| [Exploring the macro level correlations](http://muuankarski.github.io/attributions/correlations.html) | [source](http://muuankarski.github.io/attributions/correlations.md) |
 
@@
 | ------------ | -------------- |
-| [Regression analysis](regressionModelling.html) | [source](regressionModelling.Rmd) |
+| [Regression analysis](http://muuankarski.github.io/attributions/regressionModelling.html) | [source](http://muuankarski.github.io/attributions/regressionModelling.Rmd) |
 
@@
 | ------------ | -------------- |
-| [Article draft in .pdf](article2013.pdf) | [source](article2013.Rmd) |
+| [Article draft in .pdf](http://muuankarski.github.io/attributions/article2013.pdf) | [source](http://muuankarski.github.io/attributions/article2013.Rmd) |
 
@@
 
-- [HSE Moscow - November 12, 2013](moscow2013/lcsr20131112.html)
-- [Jatkokoulutusseminaari Turku November 7, 2013](turku2013/slidesTkuJatko20131107.html)
-- [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)
-
-# Misc links
-
-- [Article code in github](https://github.com/muuankarski/attributions)
-- [rustfare-package](http://markuskainu.fi/rustfare/) *- toolbox for R language for open welfare data from Russian Federation*
-- [rOpenGov](http://ropengov.github.io/) *-  R Ecosystem for Open Government Data and Computational Social Science*
-- [markuskainu.fi](http://markuskainu.fi) *- personal website*
-
------
+- [slides](http://muuankarski.github.io/attributions/attrib_slides.html)

### 2014-02-12 article2013.Rmd
Poistin eriarvoisuus+maaryhmaläpän ja lisäsin melko lopulliset kuviot&taulukot. Esitän tän Mikolle
lisäsin esimerkkitekstikapplaeen

@@
 ==============================================
+
+This site gathers together all the data analysis behind the actual paper.
+Data analysis is broken down in three parts and each part is presented in .html format for human eyes and in text source including the R-code required for reproducing the analysis.
+As both the micro data and macro data are available as open data, anyone can rerun these analysis.
 

### 2014-02-12 article2013.Rmd
Lisäsin teoriaan olennaiset kohdat lepianka & co paperista liittyen HDI:n

--- 2014-02-12 article2013.Rmd
@@
 
-*Eriarvoisuus lähti tästä*
+**Suora lainausta**: *Lepianka, D.A., J.P.T.M.
+Gelissen, and W.J.H.
+van Oorschot.
+“Popular Explanations of Poverty in Europe: Effects of Contextual and Individual Characteristics across 28 European Countries,” 2010.*
 
-*Maaryhmittely on tästä pois*
+**Theory s. 57**
 
-**Tähän tulisi HDI:n/elinolojen perustelut**
+>Socio-economic structure may determine the salience of poverty and inequality and thus influence poverty attributions.
+In general, higher unemployment rates and economic decline decrease respondents’ inclination to attribute poverty to individual failings and increase the likelihood of viewing material destitution as caused externally (Gallie and Paugam, 2002).
+We therefore hypothesize that the better the economic performance of a country, the greater the inclination of its inhabitants to view poverty as self-inflicted.
+
+**2. Independent variables s.60**
+
+>Direct measurements of national poverty levels for all 28 countries and relevant years were not readily found, and therefore we used the Human Development Index (HDI) as a proximal indicator.
+The HDI seeks to rank countries in terms of their level of development and combines normalized measures of life expectancy, educational attainment and GDP per capita for countries worldwide.
+There is evidence that the HDI is very strongly correlated to the more recently developed Human Poverty Index, and thus, to a large extent, measures the same construct (cf.
+Sahay et al., 2003: 90).
+Here, we use the average HDI score per country for the period 1994–1999.
+Data come from the World Bank’s World Development Reports 1994–1999; see www.worldbank.org.
+
+**Findings s. 65**
+
+>On the basis of the above bivariate contextual models, we selected the independent variables which were significantly related to popular poverty attributions – HDI, share of Catholics, average work ethic and egalitarianism – and estimated a rank-ordered logit model in which the effects of these contextual variables were estimated simultaneously (see Table 4, Model 1).
+6 The results of Model 1 indicate that as people live in more developed countries, they are more likely to rank Unluckiness ahead of Modern progress as an explanation for living in destitution.
+
+**Conclusions s. 86**
+
+>In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.
 

### 2014-02-13 article2013.Rmd
Kirjoitin kappaleen elinoloista teoriaan ja lueskelin paperia eteenpäin. Meni aamua (9-15) sitten toisen projektin kimpussa, niin huomiselle siirtyi tämän valmistelua.
Pyrin huomenna kirjoittamaan menetelmien ja muuttujien osalta täsmennyksiä ja avaamaan analyysien tuloksia.

--- 2014-02-12 article2013.Rmd
@@
 
-----
-
-*Eriarvoisuus lähti tästä*
-
-*Maaryhmittely on tästä pois*
-
-**Tähän tulisi HDI:n/elinolojen perustelut**
-
-----
-
-Besides the so called regime-effect, national-level economic conditions have been proved to be associated with welfare attitudes.
+Besides the so called good government, national-level economic conditions have been proved to be associated with welfare attitudes.
 This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence.
@@
 Thus, we can assume that changes in country’s economic performance are associated with attributions for poverty regardless of country grouping.
+
+In addition to macro level economic performance, the level overall of well-being and living conditions are found to be linked with poverty attributions.
+Lepianka et al.
+(2010:65,86) found out that people living in more developed countries they tend to attribute living in destitution rather to individual unluckiness than to modern progress, but they found no link between development and either type of blame explanations.
+In contrast to previous studies, variation in human development within our selection of countries if far greater.
+Therefore we are able to poor countries of Central Asia with well-off Slovenia and Czech Republic.
+Given the great differences in living conditions within the post-socialist space, we assume that in countries of low human development and fewer welfare institutions people tend to attribute poverty more with individual reasons than social.
+At the same time people experience of transition with welfare state retrenchment and increasing social and health problems may well have produced support for structural explanations, too.
+
+`Joku toinen lähde olisi paikallaan tähän liittyen. Ja tuo ristiriitaisen hypoteesin voisi oikoa..`
 
@@
 It is therefore difficult to form any solid hypotheses as to whether low education as a social risk is associated with perceptions of the causes of poverty in a similar vein than other risk factors discussed above.
+
+\newpage
 
@@
 
-Human Development Index           Measures life expectancy, education, and income indices as a composite statistic
-UNDP
+Human Development Index           Measures life expectancy, education, and income indices as a composite statistics                             UNDP
 
@@
 (see Appendix A)
+
+`Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media.
+[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf) - *leikkasin tulos-luvusta tän kappaleen*`
+
+`From Lepianka 2010  Independent variables s.60`
+
+`Direct measurements of national poverty levels for all 28 countries and relevant years were not readily found, and therefore we used the Human Development Index (HDI) as a proximal indicator.
+The HDI seeks to rank countries in terms of their level of development and combines normalized measures of life expectancy, educational attainment and GDP per capita for countries worldwide.
+There is evidence that the HDI is very strongly correlated to the more recently developed Human Poverty Index, and thus, to a large extent, measures the same construct (cf.
+Sahay et al., 2003: 90).
+Here, we use the average HDI score per country for the period 1994–1999.
+Data come from the World Bank’s World Development Reports 1994–1999; see www.worldbank.org.`
 
@@
 
---------
-
-*A more throughout description of modelling will be added as the final details in modelling will be decided*
-
 Results
 ==============================================
-
-*Results and conclusions are not yet reshaped to match the recent changes in the analysis...*
 
@@
 
-HERE WE NEED AN INTREPETATIIONS OF HDI-FIGURE. BUT THE RESOURCES FIRST
+`On the basis of the above bivariate contextual models, we selected the independent variables which were significantly related to popular poverty attributions – HDI, share of Catholics, average work ethic and egalitarianism – and estimated a rank-ordered logit model in which the effects of these contextual variables were estimated simultaneously (see Table 4, Model 1).
+6 The results of Model 1 indicate that as people live in more developed countries, they are more likely to rank Unluckiness ahead of Modern progress as an explanation for living in destitution.`
+
+`From Lepianka 2010 Conclusions s. 86`
+
+`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`
 
 ![World Bank's Voice and Accountability index in 2011 and support for social (top) and individual (bottom) blame types of attributions for poverty](figure/scatter_wbgi.pdf)
-
->Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media.
-[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf)
 

### 2014-02-13 finalPlots.Rmd
Pientä päivtystä. Huomenna iso remmi päälle
Kirjoitin kappaleen elinoloista teoriaan ja lueskelin paperia eteenpäin. Meni aamua (9-15) sitten toisen projektin kimpussa, niin huomiselle siirtyi tämän valmistelua.
Pyrin huomenna kirjoittamaan menetelmien ja muuttujien osalta täsmennyksiä ja avaamaan analyysien tuloksia.

@@
   geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
-  coord_map(project="orthographic") +
+  coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(40,60)) +
   theme_minimal() + 
@@
 
-pdf("figure/mapplot.pdf", width=14/2.54, height=12/2.54)
+pdf("figure/mapplot.pdf", width=12/2.54, height=7/2.54)
 mapplot

### 2014-02-14 article2013.Rmd
Kirjoitin teoriaa ja tutkimusasetelmaa ajantasaiseksi. Jatkan lounaan jälkeen tuloksista

@@
 author: [Markus Kainu & Mikko Niemelä]
-date: February 12, 2014
+date: February 14, 2014
 abstract: Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
@@
 Therefore, as previous studies have suggested (Kallio and Niemelä, 2013; Lepianka et al., 2010), there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space.
-Hence, the general purpose of the study is to analyze public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
+Hence, the general purpose of the study is to analyze public perceptions of the causes of poverty in Europe and Central Asia with special emphasis on post-socialist countries.
 
@@
 
-As country-level determinants of perceptions, we will analyze, whether this variation can be explained by different regimes within post-socialist countries and the post-socialist legacy.
-The study divides post-socialist countries into two groups: Central Eastern Europe (CEE) and Commonwealth of Independent States (CIS).
-CEE includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania; and CIS[^****] Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia.
-The survey data enables us also to compare post-socialist countries to a selection of Western European nations (France, Germany, Italy and Sweden).
-Thus, in order to examine the distinctiveness of post-socialist perception, Western European countries are included in the descriptive analyses.
-(See Figure 1.)
+We target our analyses at 22 post-socialist countries and use five Western European countries as references in descriptive analyses (see figure 1.).
+Our selection on 22 post-socialist countries includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania from Central Eastern Europe (CEE), and Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia from Commonwealth of Independent States (CIS)[^***].
+Of Western European countries we have included France, Germany, Italy, United Kingdom and Sweden as references.
 
@@
 
-[^****]:Georgia withdrawn its membership from CIS in 2008
+[^***]:Georgia withdrawn its membership from CIS in 2008
 
-Besides the country groups we will also analyze the role of changes in country’s economic performance during the period of global financial crises, income inequality and the state of democracy.
+As contextual level predictors we use country’s economic performance during the period of global financial crises, living conditions and the state of democracy.
 As described in previous sections, we assume that in those countries where the economic crises has meant declining economic conditions, public are more likely to endorse social explanations for poverty than in countries where national economy has continued to grow despite the global economic turbulence.
 Based on the theoretical discussion on civil liberties, political rights and institutional trust, we assume that the public support to social blame type of explanation is high in those countries where the state of democracy is low.
-Due to contradictory findings on the direction of the effect of income inequality, it is more difficult to provide any straightforward hypothesis on the importance of income inequality.
+The case with human deveplopment in less clear, but we assume, against the hypothesis by Lepianka et al.
+(2010), that lower level of human development and lack of welfare institutions may well be connected with higher support for individual blame type of explanation.
 
@@
 
-Effect of financial crises        Whether respondents perceives that their household has suffered the financial crisis great or fair amount     list
+Effect of financial crises        Whether respondents perceives that their household has suffered the financial crisis great or fair amount     lits
 
@@
 
-An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of four variables: 1) change in total GDP in 2007 - 2010, 2) gini coefficient of economic inequality, 3) 2011 democracy index and 4) country grouping.
-Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), gini coefficients from [UNU-WIDER World Income Inequality Database](http://www.wider.unu.edu/research/Database/en_GB/database/) and Democracy Index from the [Economist Intelligence Unit](http://www.eiu.com/Default.aspx).
-Gini coefficient for each country is the latest available varying from 2002 to 2006.
-(see Appendix A)
+An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of three variables: 1) change in total GDP in 2007 - 2010, 2) Human Deveplopment Index in 2010 and 3) Voice and accountability variable.
+Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), Human Development Index from [United NAtions Development Programme](http://hdr.undp.org/en/data) and Voice and Accountability from the World Bank's [Worldwide Governance Indicators](http://info.worldbank.org/governance/wgi/index.aspx#home).
+All macro level indicators are obtained from [Quality of Government Institutes](http://www.qog.pol.gu.se/) data using [rQog](http://markuskainu.fi/rqog/)-package (Kainu 2014).
 
-`Voice and accountability captures perceptions of the extent to which a country's citizens are able to participate in selecting their government, as well as freedom of expression, freedom of association, and a free media.
-[info.worldbank.org/governance/wgi/pdf/va.pdf](http://info.worldbank.org/governance/wgi/pdf/va.pdf) - *leikkasin tulos-luvusta tän kappaleen*`
+<!-- seuraava kappale piilossa
 
@@
 
+-->
+
 The empirical section starts by presenting the results of descriptive analyses.
-As we assume that individual level perceptions are dependent both on individual and country-level contextual factors, it is recommended to use multilevel regression analysis for this task (Snijders \& Bosker, 2011).
-Multilevel analysis requires hierarchical dataset and allows the estimation of country-level effects on individual level outcomes.
-The analysis is performed using lmer-function from lme4 package in R.
+As we assume that individual level perceptions are dependent both on individual and country-level contextual factors and we have a hierarchical two level dataset, it is recommended to use multilevel regression analysis for this task (Snijders & Bosker, 2011) as multilevel regression analysis allows the estimation of country-level effects on individual level outcomes.
+The analysis is performed using glmer-function from lme4 package in R.
 All of the models employed are so called multilevel logistic random intercept models at the level of individuals and countries.
 
-Our multilevel modelling focuses first in bivariate connections between two dependent variables and both individual and country level independent variables(See Appendix B).
-At the second step we include all individual level variables to a single model (models sb1 and ib1).
-Then we add each of the four country level variables one at the time (models sb2/ib2 - sb5/ib5).
-With limited number of countries we are obliged to use only few contextual variables.
-Tables will include both a fixed effects and a random effects part.
-The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups.
+We are modelling the date in stepwise manner first running the empty models for both social blame (table 3) and individual blame (table 4) type of explanation.
+At second stage we include all individual level predictors and in third stage we include only the macro level predictors.
+In final models we include both the individual and macro level variables.
+We assume that our dependent variables are independent, but to confirm this we run a further model where we predict support for social blame against the support for individual blame using both the individual and macro level variables.
+This model is compared with full models from tables 3 and 4 in table 5.
+
+`Tables will include both a fixed effects and a random effects part.
+The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups.`
 

### 2014-02-14 article2013.Rmd
Kirjoitin kuvioita vielä vähän auki. Tulen huomenna päivällä pariksi tunniksi ja viilaan paperin mallien ja johtopäätösten osalta Moskovan kuntoon

@@
 
-Table 3 shows the proportion of the total number of respondents in a country which chose a particular explanation of poverty.
+Table 2 shows the proportion of the total number of respondents in a country which chose a particular explanation of poverty.
 We assumed that citizens in post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations.
@@
 injustice in a society and laziness of the poor, are the most popular poverty attributions.
-The highest support for social blame is found in Ukraine, Latvia, Lithuania, Hungary, Russia, Slovenia, Armenia, Moldova and Azerbaijan.
+The highest support for social blame is found in Ukraine, Lithuania, Slovenia, Hungary,  Russia and Latvia.
 In these countries majority of population blames the injustice in a society for poverty.
-Totally, in 19 of 21 examined post-socialist countries injustice in a society receives the highest support.
-
-\newpage
+Totally, in 18 of 21 examined post-socialist countries injustice in a society receives the highest support.
 
@@
 print.xtable(xtable(tbl5,digits=1,
-                    caption="Weighted population shares of dependent variable by country",), #table.placement="H",
+                    caption="Weighted population shares of dependent variable by country",), 
+             #table.placement="H",
              type="latex",
@@
 Also, in most post-socialist countries the individual blame is the second most popular explanation of poverty.
-Exceptions here are Slovenia, Estonia and Georgia where citizens endorse societal explanations more likely than individual explanations.
-In Czech Republic social explanations as well as individual blame explanation receive public support quite evenly.
+Exceptions here are Ukraine, Slovenia, Estonia and Georgia where citizens endorse societal explanations more likely than individual explanations.
+In Armenia and Czech Republic social explanations as well as individual blame explanation receive public support quite evenly.
 Finally, the clearest exception in post-socialist regime is Uzbekistan where poverty is attributed strongly with internal terms.
@@
 
-\newpage
-
 Determinants of perceptions
@@
 
-Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 2, 3 and 4).
-Correlation coefficients at country group1 level are shown on top of each regression line.Figure 2 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010).
+Analyses regarding the determinants of poverty explanations starts by bivariate associations between the dependent variables and three contextual variables (Figures 3, 4 and 5).
+Correlation coefficients within the post-socialist countries are shown on top of each regression line.
+Figure 3 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010).
 In regard to post-socialist countries there is substantial variation between countries in economic change.
-The general trend is quite clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.35) and more to individual blame (.48).
+The trend is very clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.62) and more to individual blame (.71).
 In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries.
@@
 
-`On the basis of the above bivariate contextual models, we selected the independent variables which were significantly related to popular poverty attributions – HDI, share of Catholics, average work ethic and egalitarianism – and estimated a rank-ordered logit model in which the effects of these contextual variables were estimated simultaneously (see Table 4, Model 1).
-6 The results of Model 1 indicate that as people live in more developed countries, they are more likely to rank Unluckiness ahead of Modern progress as an explanation for living in destitution.`
-
-`From Lepianka 2010 Conclusions s. 86`
-
-`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`
+There is also a substantial variation in the level of human development among post-socialist countries.
+Figure 4 shows that the higher of human development is associated with higher level of social blame (.34).
+Association holds also in the case of individual blame meaning that the higher the human deveploment is the lower support for support for individual blame explanation (-.48).
+In this comparison Western countries are better-off in terms of human development and due to substantial variation in dependent variables it is harder to confirm our hypothesis than in the case of post-socialist countries.
 
@@
 
-In regard to the level of democracy we can see great differences between and within groups of countries.
-Overall association is weak (.19 for social blame and -.44 for individual blame), but in CIS countries the higher level of democracy is clearly indicating higher support for social blame and lower support for individual blame type of explanation.
-In CEE and in Western Europe the variation in the state of democracy is much smaller than in CIS countries.
+In regard to the level of goodness of government we can see major differences are great as expected.
+Overall association is substantial (.41 for social blame and -.47 for individual blame), and pointing that higher level of democracy is indicating higher support for social blame, whereas lower support for individual blame type of explanation is linked with low figure in good governance indicator.
+As expected , the variation in the goodness of government in Western Europe is much smaller than in post-socialist countries.
 In these countries the support for blame types of explanations seems to vanish when the level of democracy increases.
@@
 Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
+
+`On the basis of the above bivariate contextual models, we selected the independent variables which were significantly related to popular poverty attributions – HDI, share of Catholics, average work ethic and egalitarianism – and estimated a rank-ordered logit model in which the effects of these contextual variables were estimated simultaneously (see Table 4, Model 1).
+6 The results of Model 1 indicate that as people live in more developed countries, they are more likely to rank Unluckiness ahead of Modern progress as an explanation for living in destitution.`
+
+`From Lepianka 2010 Conclusions s. 86`
+
+`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`
 
@@
         custom.coef.names = coef.names.multi,
-        naive = TRUE,
+        #naive = TRUE,
         #use.packages=TRUE,
@@
 
-Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterly 37 (3), 413–428.
+Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterl

### 2014-02-14 code/dv_tables.R

@@ -42,7 +42,7 @@ t2 <- t2[order(-t2$socialBlame),]
 ########
 y <- data.frame(prop.table(svytable(~group_general+q309, d.df), 1)*100)
 y$Freq <- round(y$Freq, 2)
-y$country <- c("CEE Total","CIS Total","Western Europe Total")
+y$country <- c("CEE mean","CIS mean","Western Europe mean")
 y2 <-  dcast(y, group_general+country ~ q309, value.var="Freq")
 names(y2) <-  c("group_general","country","socialBlame","individualBlame",
                 "individualFate","socialFate","notStated","dontKnow")

### 2014-02-15 article2013.Rmd
JA vielä lisää spämmiä!

@@
 author: [Markus Kainu & Mikko Niemelä]
-date: February 14, 2014
+date: February 15, 2014
 abstract: Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
@@
 ==============================================
-
-This site gathers together all the data analysis behind the actual paper.
-Data analysis is broken down in three parts and each part is presented in .html format for human eyes and in text source including the R-code required for reproducing the analysis.
-As both the micro data and macro data are available as open data, anyone can rerun these analysis.
 
@@
 
-Wilson, George (1996) Toward a Revised Framework for Examining Beliefs about the Causes of Poverty. The Sociological Quarterl
+Wilson, George (1996) Toward a Revised Framework for Exa

### 2014-02-15 article2013.Rmd
Johtopäätöksiä en vielä katsonut. Mallien tulkintoja etupäässä. Palaan illalla asiaan.

@@
 ---
-title: Attributions for Poverty in Post-Socialist Countries
+title: "Attributions for Poverty in Post-Socialist Countries"
 author: [Markus Kainu & Mikko Niemelä]
@@
 
+\thispagestyle{empty}
+
 \newpage
+
+\thispagestyle{empty}
+
+\section*{Progress report}
+
+\subsection*{Summary of previous comments}
+
+- why to study welfare attitudes or attributions for poverty
+- country grouping
+- gini coefficient
+
+>(sosiaali)psykologian alalla voisi olla hyviä lähteitä tulevaisuuden odotusten ja asenteiden yhteydestä. Kommentoija puhui kontrollista ja sen menettämisen vaikutuksesta. Sitä voisi katsoa..
+
+>Toinen kysymys oli edelleen Why.
+Sofia tosin korjasi sen että sosiaalipolitiikassa tällaisen tutkiminen on tavanomaista ja tuleehan se siinä paperissa hyvin esille.
+Tästä kuitenkin vahvistui ajatus siitä, että jos tätä laittaa esim.
+alue-lehteen eikä tieteenala-lehteen niin rautalankamalli siitä miksi tutkitaan voisi olla hyvä.
+
+>Malleja ja muuttujavalintoja kehuttiin, tuloksia pidettiin "hyvinä" merkitsevyyden näkökulmasta.
+
+\subsection*{Explanation of what has been done to cope with the comments}
+
+Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Fusce porta imperdiet justo non imperdiet.
+Nam molestie ultricies lectus, vitae blandit nunc ornare nec.
+Praesent id metus nulla.
+Donec gravida eu felis sed posuere.
+Fusce bibendum nisl mauris, id tempor enim tempus et.
+Mauris id pulvinar velit, eu venenatis dui.
+Aenean ut erat sit amet diam convallis aliquam.
+Donec ultricies lobortis elit nec tempor.
+
+\newpage
+
+\setcounter{page}{1}
 
@@
 
-<!-- seuraava kappale piilossa
-
-`From Lepianka 2010  Independent variables s.60`
-
-`Direct measurements of national poverty levels for all 28 countries and relevant years were not readily found, and therefore we used the Human Development Index (HDI) as a proximal indicator.
-The HDI seeks to rank countries in terms of their level of development and combines normalized measures of life expectancy, educational attainment and GDP per capita for countries worldwide.
-There is evidence that the HDI is very strongly correlated to the more recently developed Human Poverty Index, and thus, to a large extent, measures the same construct (cf.
-Sahay et al., 2003: 90).
-Here, we use the average HDI score per country for the period 1994–1999.
-Data come from the World Bank’s World Development Reports 1994–1999; see www.worldbank.org.`
-
--->
-
 The empirical section starts by presenting the results of descriptive analyses.
@@
 All of the models employed are so called multilevel logistic random intercept models at the level of individuals and countries.
+Model summary tables include both a fixed effects and a random effects part.
+The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups.
 
@@
 In final models we include both the individual and macro level variables.
-We assume that our dependent variables are independent, but to confirm this we run a further model where we predict support for social blame against the support for individual blame using both the individual and macro level variables.
+As we are using *logistics regression* instead of *multinomial logistic regression* with dependent variable of nominal scale, we confirm our findings by running a further model where we predict the support for social blame against the support for individual blame using both the individual and contextual level variables.
 This model is compared with full models from tables 3 and 4 in table 5.
-
-`Tables will include both a fixed effects and a random effects part.
-The fixed effects part will include odds ratios and statistical significances of independent variables, while the random effects part will indicate the country-level variance and standard deviation, deviance and number of cases and groups.`
 
@@
 
-`On the basis of the above bivariate contextual models, we selected the independent variables which were significantly related to popular poverty attributions – HDI, share of Catholics, average work ethic and egalitarianism – and estimated a rank-ordered logit model in which the effects of these contextual variables were estimated simultaneously (see Table 4, Model 1).
-6 The results of Model 1 indicate that as people live in more developed countries, they are more likely to rank Unluckiness ahead of Modern progress as an explanation for living in destitution.`
-
-`From Lepianka 2010 Conclusions s. 86`
-
-`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`
-
 ```{rarticleModelsSoc, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
@@
         include.pvalues=TRUE,
-        caption="Logistic multilevel random intercept model for social blame type of explanation", 
+        caption="Logistic multilevel random intercept model for social blame (SB) type of explanation", 
         caption.above=TRUE)
@@
         include.pvalues=TRUE,
-        caption="Logistic multilevel random intercept model for social blame type of explanation", 
+        caption="Logistic multilevel random intercept model for individual blame (IB) type of explanation", 
         caption.above=TRUE)
@@
        include.pvalues=TRUE,
-       caption="Logistic multilevel random intercept model for social and individual blame type of explanation")
+       caption="Logistic multilevel random intercept model comparing the full models for both dependent variable with model predicting social blame against individual blame")
 
@@
 
-Table 4 and 5 present figures for multilevel logistic analysis.
+Tables 3, 4 and 5 present figures for multilevel logistic regressions.
 Calculating R squared or residual standard deviation makes no sense in the case of logistic regression or other binomial models.
-Instead, we use \emph{deviance} as our primary statistical summary of logistic model fit.
+Instead, we use *deviance* as our primary statistical summary of logistic model fit.
 Deviance can be treated as similar to residual standard deviation for generalized linear models.
@@
 
-Looking at the deviance for the empty models both in tables 4 and 5, we can see that there is "less" to be explained for in the model for individual blame than for social blame.
-When adding the individual level predictors to the models, the deviance decreases by 8512 for social blame and 6636 for individual blame.
-In both model sb1 and ib1 we have four statistically significant predictors.
-Low education and worsened household income level do not contribute in explaining neither of the outcomes.
-Perceived low income level has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation.
-Coefficient of .368 can be translated so that if respondent perceives belonging to low income group he or she is 9.2\% more likely to blame society for poverty than a non-low income individual.
+Looking at the deviance for the empty models both in tables 3 and 4, we can see that there is "less" to be explained for in the model for individual blame than for social blame.
+When adding the individual level predictors to the models, the deviance decreases by 8574 for social blame and 6662 for individual blame.
+In both model *SB individual* and *IB individual* we have four statistically significant predictors.
+*Perceived low income level* has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation.
+Coefficient of .36 can be translated so that if respondent perceives belonging to low income group he or she is 8.5 % more likely to blame society for poverty than a non-low income individual.
 Expected worsening in household income and financial crises experince predict social blame almost in similar manner.
-All our individual level predictors but low education predict lower probability for supporting individual blame (model ib1).
-Being greatly affected by financial crisis makes one 11\% less likely to support for individual blame type of explanation.
+*Low education* and *worsened household income* do not contribute in explaining neither of the outcomes.
+Being greatly affected by financial crisis makes one 11 % less likely to support for individual blame type of explanation.
 
+<!--
 When adding in the country grouping (CEE/CIS) we can observe decrease close to -1 in deviance in both models sb2 and id2.
 Country grouping in itself is not significant in our models, whereas gini coefficient has a weak connection to individual blame type of explanation, so that individual blame is supported more in more unequal countries, as we assumed based on studies about the association between generalised trust and equality (Uslander 2002).
-Inclusion of change in national GDP between 2007 - 2010 substracts the deviance by 14 for social blame and 9 for individual blame model.
-Change in GDP becomes the most significant predictor of our country level variables for both type of explanation.
-The better the national economy has performed during the financial crises, the more probably is the support for individual blame type of explanation.
-Outcome is very much align with the findings of Gallie and Paugam (2002) and Kallio and Niemelä (2010).
-Our fourth contextual variable, democracy index, is significant in explaining the individual blame in a way that higher level of democracy implies lower support for individual blame type of explanation.
-Coefficient plots from both full models summarizes the estimates for logistic regression coefficients at 50\% and 90\% intervals in figure 7.
+-->
 
-As for explaining the between-country variation the inclusion of country grouping adds a very little.
-Gini coefficient reduces variation from .24 to .19 for social blame, but it is the inclusion of GDP change that has the greatest effect.
-Between-country variation  to .12 for social blame and .04 for individual blame.
-In the final models the inclusion of democracy index results country-level variances at .09 and .02 for social and individual type of explanation, respectively.
-In other words, only 9\% and 2\% of the variation in the final models are due to country differences at final stage.
+Including only the contextual level predictors, as in models *SB contextual* (table 3) and *IB contextual* (table 4), has a marginal effect on deviance when comparing with empty models, but a substantial effect on between country variance.
+Of the contectual predictors the change in GDP has the most significant impact on dependent variables in a such manner, that the better national economy has performed during the crises the less there is support for social blame type of explanation, and in opposite way in regard to individual blame.
+
+Introduction of both the individual and contextual level predictors in the models has only a minor impact on significange of the predictors.
+Same individual level predictors; *perceived low income*, *negative expectations for future*, *perception of negative impacts of economic crises* and *benefit dependency* remain significant at the individual level and *change in national GDP* at the contextual level.
+As for overall model fit the individual level predictors are in key role: the impact of contextual level variables on deviance is only -10 and -16 for social blame and individual blame type of explanation, respectively.
+Model *SB vs.
+IB* in table 5 confirms the findings on the significange and direction of the association.
+
+\newpage
 
@@
 ==============================================
+
+`From Lepianka 2010 Conclusions s. 86`
+
+`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`
 
@@
 
+Kainu, Markus (2014) rQog. Download data from the Quality of Government Institute data in R. R-package avaible at: [github.com/muuankarski/rQog](https://github.com/muuankarski/rQog)
+
 Kallio, Johanna and Niemelä, Mikko (2010) Public Perceptions of the Causes of Poverty: Multilevel Analysis of 26 European Nations. Unpublished manuscript.
@@
 
-Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of California Press.
-
-Rothstein, Bo (2000) Trust, Social Dilemmas and Collective Memories. Journal of Theoretical Politics, 12 (4), 477–501.
-
-Rothstein, Bo and Stolle, Dietlind (2008) The State and Social Capital: An Institutional Theory of Generalized Trust. Comparative Politics 40 (4), 441-459.
-
-Saunders, Peter (2002) The Ends and Means of Welfare. Coping with Economic and Social Change in Australia. Cambridge: Cambridge University Press.
-
-Schumpeter, Joseph A. (1942) Capitalism, Socialism & Democracy. New York: Harper & Brothers.
-
-Szikra, D., & Tomka, B. (2009). Social Policy in East Central Europe: Major Trends in the Twentieht Century. In Post-communist Welfare Pathways (Vol. 2009).
-
-Stephenson, Svetlana (2000) Public Beliefs in the Causes of Wealth and Poverty and Legitimization of Inequalities in Russia and Estonia. Social Justice Research 13 (2), 83–100.
-
-Svallfors, Stefan (1997) Worlds of Welfare and Attitudes to Redistribution: A Comparison of Eight Western Nations. European Sociological Review 13 (3), 283–304.
-
-Tomka, B. (2006). East Central Europe and the European Social Policy Model: A Long-Term View. East European Quarterly, 40(2), 135.
-
-van Oorschot, Wim and Halman, Loek (2000) Blame or Fate, Individual or Social? An International Comparison of Popular Explanations of Poverty. European Societies 2 (1), 1–28.
-
-Wilson, George (1996) Toward a Revised Framework for Exa
+Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of Califor

### 2014-02-15 article2013.Rmd
Poistin virheet johtopäätöksistä ja siistin muutenkin Moskovaa varten. Vielä pitää ens vkolla avata koko hoito, mutta menköön nyt näin Moskovaan.

@@
 author: [Markus Kainu & Mikko Niemelä]
+affiliation: Finland
 date: February 15, 2014
+header-includes:
+    - \usepackage{fancyhdr}
+    - \pagestyle{fancy}
+    - \fancyhead[CO,CE]{}
+    - \fancyfoot[CO,CE]{}
+    - \fancyfoot[LE,RO]{\thepage}
 abstract: Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
@@
 
+\thanks{Some thanks}
+
 \thispagestyle{empty}
@@
 
-\section*{Progress report}
+\section*{Progress report for 4th LCSR International Workshop “Social and Cultural Changes in Cross-National Perspective: Values and Modernization”  (March 28 –April 4, 2014, Moscow)}
 
@@
 
-- why to study welfare attitudes or attributions for poverty
-- country grouping
-- gini coefficient
+My discussants were somewhat confused after my presentation and I can think two reasons for that.
+Firstly, the first version of my paper that they had read was very preliminary and had no results what-so-ever.
+However, the project had proceeded since summer and my presentation was rather result oriented.
+Another thing was that I had some technical difficulties and therefore couldn't quite deliver my point in the presentation.
 
->(sosiaali)psykologian alalla voisi olla hyviä lähteitä tulevaisuuden odotusten ja asenteiden yhteydestä. Kommentoija puhui kontrollista ja sen menettämisen vaikutuksesta. Sitä voisi katsoa..
-
->Toinen kysymys oli edelleen Why.
-Sofia tosin korjasi sen että sosiaalipolitiikassa tällaisen tutkiminen on tavanomaista ja tuleehan se siinä paperissa hyvin esille.
-Tästä kuitenkin vahvistui ajatus siitä, että jos tätä laittaa esim.
-alue-lehteen eikä tieteenala-lehteen niin rautalankamalli siitä miksi tutkitaan voisi olla hyvä.
-
->Malleja ja muuttujavalintoja kehuttiin, tuloksia pidettiin "hyvinä" merkitsevyyden näkökulmasta.
+But I got good comments, though, and my discussants were mainly after better reasoning why to study welfare attitudes in the first place, and many of the comments where somehow related to that *Why-question*.
+However, I was also asked to give reasons for leaving out Western European countries and to argue why such country selection was chosen.
+One commentator directed me towards social psychological literature on the effect of feeling of losing control in life and how it could be linked with welfare attitudes.
 
@@
 
-Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Fusce porta imperdiet justo non imperdiet.
-Nam molestie ultricies lectus, vitae blandit nunc ornare nec.
-Praesent id metus nulla.
-Donec gravida eu felis sed posuere.
-Fusce bibendum nisl mauris, id tempor enim tempus et.
-Mauris id pulvinar velit, eu venenatis dui.
-Aenean ut erat sit amet diam convallis aliquam.
-Donec ultricies lobortis elit nec tempor.
+I did work a lot with my colleague in December and also now in February towards the deadline and we tried to improve the paper as a whole and prepare for first submission.
+We decided to leave out any country grouping and to treat our selection of post-socialist countries as a whole and to study contextual variables at the country level.
+We tried to become more clear with why to study such issue, but it may still be somewhat unclear for someone not familiar with welfare attitude research in social policy literature.
+Also, we worked a lot on our analysis and for example, replaced gini coefficient with Human Development Index in our contextual variables.
+Previous findings on gini coefficient we somewhat contradictory and uninteresting and we thought actual living conditions could perhaps be better suited for our analysis.
+HDI was most reliable measure of living conditions as national level poverty figures are hard to compare between countries as in our selection.
+
+- The analysis can be browsed in here: [muuankarski.github.io/attributions/](https://muuankarski.github.io/attributions/)
+- Changes since Novermber can be seen through git history here: [https://github.com/muuankarski/attributions/commits/master/article2013.Rmd](https://github.com/muuankarski/attributions/commits/master/article2013.Rmd)
 
@@
 
-van Oorschot and Halman (2000) have suggested a four-tier typology (Table 1) in which individual blame emphasizes internal factors, such as undesirable behaviour of the poor, moral looseness and laziness.
+van Oorschot and Halman (2000) have suggested a four-tier typology (Figure 1) in which individual blame emphasizes internal factors, such as undesirable behaviour of the poor, moral looseness and laziness.
 Individual fate type of explanation emphasizes that poverty can be caused by uncontrollable and inescapable internal factors.
@@
 In contrast to previous studies, variation in human development within our selection of countries if far greater.
-Therefore we are able to poor countries of Central Asia with well-off Slovenia and Czech Republic.
-Given the great differences in living conditions within the post-socialist space, we assume that in countries of low human development and fewer welfare institutions people tend to attribute poverty more with individual reasons than social.
-At the same time people experience of transition with welfare state retrenchment and increasing social and health problems may well have produced support for structural explanations, too.
-
-`Joku toinen lähde olisi paikallaan tähän liittyen. Ja tuo ristiriitaisen hypoteesin voisi oikoa..`
+Therefore we are able to compare poor countries of Central Asia with well-off Slovenia and Czech Republic.
+Given the great differences in living conditions within the post-socialist space, we assume that in countries of low human development and fewer welfare institutions people tend to attribute poverty more with social reasons than individual.
 
@@
 According to van Oorschot and Halman (2000) different statements represent a four-tier typology of poverty explanations as follows: While the unlucky refers to individual fate, the laziness and lack of willpower represents individual blame.
-In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate (see Table 1).
+In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate (see figure 1).
 Based on the findings of previous studies, we assume that in general post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations.
@@
 Variables are recoded as binomial where value 1 is coded for risk and 0 for absence of risk.
-Both contextual and individual-level variables are described in the Table 2.
+Both contextual and individual-level variables are described in the Table 1.
 
@@
 
-Ghange in total GDP 2007 – 2010   Relative change from 2007 to 2010                                                                             World Bank
+Ghange in total GDP 2007 – 2010   Indicator of performance of country's economy. Measure used is relative change from 2007 to 2010                                      World Bank
 
-Human Development Index           Measures life expectancy, education, and income indices as a composite statistics                             UNDP
+Human Development Index           Indicator of human development/living conditions. Measures life expectancy, education, and income indices as a composite statistics   UNDP
 
-Voice and Accountability          Indicators measuring the political process, civil liberties and political rights                              World Bank
+Voice and Accountability          indicator of state of democracy. Measures the political process, civil liberties and political rights                                 World Bank
 ----------------------------
 
-An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of three variables: 1) change in total GDP in 2007 - 2010, 2) Human Deveplopment Index in 2010 and 3) Voice and accountability variable.
-Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), Human Development Index from [United NAtions Development Programme](http://hdr.undp.org/en/data) and Voice and Accountability from the World Bank's [Worldwide Governance Indicators](http://info.worldbank.org/governance/wgi/index.aspx#home).
+An additional dataset of selected macro indicators were added to the LiTS2-dataset consisting of three variables: 1) change in total GDP in 2007 - 2010, 2) Human Deveplopment Index in 2010 and 3) Voice and accountability indicator.
+Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), Human Development Index from [United Nations Development Programme](http://hdr.undp.org/en/data) and Voice and Accountability from the World Bank's [Worldwide Governance Indicators](http://info.worldbank.org/governance/wgi/index.aspx#home).
 All macro level indicators are obtained from [Quality of Government Institutes](http://www.qog.pol.gu.se/) data using [rQog](http://markuskainu.fi/rqog/)-package (Kainu 2014).
@@
 
-`From Lepianka 2010 Conclusions s. 86`
-
-`In addition, people who live in countries that have a higher level of development (and presumably a lower level of poverty) – measured by the HDI – are more likely to attribute living in poverty sooner to Unluckiness rather than to Modern Progress.`
-
 To sum up, the results indicate that in general social blame type of explanation is clearly the most popular explanation in post-socialist countries.
 Social blame is followed by individual blame, social fate and individual fate types of explanations.
-Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty in the European post-socialist countries.
+Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty in the post-socialist space.
 In regard to differences between the East and the West of Europe, results show that Western Europeans emphasise more social reasons for poverty.
@@
 To begin with, we found that differences between countries are responsible for almost 30 percent of the variation in the support for social blame and 12 percent for individual blame type of explanation.
-Our contextual level predictors decreased the variance down to 9 percent for social and 2 percent for individual type of explanation.
-Yet, findings showed that the membership in the CIS and CEE country group was not significant in this respect.
+Our contextual level predictors decreased the variance down to 15 percent for social and 4 percent for individual type of explanation.
 
@@
 
+<!-- no tää ei ollut merkitsevä
 Also, the level of democracy had a substantial impact.
@@
 Social blame type of explanation was, in turn, endorsed more likely in countries with higher level of democracy.
+-->
+
+Our other contextual level predictors, level of democracy and human development, were not significant in the models.
+However, the descriptive analysis showed that contracy to our assumptions, citizens in those countries where the state of democracy and human development is low do not blame more likely society than citizens in countries where they have political freedom and power to influence socio-political issues.
+Instead, findings indicate that the lower level of democracy and low level of human develpment were associated with stronger support for individual blame type of explanation.
+Social blame type of explanation was, in turn, endorsed more likely in countries with higher level of democracy and human development.
 
@@
 
-Kainu, Markus (2014) rQog. Download data from the Quality of Government Institute data in R. R-package avaible at: [github.com/muuankarski/rQog](https://github.com/muuankarski/rQog)
+Kainu, Markus (2014) rQog. Download data from the Quality of Government Institute data in R. R-package available at: [github.com/muuankarski/rQog](https://github.com/muuankarski/rQog)
 

### 2014-02-15 index.Rmd
Poistin virheet johtopäätöksistä ja siistin muutenkin Moskovaa varten. Vielä pitää ens vkolla avata koko hoito, mutta menköön nyt näin Moskovaan.

@@
-# Attributions for Poverty in Post-Socialist Countries - Markus Kainu & Mikko Niemelä
+---
+title: "Attributions for Poverty in Post-Socialist Countries"
+author: [Markus Kainu & Mikko Niemelä]
+---
 
 ****
-
-```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=TRUE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=12)
-```
 
@@
 As both the [micro data](http://www.ebrd.com/pages/research/publications/special/transitionII.shtml) and [macro data](http://www.qog.pol.gu.se/data/) are available as open data, anyone can rerun these analysis.
-
-**Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.**
 
@@
 - [slides](http://muuankarski.github.io/attributions/attrib_slides.html)
+
+---------
+
+**Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.**

### 2014-02-15 code/compile.R

@@ -1,0 +1,12 @@
+setwd("~/workspace/lits/clone/attributions")
+#
+pandoc_md <- function(name="index") {
+    input <- paste(name,".md", sep="")
+    output <- paste(name,".html", sep="")
+    system(paste("pandoc -s -S --toc -H /home/aurelius/workspace/web/css/rmarkdown.css",input,"-o",output))
+}
+
+library(knitr)
+knit("index.Rmd")
+system("cp index.md README.md")
+pandoc_md('index')

### 2014-02-19 article2013.Rmd
Mikon muutokset!

@@
 
-The purpose of this article is to examine cross-country differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries.
+The purpose of this article is to examine cross-national differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries.
 Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and Eastern European countries (Gallie and Paugam, 2002; Kallio and Niemelä, 2013; Lepianka, 2007; van Oorschot and Halman, 2000).
@@
 Hence, it seems that unlike in Western Europe traditionally opposite attitudinal dimensions, i.e.
-structural and individualistic explanations of poverty coincide with each other in East Europeans perceptions.
+structural and individualistic explanations of poverty coincide with each other in East Europeans' perceptions.
 However, there is shortage of comparative studies on attributions for poverty in particular focusing on the East Central European countries or on the whole post-socialist space.
@@
 The article explores to what extent specific risks related to household financial circumstances at the individual level and the state of democracy, income inequality and changes in national economy at the contextual level are associated with different types of lay explanations of poverty.
-Finally, extensive country-selection of our data provides a novel opportunity to examine the effects of the financial crises both at the individual and contextual level and their association with attributions for poverty.
+Finally, extensive country-selection of our data provides a novel opportunity to examine the effects of the financial crisis both at the individual and contextual level and their association with attributions for poverty.
 At the individual level, the study focuses not only on the individuals’ current financial situation.
@@
 The article is structured as follows: A brief overview of theoretical discourse on explanations of poverty and prior findings on cross-national differences in and contextual level determinants of attributions for poverty is presented in Section 2.
-Section 3 focuses on the self-interest approach as an individual level explanatory factor of the perceptions with particular interest of risk factors related to individual’s financial situation.
+Section 3 focuses on the self-interest approach as an individual-level explanatory factor of the perceptions with particular interest of risk factors related to individual’s financial situation.
 Research design, survey data and methods are described in detail in Section 4.
@@
 In Russia, on the other hand, the economic system is blamed even more than the individuals themselves (Gorshkov and Tikhonova, 2006; Stephenson, 2000).
-These results are consistent with Lepianka's (2007: 127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia.
+These results are consistent with Lepianka's (2007: 127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia as well as with Habibov's (2011) comparative evidence from the Caucasus, Central Asia, Russia and Ukraine.
 Interesting result is that social blame coincides strongly with individual blame.
@@
 
-Based on qualitative poverty studies in Central Asia, Caucasus and Ukraine and Moldova Dudwick et al.
+Based on qualitative poverty studies in Central Asia, the Caucasus, Ukraine and Moldova, Dudwick et al.
 (2003: 23) states that, generally, people blamed their own poverty on the failure of the Soviet state and the corruption, indifference, and incompetence of their new leaders.
@@
 This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence.
-In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crises in distinct ways.
+In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crisis in distinct ways.
 While many Western and Eastern European countries have witnessed economic decline, the economic performance in Western Asia has been strong.
-Hence, by analyzing the role of changes in economic performance during the period of global financial crises, the research design of this article provides empirical evidence on the effects of crises to public perceptions of the causes of poverty.
+Hence, by analyzing the role of changes in economic performance during the period of global financial crisis, the research design of this article provides empirical evidence on the effects of crisis to public perceptions of the causes of poverty.
 Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007).
-In regard to explanations of poverty, Gallie and Paugam (2002:21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved.
+In regard to explanations of poverty, Gallie and Paugam (2002: 21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved.
 This result is confirmed by Kallio and Niemelä (2013).
@@
 
-In addition to macro level economic performance, the level overall of well-being and living conditions are found to be linked with poverty attributions.
-Lepianka et al.
-(2010:65,86) found out that people living in more developed countries they tend to attribute living in destitution rather to individual unluckiness than to modern progress, but they found no link between development and either type of blame explanations.
-In contrast to previous studies, variation in human development within our selection of countries if far greater.
-Therefore we are able to compare poor countries of Central Asia with well-off Slovenia and Czech Republic.
-Given the great differences in living conditions within the post-socialist space, we assume that in countries of low human development and fewer welfare institutions people tend to attribute poverty more with social reasons than individual.
+In addition to macro-level economic performance, living conditions in general are found to be linked with attributions for poverty.
+People living in countries with high unemployment are more likely to endorse social than individualistic explanations (Albrekt Larsen, 2006: 74-79).
+In addition, the increased incidence of poverty reduces the likelihood of choosing individual laziness as an explanation of poverty (Lepianka, 2007: 118-121).
+However, there are to some extent contradictory findings as well.
+While Kallio and Niemelä (2013) did not find association between individual blame type of explanation and incidence of poverty, Lepianka et al.
+(2010) found out that people living in more developed countries tend to attribute poverty rather to individual and social fate than blame types of explanations.
+Yet, in contrast to previous studies, the variation between post-socialist countries is far larger than in Western Europe.
+Consequently, by comparing for instance poorer countries of Central Asia with more well-off Slovenia and Czech Republic, this study provides a possibility to analyze whether the association between attributions for poverty and living conditions is also stronger in the post-socialist countries.
+This study examines human development index (HDI) as a proxy for living conditions in a given country.
+Given the great differences in living conditions within the post-socialist space, we assume that in countries of low level of HDI people tend to attribute poverty more with social than individual reasons.
 
@@
 In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalized by the individuals in the society in question (Taylor-Gooby, 1985; Svallfors, 1995).
-More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice, 2001) and resource/class-based risks (Cusack et al.
-2006; Blomberg et al.
-2012; Finseraas and Ringdal, 2012) as key to welfare attitudes.
+More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice, 2001) and resource/class-based risks (Cusack et al., 2006; Blomberg et al., 2012; Finseraas and Ringdal, 2012) as key to welfare attitudes.
 Thus, studies have assumed – and to some extent proved – that groups that have a greater risk of facing social problems or being dependent on welfare state benefits might perceive the government’s role in welfare issues in a distinct way compared to those who are not exposed to these risks.
 
-Inspired by the studies which emphasise risks as the key mechanisms, this article tests to what extent factors related to financial hardship are associated – beyond the welfare regime – with perceptions of the causes of poverty.
+Inspired by the studies which emphasize risks as the key mechanisms, this article tests to what extent factors related to financial hardship are associated with perceptions of the causes of poverty.
 Following this line of reasoning, we are interested in whether those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.
-Thus, the research design in regard to determinants of poverty attributions enables us to study from a fairly unexplored perspective the different mechanisms through which individuals who have a greater risk to experience material destitution perceive the causes of poverty.
+Thus, the research design in regard to determinants of poverty attributions enables us to study from a fairly unexplored perspective to the different mechanisms through which individuals who have a greater risk to experience material destitution perceive the causes of poverty.
 
@@
 We assume that economic crisis have affected especially those who are the most vulnerable groups in the society.
-Therefore, we will analyse whether those who report that the economic crisis has affected their household at least a fair amount perceive the causes of poverty differently than those who report that it has had no or only a little effect to their household.
+Therefore, we will analyze whether those who report that the economic crisis has affected their household at least a fair amount perceive the causes of poverty differently than those who report that it has had no or only a little effect to their household.
 According to self-interest hypothesis, we assume that those whose economic situation has declined and those whose household has affected by economic crisis are more inclined to blame society for poverty than those who has not experienced economic problems.
@@
 According to van Oorschot and Halman (2000) different statements represent a four-tier typology of poverty explanations as follows: While the unlucky refers to individual fate, the laziness and lack of willpower represents individual blame.
-In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate (see figure 1).
+In addition, injustice in a society indicates social blame, whereas the view that poverty is simply part of modern life refers to social fate (see Figure 1).
 Based on the findings of previous studies, we assume that in general post-socialist countries endorse more likely individual blame and social blame types of explanations than individual and social fate types of explanations.
@@
 
-We target our analyses at 22 post-socialist countries and use five Western European countries as references in descriptive analyses (see figure 1.).
+We target our analyses at 22 post-socialist countries and use five Western European countries as references in descriptive analyses (see Figure 2).
 Our selection on 22 post-socialist countries includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania from Central Eastern Europe (CEE), and Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia from Commonwealth of Independent States (CIS)[^***].
@@
 
-As contextual level predictors we use country’s economic performance during the period of global financial crises, living conditions and the state of democracy.
-As described in previous sections, we assume that in those countries where the economic crises has meant declining economic conditions, public are more likely to endorse social explanations for poverty than in countries where national economy has continued to grow despite the global economic turbulence.
+As contextual level predictors we use country’s economic performance during the period of global financial crisis, living conditions and the state of democracy.
+As described in previous sections, we assume that in those countries where the economic crisis has meant declining economic conditions, public are more likely to endorse social explanations for poverty than in countries where national economy has continued to grow despite the global economic turbulence.
 Based on the theoretical discussion on civil liberties, political rights and institutional trust, we assume that the public support to social blame type of explanation is high in those countries where the state of democracy is low.
-The case with human deveplopment in less clear, but we assume, against the hypothesis by Lepianka et al.
-(2010), that lower level of human development and lack of welfare institutions may well be connected with higher support for individual blame type of explanation.
+The case with human deveplopment is to some extent less clear, but we assume that the lower level of human development is connected with stronger support for social than individual explanation.
 
@@
 Hereby, the research design aims at contributing to the theoretical reasoning regarding the importance of self-interest for attributions for poverty which assumes that those who have financial difficulties or those who have a greater risk to be poor perceive the causes of poverty differently than those who are not exposed to these poverty-related risks.
-Analyses include six individual-level predictors, namely transfer dependency, low education, perceived low income level, income-level compared to the past and expected change in income in the future as well as perceived effect of the financial crises.
+Analyses include six individual-level predictors, namely transfer dependency, low education, perceived low income level, income-level compared to the past and expected change in income in the future as well as perceived effect of the financial crisis.
 Variables are recoded as binomial where value 1 is coded for risk and 0 for absence of risk.
@@
 
-Effect of financial crises        Whether respondents perceives that their household has suffered the financial crisis great or fair amount                             lits
+Effect of financial criris        Whether respondents perceives that their household has suffered the financial crisis great or fair amount                             lits
 
@@
 Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), Human Development Index from [United Nations Development Programme](http://hdr.undp.org/en/data) and Voice and Accountability from the World Bank's [Worldwide Governance Indicators](http://info.worldbank.org/governance/wgi/index.aspx#home).
-All macro level indicators are obtained from [Quality of Government Institutes](http://www.qog.pol.gu.se/) data using [rQog](http://markuskainu.fi/rqog/)-package (Kainu 2014).
+All macro level indicators are obtained from [Quality of Government Institutes](http://www.qog.pol.gu.se/) data using [rQog](http://markuskainu.fi/rqog/)-package (Kainu, 2014).
 
 The empirical section starts by presenting the results of descriptive analyses.
-As we assume that individual level perceptions are dependent both on individual and country-level contextual factors and we have a hierarchical two level dataset, it is recommended to use multilevel regression analysis for this task (Snijders & Bosker, 2011) as multilevel regression analysis allows the estimation of country-level effects on individual level outcomes.
+As we assume that individual level perceptions are dependent both on individual and country-level contextual factors and we have a hierarchical two level dataset, it is recommended to use multilevel regression analysis for this task (Snijders and Bosker, 2011) as multilevel regression analysis allows the estimation of country-level effects on individual level outcomes.
 The analysis is performed using glmer-function from lme4 package in R.
@@
 
-We are modelling the date in stepwise manner first running the empty models for both social blame (table 3) and individual blame (table 4) type of explanation.
-At second stage we include all individual level predictors and in third stage we include only the macro level predictors.
-In final models we include both the individual and macro level variables.
+We are modelling the date in stepwise manner first running the empty models for both social blame (Table 3) and individual blame (Table 4) type of explanation.
+At second stage we include all individual-level predictors and in third stage we include only the macro-level predictors.
+In final models we include both the individual and macro-level variables.
 As we are using *logistics regression* instead of *multinomial logistic regression* with dependent variable of nominal scale, we confirm our findings by running a further model where we predict the support for social blame against the support for individual blame using both the individual and contextual level variables.
-This model is compared with full models from tables 3 and 4 in table 5.
+This model is compared with full models from tables 3 and 4 in Table 5.
 
@@
 In regard to differences between the East and the West of Europe, results show that there are some differences.
-Yet there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al., 2010; Kallio and Niemelä, 2010).
+Yet, there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al., 2010; Kallio and Niemelä, 2010).
 Clearest differences between the West and the East is the stronger support for social fate in Western Europe.
@@
 
-Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty.
+Hence, the results are in line with our assumptions and prior research (Habibov, 2011; Lepianka et al., 2010) regarding the general support for different explanations of poverty.
 Overall, the results indicate that for the most of the post-socialist countries two most popular explanations are individual and social blame types of explanations.Yet, the differences between social fate and individual blame types of explanations are in some countries very small.
 In general, Western Europeans (except Britons) do not support individual blame type of explanation as much as citizens in the East.
-Instead they emphasise more social reasons for poverty.
-Results also clearly emphasise that the idea that poverty is caused by just individual bad luck is the least popular perception of the reason for poverty.
+Instead they emphasize more social reasons for poverty.
+Results also clearly indicate that the idea that poverty is caused by just individual bad luck is the least popular perception of the reason for poverty.
 

### 2014-02-20 article2013.Rmd
Alaluvun 5.2. viilaukset

@@
 
-Transfer dependency               Social or private transfers as main source of income                                                                                  lits
+Transfer dependency               Social or private transfers as a main source of income                                                                                  lits
 
-Low education                     No or only compulsory level education                                                                                                 lits
+Low education                     No or only compulsory level of education                                                                                                 lits
 
@@
 
-Voice and Accountability          indicator of state of democracy. Measures the political process, civil liberties and political rights                                 World Bank
+Voice and Accountability          Indicator of state of democracy. Measures the political process, civil liberties and political rights                                 World Bank
 ----------------------------
@@
 The results above indicate that there are cross-national differences between countries and within country groups which means that there are also some other contextual and individual-level factors which should explain the cross-national differences.
-Therefore, we will also analyse the importance of country-level economic, social and political conditions and individual-level risk factors related to financial hardship as determinants for attributions for poverty.
+Therefore, we will also analyze the importance of country-level economic, social and political conditions and individual-level risk factors related to financial hardship as determinants for attributions for poverty.
 Because the descriptive results showed that the majority of citizens in the post-socialist countries endorsed social and individual blame types of explanations, the subsequent analyses will focus only on these two explanations.
@@
 Correlation coefficients within the post-socialist countries are shown on top of each regression line.
-Figure 3 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crises (2007-2010).
+Figure 3 shows the association between attributions for poverty and the change in the size of economy during the period of global financial crisis (2007-2010).
 In regard to post-socialist countries there is substantial variation between countries in economic change.
-The trend is very clear: the less country has suffered from the financial crises the less there is support to social blame type of explanation (-.62) and more to individual blame (.71).
+The trend is very clear: the less country has suffered from the financial crisis the less there is support to social blame type of explanation (-.62) and more to individual blame (.71).
+Thus, the results are in line with previous studies which suggest positive correlation between the economic growth and support for the individualistic perception of the causes of poverty (Gallie and Paugam, 2002; Kallio and Niemelä, 2013).
 In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries.
@@
 
-There is also a substantial variation in the level of human development among post-socialist countries.
-Figure 4 shows that the higher of human development is associated with higher level of social blame (.34).
-Association holds also in the case of individual blame meaning that the higher the human deveploment is the lower support for support for individual blame explanation (-.48).
-In this comparison Western countries are better-off in terms of human development and due to substantial variation in dependent variables it is harder to confirm our hypothesis than in the case of post-socialist countries.
+In regard to country-level living conditions, we assumed that the high level of human development is associated more with individual than social blame.
+However, the results indicate that the hypothesis is not supported.
+Figure 4 shows that higher level of human development is associated with higher level of social blame (.34).
+Association holds also in the case of individual blame, meaning that the higher the human deveploment is the lower is the support for individual blame explanation (-.48).
+Western European countries, on the other hand, are more better-off in terms of human development.
+It seems that there is not any significant association between human development and blame types of explanations in Western European countries (also Kallio and Niemelä, 2013; Lepianka et al.
+2010).
 
@@
 
-In regard to the level of goodness of government we can see major differences are great as expected.
-Overall association is substantial (.41 for social blame and -.47 for individual blame), and pointing that higher level of democracy is indicating higher support for social blame, whereas lower support for individual blame type of explanation is linked with low figure in good governance indicator.
-As expected , the variation in the goodness of government in Western Europe is much smaller than in post-socialist countries.
-In these countries the support for blame types of explanations seems to vanish when the level of democracy increases.
-Consequently, both social and individual fate types of  explanations gain higher support in more democratic societies.
+Finally, there also large variation between countries in the level of goodness of government.
+Overall, the association between attributions for poverty and the level of good governance is substantial (.41 for social blame and -.47 for individual blame).
+Regarding post-socialist countries, however, the direction of association is also here opposite as expected.
+On the one hand, Figure 5 shows that higher level of democracy is indicating higher support for social blame.
+On the other hand, individual blame type of explanation is supported more likely in countries with low degree of good governance.
+In regard to Western European countries, the variation in the goodness of government is much smaller than in post-socialist countries.
+Also the association between perceptions and the good governance differs.
+Unlike in post-socialist countries, the support for blame types of explanations seems to vanish when the level of good governance increases.
 

### 2014-02-21 article2013.Rmd
5.3. muutoksia alkuun, jatkuu vielä tänään

@@
 
-As the main controbution of our study relates to cross-national variation of attributions for poverty within post-socialist countries, we shall largely focus on the effects of contextual variables.
-Nevertheless, as we are covering several countries that haven't been included in previous studies of individual level determinants, we shall also look carefully at the individual level.
+Descriptive analyses presented above would posit that there are rather clear, though to some extent unexpected, associations between country-level contextual characteristics and the perceptions of the causes of poverty.
+In order to explore the effects of both contextual and individual-level determinants in a more detailed statistical manner, the analytsis continues with multilevel logistic regressions (Tables 3, 4 and 5).
+
+Between-country variances in Tables 3 and 5 are indicators for the variance in dependent variable that is explained by country differences.
+Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation.
+Inclusion of individual-level predictors decreases the variance in both models, but only moderately.
+
+In addition, while calculating R squared or residual standard deviation makes no sense in the case of binomial models, our primary statistical summary of logistic model fit is *deviance*.
+Deviance can be treated as similar to residual standard deviation for generalized linear models.
+Basically, deviance is a measure of error, and lower deviance implicates better fit to data.
+When a random predictor is added to model, we expect deviance to decrease, on average, by 1.
+When predictor is more informative deviance should decrease more than 1.
+(Gelman and Hill, 2007).
+As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.
+The deviance for the empty models both in Tables 3 and 4 indicates that there is "less" to be explained for in the model for individual blame than for social blame.
+When adding the individual-level predictors to the models, the deviance decreases by 8574 for social blame and 6662 for individual blame.
 
@@
 
-Tables 3, 4 and 5 present figures for multilevel logistic regressions.
-Calculating R squared or residual standard deviation makes no sense in the case of logistic regression or other binomial models.
-Instead, we use *deviance* as our primary statistical summary of logistic model fit.
-Deviance can be treated as similar to residual standard deviation for generalized linear models.
-Basically, deviance is a measure of error, and lower deviance implicates better fit to data.
-When a random predictor is added to model, we expect deviance to decrease, on average, by 1.
-When predictor is more informative deviance should decrease more than 1.
-(Gelman and Hill, 2007).
-As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.
-
-Between-country variances are on the first rows in random effects, meaning the variance in dependent variable that is explained by country differences.
-Variance in empty models is .29 in social blame and .12 in individual blame indicating that there are differences between countries, and that differences are greater in respect to support for social blame type of explanation.
-Inclusion of individual level predictors decreases the variance in both models, but only moderately.
-
-Looking at the deviance for the empty models both in tables 3 and 4, we can see that there is "less" to be explained for in the model for individual blame than for social blame.
-When adding the individual level predictors to the models, the deviance decreases by 8574 for social blame and 6662 for individual blame.
  In both model *SB individual* and *IB individual* we have four statistically significant predictors.

### 2014-02-22 article2013.Rmd
5.3. Individual-level tulokset
Jatkan vielä tästä macro-leveleitä

@@
 
-Effect of financial criris        Whether respondents perceives that their household has suffered the financial crisis great or fair amount                             lits
+Effect of financial crisis        Whether respondents perceives that their household has suffered the financial crisis great or fair amount                             lits
 
@@
 When a random predictor is added to model, we expect deviance to decrease, on average, by 1.
-When predictor is more informative deviance should decrease more than 1.
-(Gelman and Hill, 2007).
+When predictor is more informative deviance should decrease more than 1 (Gelman and Hill, 2007).
 As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.
@@
 
- In both model *SB individual* and *IB individual* we have four statistically significant predictors.
-*Perceived low income level* has the highest coefficient of our binary individual level predictors for explaining social blame type of explanation.
-Coefficient of .36 can be translated so that if respondent perceives belonging to low income group he or she is 8.5 % more likely to blame society for poverty than a non-low income individual.
-Expected worsening in household income and financial crises experince predict social blame almost in similar manner.
-*Low education* and *worsened household income* do not contribute in explaining neither of the outcomes.
-Being greatly affected by financial crisis makes one 11 % less likely to support for individual blame type of explanation.
+The self-interest hypothesis assumes that those who have financial difficulties or those who have a greater risk to be poor are more inclined to support societal reasons for poverty than those who are not exposed to poverty-related risks.
+Both Table 3 and Table 4 show four statistically significant individual-level predictors.
+*Perceived low income level* has the highest coefficient of our binary individual-level predictors for explaining social blame type of explanation.
+Coefficient of -.36 can be translated so that if respondent perceives belonging to low income group he or she is 8.5 % more likely to blame society for poverty than a non-low income individual.
+Dependency on public or private transfers, expected worsening in household income in the future and experienced negative effect of financial crisis to household's economy predict the support for social blame explanation almost in similar manner (Table 3).
+The self-interest hypothesis is also supported in Table 4 which shows that negative experience of financial crisis, transfer dependency, perceived low income level and expected worsening of household income level are statistically associated with individualistic perception of the causes of poverty.
+For instance, being greatly affected by financial crisis makes one 11 % less likely to support for individual blame type of explanation.
+
+However, there are also individual-level risk factors which are not connected with perceptions of the causes of poverty.
+First, analyses indicate to some extent equivocal results of the explanatory effect of past experiences: unlike the perceived negative effect of financial crisis, the subjective assessment of negative changes in income is not associated with perceptions.
+Second, low level of education does not contribute in explaining neither social nor individualistic perception of the causes of poverty.
 

### 2014-02-27 article2013.Rmd
versio welfare clusterin lukupiiriin. poistin moskovan selittelyt alusta

@@
 
-\thanks{Some thanks}
-
 \thispagestyle{empty}
-
-\newpage
-
-\thispagestyle{empty}
-
-\section*{Progress report for 4th LCSR International Workshop “Social and Cultural Changes in Cross-National Perspective: Values and Modernization”  (March 28 –April 4, 2014, Moscow)}
-
-\subsection*{Summary of previous comments}
-
-My discussants were somewhat confused after my presentation and I can think two reasons for that.
-Firstly, the first version of my paper that they had read was very preliminary and had no results what-so-ever.
-However, the project had proceeded since summer and my presentation was rather result oriented.
-Another thing was that I had some technical difficulties and therefore couldn't quite deliver my point in the presentation.
-
-But I got good comments, though, and my discussants were mainly after better reasoning why to study welfare attitudes in the first place, and many of the comments where somehow related to that *Why-question*.
-However, I was also asked to give reasons for leaving out Western European countries and to argue why such country selection was chosen.
-One commentator directed me towards social psychological literature on the effect of feeling of losing control in life and how it could be linked with welfare attitudes.
-
-\subsection*{Explanation of what has been done to cope with the comments}
-
-I did work a lot with my colleague in December and also now in February towards the deadline and we tried to improve the paper as a whole and prepare for first submission.
-We decided to leave out any country grouping and to treat our selection of post-socialist countries as a whole and to study contextual variables at the country level.
-We tried to become more clear with why to study such issue, but it may still be somewhat unclear for someone not familiar with welfare attitude research in social policy literature.
-Also, we worked a lot on our analysis and for example, replaced gini coefficient with Human Development Index in our contextual variables.
-Previous findings on gini coefficient we somewhat contradictory and uninteresting and we thought actual living conditions could perhaps be better suited for our analysis.
-HDI was most reliable measure of living conditions as national level poverty figures are hard to compare between countries as in our selection.
-
-- The analysis can be browsed in here: [muuankarski.github.io/attributions/](https://muuankarski.github.io/attributions/)
-- Changes since Novermber can be seen through git history here: [https://github.com/muuankarski/attributions/commits/master/article2013.Rmd](https://github.com/muuankarski/attributions/commits/master/article2013.Rmd)
 

### 2014-02-27 article2013.Rmd
Changed the date too

@@
 affiliation: Finland
-date: February 15, 2014
+date: February 27, 2014
 header-includes:

### 2014-03-02 article2013.Rmd
5.3 results ok
Muutin tekstissä taulukon 5 liitetaulukoksi 1. Se jää nykyisellään niin pieneen rooliin että sen voisi siirtää liitetaulukoksi. Mikäli sen haluaa pitään taulukko 5:na niin pitäisi kirjoitella vähän enemmän lihaa luiden ympärille eli avata sitä enemmän.

@@
 As we are using *logistics regression* instead of *multinomial logistic regression* with dependent variable of nominal scale, we confirm our findings by running a further model where we predict the support for social blame against the support for individual blame using both the individual and contextual level variables.
-This model is compared with full models from tables 3 and 4 in Table 5.
+This model is compared with full models from Tables 3 and 4 in Appendix Table 1.
 
@@
 
-Including only the contextual level predictors, as in models *SB contextual* (table 3) and *IB contextual* (table 4), has a marginal effect on deviance when comparing with empty models, but a substantial effect on between country variance.
-Of the contectual predictors the change in GDP has the most significant impact on dependent variables in a such manner, that the better national economy has performed during the crises the less there is support for social blame type of explanation, and in opposite way in regard to individual blame.
+Descriptive figures indicated rather clear associations between country-level contextual factors and perceptions of the causes of poverty.
+Tables 3 and 4 show that inclusion of only the contextual level determinants to the models (SB contextual and IB contextual respectively) has a marginal effect on deviance when the models are compared to empty models, but a substantial effect on between country variance.
+The change in GDP has the most significant impact on attributions for poverty.
+In line with our hypothesis, the larger the economic growth in a country is, more likely people are to blame the poor themselves and less likely they endorse injustice in society as a reason for poverty.
 
-Introduction of both the individual and contextual level predictors in the models has only a minor impact on significange of the predictors.
-Same individual level predictors; *perceived low income*, *negative expectations for future*, *perception of negative impacts of economic crises* and *benefit dependency* remain significant at the individual level and *change in national GDP* at the contextual level.
-As for overall model fit the individual level predictors are in key role: the impact of contextual level variables on deviance is only -10 and -16 for social blame and individual blame type of explanation, respectively.
-Model *SB vs.
-IB* in table 5 confirms the findings on the significange and direction of the association.
+Interesting enough, however, results show that unlike descriptive statistics indicated neither country-level living conditions (measured by Human Development Index) nor the level of goodness of government are significantly associated with attributions for poverty.
+In addition, final models in Tables 3 and 4 show that inclusion of both the individual and contextual level predictors to the models has only a minor impact on significance of the predictors.
+Consequently, at the individual-level *perceived low income*, *negative expectations for future*, *perception of negative impacts of economic crisis* and *benefit dependency* and at the contextual level *change in national GDP* remain significant determinants for the perceptions of the causes of poverty.
+The findings on the significance and direction of the association are confirmed in Appendix Table 1 which examines the predictive power of the support for social blame against the support for individual blame using both the individual and contextual level variables.
 

### 2014-03-03 article2013.Rmd
Mikon muutokset pedariin. Palaan toimistolla asiaan liitetaulukon yms. osalta

@@
 affiliation: Finland
-date: February 27, 2014
+date: March 3, 2014
 header-includes:

### 2014-03-05 article2013.Rmd
päivitin slaidit iltapäivän ja moskovan esiyksiin

@@
 
-```{rarticleModelsFull, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
-load("data/finalModels.RData")
-
-model.names.multi.2 = c("SB vs. not",
-                        "IB vs. not",
-                        "SB vs. IB")
-
-coef.names.multi = c("(intercept)",
-                     "Dependent",
-                     "Low education",
-                     "Low income",
-                     "Income has worsened",
-                     "Income will worsen",
-                     "Has affected great or fair amount",
-                     "Human Developmet Index",
-                     "Change in GDP between 2007 to 2010",
-                     "Voice and Accountability")
-
-library(texreg)
-texreg(list(soc.4,ind.4,socVsInc),
-       #float.pos="H",
-       #use.packages=TRUE,
-       #booktabs=TRUE,
-       #dcolumn=TRUE,
-       scriptsize=TRUE,
-       naive = TRUE,
-       caption.above=TRUE,
-       doctype=FALSE,
-       custom.model.names = model.names.multi.2,
-       custom.coef.names = coef.names.multi,
-       include.pvalues=TRUE,
-       caption="Logistic multilevel random intercept model comparing the full models for both dependent variable with model predicting social blame against individual blame")
-
-```
-
 The self-interest hypothesis assumes that those who have financial difficulties or those who have a greater risk to be poor are more inclined to support societal reasons for poverty than those who are not exposed to poverty-related risks.
@@
 Social blame is followed by individual blame, social fate and individual fate types of explanations.
-Hence, the results are in line with our assumptions and prior research (Lepianka et al., 2010) regarding the general support for different explanations of poverty in the post-socialist space.
-In regard to differences between the East and the West of Europe, results show that Western Europeans emphasise more social reasons for poverty.
+Hence, the results are in line with our assumptions and prior research (Habibov, 2011; Lepianka et al., 2010) regarding the general support for different explanations of poverty in the post-socialist space.
+In regard to differences between the East and the West of Europe, results show that Western Europeans emphasize more social reasons for poverty.
 
-As assumed results also emphasised large between-country differences among post-socialist countries.
+As assumed results also emphasized large between-country differences among post-socialist countries.
 By employing a multilevel approach we wanted to shed light especially on this issue.
@@
 
-As we hypothesised changes in country’s economic performance during the period of global financial crises were associated with attributions for poverty.
+As we hypothesized changes in country’s economic performance during the period of global financial crisis were associated with attributions for poverty.
 In fact, changes in economic performance was the most significant predictor of our country-level variables for both social and individual blame types of explanations: the better the economy had performed the more support gained the individual blame type of explanation.
-And on the other hand, increased financial strain is associated with stronger support for the view that poverty is caused injustices of society.
+And on the other hand, increased financial strain is associated with stronger support for the view that poverty is caused by injustices of society.
 
@@
 Results of the individual-level determinants of perceptions supported the self-interest hypothesis which states that those who have financial difficulties perceive the causes of poverty differently than those who are not exposed to poverty-related risks.
-Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crises has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty.
+Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crisis has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty.
 Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).
@@
 There are, of course, other variables which might be associated with attributions even more than risk factors.
-For instance, another line of research have emphasised the role of other beliefs, attitudes and values rather than interest (e.g.
+For instance, another line of research have emphasized the role of other beliefs, attitudes and values rather than interest (e.g.
 Cozzarelli et al., 2001).
@@
 As a methodological contribution, findings showed interesting differences between the East and the West regarding the effects of contextual variables.
-Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and income inequality are quite modest compared especially to CIS countries.
+Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and living conditions are quite modest compared especially to CIS countries.
 Consequently our results showed more modest and in some cases opposite associations between dependent and contextual variables among Western European countries than among post-socialist group.
@@
 
+Habibov, Nazim N.
+(2011) Public Beliefs Regarding the Causes of Poverty during Transition.
+Evidence from the Caucasus, Central Asia, Russia, and Ukraine.
+International Journal of Sociology and Social Policy 31 (1-2), 53-74.
+
 Hunt, Matthew O. (1996) The Individual, Society, or Both? A Comparison of Black, Latino, and White Beliefs about the Causes of Poverty. Social Forces 75 (1), 293–322.
@@
 Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of Califor
+
+\newpage
+
+Appendix
+=======================================================
+
+\setcounter{table}{0}
+\renewcommand\thetable{\arabic{table}}
+
+```{rarticleModelsFull, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
+load("data/finalModels.RData")
+
+model.names.multi.2 = c("SB vs. not",
+                        "IB vs. not",
+                        "SB vs. IB")
+
+coef.names.multi = c("(intercept)",
+                     "Dependent",
+                     "Low education",
+                     "Low income",
+                     "Income has worsened",
+                     "Income will worsen",
+                     "Has affected great or fair amount",
+                     "Human Developmet Index",
+                     "Change in GDP between 2007 to 2010",
+                     "Voice and Accountability")
+
+library(texreg)
+texreg(list(soc.4,ind.4,socVsInc),
+       float.pos="H",
+       #use.packages=TRUE,
+       #booktabs=TRUE,
+       #dcolumn=TRUE,
+       scriptsize=TRUE

### 2014-03-05 finalPlots.Rmd
päivitin slaidit iltapäivän ja moskovan esiyksiin

@@
 library(grid)
-map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
+map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
                                 layer="lits_all_simple"))
@@
 ppi <- 300
-png("figure/mapplot.png", width=14/2.54*ppi, height=12/2.54*ppi, res=ppi)
+png("figure/mapplot.png", width=15/2.54*ppi, height=9/2.54*ppi, res=ppi)
 mapplot

### 2014-03-05 mapPlots.Rmd
päivitin slaidit iltapäivän ja moskovan esiyksiin

@@
 load("~/workspace/lits/attributions/data/lits.RData")
-map <- suppressWarnings(readOGR(dsn="/home/aurelius/workspace/data/shapefiles/lits/", 
+map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
                                 layer="lits_all_simple"))
@@
 
-```{rmapPlotsGroup1}
+```{rmapPlotsGroup1, fig.height=13, fig.width=16}
 ggplot(map.df, aes(long,lat,group=group)) +
@@
   geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
-  coord_map(project="orthographic") +
+  coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(40,60)) +
   theme(legend.position="top")
@@
 plotMap <- merge(map.df,macro,by.x="NAME",by.y="cntry")
+plotMap <- plotMap[plotMap$year == 2010,]
 plotMap <- plotMap[order(plotMap$order),]
@@
 
-### Perceived Gini coefficient
+### Human Development Index
 
@@
   geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
-  geom_polygon(aes(fill = perGini)) +
+  geom_polygon(aes(fill = undp_hdi)) +
   geom_polygon(data = plotMap, aes(long,lat), 
@@
   coord_map(project="orthographic") +
-  scale_fill_gradient2(low=muted("blue"), high=muted("red"), midpoint=33.55) +
+  scale_fill_gradient2(low=muted("red"), high=muted("blue"), midpoint=0.7570) +
   theme(legend.position="top") +

### 2014-03-08 article2013.Rmd
Lähteet ja pieniä viilauksia
Tsekkasin tekstin sisäiset viitemerkinnät ja lähdeluettelon lehden asetusten mukaisesti. Lähdeluettelossa sulle vielä muutama lisättävä, ne on merkattu lähdeluetteloon.

@@
 
-As Joseph Schumpeter (1942: 12) once argued, "attitudes are coins that do not readily melt".
+As Joseph Schumpeter (1942, p.
+12) once argued, "attitudes are coins that do not readily melt".
 This means that normative expectations, justice beliefs or welfare perceptions are very hard to change.
-In this way, welfare attitudes can be associated with policy changes by highlighting public sentiments, which are normative background assumptions or ideas that constrain decision-making and institutional change by limiting the range of programme that decision-making elites are likely to perceive as acceptable and legitimate to both their constituents and themselves (Campbell, 2004: 96-100).
-On the other hand, "the worlds on welfare" literature on welfare attitudes has in turn emphasized that type of welfare policy affects public attitudes (Andress and Heien, 2001; Arts and Gelissen, 2001; Jaeger, 2006; Svallfors, 1997).
+In this way, welfare attitudes can be associated with policy changes by highlighting public sentiments, which are normative background assumptions or ideas that constrain decision-making and institutional change by limiting the range of programme that decision-making elites are likely to perceive as acceptable and legitimate to both their constituents and themselves (Campbell 2004, pp.
+96-100).
+On the other hand, "the worlds on welfare" literature on welfare attitudes has in turn emphasized that type of welfare policy affects public attitudes (Andress and Heien 2001; Arts and Gelissen 2001; Jaeger 2006; Svallfors 1997).
 In regard to public perceptions of the causes of poverty, in the marginal welfare-policy dominated liberal regime, the poor will be asked to fulfil much harder deservingness criteria than in a universal social democratic regime (van Oorschot 2006).
-In countries, in which welfare programmes are more marginal and less generous, explanations of poverty are more often based on individual morality, reflecting the strength of the dominant individualistic ideology, than they are in the Continental and Nordic welfare states – where external and structural explanations prevail (Albrekt Larsen, 2006: 47-56; Alesina and Gleaser, 2004).
+In countries, in which welfare programmes are more marginal and less generous, explanations of poverty are more often based on individual morality, reflecting the strength of the dominant individualistic ideology, than they are in the Continental and Nordic welfare states – where external and structural explanations prevail (Albrekt Larsen 2006, pp.
+47-56; Alesina and Gleaser 2004).
 
@@
 Since the early 1990s, the development of post-socialist welfare state has received abundant scholarly attention (e.g.
-Cerami and Vanhuysse, 2009, Inglot, 2008, Haggard and Kauffman, 2008; Cook, 2007; Aidukaite, 2004; Deacon 1992).
+Cerami and Vanhuysse 2009; Inglot 2008; Haggard and Kauffman 2008; Cook 2007; Aidukaite 2004; Deacon 1992).
 As the decades under socialist regime have had an undeniable influence on societies, these studies have emphasized the role of institutional differences within region as well as pre-socialist historical origins of the welfare states.
-Now after 20 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka, 2009, Tomka, 2006, Cerami and Vanhuysse, 2009, Inglot, 2008).
+Now after 20 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka 2009; Tomka 2006; Cerami and Vanhuysse 2009; Inglot 2008).
 
 The purpose of this article is to examine cross-national differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries.
-Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and Eastern European countries (Gallie and Paugam, 2002; Kallio and Niemelä, 2013; Lepianka, 2007; van Oorschot and Halman, 2000).
+Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and Eastern European countries (Gallie and Paugam 2002; Kallio and Niemelä 2014; Lepianka 2007; van Oorschot and Halman 2000).
 East Europeans are more inclined to blame either the individual behavior or the flaws of the economic system of country than their Western counterparts.
@@
 In addition, the study applies multilevel techniques in order to analyze the role of individual and contextual level explanatory factors as determinants of poverty attributions.
-The article explores to what extent specific risks related to household financial circumstances at the individual level and the state of democracy, income inequality and changes in national economy at the contextual level are associated with different types of lay explanations of poverty.
+The article explores to what extent specific risks related to household financial circumstances at the individual-level and the state of democracy, income inequality and changes in national economy at the contextual level are associated with different types of lay explanations of poverty.
 Finally, extensive country-selection of our data provides a novel opportunity to examine the effects of the financial crisis both at the individual and contextual level and their association with attributions for poverty.
-At the individual level, the study focuses not only on the individuals’ current financial situation.
+At the individual-level, the study focuses not only on the individuals’ current financial situation.
 Instead, the article provides an empirical example of how past experiences, current financial position as well as future expectations concerning individual’s economic situation are associated with perceptions of the causes of poverty.
@@
 Later, many factor analytic studies have given empirical support to Feagin’s categorization (e.g.
-Feather, 1974; Furnham, 1982; Hunt, 1996; Morçöl, 1997; Nasser et al., 2002; Niemelä, 2008).
+Feather 1974; Furnham 1982; Hunt 1996; Morçöl 1997; Nasser et al.
+2002; Niemelä 2008).
 Yet some studies have expanded our understanding by incorporating more contemporary beliefs into the attributional scales.
-These works emphasize especially the relevance of cultural attributions for poverty (Bullock et al., 2003; Cozzarelli et al., 2001).
+These works emphasize especially the relevance of cultural attributions for poverty (Bullock et al.
+2003; Cozzarelli et al.
+2001).
 
@@
 Western Europeans are more inclined to support social explanations for poverty than their Eastern counterparts.
-On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä, 2013; Kreidl, 2000; Lepianka, 2007; Lepianka et al., 2010; van Oorschot and Halman, 2000).
+On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä 2014; Kreidl 2000; Lepianka 2007; Lepianka et al.
+2010; van Oorschot and Halman 2000).
 Earlier results suggest, however, that there are variations both within and across groups of countries.
-Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä, 2013).
+Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä 2014).
 Also Lepianka et al.
@@
 They showed that individual fate type of explanation is supported relatively less frequently in the East of Europe than in Western European countries.
-Regarding the variation within the Eastern European countries, they found that in some countries individual blame is strongly endorsed: especially the citizens of Czech Republic attribute poverty in internal terms (see also Lepianka, 2007).
+Regarding the variation within the Eastern European countries, they found that in some countries individual blame is strongly endorsed: especially the citizens of Czech Republic attribute poverty in internal terms (see also Lepianka 2007).
 On the other hand, there are countries like Lithuania, Poland, Romania and Croatia in which social blame type of explanation is strongly supported.
@@
 Estonians are as likely to blame the flaws of the economic system as the individualistic reasons.
-In Russia, on the other hand, the economic system is blamed even more than the individuals themselves (Gorshkov and Tikhonova, 2006; Stephenson, 2000).
-These results are consistent with Lepianka's (2007: 127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia as well as with Habibov's (2011) comparative evidence from the Caucasus, Central Asia, Russia and Ukraine.
+In Russia, on the other hand, the economic system is blamed even more than the individuals themselves (Gorshkov and Tikhonova 2006; Stephenson 2000).
+These results are consistent with Lepianka's (2007, pp.
+127–135) comparative analysis of Hungary, Russia, Czech Republic and Estonia as well as with Habibov's (2011) comparative evidence from the Caucasus, Central Asia, Russia and Ukraine.
 Interesting result is that social blame coincides strongly with individual blame.
-This illustrates the split-consciousness of public perceptions, which means the coexistence of both dominant and potentially challenging beliefs, where the former represents dominant stratification ideology or value structure of a given country and the latter everyday stratification-related experience of an individual (Kluegel et al., 1995).
+This illustrates the split-consciousness of public perceptions, which means the coexistence of both dominant and potentially challenging beliefs, where the former represents dominant stratification ideology or value structure of a given country and the latter everyday stratification-related experience of an individual (Kluegel et al.
+1995).
 
 Based on qualitative poverty studies in Central Asia, the Caucasus, Ukraine and Moldova, Dudwick et al.
-(2003: 23) states that, generally, people blamed their own poverty on the failure of the Soviet state and the corruption, indifference, and incompetence of their new leaders.
+(2003, p.
+23) states that, generally, people blamed their own poverty on the failure of the Soviet state and the corruption, indifference, and incompetence of their new leaders.
 At the same time, however, they often attributed the poverty of others to individual failure, such as laziness, alcoholism, having too many children to support, or having too few children to provide for their old age.
 Emphazising the positive association between horizontal and vertical trust (see e.g.
-Rothstein, 2000), they also reported that many of the respondents had become very distrustful of government officials and many suspected that government officials of exploiting the political and economic turmoil for their own gain, at the expense of ordinary citizens.
+Rothstein 2000), they also reported that many of the respondents had become very distrustful of government officials and many suspected that government officials of exploiting the political and economic turmoil for their own gain, at the expense of ordinary citizens.
 With extensive country selection, the data utilized in this study provides a marvellous possibility to examine the association between attributions for poverty and the context regarding political process, civil liberties and political rights.
@@
 Hence, by analyzing the role of changes in economic performance during the period of global financial crisis, the research design of this article provides empirical evidence on the effects of crisis to public perceptions of the causes of poverty.
-Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune, 2007).
-In regard to explanations of poverty, Gallie and Paugam (2002: 21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved.
-This result is confirmed by Kallio and Niemelä (2013).
+Prior studies have shown that increased financial strain is associated with stronger support for state responsibility for economic provision (Blekesaune 2007).
+In regard to explanations of poverty, Gallie and Paugam (2002, pp.
+21-24) found that in the majority of European countries, there was a marked decline of support for the individual blame explanation between 1989 and 1993 as economic conditions deteriorated, followed  by an increase between 1993 and 2001 when economic conditions improved.
+This result is confirmed by Kallio and Niemelä (2014).
 They found that the larger the economic growth in a country is, more likely people are to blame the poor themselves.
@@
 In addition to macro-level economic performance, living conditions in general are found to be linked with attributions for poverty.
-People living in countries with high unemployment are more likely to endorse social than individualistic explanations (Albrekt Larsen, 2006: 74-79).
+People living in countries with high unemployment are more likely to endorse social than individualistic explanations (Albrekt Larsen 2006, pp.
+74-79).
 In addition, the increased incidence of poverty reduces the likelihood of choosing individual laziness as an explanation of poverty (Lepianka, 2007: 118-121).
 However, there are to some extent contradictory findings as well.
-While Kallio and Niemelä (2013) did not find association between individual blame type of explanation and incidence of poverty, Lepianka et al.
+While Kallio and Niemelä (2014) did not find association between individual blame type of explanation and incidence of poverty, Lepianka et al.
 (2010) found out that people living in more developed countries tend to attribute poverty rather to individual and social fate than blame types of explanations.
@@
 Consequently, by comparing for instance poorer countries of Central Asia with more well-off Slovenia and Czech Republic, this study provides a possibility to analyze whether the association between attributions for poverty and living conditions is also stronger in the post-socialist countries.
-This study examines human development index (HDI) as a proxy for living conditions in a given country.
+This study examines Human Development Index (HDI) as a proxy for living conditions in a given country.
 Given the great differences in living conditions within the post-socialist space, we assume that in countries of low level of HDI people tend to attribute poverty more with social than individual reasons.
@@
 
-In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalized by the individuals in the society in question (Taylor-Gooby, 1985; Svallfors, 1995).
-More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice, 2001) and resource/class-based risks (Cusack et al., 2006; Blomberg et al., 2012; Finseraas and Ringdal, 2012) as key to welfare attitudes.
+In regard to individual-level determinants, popular attitudes toward the welfare state have traditionally been assumed to be dependent on long-term class-related interest, short-term self-interest and the values and norms that have been internalized by the individuals in the society in question (Taylor-Gooby 1985; Svallfors 2012).
+More recent directions of comparative literature on welfare attitudes tend to also focus on the distribution of marketable capacities such as skill specificity (Iversen and Soskice 2001) and resource/class-based risks (Cusack et al.
+2006; Blomberg et al.
+2012; Finseraas and Ringdal 2012) as key to welfare attitudes.
 Thus, studies have assumed – and to some extent proved – that groups that have a greater risk of facing social problems or being dependent on welfare state benefits might perceive the government’s role in welfare issues in a distinct way compared to those who are not exposed to these risks.
@@
 The previous studies on individual-level determinants of poverty perceptions have focused on individual’s socio-demographic characteristics (e.g.
-Bullock, 1999; Cozzarelli et al., 2001; Furnham, 1982; Morcöl, 1997; Niemelä, 2008).
+Bullock 1999; Cozzarelli et al.
+2001; Furnham 1982; Morçöl 1997; Niemelä 2008).
 Results have, however, shown that socio-demographic variables explain very little and they do not reveal clear or consistent patterns.
-On the other hand, there are studies which support the self-interest hypothesis that those who perceive themselves to be poor are more likely to blame society and social structures for poverty than people well above the poverty line (Niemelä, 2008; Saunders, 2002: 155–156).
+On the other hand, there are studies which support the self-interest hypothesis that those who perceive themselves to be poor are more likely to blame society and social structures for poverty than people well above the poverty line (Niemelä 2008; Saunders 2003).
 Yet, there are also contradictory findings indicating that personal experience of poverty might lead to individualistic explanations.
-For instance, claimants of public welfare seem to hold negative views about other recipients (Bullock, 1999; 2004; Golding and Middleton, 1982: 178).
+For instance, claimants of public welfare seem to hold negative views about other recipients (Bullock 1999; 2004; Golding and Middleton 1982, p.
+178).
 
 Studies which have explored the importance of individual’s economic situation on attributions for poverty have focused on individual’s cross-sectional situation of financial circumstances.
-However, as suggested in the literature on welfare attitudes, not only the risk patterns but also the way people perceive their future risks might have an effect on attitudes toward the welfare state (Andress and Heien, 2001; Boeri et al., 2001; Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).
+However, as suggested in the literature on welfare attitudes, not only the risk patterns but also the way people perceive their future risks might have an effect on attitudes toward the welfare state (Andress and Heien 2001; Boeri et al.
+2001; Blomberg et al.
+2012; Nygaard Andersen and Ringdal 2012).
 In their European comparison Blomberg et al.
@@
 Not only the current situation or the perceived future but also the past experiences might be associated with individual’s evaluations (e.g.
-Kumlin, 2007).
+Kumlin 2007).
 In regard to attributions for poverty, the effects of past experiences are – according to our knowledge – totally unexplored area of research.
-In this article we will analyse past experiences by two ways.
+In this article we will analyze past experiences by two ways.
 First, we are interested in whether those whose current economic situation is lower than four years ago perceive the causes of poverty distinct way compared to those whose economic situation has improved or remained in the same level.
@@
 In addition to economic situation, other indicators which measure a greater than average risk of facing more permanent social problems, are "transfer dependency" and low education.
-Transfer dependency refers to a situation that the respondent’s main source of income relies on private or social transfers (Blomberg et al., 2012; Fridberg and Kangas, 2008).
+Transfer dependency refers to a situation that the respondent’s main source of income relies on private or social transfers (Blomberg et al.
+2012; Fridberg and Kangas 2008).
 Because of their close relation with welfare system, it can be assumed that welfare recipients perceive more likely social blame than indivual blame type of explanation of poverty compared to the rest of the population.
@@
 Prior findings do not reveal consistent patterns regarding the effects of education.
-On the one hand, some studies have found the so-called inverted U-relationship which means that people with a middle level of education support the individualistic explanation, whereas people with a low or a high level of education are more likely to endorse structural explanations emphazising social inequalities and the flaws of the economic system of a given country (Feagin, 1972; cf.
-Furnham, 1982).
+On the one hand, some studies have found the so-called inverted U-relationship which means that people with a middle level of education support the individualistic explanation, whereas people with a low or a high level of education are more likely to endorse structural explanations emphazising social inequalities and the flaws of the economic system of a given country (Feagin 1972; cf.
+Furnham 1982).
 On the other hand, there are studies which suggest that people with a lower level of education are more likely to favour individualistic explanations.
-Moreover, when the educational level increases, the likelihood of perceiving causes of poverty in individualistic terms decreases (Niemelä, 2008).
+Moreover, when the educational level increases, the likelihood of perceiving causes of poverty in individualistic terms decreases (Niemelä 2008).
 It is therefore difficult to form any solid hypotheses as to whether low education as a social risk is associated with perceptions of the causes of poverty in a similar vein than other risk factors discussed above.
@@
 In addition, most of the prior research is based on attitude surveys from the 1990s and most of them have restricted country selection.
-Therefore, as previous studies have suggested (Kallio and Niemelä, 2013; Lepianka et al., 2010), there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space.
+Therefore, as previous studies (Kallio and Niemelä 2014; Lepianka et al.
+2010) have suggested, there is a need for detailed and more up-to-date analysis which would focus on the attributions for poverty within the post-socialist space.
 Hence, the general purpose of the study is to analyze public perceptions of the causes of poverty in Europe and Central Asia with special emphasis on post-socialist countries.
 
-The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank (WB) in 2010.
+The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank in 2010.
 The analyses cover 22 post-socialist countries and 4 Western European countries.
-The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see ERBD, 2011).
+The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see ERBD 2011).
 The survey had a special focus on how people’s lives were affected by the global economic crises and its aftermath.
@@
 
-The respondents were asked the following question which measures public perceptions of the causes of poverty: **In your opinion, what is the main reason why there are some people in need in our country today?**
+The respondents were asked the following question which measures public perceptions of the causes of poverty: In your opinion, what is the main reason why there are some people in need in our country today?
 
@@
 
-Transfer dependency               Social or private transfers as a main source of income                                                                                  lits
+Transfer dependency               Social or private transfers as a main source of income                                                                                  LiTS2
 
-Low education                     No or only compulsory level of education                                                                                                 lits
+Low education                     No or only compulsory level of education                                                                                                 LiTS2
 
-Perceived low income level        Perceived income level below the country median                                                                                       lits
+Perceived low income level        Perceived income level below the country median                                                                                       LiTS2
 
-Income compared to past           Perceived change in income compared to situation in four years ago                                                                    lits
+Income compared to past           Perceived change in income compared to situation in four years ago                                                                    LiTS2
 
-Income compared to future         Expected change in income over the next four years                                                                                    lits
+Income compared to future         Expected change in income over the next four years                                                                                    LiTS2
 
-Effect of financial crisis        Whether respondents perceives that their household has suffered the financial crisis great or fair amount                             lits
+Effect of financial crisis        Whether respondents perceives that their household has suffered the financial crisis great or fair amount                             LiTS2
 
@@
 Data for change in GDP is from World Bank's [World Development Indicators](http://data.worldbank.org/data-catalog/world-development-indicators), Human Development Index from [United Nations Development Programme](http://hdr.undp.org/en/data) and Voice and Accountability from the World Bank's [Worldwide Governance Indicators](http://info.worldbank.org/governance/wgi/index.aspx#home).
-All macro level indicators are obtained from [Quality of Government Institutes](http://www.qog.pol.gu.se/) data using [rQog](http://markuskainu.fi/rqog/)-package (Kainu, 2014).
+All macro level indicators are obtained from [Quality of Government Institutes](http://www.qog.pol.gu.se/) data using [rQog](http://markuskainu.fi/rqog/)-package (Kainu 2014).
 
 The empirical section starts by presenting the results of descriptive analyses.
-As we assume that individual level perceptions are dependent both on individual and country-level contextual factors and we have a hierarchical two level dataset, it is recommended to use multilevel regression analysis for this task (Snijders and Bosker, 2011) as multilevel regression analysis allows the estimation of country-level effects on individual level outcomes.
+As we assume that individual level perceptions are dependent both on individual and country-level contextual factors and we have a hierarchical two level dataset, it is recommended to use multilevel regression analysis for this task (Snijders and Bosker 2011) as multilevel regression analysis allows the estimation of country-level effects on individual level outcomes.
 The analysis is performed using glmer-function from lme4 package in R.
@@
 We are modelling the date in stepwise manner first running the empty models for both social blame (Table 3) and individual blame (Table 4) type of explanation.
-At second stage we include all individual-level predictors and in third stage we include only the macro-level predictors.
+At the second stage we include all individual-level predictors and in the third stage we include only the macro-level predictors.
 In final models we include both the individual and macro-level variables.
@@
 In regard to differences between the East and the West of Europe, results show that there are some differences.
-Yet, there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al., 2010; Kallio and Niemelä, 2010).
+Yet, there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al.
+2010; Kallio and Niemelä 2014).
 Clearest differences between the West and the East is the stronger support for social fate in Western Europe.
@@
 On the other hand, attributions for poverty in France and Italy seems to be quite similar with post-socialist countries - blame types of explanations are endorsed more likely than fate types of explanations.
-Finally, the figures of Great Britain are the example of selective welfare policy dominated liberal regime in which the individual blame type of explanations is strongly supported (see also Albrekt Larsen, 2006).
+Finally, the figures of Great Britain are the example of selective welfare policy dominated liberal regime in which the individual blame type of explanations is strongly supported (see also Albrekt Larsen 2006).
 
-Hence, the results are in line with our assumptions and prior research (Habibov, 2011; Lepianka et al., 2010) regarding the general support for different explanations of poverty.
+Hence, the results are in line with our assumptions and prior research (Habibov 2011; Lepianka et al.
+2010) regarding the general support for different explanations of poverty.
 Overall, the results indicate that for the most of the post-socialist countries two most popular explanations are individual and social blame types of explanations.Yet, the differences between social fate and individual blame types of explanations are in some countries very small.
@@
 The trend is very clear: the less country has suffered from the financial crisis the less there is support to social blame type of explanation (-.62) and more to individual blame (.71).
-Thus, the results are in line with previous studies which suggest positive correlation between the economic growth and support for the individualistic perception of the causes of poverty (Gallie and Paugam, 2002; Kallio and Niemelä, 2013).
+Thus, the results are in line with previous studies which suggest positive correlation between the economic growth and support for the individualistic perception of the causes of poverty (Gallie and Paugam 2002; Kallio and Niemelä 2014).
 In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries.
@@
 Western European countries, on the other hand, are more better-off in terms of human development.
-It seems that there is not any significant association between human development and blame types of explanations in Western European countries (also Kallio and Niemelä, 2013; Lepianka et al.
+It seems that there is not any significant association between human development and blame types of explanations among Western European countries (also Kallio and Niemelä 2014; Lepianka et al.
 2010).
@@
 
-Finally, there also large variation between countries in the level of goodness of government.
-Overall, the association between attributions for poverty and the level of good governance is substantial (.41 for social blame and -.47 for individual blame).
+Finally, there also large variation between countries in the level of democracy.
+Overall, the association between attributions for poverty and the level of democracy is substantial (.41 for social blame and -.47 for individual blame).
 Regarding post-socialist countries, however, the direction of association is also here opposite as expected.
@@
 On the other hand, individual blame type of explanation is supported more likely in countries with low degree of good governance.
-In regard to Western European countries, the variation in the goodness of government is much smaller than in post-socialist countries.
-Also the association between perceptions and the good governance differs.
-Unlike in post-socialist countries, the support for blame types of explanations seems to vanish when the level of good governance increases.
+In regard to Western European countries, the variation in democracy is much smaller than in post-socialist countries.
+Also it's association with perceptions differs.
+Unlike in post-socialist countries, the support for blame types of explanations seems to vanish when the level of democracy increases.
 
@@
 Descriptive analyses presented above would posit that there are rather clear, though to some extent unexpected, associations between country-level contextual characteristics and the perceptions of the causes of poverty.
-In order to explore the effects of both contextual and individual-level determinants in a more detailed statistical manner, the analytsis continues with multilevel logistic regressions (Tables 3, 4 and 5).
+In order to explore the effects of both contextual and individual-level determinants in a more detailed statistical manner, the analysis continues with multilevel logistic regressions (Tables 3, 4 and 5).
 
@@
 When a random predictor is added to model, we expect deviance to decrease, on average, by 1.
-When predictor is more informative deviance should decrease more than 1 (Gelman and Hill, 2007).
+When predictor is more informative deviance should decrease more than 1 (Gelman and Hill 2007).
 As for getting sensible interpretation of the regression coefficients in logistic regression we can divide the coefficients by 4 to get the probability of certain covariate.
@@
 
-Interesting enough, however, results show that unlike descriptive statistics indicated neither country-level living conditions (measured by Human Development Index) nor the level of goodness of government are significantly associated with attributions for poverty.
+Interesting enough, however, results show that unlike descriptive statistics indicated neither country-level living conditions (measured by Human Development Index) nor the level of democracy are significantly associated with attributions for poverty.
 In addition, final models in Tables 3 and 4 show that inclusion of both the individual and contextual level predictors to the models has only a minor impact on significance of the predictors.
@@
 Social blame is followed by individual blame, social fate and individual fate types of explanations.
-Hence, the results are in line with our assumptions and prior research (Habibov, 2011; Lepianka et al., 2010) regarding the general support for different explanations of poverty in the post-socialist space.
+Hence, the results are in line with our assumptions and prior research (Habibov 2011; Lepianka et al.
+2010) regarding the general support for different explanations of poverty in the post-socialist space.
 In regard to differences between the East and the West of Europe, results show that Western Europeans emphasize more social reasons for poverty.
@@
 
-Our other contextual level predictors, level of democracy and human development, were not significant in the models.
+Our other contextual level predictors, level of democracy and human development, were not significantly associated with public perceptions of the causes of poverty.
 However, the descriptive analysis showed that contracy to our assumptions, citizens in those countries where the state of democracy and human development is low do not blame more likely society than citizens in countries where they have political freedom and power to influence socio-political issues.
@@
 Findings indicate that those whose main source of income rely mainly on social or private transfers, those who perceive their income level to be lower than country’s median, those who perceive their future economic situation to be worse than now as well as those who perceive that the global financial crisis has affected negatively to their household’s economic situation endorse more likely external (social blame) than internal (individual blame) reasons for poverty.
-Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al., 2012; Nygaard Andersen and Ringdal, 2012).
+Thus, analyses showed that that it would be fruitful to include a perspective of risk groups when studying attitudes toward poverty, and also to include future expectations in the equation (also Blomberg et al.
+2012; Nygaard Andersen and Ringdal 2012).
 
@@
 For instance, another line of research have emphasized the role of other beliefs, attitudes and values rather than interest (e.g.
-Cozzarelli et al., 2001).
+Cozzarelli et al.
+2001).
 Treating the values, attitudes and other perceptions as determinants for attributions of poverty is, however, problematic because it is difficult to prove the causal relationship between attributions and other attitudes.
@@
 On the one hand, in order to examine the effects of contextual factors, comparative social research should try to utilize comparative data encompassing more diverse cases than, for instance, conventional EU17 or OECD countries.
-On the other hand, the well-known limitation of this kind of variable-oriented approach is that larger number of cases may pose also more obstacles to making interpretive statements about specific cases or even about categories of cases (Ragin, 1987).
+On the other hand, the well-known limitation of this kind of variable-oriented approach is that larger number of cases may pose also more obstacles to making interpretive statements about specific cases or even about categories of cases (Ragin 1987).
 For instance, in regard to policy implications of research on welfare attitudes, broadening the country selection from Western democracies (in some cases) to more authoritarian societies raises a question whether public attitudes have similar constraining role in policy-making that have been assumed and to some extent proved to be in the policy-making of Western democracies.
@@
 
-Aidukaite, Jolanta (2004) The Emergence of the Post-Socialist Welfare State - The Case of the Baltic States: Estonia, Latvia and Lithuania, Södertörn University College, Sweden.
+Aidukaite, J. (2004) The Emergence of the Post-Socialist Welfare State - The Case of the Baltic States: Estonia, Latvia and Lithuania (Huddinge, Södertörn University).
 
-Albrekt Larsen, Christian (2006) The Institutional Logic of Welfare Attitudes. How Welfare Regimes Influence Public Support. Hampshire: Ashgate.
+Albrekt Larsen, C. (2006) The Institutional Logic of Welfare Attitudes. How Welfare Regimes Influence Public Support (Hampshire, Ashgate).
 
-Alesina, Alberto and Glaeser, Edward L. (2004) Fighting Poverty in the US and Europe. A World of Difference. Oxford: Oxford University Press.
+Alesina, A. & Glaeser, E. L. (2004) Fighting Poverty in the US and Europe. A World of Difference (Oxford, Oxford University Press).
 
-Andress, Hans-Jürgen and Heien, Thorsten (2001) Four Worlds of Welfare State Attutudes? A Comparison of Germany, Norway, and the United States. European Sociological Review 17 (4), 337–356.
+Andress, H-J. & Heien, T. (2001) 'Four Worlds of Welfare State Attutudes? A Comparison of Germany, Norway, and the United States', European Sociological Review, 17, 4, 337–356.
 
-Arts, Wil and Gelissen, John (2001) Welfare States, Solidarity and Justice Principle: Does the Type Really Matter. Acta Sociologica 44 (4), 283–299.
+Arts, W. & Gelissen, J. (2001) 'Welfare States, Solidarity and Justice Principle: Does the Type Really Matter', Acta Sociologica, 44, 4, 283–299.
 
-Blomberg, Helena and Kallio, Johanna and Kangas, Olli and Kroll, Christian and Niemelä, Mikko (2012) Attitudes among High-Risk Groups.
-In Svallfors, Stefan (ed.) Contested Welfare States.
-Welfare Attitudes in Europe and Beyond.
-Stanford University Press: Stanford, CA, 58–80.
+Blekesaune, M. (2007) 'Economic Conditions and Public Attitudes to Welfare Policies', European Sociological Review, 23, 3, 393-403.
 
-Boeri, T., Börsch-Supan, A. and Tabellini, G. (2001) Would You Like to Shrink the Welfare State? A Survey of European Citizens. Economic Policy 16, 7–50.
+Blomberg, H., Kallio, J., Kangas, O., Kroll, C.
+& Niemelä, M.
+(2012) 'Attitudes Among High-Risk Groups', in Svallfors, S.
+(ed) Contested Welfare States.
+Welfare Attitudes in Europe and Beyond (Stanford, Stanford University Press).
 
-Bullock, Heather E. (1999) Attributions for Poverty: A Comparison of Middle-Class and Welfare Recipient Attitudes. Journal of Applied Social Psychology 29 (10), 2059–2082.
+Boeri, T., Börsch-Supan, A. & Tabellini, G. (2001) 'Would You Like to Shrink the Welfare State? A Survey of European Citizens', Economic Policy, 16, 7–50.
 
-Bullock, Heather E. (2004) From the Front Lines of Welfare Reform: An Analysis of Social Worker and Welfare Recipient Attitudes. The Journal of Social Psychology 144 (6), 571-588.
+Bullock, H. E. (1999) 'Attributions for Poverty: A Comparison of Middle-Class and Welfare Recipient Attitudes', Journal of Applied Social Psychology, 29, 10, 2059–2082.
 
-Bullock, Heather E.
-and Williams, Wendy R.
-and Limbert, Wendy M.
-(2003) Predicting Support for Welfare Policies: The Impact of Attributions and Beliefs about Inequality.
-Journal of Poverty 7 (3), 35–56.
+Bullock, H. E. (2004) 'From the Front Lines of Welfare Reform: An Analysis of Social Worker and Welfare Recipient Attitudes', The Journal of Social Psychology, 144, 6, 571-588.
 
-Campbell, John L. (2004) Institutional Change and Globalization. Princeton and Oxford: Princeton University Press.
+Bullock, H. E., Williams, W. R. & Limbert, W. M. (2003) 'Predicting Support for Welfare Policies: The Impact of Attributions and Beliefs about Inequality', Journal of Poverty, 7, 3, 35–56.
 
-Cerami, Aifio and Vanhuysse, Pieter (2009) Post-Communist Welfare Pathways. Theorizing Social Policy Transformations in Central and Eastern Europe. Hampshire: Palgrave Macmillan.
+Campbell, J. L. (2004) Institutional Change and Globalization (Princeton and Oxford, Princeton University Press).
 
-Cook, L. J. (2007) Postcommunist Welfare States: Reform Politics in Russia and Eastern Europe, Cornell University Press.
+Cerami, A. & Vanhuysse, P. (2009) Post-Communist Welfare Pathways. Theorizing Social Policy Transformations in Central and Eastern Europe (Hampshire, Palgrave Macmillan).
 
-Cozzarelli, Catherine and Wilkinson, Anna V. and Tagler, Michael J. (2001) Attitudes Toward the Poor and Attributions for Poverty. Journal of Social Issues 57 (2), 207-227.
+Cook, L. J. (2007) Postcommunist Welfare States: Reform Politics in Russia and Eastern Europe (Ithaca, Cornell University Press).
 
-Cusack, T., Iversen, T. and Rehm, P. (2006) Risks at Work: The Demand and Supply Sides of Government Redistribution. Oxford Review of Economic Policy 22, 365–389.
+Cozzarelli, C., Wilkinson, A. V. & Tagler, M. J. (2001) 'Attitudes Toward the Poor and Attributions for Poverty', Journal of Social Issues, 57, 2, 207-227.
 
-Dudwick, N., Gomart, E., and Marc, A. (2003) When Things Fall Apart: Qualitative Studies of Poverty in the Former Soviet Union. World Bank Publications.
+Cusack, T., Iversen, T. & Rehm, P. (2006) 'Risks at Work: The Demand and Supply Sides of Government Redistribution', Oxford Review of Economic Policy, 22, 365–389.
 
-Economicst Intelligence Unit (2001) Democracy index 2011. Democracy under stress. Downloaded at: \href{http://www.eiu.com/democracyindex2011}{www.eiu.com/democracyindex2011}
+DEACON 1992 MARKUS OBS
 
-Feagin, Joe R. (1972) Poverty: We Still Believe That God Helps Those Who Help Themselves. Psychology Today 6 (2), 101–129.
+Dudwick, N., Gomart, E., & Marc, A. (2003) When Things Fall Apart: Qualitative Studies of Poverty in the Former Soviet Union (New York, World Bank).
 
-Feagin, Joe R. (1975) Subordinating the Poor: Welfare and American Beliefs. NJ Prentice Hall: Upper Saddle River.
+ERBD 2011 MARKUS OBS
 
-Feather, Norman T. (1974) Explanations of Poverty in Australian and American Samples: The Person, Society, or Both? Australian Journal of Psychology 26 (3), 199–216.
+Feagin, J. R. (1972) 'Poverty: We Still Believe That God Helps Those Who Help Themselves', Psychology Today, 6, 2, 101–129.
 
-Fridberg, Torben and Kangas, Olli (2008) The Welfare State, Poverty and Social Exclusion.
-In Ervasti, H., Fridberg, T., Hjerm, M.
-and Ringdal, K.
-(eds.) Nordic Social Attitudes in a European Perspective.
-Edward Elgar: Cheltenhamn, 22–47.
+Feagin, J. R. (1975) Subordinating the Poor: Welfare and American Beliefs (NJ Prentice Hall, Upper Saddle River).
 
-Furnham, Adrian (1982) Why are the poor always with us? Explanations for poverty in Great Britain. British Journal of Social Psychology 21(3), 311–322.
+Feather, N. T. (1974) 'Explanations of Poverty in Australian and American Samples: The Person, Society, or Both?', Australian Journal of Psychology, 26, 3, 199–216.
 
-Gallie, Duncan and Paugam, Serge (2002) Social Precarity and Social Integration.
-Report for the European Commission Based on Eurobarometer 56.1.
-Brussels: European Commission Directorate-General Employment.
+Finseraas, H.
+& Ringdal, K.
+(2012) 'Economic Globalization, Personal Risks and the Demand for a Comprehensive Welfare State', in Ervasti, H., Goul Andersen, J., Fridberg, T.
+& Ringdal, K.
+(eds.) The Future of the Welfare State.
+Social Policy Attitudes and Social Capital in Europe (Cheltenham, Edward Elgar).
 
-Giczi, Johanna and Sik, Endre (2009) Trust and Social Capital in Europe. In TARKI European Social Report 2009. TARKI: Budapest, 63-81.
+Fridberg, T.
+& Kangas, O.
+(2008) 'The Welfare State, Poverty and Social Exclusion', in Ervasti, H., Fridberg, T., Hjerm, M.
+& Ringdal, K.
+(eds) Nordic Social Attitudes in a European Perspective (Cheltenham, Edward Elgar).
 
-Golding, Peter and Middleton, Sue (1982) Images of Welfare. Press and Public Attitudes to Poverty. Oxford: Basil Blackwell.
+Furnham, A. (1982) 'Why Are the Poor Always With Us? Explanations for Poverty in Great Britain', British Journal of Social Psychology, 21,3, 311–322.
 
-Gorshkov, Mikhail K. and Tikhonova, Natalia E. (2006) Wealth and Poverty in the Perceptions of Russians. Sociological Research 45 (1), 27–40.
+Gallie, D. & Paugam, S. (2002) Social Precarity and Social Integration. Report for the European Commission Based on Eurobarometer 56.1 (Brussels, European Commission Directorate-General Employment).
 
-Habibov, Nazim N.
-(2011) Public Beliefs Regarding the Causes of Poverty during Transition.
-Evidence from the Caucasus, Central Asia, Russia, and Ukraine.
-International Journal of Sociology and Social Policy 31 (1-2), 53-74.
+GELMAN & HILL 2007 MARKUS OBS
 
-Hunt, Matthew O. (1996) The Individual, Society, or Both? A Comparison of Black, Latino, and White Beliefs about the Causes of Poverty. Social Forces 75 (1), 293–322.
+Golding, P. & Middleton, S. (1982) Images of Welfare. Press and Public Attitudes to Poverty (Oxford, Basil Blackwell).
 
-Inglot, T. (2008). Welfare States in East Central Europe, 1919-2004 (1st ed.). Cambridge University Press.
+Gorshkov, M. K. & Tikhonova, N. E. (2006) 'Wealth and Poverty in the Perceptions of Russians', Sociological Research, 45, 1, 27–40.
 
-Iversen, T. and Soskice, D. (2001) An Asset Theory of Social Policy Preferences. American Political Science Review 95, 875–893.
+Habibov, N.
+N.
+(2011) 'Public Beliefs Regarding the Causes of Poverty during Transition.
+Evidence from the Caucasus, Central Asia, Russia, and Ukraine', International Journal of Sociology and Social Policy, 31, 1-2, 53-74.
 
-Jaeger, Mads Meier (2006) Welfare Regime and Attitudes Towards Redistribution: The Regime Hypothesis Revisited. European Sociological Review 22 (2), 157–170.
+Haggard, S. & Kaufman, R. R. (2008) Development, Democracy, and Welfare States: Latin America, East Asia, and Eastern Europe (Princeton, Princeton University Press).
 
-Kainu, Markus (2014) rQog. Download data from the Quality of Government Institute data in R. R-package available at: [github.com/muuankarski/rQog](https://github.com/muuankarski/rQog)
+Hunt, M. O. (1996) 'The Individual, Society, or Both? A Comparison of Black, Latino, and White Beliefs about the Causes of Poverty', Social Forces, 75, 1, 293–322.
 
-Kallio, Johanna and Niemelä, Mikko (2010) Public Perceptions of the Causes of Poverty: Multilevel Analysis of 26 European Nations. Unpublished manuscript.
+Inglot, T. (2008) Welfare States in East Central Europe, 1919-2004 (Cambridge, Cambridge University Press).
 
-Haggard, S. and Kaufman, R. R. (2008) Development, Democracy, and Welfare States: Latin America, East Asia, and Eastern Europe, Princeton Univ Pr.
+Iversen, T. & Soskice, D. (2001) 'An Asset Theory of Social Policy Preferences', American Political Science Review, 95, 875–893.
 
-Kluegel, James R. and Mason, David S. and Wegener, Bernd (eds) (1995) Social Justice and Political Change: Public Opinion in Capitalist and Post-communist States, New York: Aldine De Gruyter.
+Jaeger, M. M. (2006) 'Welfare Regime and Attitudes Towards Redistribution: The Regime Hypothesis Revisited', European Sociological Review, 22, 2, 157–170.
 
-Kouvo, Antti and Kankainen, Tomi and Niemelä, Mikko (2012) Welfare Benefits and Generalized Trust in Finland and Europe.
-In Ervasti, Heikki and Goul Andersen Jörgen and Fridberg, Torben and Ringdal, Kristen (eds.) The Future of the Welfare State.
-Social Policy Attitudes and Social Capital in Europe.
-Edward Elgar: Cheltenham, 195-213.
+Kainu, M. (2014) rQog. Download data from the Quality of Government Institute data in R. R-package available at: [github.com/muuankarski/rQog](https://github.com/muuankarski/rQog)
 
-Kreidl, Martin (2000) Perceptions of Poverty and Wealth in Western and Post-Communist Countries. Social Justice Research 13 (2), 151–176.
+Kallio, J. & Niemelä, M. (2014) 'Who Blames the Poor? Multilevel Evidence of Support for and Determinants of Individualistic Explanation of Poverty in Europe', European Societies, 16, 1, 112-135.
 
-Kumlin, Staffan (2007) Overload or Undermined?
-European Welfare States in the Face of Performance Dissatisfaction.
-In Svallfors, Stefan (ed.) The Political Sociology of the Welfare State: Institutions, Social Cleavages, and Orientations.
-Stanford University Press: Stanford, CA, 80–116.
+Kluegel, J. R., Mason, D. S. & Wegener, B. (eds) (1995) Social Justice and Political Change: Public Opinion in Capitalist and Post-communist States (New York, Aldine De Gruyter).
 
-Lepianka, Dorota (2007) Are the Poor to be Blamed or Pitied? A Comparative Study of Popular Attributions in Europe. Tilburg: Tilburg University.
+Kreidl, M. (2000) 'Perceptions of Poverty and Wealth in Western and Post-Communist Countries', Social Justice Research, 13, 2, 151–176.
 
-Lepianka, Dorota and Gelissen, John and van Oorschot, Wim (2010) Popular Explanations of Poverty in Europe.
-Effects of Contextual and Individual Characteristics across 28 European Countries.
-Acta Sociologica 53 (1), 53–72.
+Kumlin, S.
+(2007) 'Overload or Undermined?
+European Welfare States in the Face of Performance Dissatisfaction', in Svallfors, S.
+(ed.) The Political Sociology of the Welfare State: Institutions, Social Cleavages, and Orientations (Stanford, Stanford University Press).
 
-Morçöl, Göktuğ (1997) Lay Explanations for Poverty in Turkey and Their Determinants. Journal of Social Psychology 137 (6) 728–738.
+Lepianka, D. (2007) Are the Poor to be Blamed or Pitied? A Comparative Study of Popular Attributions in Europe (Tilburg, Tilburg University).
 
-Murthi, M., & Tiongson, E. R. (2008). Attitudes to Equality: The Socialist Legacy Revisited. SSRN eLibrary. Retrieved from http://papers.ssrn.com/sol3/papers.cfm?abstract_id=1098627
+Lepianka, D., Gelissen, J.
+& van Oorschot, W.
+(2010) 'Popular Explanations of Poverty in Europe.
+Effects of Contextual and Individual Characteristics across 28 European Countries', Acta Sociologica, 53, 1, 53–72.
 
-Nasser, Ramzi and Abouchedid, Kamal and Khashan, Hilal (2002) Perceptions of the Causes of Poverty Comparing Three National Groups: Lebanon, Portugal, and South Africa.
-Current Research in Social Psychology 8 (7), 101–119.
+Morçöl, G. (1997) 'Lay Explanations for Poverty in Turkey and Their Determinants', Journal of Social Psychology, 137, 6, 728–738.
 
-Niemelä, Mikko (2008) Perceptions of the Causes of Poverty in Finland. Acta Sociologica, 51 (1), 23–40.
+Nasser, R., Abouchedid, K.
+& Khashan, H.
+(2002) 'Perceptions of the Causes of Poverty Comparing Three National Groups: Lebanon, Portugal, and South Africa', Current Research in Social Psychology, 8, 7, 101–119.
 
-Nilson, Lina B. (1981) Reconsidering Ideological Lines: Beliefs about Poverty in America. The Sociological Quarterly 22 (4), 531-548.
+Niemelä, M. (2008) 'Perceptions of the Causes of Poverty in Finland', Acta Sociologica, 51, 1, 23–40.
 
-Ragin, Charles C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies. Berkeley: University of Califor
+Nygaard Andersen, S.
+& Ringdal, K.
+(2012) 'Welfare Regimes and Personal Risks', in Ervasti, H., Goul Andersen, J., Fridberg, T.
+& Ringdal, K.
+(eds.) The Future of the Welfare State.
+Social Policy Attitudes and Social Capital in Europe (Cheltenham, Edward Elgar).
+
+Ragin, C. C. (1987) The Comparative Method. Moving Beyond Qualitative and Quantitative Strategies (Berkeley, University of California).
+
+Rothstein, B. (2000) 'Trust, Social Dilemmas and Collective Memories', Journal of Theoretical Politics, 12, 4, 477-501.
+
+Saunders, P. (2003) 'Stability and Change in Community Perceptions of Poverty: Evidence from Australia', Journal of Poverty, 7, 4, 1-20.
+
+Schumpeter, J. A. (1942) Capitalism, Socialism and Democracy (New York, Routledge).
+
+SNIJDERS & BOSKER 2011 MARKUS OBS
+
+Stephenson, S. (2000) 'Public Beliefs in the Causes of Wealth and Poverty and Legitimization of Inequalities in Russia and Estonia', Social Justice Research, 13, 2, 83-100.
+
+Svallfors, S. (1997) 'Worlds of Welfare and Attitudes to Redistribution: a Comparison of Eight Western Nations', European Sociological Review, 13, 3, 283-304.
+
+Svallfors, S. (2012) 'Welfare States and Welfare Attitudes', in Svallfors, S. (ed) Contested Welfare States. Welfare Attitudes in Europe and Beyond (Stanford, Stanford University Press).
+
+SZIKRA & TOMKA 2009 MARKUS OBS
+
+Taylor-Gooby, P. (1985) Public Opinion, Ideology, and State Welfare (London, Routledge & Kegan Paul).
+
+TOMKA 2006 MARKUS OBS
+
+van Oorschot, W. (2006) 'Making the Difference in Social Europe: Deservingness Perceptions Among Citizens of European Welfare States', Journal of European Social Policy, 16, 1, 23-42.
+
+van Oorschot, W. & Halman, L. (2000) 'Blame or Fate, Individual or Social?', European Societies, 2, 1, 1-28.
 

### 2014-03-27 article2013.Rmd
eastern europe viittaukset vaihdettu korrektimpiin. samoin lahteet lisatty. jatin kuvioihin viela Central Eastern Europe ryhmittelyn. eikohan se siina kelpaa. terveiset Moskovan junasta Lahden ja Kouvolan valilta. Just puski hytti tayteen venalaisia votkakassit heiluen. ihmettelevat hamletia ylahyllylla. jatkan slaideista huomenna.

@@
 affiliation: Finland
-date: March 3, 2014
+date: March 27, 2014
 header-includes:
@@
 As the decades under socialist regime have had an undeniable influence on societies, these studies have emphasized the role of institutional differences within region as well as pre-socialist historical origins of the welfare states.
-Now after 20 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka 2009; Tomka 2006; Cerami and Vanhuysse 2009; Inglot 2008).
+Now after 25 years of transition, as there has not been a visible convergence towards a common "post-socialist model", the attention has shifted even more on various institutional and historical explanations (Szikra and Tomka 2009; Tomka 2006; Cerami and Vanhuysse 2009; Inglot 2008).
 
 The purpose of this article is to examine cross-national differences in attributions for poverty in Europe and Western Asia with special emphasis on post-socialist countries.
-Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and Eastern European countries (Gallie and Paugam 2002; Kallio and Niemelä 2014; Lepianka 2007; van Oorschot and Halman 2000).
-East Europeans are more inclined to blame either the individual behavior or the flaws of the economic system of country than their Western counterparts.
+Prior findings of the comparative studies on public perceptions of the causes of poverty have emphasized clear differences between Western and post-socialist European countries (Gallie and Paugam 2002; Kallio and Niemelä 2014; Lepianka 2007; van Oorschot and Halman 2000).
+People in post-potsocialist Europe are more inclined to blame either the individual behavior or the flaws of the economic system of country than their Western counterparts.
 Hence, it seems that unlike in Western Europe traditionally opposite attitudinal dimensions, i.e.
-structural and individualistic explanations of poverty coincide with each other in East Europeans' perceptions.
-However, there is shortage of comparative studies on attributions for poverty in particular focusing on the East Central European countries or on the whole post-socialist space.
+structural and individualistic explanations of poverty coincide with each other in post-socialist Europeans' perceptions.
+However, there is shortage of comparative studies on attributions for poverty in particular focusing on the post-socialist European countries or on the whole post-socialist space.
 There is therefore a need for a more detailed and up-to-date analysis of the public perceptions of the causes of poverty in post-socialist countries.
@@
 The novelty of this study is that it covers wide range of countries from Europe and Western Asia.
-Thus, compared to previous studies, it provides more valid evidence on the differences in perceptions between the East and the West and within post-socialist regime as well as on the importance of the post-socialist legacy.
+Thus, compared to previous studies, it provides more valid evidence on the differences in perceptions between the post-socialist countries and the West and within post-socialist space.
 In addition, the study applies multilevel techniques in order to analyze the role of individual and contextual level explanatory factors as determinants of poverty attributions.
@@
 
-Prior empirical findings on attributions for poverty in Europe have shown that there are some differences between the West and the East of Europe.
-Western Europeans are more inclined to support social explanations for poverty than their Eastern counterparts.
-On the other hand, individual blame type of explanation is generally more popular in the East than in the West of Europe (Kallio and Niemelä 2014; Kreidl 2000; Lepianka 2007; Lepianka et al.
+Prior empirical findings on attributions for poverty in Europe have shown that there are some differences between the Western and the post-socialist Europe.
+Western Europeans are more inclined to support social explanations for poverty than their post-socialist Europeans counterparts.
+On the other hand, individual blame type of explanation is generally more popular in the post-socialist Europe than in the West of Europe (Kallio and Niemelä 2014; Kreidl 2000; Lepianka 2007; Lepianka et al.
 2010; van Oorschot and Halman 2000).
 Earlier results suggest, however, that there are variations both within and across groups of countries.
-Interestingly enough, the variation is highest within the East Central European countries (Kallio and Niemelä 2014).
+Interestingly enough, the variation is highest within the post-socialist European countries (Kallio and Niemelä 2014).
 Also Lepianka et al.
 (2010) come to the conclusion that further investigation is required due to rather large variation across countries.
-They showed that individual fate type of explanation is supported relatively less frequently in the East of Europe than in Western European countries.
-Regarding the variation within the Eastern European countries, they found that in some countries individual blame is strongly endorsed: especially the citizens of Czech Republic attribute poverty in internal terms (see also Lepianka 2007).
+They showed that individual fate type of explanation is supported relatively less frequently in the post-socialist Europe than in Western European countries.
+Regarding the variation within the post-socialist European countries, they found that in some countries individual blame is strongly endorsed: especially the citizens of Czech Republic attribute poverty in internal terms (see also Lepianka 2007).
 On the other hand, there are countries like Lithuania, Poland, Romania and Croatia in which social blame type of explanation is strongly supported.
@@
 This is particularly interesting in these times when the Western world is meeting the increasing economic turbulence.
-In regard to East-West differences as well as variation within post-socialist countries, countries examined in this study have faced global economic crisis in distinct ways.
-While many Western and Eastern European countries have witnessed economic decline, the economic performance in Western Asia has been strong.
+In regard to differences between post-socialist countries and Western Europe as well as variation within post-socialist countries, countries examined in this study have faced global economic crisis in distinct ways.
+While many Western and post-socialist European countries have witnessed economic decline, the economic performance in Western Asia has been strong.
 Hence, by analyzing the role of changes in economic performance during the period of global financial crisis, the research design of this article provides empirical evidence on the effects of crisis to public perceptions of the causes of poverty.
@@
 
-Earlier studies have shown that attributions for poverty differ in general between the West and the East of Europe but the variation between East Central European countries is relatively large.
+Earlier studies have shown that attributions for poverty differ in general between the Western and the post-socialist Europe but the variation between post-socialist European countries is relatively large.
 There is, however, a shortage of comparative studies which have focused particularly on differences within post-socialist countries.
@@
 The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank in 2010.
-The analyses cover 22 post-socialist countries and 4 Western European countries.
-The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see ERBD 2011).
+The analyses cover 22 post-socialist countries and 5 Western European countries.
+The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see EBRD 2011).
 The survey had a special focus on how people’s lives were affected by the global economic crises and its aftermath.
@@
 
-In regard to differences between the East and the West of Europe, results show that there are some differences.
+In regard to differences between the Western and the post-socialist Europe, results show that there are some differences.
 Yet, there are also substantial variation within Western Europe, which correspond earlier findings on cross-national differences in Western Europe (Lepianka et al.
 2010; Kallio and Niemelä 2014).
-Clearest differences between the West and the East is the stronger support for social fate in Western Europe.
+Clearest differences between the Western and the post-socialist Europe is the stronger support for social fate in Western Europe.
 In Germany and Sweden external reasons for poverty receive the majority of support.
@@
 Thus, the results are in line with previous studies which suggest positive correlation between the economic growth and support for the individualistic perception of the causes of poverty (Gallie and Paugam 2002; Kallio and Niemelä 2014).
-In regard to East-West differences, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries.
+In regard to differences between Western Europe and the post-socialist countries, we can see that the variation in economic situation is much smaller in Western European countries, and the association between economic change and attributions for poverty is opposite to the post-socialist countries.
 However, this is greatly due to economic decline in Great Britain.
@@
 2010) regarding the general support for different explanations of poverty in the post-socialist space.
-In regard to differences between the East and the West of Europe, results show that Western Europeans emphasize more social reasons for poverty.
+In regard to differences between the Western and the post-socialist Europe, results show that Western Europeans emphasize more social reasons for poverty.
 
@@
 
-As a methodological contribution, findings showed interesting differences between the East and the West regarding the effects of contextual variables.
+As a methodological contribution, findings showed interesting differences between the Western Europe and the post-socialist countries regarding the effects of contextual variables.
 Even though the selection of Western European countries in this study included countries from different welfare regimes, the differences between countries in economic change, democracy and living conditions are quite modest compared especially to CIS countries.
@@
 
-DEACON 1992 MARKUS OBS
+Deacon, B., Castle-Kanerova, M., Manning, N., Millard, F., Orosz, E., Szalai, J.
+& Vidinova, A.
+(1992) 'The New Eastern Europe: social policy past, present and future', Sage Publications London/Newbury Park.
 
@@
 
-ERBD 2011 MARKUS OBS
+EBRD (2011) 'Transition Report 2011. Crisis and Transition: The People's Perspective', European Bank of Reconstruction and Development.
 
@@
 
-GELMAN & HILL 2007 MARKUS OBS
+Gelman, A. & Hill, J. (2007) 'Data Analysis Using Regression And Multilevel/Hierarchical Models', Cambridge University Press.
 
@@
 
-Schumpeter, J. A. (1942) Capitalism, Socialism and Democracy (New York, Routledge).
+Schumpeter, J. A. (1942) 'Capitalism, Socialism and Democracy'. (New York, Routledge).
 
-SNIJDERS & BOSKER 2011 MARKUS OBS
+Snijders, T. A. B. & Bosker, R. (2011) 'Multilevel Analysis: An Introduction to Basic and Advanced Multilevel Modeling', Second Edition. Sage Publications Ltd.
 
@@
 
-SZIKRA & TOMKA 2009 MARKUS OBS
+Szikra, D.
+& Tomka B.
+(2009) 'Social Policy in East Central Europe: Major Trends in the Twentieht Century', In Cerami, A, ja Vanhuysse, P.
+(2009) 'Post-Communist Welfare Pathways: Theorizing Social Policy Transformations in Central and Eastern Europe', Palgrave Macmillan,
 
@@
 
-TOMKA 2006 MARKUS OBS
+Tomka, B. (2006) 'East Central Europe and the European Social Policy Model: A Long-Term View'. East European Quarterly 40, nro 2: 135.
 
@@
        #dcolumn=TRUE,
-       scriptsize=TRUE
+       scriptsize=TRUE,
+        custom.model.names = model.names.multi.2,
+        custom.coef.names = coef.names.multi,
+       include.pvalues=TRUE,
+        caption="Logistic multilevel random intercept model comparing the full models for both dependent variable with model predicting social blame against individual blame", 
+        caption.above=TRUE)
+```

### 2014-03-28 index.Rmd
added tables to presentation

--- 2013-12-01 index.Rmd
@@
 
+### Misc analysis
+
+| html-ouput | source code |
+| ------------ | -------------- |
+| [Comparing waves 2006 and 2011](compareWaves.html) | [source](compareWaves.Rmd) |
+| [Exploring the macro level correlations](correlations.html) | [source](correlations.md) |
+
 ## Modelling the data
@@
 
-| pdf-ouput | source code |
-| ------------ | -------------- |
-| [Article draft in .pdf](article2013.pdf) | [source](article2013.Rmd) |
+[Article draft in .pdf](article2013.pdf)
 
@@
 
-- [HSE Moscow - November 12, 2013](moscow2013/lcsr20131112.html)
-- [Jatkokoulutusseminaari Turku November 7, 2013](turku2013/slidesTkuJatko20131107.html)
-- [ESA Torino - August 29, 2013](torino2013/TorinoSlides2013_final.html)
+| html-ouput | source code |
+| ------------ | -------------- |
+| [slides](attrib_slides.html) | [source](attrib_slides.Rpres) |
 
-# Misc links
+---------
 
-- [Article code in github](https://github.com/muuankarski/attributions)
-- [rustfare-package](http://markuskainu.fi/rustfare/) *- toolbox for R language for open welfare data from Russian Federation*
-- [rOpenGov](http://ropengov.github.io/) *-  R Ecosystem for Open Government Data and Computational Social Science*
-- [markuskainu.fi](http://markuskainu.fi) *- personal website*
-
------
+**Analytical code is licensed under open [FreeBSD](http://en.wikipedia.org/wiki/BSD_licenses#2-clause_license_.28.22Simplified_BSD_License.22_or_.22FreeBSD_License.22.29)-licence.**

### 2014-03-28 code/compile.R

@@ -3,7 +3,7 @@ setwd("~/workspace/lits/clone/attributions")
 pandoc_md <- function(name="index") {
     input <- paste(name,".md", sep="")
     output <- paste(name,".html", sep="")
-    system(paste("pandoc -s -S --toc -H /home/aurelius/workspace/web/css/rmarkdown.css",input,"-o",output))
+    system(paste("pandoc -s -S -H /home/aurelius/workspace/web/css/rmarkdown.css",input,"-o",output))
 }
 
 library(knitr)

### 2014-03-29 article2013.Rmd
Ranska puuttui kuvailevistagit add . Lisäsin ja viilasin kuviot ja analyysit..

@@
 affiliation: Finland
-date: March 27, 2014
+date: March 29, 2014
 header-includes:

### 2014-03-29 finalPlots.Rmd
Ranska puuttui kuvailevistagit add . Lisäsin ja viilasin kuviot ja analyysit..

@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=FALSE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
+opts_chunk$set(echo=FALSE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
 ```
@@
 #########################################
-setwd("~/workspace/lits/attributions")
+setwd("~/workspace/lits/clone/attributions")
 #########################################
@@
 #########################################
-load("~/workspace/lits/attributions/data/lits.RData")
+load("~/workspace/lits/clone/attributions/data/lits.RData")
 
@@
 map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
-                                layer="lits_all_simple"))
+                                layer="lits_all_simple_2014_crimea"))
 
@@
 
+mapplot_png <- ggplot(map.df, aes(long,lat,group=group)) +
+  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
+  scale_fill_manual(values=cbPalette) +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white", size=.3) +
+  geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
+  coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(35,65)) +
+  theme_minimal() + 
+  theme(legend.position="top") +
+  theme(legend.title=element_blank()) +
+  theme(axis.title.x = element_blank(),
+        axis.title.y = element_blank()) +
+  theme(axis.text.x = element_blank(),
+        axis.text.y = element_blank()) +
+  theme(axis.ticks.x = element_blank(),
+        axis.ticks.y = element_blank()) +
+  theme(legend.key.size = unit(3, "mm"))
+
 ppi <- 300
-png("figure/mapplot.png", width=15/2.54*ppi, height=9/2.54*ppi, res=ppi)
-mapplot
+png("figure/mapplot_png.png", width=15/2.54*ppi, height=10/2.54*ppi, res=ppi)
+mapplot_png
 dev.off()

### 2014-03-29 loadClean.Rmd
Ranska puuttui kuvailevistagit add . Lisäsin ja viilasin kuviot ja analyysit..

@@
 
-load("~/workspace/data/lits/lits1.RData")
+load("~/data/lits/lits1.RData")
 # FROM WEB
@@
 # FROM LOCAL DISK
-load("~/workspace/data/lits/lits2.RData")
+load("~/data/lits/lits2.RData")
 ```
@@
 
-summary(lits2$group_analysis)
+table(lits2$group_analysis,lits2$cntry)
 
@@
 
+library(devtools)
+install_github(repo = "rqog", username = "muuankarski")
+library(rqog)
+
+dat <- read_qog(which.data = "basic")
+# Subset the data
+dat2 <- dat[dat$cname %in% c('Armenia','Azerbaijan','Georgia','Russia',
+                   'Ukraine','Moldova','Belarus','Kyrgyzstan',
+                   'Kazakhstan','Tajikistan',
+                   'Uzbekistan','Estonia','Latvia','Lithuania',
+                   'Poland','Czech Republic','Slovakia','Hungary',
+                   'Slovenia','Romania','Bulgaria',
+                   'Italy','Germany','United Kingdom',
+                   'Sweden','France (1963-)'), ]
+dat2 <- dat2[,c("cname","year","wbgi_vae", # World Bank - Voice and Accountability
+                     "undp_hdi")]
+datQogSum <- dat2[dat2$year %in% c(2006,2010), ]
+datQogSum$cname <- as.character(datQogSum$cname)
+datQogSum$cname[datQogSum$cname == "France (1963-)"] <- "France"
+datQogSum$cname <- factor(datQogSum$cname)
+
 # datQog <- datQog[datQog$year >= 2000, ]
@@
 
-datQogSum.w <- dcast(datQogSum, 
-                   cname+year ~ indicator, 
-                       value.var="value")
+datQogSum.w <- datQogSum
 
@@
 
-save.image("~/workspace/lits/attributions/data/lits.RData")
+save.image("~/workspace/lits/clone/attributions/data/lits.RData")
 #load("~/workspace/lits/attributions/data/lits.RData")

### 2014-03-29 mapPlots.Rmd
Ranska puuttui kuvailevistagit add . Lisäsin ja viilasin kuviot ja analyysit..

@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=FALSE,eval=TRUE, cache=TRUE,message=FALSE,warning=FALSE,fig.width=15,fig.height=11)
+opts_chunk$set(echo=FALSE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=5)
+
 ```
@@
 ```{rmapPlotsLoadData}
+
 library(RColorBrewer)
@@
 library(rgdal)
-load("~/workspace/lits/attributions/data/lits.RData")
+library(rgeos)
+library(grid)
+load("~/workspace/lits/clone/attributions/data/lits.RData")
 map <- suppressWarnings(readOGR(dsn="/home/aurelius/data/shapefiles/lits/", 
-                                layer="lits_all_simple"))
+                                layer="lits_all_simple_2014_crimea"))
 
@@
 ```{rmapPlotsGroup1, fig.height=13, fig.width=16}
+library(grid)
 ggplot(map.df, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = group_analysis)) + # or group_fine !!
   scale_fill_manual(values=cbPalette) +
-  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_text(data=cnames, aes(long, lat, label = NAME, group=NAME), size=2, color="black") +
   coord_map(project="orthographic", xlim=c(-5,80),
-                ylim=c(40,60)) +
-  theme(legend.position="top")
+                ylim=c(40,62)) +
+  theme(legend.position="right")
 
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = socialBlame)) +
@@
                fill=NA, color = "white", size=0.1) +
-  coord_map(project="orthographic") +
-  scale_fill_gradient2(low="white", high=muted("blue")) +
-  theme(legend.position="top") +
-  theme(legend.text = element_text(size=14))
+      coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(37,65)) +
+  theme(legend.key.size = unit(10, "mm")) +
+  theme(legend.title = element_text(size=16)) +
+  theme(legend.position="right") +
+  theme(legend.text = element_text(size=16)) +
+  scale_fill_gradient(low = "white", high = "orange")
 
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = individualBlame)) +
@@
                fill=NA, color = "white", size=0.1) +
-  coord_map(project="orthographic") +
-  scale_fill_gradient2(low="white", high=muted("blue")) +
-  theme(legend.position="top") +
-  theme(legend.text = element_text(size=14))
+      coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(40,62)) +
+  theme(legend.key.size = unit(10, "mm")) +
+  theme(legend.title = element_text(size=16)) +
+  theme(legend.position="right") +
+  theme(legend.text = element_text(size=16)) +
+  scale_fill_gradient(low = "white", high = "orange")
 ```
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = socialFate)) +
@@
                fill=NA, color = "white", size=0.1) +
-  coord_map(project="orthographic") +
-  scale_fill_gradient2(low="white", high=muted("blue")) +
-  theme(legend.position="top") +
-  theme(legend.text = element_text(size=14))
+      coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(40,62)) +
+  theme(legend.key.size = unit(10, "mm")) +
+  theme(legend.title = element_text(size=16)) +
+
+  theme(legend.position="right") +
+  theme(legend.text = element_text(size=16)) +
+  scale_fill_gradient(low = "white", high = "orange")
 ```
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = individualFate)) +
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = notStated)) +
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = dontKnow)) +
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = gdpChange)) +
@@
                fill=NA, color = "white", size=0.1) +
-  coord_map(project="orthographic") +
-  scale_fill_gradient2(low=muted("red"), high=muted("blue")) +
-  theme(legend.position="top") +
-  theme(legend.text = element_text(size=14))
+      coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(40,62)) +
+  theme(legend.key.size = unit(10, "mm")) +
+  theme(legend.title = element_text(size=16)) +
+  scale_fill_gradient2(low=muted("red"), high=muted("green")) +
+  theme(legend.position="right") +
+  theme(legend.text = element_text(size=16))
 ```
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = undp_hdi)) +
@@
                fill=NA, color = "white", size=0.1) +
-  coord_map(project="orthographic") +
-  scale_fill_gradient2(low=muted("red"), high=muted("blue"), midpoint=0.7570) +
-  theme(legend.position="top") +
-  theme(legend.text = element_text(size=14))
+      coord_map(project="orthographic", xlim=c(-5,80),
+                ylim=c(40,62)) +
+  theme(legend.key.size = unit(10, "mm")) +
+  theme(legend.title = element_text(size=16)) +
+  scale_fill_gradient2(low=muted("red"), high=muted("green"), midpoint=0.7570) +
+  theme(legend.position="right") +
+  theme(legend.text = element_text(size=16))
 ```
@@
 ggplot(plotMap, aes(long,lat,group=group)) +
-  geom_polygon(data = map.df, aes(long,lat), fill="gray90", color = "white") +
+  geom_polygon(data = map.df, aes(long,lat), fill=NA, color = "white") +
   geom_polygon(aes(fill = wbgi_vae)) +

### 2014-03-29 code/dv_tables.R

@@ -1,4 +1,4 @@
-load("~/workspace/lits/attributions/data/lits.RData")
+load("~/workspace/lits/clone/attributions/data/lits.RData")
 library(survey)
 d.df <- svydesign(id = ~SerialID, 
                   weights = ~weight, 

### 2014-03-29 article2013.Rmd
malleja en jaksanut avata, tein pari uutta kuvaa ja tiivistin muita. katotaan mita ehtii 20 min tehda

@@
 ```{rattribArticleDescTbl1, results='asis', eval=TRUE, echo=FALSE, cache=TRUE, message=FALSE}
+
 source("code/dv_tables.R")
 library(xtable)
+
 print.xtable(xtable(tbl5,digits=1,

### 2014-03-29 mapPlots.Rmd
malleja en jaksanut avata, tein pari uutta kuvaa ja tiivistin muita. katotaan mita ehtii 20 min tehda

@@
   coord_map(project="orthographic", xlim=c(-5,80),
-                ylim=c(40,62)) +
+                ylim=c(37,65)) +
   theme(legend.position="right")
@@
       coord_map(project="orthographic", xlim=c(-5,80),
-                ylim=c(40,62)) +
+                ylim=c(37,65)) +
   theme(legend.key.size = unit(10, "mm")) +
@@
       coord_map(project="orthographic", xlim=c(-5,80),
-                ylim=c(40,62)) +
+                ylim=c(37,65)) +
   theme(legend.key.size = unit(10, "mm")) +
   theme(legend.title = element_text(size=16)) +
-
   theme(legend.position="right") +
@@
 
-```{rmapPlotsIvMacro1}
+```{rmapPlotsIvMacro1,fig.width=11,fig.height=3}
 ggplot(plotMap, aes(long,lat,group=group)) +
@@
       coord_map(project="orthographic", xlim=c(-5,80),
-                ylim=c(40,62)) +
+                ylim=c(40,55)) +
   theme(legend.key.size = unit(10, "mm")) +
@@
 
-```{rmapPlotsIvMacro2}
+```{rmapPlotsIvMacro2,fig.width=11,fig.height=3}
 ggplot(plotMap, aes(long,lat,group=group)) +
@@
 
-```{rmapPlotsIvMacro3}
+```{rmapPlotsIvMacro3,fig.width=11,fig.height=3}
 ggplot(plotMap, aes(long,lat,group=group)) +

### 2014-03-30 article2013.Rmd
last checks. korjasin postsos maiden lkm 21een 22sta

@@
 The study utilizes the data from the second wave of Life in Transition Survey[^***2] (LiTS2), conducted by European Bank of Development and Reconstruction (EBRD) and  World Bank in 2010.
-The analyses cover 22 post-socialist countries and 5 Western European countries.
+The analyses cover 21 post-socialist countries and 5 Western European countries.
 The survey was collected using a two-stage clustered stratified sampling procedure to select the households that were included in the sample (for more detailed description of sampling and data collection, see EBRD 2011).
@@
 
-We target our analyses at 22 post-socialist countries and use five Western European countries as references in descriptive analyses (see Figure 2).
-Our selection on 22 post-socialist countries includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania from Central Eastern Europe (CEE), and Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia from Commonwealth of Independent States (CIS)[^***].
+We target our analyses at 21 post-socialist countries and use five Western European countries as references in descriptive analyses (see Figure 2).
+Our selection on 21 post-socialist countries includes Czech Republic, Estonia, Hungary, Bulgaria, Latvia, Lithuania, Poland, Slovakia, Slovenia, Romania from Central Eastern Europe (CEE), and Armenia, Azerbaijan, Belarus, Georgia, Kazakhstan, Kyrgyzstan, Moldova, Tajikistan, Ukraine, Uzbekistan and Russia from Commonwealth of Independent States (CIS)[^***].
 Of Western European countries we have included France, Germany, Italy, United Kingdom and Sweden as references.

### 2014-03-30 article2013.Rmd
last checks. korjasin postsos maiden lkm 21een 22sta ja paper updated

@@
 affiliation: Finland
-date: March 29, 2014
+date: March 30, 2014
 header-includes:

### 2014-06-02 article2013.Rmd
moskovan korjauksia maalis-huhtkikuussa 2014. committaus ennen tutzing videosemman muutoksia kesakuun alussa

@@
 affiliation: Finland
-date: March 30, 2014
+date: April 1, 2014
 header-includes:
@@
     - \fancyfoot[LE,RO]{\thepage}
-abstract: Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in Europe and Western Asia with special emphasis on post-socialist countries.
+abstract: |
+  Utilising The Life in Transition Survey (LiTS) conducted in 2010, the purpose of the paper is to examine public perceptions of the causes of poverty in post-socialist countries of Europe and of Central and Western Asia.
+Citizens' attitudes towards the poor can be seen as reflecting the dominant ideology of country as well as the moral economy of the welfare state.
+Prior research has also shown that attributions for poverty have implications for the legitimacy of welfare system and for the viability of and support for anti-poverty policies.
+
   The study applies multilevel techniques in order to analyse the role of individual and country-level explanatory factors as determinants of poverty attributions.
-As country-level determinants, paper analyses the changes in country's economic performance during the period of global financial crises, income inequality and the state of democracy.
-As individual-level determinants, the paper tests to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.
+As country-level determinants, paper analyses the changes in country's economic performance during the period of global financial crises, human development and the state of democracy.
+As individual-level determinants, the paper tests so called self-interest hypothesis that is, to what extent risk factors related to financial hardship are associated with perceptions of the causes of poverty.
+
+  The results indicate that in general social blame type of explanation is clearly the most popular explanation in post-socialist countries.
+Social blame is followed by individual blame, social fate and individual fate types of explanations.
+In regard to differences between the post-socialist Europe and the West of Europe, results show that Western Europeans emphasize more social reasons for poverty.
+Results of the individual-level determinants of perceptions supported the self-interest hypothesis which states that those who have financial difficulties perceive the causes of poverty differently than those who are not exposed to poverty-related risks.
 keywords: poverty

### 2014-06-03 article2013.Rmd
tutzing videoluento purkissa..

@@
 affiliation: Finland
-date: April 1, 2014
+date: June 2, 2014
 header-includes:

### 2015-01-14 descriptive_analysis.Rmd
malleja en jaksanut avata, tein pari uutta kuvaa ja tiivistin muita. katotaan mita ehtii 20 min tehda
some polisihing for the analysis

@@
 ---
-title: Descriptive visualizations
+title: "Descriptive visualizations"
 output:
   html_document:
-    toc: true
+    number_sections: yes
     theme: flatly
+    toc: yes
 ---
@@
 ```{rdescplot1_1}
-load("~/workspace/lits/attributions/data/lits.RData")
+load("~/workspace/lits/clone/attributions/data/lits.RData")
 library(ggplot2)
@@
 
-df$cntry <- factor(df$cntry)
-tbl <- data.frame(prop.table(table(df$cntry,df$q227),1)*100)
-reg <- as.data.frame(table(df$cntry,df$group_general))
-reg <- subset(reg, Freq > 0)
-reg <- reg[with(reg, order(Var2)), ]
-x <- merge(tbl,reg[,1:2],by="Var1")
-names(x) <- c("cntry","cond","freq","group")
-
-x <- x[with(x, order(cond)), ]
-x$cond <- as.numeric(levels(x$cond))[x$cond]
-x <- subset(x, cond >= -1)
-x$cntry <- factor(x$cntry, levels=reg$Var1)
-
-ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
-  geom_bar(stat="identity") +
-  facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
+# 
+# 
+# df$cntry <- factor(df$cntry)
+# tbl <- data.frame(prop.table(table(df$cntry,df$q227),1)*100)
+# reg <- as.data.frame(table(df$cntry,df$group_general))
+# reg <- subset(reg, Freq > 0)
+# reg <- reg[with(reg, order(Var2)), ]
+# x <- merge(tbl,reg[,1:2],by="Var1")
+# names(x) <- c("cntry","cond","freq","group")
+# 
+# x <- x[with(x, order(cond)), ]
+# x$cond <- as.numeric(levels(x$cond))[x$cond]
+# x <- subset(x, cond >= -1)
+# x$cntry <- factor(x$cntry, levels=reg$Var1)
+# 
+# ggplot(x, aes(x=cntry, y=freq, fill=cond)) +
+#   geom_bar(stat="identity") +
+#   facet_grid(group~.) + theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
 ```
@@
 
-```{rdescplot2_1}
+```{rdescplot2_1, eval=FALSE}
 df$cntry <- factor(df$cntry)
@@
 names(x) <- c("cntry","cond","freq","group")
-x$cond <- factor(x$cond, levels=c("Dependent","Not.dependent"))
+x$cond <- factor(x$cond, levels=c("dependent","notdependent"))
 
@@
 
-```{rdescplot3_1}
+```{rdescplot3_1, eval=FALSE}
 df$cntry <- factor(df$cntry)
@@
 
-```{rdescplot4_1}
+```{rdescplot4_1, eval=FALSE}
 df$cntry <- factor(df$cntry)
@@
 
-```{rdescplot5_1}
+```{rdescplot5_1, eval=FALSE}
 df$cntry <- factor(df$cntry)
@@
 
-```{rdescplot6_1}
+```{rdescplot6_1, eval=FALSE}
 df$cntry <- factor(df$cntry)
@@
 
-```{rdescplot7}
-df <- macro[with(macro, order(uw_gini)), ]
-df$cntry <- factor(df$cntry, levels=df$cntry)
-ggplot(df, aes(x=cntry,y=uw_gini, colour=group)) +
+```{rdescplot7, eval=FALSE}
+df <- macro[with(macro, order(perGini)), ]
+#df$cntry <- factor(df$cntry, levels=df$cntry)
+ggplot(df, aes(x=cntry,y=perGini, colour=group_general)) +
   geom_point(size=5) +
@@
 ```{rdescplot8}
-df <- macro[with(macro, order(gdpchange07.10)), ]
-df$cntry <- factor(df$cntry, levels=df$cntry)
-ggplot(df, aes(x=cntry,y=gdpchange07.10, colour=group)) +
+df <- macro[with(macro, order(gdpChange)), ]
+#df$cntry <- factor(df$cntry, levels=df$cntry)
+ggplot(df, aes(x=cntry,y=gdpChange, colour=group_general)) +
   geom_point(size=5) +
@@
 df <- macro[with(macro, order(wbgi_vae)), ]
-df$cntry <- factor(df$cntry, levels=df$cntry)
+#df$cntry <- factor(df$cntry, levels=df$cntry)
 
-ggplot(df, aes(x=cntry,y=wbgi_vae, colour=group)) +
+ggplot(df, aes(x=cntry,y=wbgi_vae, colour=group_general)) +
   geom_point(size=5) +
@@
 library(GGally)
-ggpairs(macro[, c("uw_gini", "gdpchange07.10", "wbgi_vae","socialBlame","individualBlame")])
+ggpairs(macro[, c("perGini", "gdpChange", "wbgi_vae","socialBlame","individualBlame")])
 

### 2015-01-14 finalPlots.Rmd
some polisihing for the analysis

@@
-# Final graphics for the paper
+---
+title: "Final graphics for the paper"
+output:
+  html_document:
+    number_sections: yes
+    theme: flatly
+    toc: yes
+---
 
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=FALSE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
+library(knitr)
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=10)
 ```

### 2015-01-14 mapPlots.Rmd
some polisihing for the analysis

@@
 ---
-title: Spatial visualization
+title: "Spatial visualization"
 output:
   html_document:
-    toc: true
+    number_sections: yes
     theme: flatly
+    toc: yes
 ---
@@
 ```{roptsattrib, echo=FALSE}
-opts_chunk$set(echo=FALSE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=5)
+library(knitr)
+opts_chunk$set(echo=TRUE,eval=TRUE, cache=FALSE,message=FALSE,warning=FALSE,fig.width=12,fig.height=5)
 
